---
title: "Tuberculosis Prevalence among persons living with HIV in Low- and Middle-income Countries: A Systematic Review and Meta-analysis" 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

ABOUT THIS SCRIPT

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multivariate multilevel model on the studies that report sex stratified prevalence results. This model aims to estimate sex stratified prevalence adjusted for geographic factors. 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
library(tidybayes)
library(wesanderson)
library(ggplot2)
library(maps)


expit <- function (x) {exp(x)/(1 + exp(x))}

myPal <- c("black", "mediumturquoise")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))
##############################################################################|
##### All surveys that we extracted HIV data from
##############################################################################|
allHIV <- readRDS(here("data/fullDataClean.rds")) %>% 
          filter(report.hiv == "Yes")

##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING HIV status for 
##### QUANTITATIVE ANALYSIS                                  #################

cleanHivDF0 <- bactPostIndicator("hiv") %>% 
##### Manually move the two South African studies that report "prevalent TB" which 
##### meets our bacteriologically-confirmed TB threshold back into the analysis. 
               mutate(hiv.analysis.indicator = ifelse(covidence.id==26953, 
                                                      "prev100k.ci.prevalent.tb", 
                                                      ifelse(covidence.id == 26911, 
                                                             "n.prevalent.tb", 
                                                             hiv.analysis.indicator))) %>% 
               filter(hiv.analysis.indicator !="none",
##### Remove duplicate survey 
                      covidence.id != 26886)
### Calculate mean publication lag 
meanLag <- round(mean(cleanHivDF0$publication.year - cleanHivDF0$study.end.year, na.rm=TRUE))
```

### Figure of countries of all identified surveys

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 35}
nCountryAll <- allHIV %>% 
            count(study.country) %>% 
            mutate(n=as.factor(n))

world_map <- map_data("world")
world_map <- world_map[world_map$region != "Antarctica", ]

nCountryMap <- left_join(world_map, nCountryAll %>% rename(region=study.country), by="region")

ggplot(nCountryMap, aes(x=long, y=lat)) +
          geom_polygon(aes(group = region, subgroup = group), 
                       fill="lightgrey")  +
      geom_polygon(aes(group = region, subgroup = group,
                       fill = n), colour = "white")  +
    scale_fill_manual(values= c("lightblue", "dodgerblue3", "blue4"),
                      name = "Number of\nprevalence surveys",
                      na.translate = FALSE) +
    theme_void() + theme(legend.position = "inside",
                            legend.position.inside = c(.47,.35))


ggsave(here("output/allHivWorld.png"), plot=last_plot(), width = 15, height = 8.5, units = "in")

rm(allHIV)
```

```{r, echo = FALSE,  warning = FALSE, message=FALSE}
##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanHivDF <-  cleanHivDF0 %>% 
               mutate("id" = row_number()) %>% 
    ### Scale all prevalence estimates back to decimals 
    mutate_at(vars(matches('prev100k.bacteriological.tb.|prev100k.ci.([a-z]{3})er.bacteriological.tb.|prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.|prev100k.prevalent.tb.|prev100k.ci.([a-z]{3})er.prevalent.tb.')),  ~ (. / 1e5)) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           study.rurality, 
           WHO.region,
           screening.tests,
           hiv.status.definition,
           participants.sputum.collected,
           bacteriological.tb.definition,
           study.start.year, 
           study.end.year, 
           publication.year,
           study.quality.summary,
           n.participants.male, n.participants.female,
           starts_with("n.participants.age.grp"), 
           ends_with("hiv.positive"), ends_with("hiv.negative"), ends_with("hiv.total"),
           hiv.analysis.indicator) %>%
    
    mutate("tempPrevHivPos" =  case_when(hiv.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.positive,
                                       hiv.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive, 
                                        hiv.analysis.indicator == 
                                           "prev100k.ci.prevalent.tb" ~ prev100k.prevalent.tb.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.hiv.positive/n.participants.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.positive/n.participants.hiv.positive,
                                       hiv.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.hiv.positive/n.participants.hiv.positive,
                                       hiv.analysis.indicator == 
                                           "n.prevalent.tb" ~ n.prevalent.tb.hiv.positive/n.participants.hiv.positive,
                                        hiv.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive),
           "tempPrevHivNeg" = case_when(hiv.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.hiv.negative, 
                                        hiv.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.negative,
                                        hiv.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.hiv.negative, 
                                        hiv.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative, 
                                        hiv.analysis.indicator == 
                                           "prev100k.ci.prevalent.tb" ~ prev100k.prevalent.tb.hiv.negative, 
                                       hiv.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.hiv.negative/n.participants.hiv.negative, 
                                       hiv.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.negative/n.participants.hiv.negative,
                                       hiv.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.hiv.negative/n.participants.hiv.negative,
                                       hiv.analysis.indicator == 
                                           "n.prevalent.tb" ~ n.prevalent.tb.hiv.negative/n.participants.hiv.negative,
                                       hiv.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative)) %>%
    pivot_longer(cols = c(tempPrevHivNeg,tempPrevHivPos), 
                 names_to = "HIVStatus", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivPos" ~ (adj.prev100k.ci.upper.bacteriological.tb.hiv.positive-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivNeg" ~ (adj.prev100k.ci.upper.bacteriological.tb.hiv.negative-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivPos" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivNeg" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivPos" ~ (prev100k.ci.upper.bacteriological.tb.hiv.positive-
                                                                 prev100k.ci.lower.bacteriological.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivNeg" ~ (prev100k.ci.upper.bacteriological.tb.hiv.negative-
                                                                   prev100k.ci.lower.bacteriological.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivPos" ~ (prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivNeg" ~ (prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.prevalent.tb" & 
                                            HIVStatus == "HivPos" ~ (prev100k.ci.upper.prevalent.tb.hiv.positive-
                                                                 prev100k.ci.lower.prevalent.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.prevalent.tb" & 
                                            HIVStatus == "HivNeg" ~ (prev100k.ci.upper.prevalent.tb.hiv.negative-
                                                                   prev100k.ci.lower.prevalent.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", "n.prevalent.tb",
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & HIVStatus == "HivPos" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.positive),
                                        hiv.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", "n.prevalent.tb",
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & HIVStatus == "HivNeg" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.negative)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  HIVStatus == "HivPos" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.hiv.positive/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.hiv.positive)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.hiv.positive/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.hiv.positive))) /3.92,
                                              hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  HIVStatus == "HivNeg" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.hiv.negative/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.hiv.negative)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.hiv.negative /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.hiv.negative))) /3.92,
                                              hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  HIVStatus == "HivPos" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                              hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  HIVStatus == "HivNeg" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92,
                                              hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 HIVStatus == "HivPos" ~ (log(prev100k.ci.upper.bacteriological.tb.hiv.positive/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.hiv.positive)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.hiv.positive/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.hiv.positive))) /3.92,
                                                 hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 HIVStatus == "HivNeg" ~ (log(prev100k.ci.upper.bacteriological.tb.hiv.negative/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.hiv.negative)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.hiv.negative /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.hiv.negative))) /3.92,
                                                 hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     HIVStatus == "HivPos" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                                 hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     HIVStatus == "HivNeg" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92, 
                                                hiv.analysis.indicator == "prev100k.ci.prevalent.tb" &
                                                     HIVStatus == "HivPos" ~ (log(prev100k.ci.upper.prevalent.tb.hiv.positive/
                                                                          (1-prev100k.ci.upper.prevalent.tb.hiv.positive)) -
                                                                      log(prev100k.ci.lower.prevalent.tb.hiv.positive /
                                                                          (1-prev100k.ci.lower.prevalent.tb.hiv.positive)))/3.92,
                                                 hiv.analysis.indicator == "prev100k.ci.prevalent.tb" &
                                                     HIVStatus == "HivNeg" ~ (log(prev100k.ci.upper.prevalent.tb.hiv.negative/
                                                                            (1-prev100k.ci.upper.prevalent.tb.hiv.negative)) -
                                                                        log(prev100k.ci.lower.prevalent.tb.hiv.negative /
                                                                            (1-prev100k.ci.lower.prevalent.tb.hiv.negative)))/3.92, 
                                            hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", "n.prevalent.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & HIVStatus == "HivPos" ~ 
                                            (1/(n.participants.hiv.positive*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        hiv.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", "n.prevalent.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & HIVStatus == "HivNeg" ~ 
                                            (1/(n.participants.hiv.negative*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(is.na(study.geography), 
                                                       NA, 
                                                       ifelse(study.geography == "Nationally representative", 
                                                       "Yes", "No")),

                                       year_z = ifelse(!is.na(study.end.year), 
                                                       study.end.year - round(mean(study.end.year, na.rm=TRUE)),
                                                       (publication.year - meanLag) - round(mean((publication.year - meanLag), na.rm=TRUE))),                                       
                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       
                                       symptomScreen = ifelse(is.na(screening.tests), 
                                                                    NA, 
                                                                    ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No")),
                                       symptomsRequired = ifelse(is.na(participants.sputum.collected), 
                                                                       NA, 
                                                                       ifelse(!grepl("x-ray|all participants|unknown|other",
                                                                        participants.sputum.collected,ignore.case=TRUE),
                                                                 "Yes", "No")),
                                       XpertCultRequired = ifelse(is.na(bacteriological.tb.definition),
                                                                  NA, ifelse(!grepl("smear|micros|field|NTP", 
                                                                              bacteriological.tb.definition,
                                                                              ignore.case=TRUE), 
                                                                  "Yes", 
                                                                   "No")),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))

##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("hiv")

bactModelHivDF <- cleanHivDF %>%
    select(covidence.id, id, WHO.region, study.country, study.rurality,
           HIVStatus, year_z, hiv.status.definition, study.quality.summary,
           natRep, bias, symptomScreen, symptomsRequired, 
           XpertCultRequired, #`Adjusted Prevalence`, `Standard Error`,
           LogOdds, LogOddsStandardError, hiv.analysis.indicator, 
           n.participants.female, n.participants.male, 
           n.participants.hiv.positive, n.participants.hiv.negative, 
           starts_with("n.participants.age.grp"),
           publication.year, study.end.year) %>%
    pivot_wider(names_from = HIVStatus,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(logOR = LogOdds_HivPos - LogOdds_HivNeg,
           unadj_logOR_se = sqrt(LogOddsStandardError_HivPos^2 + LogOddsStandardError_HivNeg^2), 
           logOR_se = ifelse(grepl("adj", hiv.analysis.indicator), 
                                     unadj_logOR_se,
                                     ifelse(grepl("smear", hiv.analysis.indicator),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Bact Variance Median"])))) %>% filter(hiv.status.definition == "HIV test")

```

### Figure of countries used in quantitative analysis 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 35}
nCountry <- bactModelHivDF %>% 
            count(study.country) %>% 
            mutate(n=as.factor(n))

world_map <- map_data("world", region = c("Morocco", "Algeria", "Tunisia", "Libya",
             "Egypt", "Mauritania", "Mali", "Burkina Faso",
             "Niger", "Chad", "Sudan", "Ethiopia", "Eritrea",
             "Djibouti", "South Sudan", "Somalia", "Senegal",
             "Cape Verde", "Gambia", "Guinea", "Guinea-Bissau",
             "Liberia", "Sierra Leone", "Ivory Coast", "Ghana",
             "Togo", "Benin", "Nigeria", "Cameroon", 
             "Central African Republic", "Gabon", "Equatorial Guinea",
             "Sao Tome and Principe", "Uganda", "Republic of Congo", 
             "Democratic Republic of the Congo", "Rwanda", "Burundi",
             "Kenya", "Tanzania", "Angola", "Mozambique", "Malawi", 
             "Mauritius", "Seychelles", "Comoros", 
             "Botswana", "Zimbabwe", "Zambia", "Namibia", "Swaziland", 
             "South Africa", "Lesotho", "Western Sahara"))

nCountryMap <- left_join(world_map, nCountry %>% rename(region=study.country), by="region")

ggplot(nCountryMap, aes(x=long, y=lat)) +
          geom_polygon(aes(group = region, subgroup = group), 
                       fill="lightgrey")  +
      geom_polygon(aes(group = region, subgroup = group, 
                       fill = n), colour = "white")  +
  coord_map(projection = "mercator",
            xlim = c(-26,64), ylim = c(-35,38)) +
    # # scale_fill_gradient(low = "lightblue", high = "dodgerblue4", 
    # #                     name = "Number of\nprevalence surveys") +
    scale_fill_manual(values= c("lightblue", "dodgerblue3"), 
                      name = "Number of\nprevalence surveys",
                      na.translate = FALSE) +
    theme_void() + theme(legend.position = "none",
                            legend.position.inside = c(.3,.4))


ggsave(here("output/hivWorld.png"), plot=last_plot(), width = 15, height = 8.5, units = "in")

```

# Model 1: Main effect model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|study.country) + (1|id),
                       data = bactModelHivDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectMod)
plot(mainEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- bactModelHivDF %>%
      select(study.country, LogOdds_HivNeg) %>%
      mutate(logOR_se=1)

mainEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=nd,
                                 re_formula = ~ 1, 
                                 allow_new_levels = TRUE) %>% 
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectModEst %>% ungroup %>% mean_qi(hivRatio)
```

Survey-level effects

```{r, cache=TRUE}
surveyEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=bactModelHivDF,
                                 re_formula = ~ 1  + (1|study.country) + (1|id),
                                 allow_new_levels = TRUE) %>%
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

surveyEffectModEst %>% mean_qi(hivRatio)

nd_country <- bactModelHivDF %>% select(study.country) %>% unique() %>% 
              mutate(logOR_se = 1)

countryEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=nd_country,
                                 re_formula = ~ 1  + (1|study.country),
                                 allow_new_levels = TRUE) %>%
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

countryEffectModEst %>% mean_qi(hivRatio)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
### Get the empirical values 
empirical_prev <- bactModelHivDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        bias = bias,
        symptomScreen = symptomScreen, 
        p_hivneg_empirical = plogis(LogOdds_HivNeg),
        p_hivpos_empirical = plogis(LogOdds_HivPos),
        or_empirical = p_hivpos_empirical / p_hivneg_empirical,
        or_empirical = (p_hivpos_empirical / (1 - p_hivpos_empirical)) / 
            (p_hivneg_empirical / (1 - p_hivneg_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific TB prevalence")
    
```

Bring estimates from study model together to make a forest plot for the manuscript.

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.width = 20, fig.height = 28}

# Join survey labels
names <- cleanHivDF0 %>%
    mutate(id = row_number()) %>%
    select(id, figure.id.yr) %>%
    mutate(figure.id.yr = recode(figure.id.yr, 
                   "South Africa (Western Cape Province, c) 2005" = 
                   "South Africa\n(Western Cape Province) 2005",
                   "South Africa (Western Cape Province) 2008" = 
                   "South Africa\n(Western Cape Province) 2008", 
                   "Malawi (Blantyre) 2019-2020"=
                   "Malawi\n(Blantyre) 2019-2020", 
                   "Zimbabwe (Harare) 2005-2006" =
                   "Zimbabwe\n(Harare) 2005-2006", 
                   "Lesotho (National) 2019" =
                    "Lesotho\n(National) 2019", 
                   "Zambia (Lusaka Province) 2005"=
                       "Zambia\n(Lusaka Province) 2005", 
                   "Zambia (National)  2013-2014"=
                       "Zambia\n(National)  2013-2014"
                   ))

# Survey-level estimates
survey_sum <- surveyEffectModEst %>%
    group_by(id) %>%
    mean_qi(hivRatio) %>%
    mutate(grouping = "survey") %>%
    left_join(names, by = "id") %>%
    left_join(bactModelHivDF %>% select(id, WHO.region), by = "id")

# Global summary
global_sum <- mainEffectModEst %>%
    ungroup() %>%
    mean_qi(hivRatio) %>%
    mutate(grouping = "Overall",  
           figure.id.yr = "Overall")

# Combine all
forestDF <- bind_rows(survey_sum, global_sum)

# Fill display label and facet group
forestDF <- forestDF %>%
    mutate(
        display_label_raw = ifelse(is.na(figure.id.yr), grouping, figure.id.yr),
        facet_group = case_when(
            grouping == "survey" ~ WHO.region,
            # grouping == "regional" ~ figure.id.yr,
            grouping == "Overall" ~ "Overall"
        ), 
        type="Posterior mean and 95% credible interval"
    )

# Define facet order
group_order <- c(
    # "Africa region", 
    # "Americas region",
    # "Eastern Mediterranean region", 
    # "South-East Asia region", 
    # "Western Pacific region", 
    "Overall"
)

# Separate surveys and summaries
surveys_only <- forestDF %>%
    filter(grouping == "survey") %>%
    # group_by(facet_group) %>%
    arrange(desc(hivRatio), .by_group = TRUE) %>%
    ungroup()

summaries_only <- forestDF %>%
    filter(grouping != "survey")

# Recombine with surveys first, then summaries (preserves facet-wise order)
forestDF_plot <- bind_rows(surveys_only, summaries_only) %>%
    # mutate(facet_group = factor(facet_group, levels = group_order)) %>%
    mutate(display_label = case_when(
        grouping == "Overall" ~ "Overall",
        # grouping == "regional" ~ paste0("<b>", display_label_raw, "</b>"),
        TRUE ~ display_label_raw
    )) %>%
    mutate(display_label = factor(display_label, levels = rev(unique(display_label))))

# Prepare empirical data
empirical_prev_labels <- left_join(empirical_prev, survey_sum) %>%
    rename(display_label_raw = figure.id.yr) %>%
    select(display_label_raw, or_empirical) %>%
    mutate(
        grouping = "survey",
        display_label = display_label_raw
    ) %>%
    mutate(
           type = "Ratios calculated from reported prevalence")

# Plot
forestRatio <- ggplot(forestDF_plot) +
    geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
        geom_point(data = empirical_prev_labels,
           aes(x = display_label, y = or_empirical, color=type), alpha=0.8, size=8) +
    geom_pointinterval(aes(x = display_label, y = hivRatio, ymin = .lower, ymax = .upper, fill=type),
                       size = 15) +
    coord_flip() +
    scale_color_manual(values = myPal[c(2,1)]) +
    scale_y_log10(
        breaks = c(0.25, 0.5, 1, 2, 4, 8, 16),
        labels = c("0.25", "0.5", "1", "2", "4", "8", "16"),
    ) +
    # scale_x_discrete(labels = label_wrap(30)) + 
    labs(
        x = NULL,
        y = "HIV positive-to-HIV negative ratio\nof bacteriologically-confirmed TB prevalence"
    ) +
    theme_minimal(base_size = 18) +
    expand_limits(y=c(0,16)) +
    theme(
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        # panel.spacing.y = unit(1, "lines"),
        # axis.text.y = ggtext::element_markdown(hjust = 1),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.background = element_rect(fill = "white"),
        legend.position = "bottom", legend.title=element_blank(),
        legend.spacing.y = unit(-.25, 'cm'),
        legend.box.margin=margin(-10,-10,-6,-10),
        panel.spacing = unit(0.18, "cm"),
        plot.margin = unit(c(.1, .7, 0, .1), "cm")
    ) ; forestRatio

# Save
ggsave(plot = forestRatio, "output/hivFigure2.png", height = 8, width=12, units = "in")
```

### Covariate analysis 

# Model 2: Covariate analysis
```{r, warning = FALSE, message=FALSE, cache=TRUE}
artCoverage <- read.csv(here("data/hivAnalysis/unaids-2025-adult-art-coverage.csv")) %>% rename(study.country=spectrum_country, 
           study.end.year = year)
bactModelHivCovDF <- left_join(bactModelHivDF, artCoverage %>% 
                                   select(study.country,
                                          study.end.year, 
                                          art_coverage)) %>%
                     mutate(art_coverage = art_coverage * 100)

```

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
artEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                           1 + art_coverage + (1|study.country) + (1|id),
                       data = bactModelHivCovDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check

```{r, warning = FALSE, message=FALSE, cache=TRUE}
summary(artEffectMod)
plot(artEffectMod)
```

### Prevalence to notification comparison 

```{r}
### Population with HIV data from UNAIDS
pop <- artCoverage

### Read in TB notification data from WHO
notif0 <- read.csv(here("data/TB_notifications_2025-07-18.csv"))

### Calculate the HIV-status stratified TB notification 
### rate and transform to log odds space
notif0 <- notif0 %>% 
          ## Add 2007 because 2006 data for Zimbabwe does not exist
          filter(year %in% c(unique(bactModelHivDF$study.end.year),2007),  
                country %in% unique(bactModelHivDF$study.country)) %>% 
          select(country, g_whoregion, 
                year, 
                newrel_m15plus, 
                newrel_f15plus,
                newrel_hivpos_014,
                new_sp, new_sn, 
                newrel_hivpos, 
                hivtest_pos, 
                hivtest) %>%
         rename(study.country = country, 
                study.end.year = year) %>%
         left_join(pop, by=c("study.country", "study.end.year")) %>%
         mutate(notifRateHivPos = ifelse(study.end.year < 2019, 
                                     hivtest_pos / hivpop, 
                                     ifelse(is.na(newrel_hivpos_014), 
                                            newrel_hivpos / hivpop,
                                           (newrel_hivpos - newrel_hivpos_014)/hivpop)), 
                notifRateHivNeg = ifelse(study.end.year == 2014, 
                                     (newrel_m15plus + newrel_f15plus - hivtest_pos) / (totpop - hivpop), 
                                     ifelse(study.end.year < 2014, 
                                     (new_sp + new_sn - hivtest_pos) / (totpop-hivpop), 
                                     ifelse(is.na(newrel_hivpos_014), 
                                            (newrel_m15plus + newrel_f15plus - newrel_hivpos) / (totpop-hivpop), 
                                            (newrel_m15plus + newrel_f15plus - newrel_hivpos - newrel_hivpos_014) / (totpop-hivpop))))) %>%
         pivot_longer(cols=starts_with("notifRate"),
                      names_to = "HIVStatus",
                      names_prefix = "notifRate",
                      values_to = "notifRate",
                      values_drop_na = TRUE) %>%
        select(study.country, study.end.year, `HIVStatus`, notifRate, totpop, hivpop) %>%
        mutate(notifRate = ifelse(notifRate==0, 
                                  NA, notifRate), 
               "Standard Error" = case_when(HIVStatus == "HivPos" ~ 
                                                sqrt((notifRate*(1-notifRate))/hivpop), 
                                            HIVStatus == "HivNeg" ~ 
                                                sqrt((notifRate*(1-notifRate))/(totpop-hivpop))),
               "LogOdds_notif" = log(notifRate/(1-notifRate)),
               "LogOddsStandardError_notif" = case_when(HIVStatus == "HivPos" ~ 
                                                  (1/(hivpop*notifRate*(1-notifRate)))^2,
                                                        HIVStatus == "HivNeg" ~ 
                                                  (1/((totpop-hivpop)*notifRate*(1-notifRate)))^2))
       
### Widen and calculate the log odds ratio and associated standard error
notif  <- notif0 %>%
          select(study.country, study.end.year, HIVStatus, LogOdds_notif, LogOddsStandardError_notif) %>%
          pivot_wider(names_from = HIVStatus,
                     values_from = c(LogOdds_notif, LogOddsStandardError_notif),
                     names_sep = "_") %>%
          mutate(logOR_notif = LogOdds_notif_HivPos - LogOdds_notif_HivNeg,
                 logOR_se_notif = sqrt(LogOddsStandardError_notif_HivPos^2 + LogOddsStandardError_notif_HivNeg^2)) %>%
          right_join(bactModelHivDF %>% 
                     mutate(study.end.year = recode(study.end.year, `2006` = 2007)), by=c("study.country", "study.end.year")) %>% 
        mutate(study.end.year = recode(study.end.year, `2007` = 2006))

###############################################################################
### Run a multivariate model
#prevalence model:
prev_model <- bf(
  logOR | se(logOR_se, sigma = TRUE) ~ 1 + (1 |p| id) + (1 | study.country)
)
 
#notifications model:
notif_model <- bf(
  logOR_notif | se(logOR_se_notif, sigma = TRUE) ~ 1 + (1 |p| id) + (1 | study.country)
)
 
notif_model_res <- brm(
  formula = notif_model,
  data = notif,
  family = gaussian(), 
  # sample_prior = "only",
  prior = c(
    prior(student_t(7, 0, 1.5), class = Intercept),
    prior(exponential(2), class = sd)
  ),
  control = list(adapt_delta = 0.99),
  chains = 4,
  cores = 4,
  iter = 4000
)

summary(notif_model_res)
plot(notif_model_res)
```

```{r}
#fit the model
p_to_n_hiv_model <- brm(
  formula = prev_model + notif_model + set_rescor(FALSE),
  data = notif,
  family = gaussian(), 
  # sample_prior = "only",
  prior = c(
    prior(student_t(7, 0, 1.5), class = Intercept, resp = "logOR"),
    prior(student_t(7, 0, 1.5), class = Intercept, resp = "logORnotif"),
    prior(exponential(2), class = sd, resp = "logOR"),
    prior(exponential(2), class = sd, resp = "logORnotif")
  ),
  control = list(adapt_delta = 0.99),
  chains = 4,
  cores = 4,
  iter = 4000
)

summary(p_to_n_hiv_model)

nd <- notif %>% select(study.country) %>%
        unique() %>%
        mutate(logOR_se_notif=1, logOR_se =1)

mainEffectPNEst <- add_epred_draws(object = p_to_n_hiv_model,
                                      newdata=nd,
                                      re_formula =   ~ 1 + (1 | study.country),
                                 allow_new_levels = TRUE) %>%
                    mutate(#ratio = exp(.epred),
                           type="Posterior mean and 95% credible interval")


mainEffectPNEstSummary <- mainEffectPNEst %>%
                             group_by(study.country, .category) %>%
                             mean_qi(exp(.epred), na.rm=T); mainEffectPNEstSummary

mainEffectPNEstSummary <- mainEffectPNEst %>%
                             group_by(.category) %>%
                             mean_qi(exp(.epred), na.rm=T); mainEffectPNEstSummary
```

```{r}
mainEffectPNEst %>% pivot_wider(names_from = .category,
                     values_from = c(.epred),
                     names_sep = "_") %>%
                     group_by(study.country) %>%
                     mean_qi(expit(logOR-logORnotif), na.rm=T)

mainEffectPNEst %>% pivot_wider(names_from = .category,
                     values_from = c(.epred),
                     names_sep = "_") %>% 
                     ungroup() %>%
                     mean_qi(expit(logOR-logORnotif), na.rm=T)
```

# Smear positive analysis 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
cleanHivSmrDF0 <- smrPosIndicator("hiv") %>% filter(hiv.smr.analysis.indicator !="none")

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanHivSmrDF <- cleanHivSmrDF0 %>% 
                 filter(hiv.smr.analysis.indicator != "none") %>% 
                 mutate("id" = row_number()) %>%
                 ### Scale all prevalence estimates back to decimals 
                 mutate_at(vars(matches('prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>% 
                 select(id,
                       covidence.id, 
                       figure.id.yr, 
                       study.geography, 
                       title.extracted,
                       study.country,
                       WHO.region,
                       study.start.year, 
                       study.end.year, 
                       ends_with("hiv.positive"), ends_with("hiv.negative"), ends_with("hiv.total"),
                       hiv.smr.analysis.indicator, 
                       hiv.status.definition) %>%
    
    mutate("tempPrevHivPos" =  case_when(hiv.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.positive,
                                       hiv.smr.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive, 
                                       hiv.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.positive/n.participants.hiv.positive,
                                        hiv.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive),
           "tempPrevHivNeg" = case_when(hiv.smr.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.negative,
                                        hiv.smr.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative, 
                                       hiv.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.negative/n.participants.hiv.negative,
                                       hiv.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative)) %>%
    pivot_longer(cols = c(tempPrevHivNeg,tempPrevHivPos), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivPos" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivNeg" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                       
                                        hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivPos" ~ (prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivNeg" ~ (prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                        hiv.smr.analysis.indicator %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "HivPos" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.positive),
                                        hiv.smr.analysis.indicator  %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "HivNeg" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.negative)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "HivPos" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                              hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "HivNeg" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92,
                                                 hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "HivPos" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                                 hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "HivNeg" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92, 
                                            hiv.smr.analysis.indicator %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "HivPos" ~ 
                                            log(1/(n.participants.hiv.positive*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        hiv.smr.analysis.indicator  %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "HivNeg" ~ 
                                            log(1/(n.participants.hiv.negative*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       # year_z = ifelse(!is.na(study.end.year), 
                                       #                 study.end.year - round(mean(study.end.year, na.rm=TRUE))
                                       #                 (publication.year - meanLag) - round(mean((publication.year - meanLag), na.rm=TRUE))),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))  %>% filter(LogOddsStandardError < Inf)

smrModelDF <- cleanHivSmrDF %>%
    select(covidence.id, id, WHO.region, study.country, Sex, natRep, LogOdds, LogOddsStandardError, 
           n.participants.hiv.positive, n.participants.hiv.negative, hiv.status.definition, title.extracted) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_HivPos - LogOdds_HivNeg,
        logOR_se = sqrt(LogOddsStandardError_HivPos^2 + LogOddsStandardError_HivNeg^2) 
    ) %>% filter( hiv.status.definition == "HIV test")
```

# Smear positive descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(smrModelDF)[1], "unique surveys with sex-stratified smear positive TB"))

# sort(unique(cleanHivDF0$title.extracted))
print(paste(length(unique(smrModelDF$title.extracted)), "unique publications"))

print(paste(sum(!is.na(smrModelDF$n.participants.hiv.negative)), "studies report participant numbers totaling", 
            sum(smrModelDF$n.participants.hiv.negative, smrModelDF$n.participants.hiv.positive, na.rm = TRUE), "participants"))

print(paste0(round(sum(smrModelDF$n.participants.hiv.positive, na.rm=TRUE)/sum(smrModelDF$n.participants.hiv.negative, smrModelDF$n.participants.hiv.positive, na.rm = TRUE)*100, 2), "% of participants were hivpos" ))
```
More descriptives 

```{r}
table(smrModelDF$WHO.region)
smrModelDF %>% reframe(Participants = sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE), .by=WHO.region)
smrModelDF %>% reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=WHO.region)
```

# Model X: Smear positive main effect 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | study.country) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectSmrMod)
plot(mainEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- smrModelDF %>%
      select(study.country, LogOdds_HivNeg) %>%
      mutate(logOR_se=1)


mainEffectSmrModEst <- add_epred_draws(object = mainEffectSmrMod,
                                 newdata=nd,
                                 re_formula = ~ 1) %>% 
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectSmrModEst %>% ungroup %>% mean_qi(hivRatio)
```

Survey specific estimate
```{r}
surveyEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=bactModelHivDF,
                                 re_formula = ~ 1  + (1|study.country) + (1|id),
                                 allow_new_levels = TRUE) %>%
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

surveyEffectModEst %>% mean_qi(hivRatio) %>% 
    select(id, study.country, hivRatio, .lower, .upper)
```


#### Comparison of bacteriologically confirmed estimates from studies above

```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectBCSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | study.country) + (1|id),
                       data = bactModelHivDF %>% 
                              filter(covidence.id %in% smrModelDF$covidence.id),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectBCSmrMod)
plot(mainEffectBCSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- smrModelDF %>% 
      filter(covidence.id %in% smrModelDF$covidence.id) %>%
      select(study.country, LogOdds_HivNeg) %>%
      mutate(logOR_se=1) 


mainEffectBCSmrModEst <- add_epred_draws(object = mainEffectBCSmrMod,
                                 newdata=nd,
                                 re_formula = ~ 1) %>% 
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectBCSmrModEst %>% ungroup %>% mean_qi(hivRatio)
```
