---
title: "Tuberculosis Prevalence among persons living with HIV in Low- and Middle-income Countries: A Systematic Review and Meta-analysis" 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

ABOUT THIS SCRIPT

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multivariate multilevel model on the studies that report sex stratified prevalence results. This model aims to estimate sex stratified prevalence adjusted for geographic factors. 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
# library(posterior)
library(tidybayes)
library(wesanderson)
library(ggplot2)

expit <- function (x) {exp(x)/(1 + exp(x))}

myPal <- c("black", "mediumturquoise")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))

##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################


cleanHivDF0 <- bactPostIndicator("hiv") %>% filter(hiv.analysis.indicator !="none")
### Calculate mean publication lag 
meanLag <- round(mean(cleanHivDF0$publication.year - cleanHivDF0$study.end.year, na.rm=TRUE))
```

# Descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(cleanHivDF0)[1], "unique surveys with HIV-stratified bacteriological positive TB"))

print(paste(length(unique(cleanHivDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanHivDF0$n.participants.hiv.negative)), "studies report participant numbers totaling", 
            sum(cleanHivDF0$n.participants.hiv.negative, cleanHivDF0$n.participants.hiv.positive, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanHivDF0$n.participants.hiv.positive, na.rm=TRUE)/sum(cleanHivDF0$n.participants.hiv.positive, cleanHivDF0$n.participants.hiv.negative, na.rm = TRUE)*100, 2), "% of participants were HIV positive" ))
```

# Risk of bias statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE}
table(cleanHivDF0$study.quality.summary)
```

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanHivDF <-  cleanHivDF0 %>% 
               filter(hiv.analysis.indicator != "none") %>% 
               mutate("id" = row_number()) %>% 
    ### Scale all prevalence estimates back to decimals 
    mutate_at(vars(matches('prev100k.bacteriological.tb.|prev100k.ci.([a-z]{3})er.bacteriological.tb.|prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           study.rurality, 
           WHO.region,
           screening.tests,
           participants.sputum.collected,
           bacteriological.tb.definition,
           study.start.year, 
           study.end.year, 
           publication.year,
           study.quality.summary,
           ends_with("hiv.positive"), ends_with("hiv.negative"), ends_with("hiv.total"),
           hiv.analysis.indicator) %>%
    
    mutate("tempPrevHivPos" =  case_when(hiv.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.positive,
                                       hiv.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.hiv.positive/n.participants.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.positive/n.participants.hiv.positive,
                                       hiv.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.hiv.positive/n.participants.hiv.positive,
                                        hiv.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive),
           "tempPrevHivNeg" = case_when(hiv.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.hiv.negative, 
                                        hiv.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.negative,
                                        hiv.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.hiv.negative, 
                                        hiv.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative, 
                                       hiv.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.hiv.negative/n.participants.hiv.negative, 
                                       hiv.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.negative/n.participants.hiv.negative,
                                       hiv.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.hiv.negative/n.participants.hiv.negative,
                                       hiv.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative)) %>%
    pivot_longer(cols = c(tempPrevHivNeg,tempPrevHivPos), 
                 names_to = "HIVStatus", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivPos" ~ (adj.prev100k.ci.upper.bacteriological.tb.hiv.positive-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivNeg" ~ (adj.prev100k.ci.upper.bacteriological.tb.hiv.negative-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivPos" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivNeg" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivPos" ~ (prev100k.ci.upper.bacteriological.tb.hiv.positive-
                                                                 prev100k.ci.lower.bacteriological.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivNeg" ~ (prev100k.ci.upper.bacteriological.tb.hiv.negative-
                                                                   prev100k.ci.lower.bacteriological.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivPos" ~ (prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivNeg" ~ (prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & HIVStatus == "HivPos" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.positive),
                                        hiv.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & HIVStatus == "HivNeg" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.negative)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  HIVStatus == "HivPos" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.hiv.positive/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.hiv.positive)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.hiv.positive/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.hiv.positive))) /3.92,
                                              hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  HIVStatus == "HivNeg" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.hiv.negative/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.hiv.negative)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.hiv.negative /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.hiv.negative))) /3.92,
                                              hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  HIVStatus == "HivPos" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                              hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  HIVStatus == "HivNeg" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92,
                                              hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 HIVStatus == "HivPos" ~ (log(prev100k.ci.upper.bacteriological.tb.hiv.positive/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.hiv.positive)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.hiv.positive/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.hiv.positive))) /3.92,
                                                 hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 HIVStatus == "HivNeg" ~ (log(prev100k.ci.upper.bacteriological.tb.hiv.negative/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.hiv.negative)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.hiv.negative /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.hiv.negative))) /3.92,
                                                 hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     HIVStatus == "HivPos" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                                 hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     HIVStatus == "HivNeg" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92, 
                                            hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & HIVStatus == "HivPos" ~ 
                                            log(1/(n.participants.hiv.positive*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        hiv.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & HIVStatus == "HivNeg" ~ 
                                            log(1/(n.participants.hiv.negative*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(is.na(study.geography), 
                                                       NA, 
                                                       ifelse(study.geography == "Nationally representative", 
                                                       "Yes", "No")),

                                       year_z = ifelse(!is.na(study.end.year), 
                                                       study.end.year - round(mean(study.end.year, na.rm=TRUE)),
                                                       (publication.year - meanLag) - round(mean((publication.year - meanLag), na.rm=TRUE))),                                       
                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       
                                       symptomScreen = ifelse(is.na(screening.tests), 
                                                                    NA, 
                                                                    ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No")),
                                       symptomsRequired = ifelse(is.na(participants.sputum.collected), 
                                                                       NA, 
                                                                       ifelse(!grepl("x-ray|all participants|unknown|other",
                                                                        participants.sputum.collected,ignore.case=TRUE),
                                                                 "Yes", "No")),
                                       XpertCultRequired = ifelse(is.na(bacteriological.tb.definition),
                                                                  NA, ifelse(!grepl("smear|micros|field|NTP", 
                                                                              bacteriological.tb.definition,
                                                                              ignore.case=TRUE), 
                                                                  "Yes", 
                                                                   "No")),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))

##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("hiv")

bactModelDF <- cleanHivDF %>%
    select(id, WHO.region, study.country, study.rurality,
           HIVStatus, year_z, 
           natRep, bias, symptomScreen, symptomsRequired, 
           XpertCultRequired, 
           LogOdds, LogOddsStandardError, hiv.analysis.indicator, 
           n.participants.hiv.positive, n.participants.hiv.negative, 
           publication.year, study.end.year) %>%
    pivot_wider(names_from = HIVStatus,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(logOR = LogOdds_HivPos - LogOdds_HivNeg,
           unadj_logOR_se = sqrt(LogOddsStandardError_HivPos^2 + LogOddsStandardError_HivNeg^2), 
           logOR_se = ifelse(grepl("adj", hiv.analysis.indicator), 
                                     unadj_logOR_se,
                                     ifelse(grepl("smear", hiv.analysis.indicator),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Bact Variance Median"])))) %>% filter(id !=4)

```
More descriptives 

```{r}
table(bactModelDF$WHO.region)
bactModelDF %>% reframe(Participants = sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE), .by=WHO.region)
bactModelDF %>% reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=WHO.region)
table(bactModelDF$natRep)
bactModelDF %>% reframe(Participants = sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE), .by=natRep)
bactModelDF %>% reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=natRep)
table(bactModelDF$bias)
bactModelDF %>% reframe(Participants = sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE), .by=bias)
bactModelDF %>% reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=bias)
table(bactModelDF$symptomScreen)
bactModelDF %>% reframe(Participants = sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE), .by=symptomScreen)
bactModelDF %>% reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=symptomScreen)

bactModelDF %>% reframe(Participants = sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE), .by=symptomsRequired)
bactModelDF %>% reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=symptomsRequired)

bactModelDF %>% reframe(Participants = sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE), .by=XpertCultRequired)
bactModelDF %>% reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=XpertCultRequired)

bactModelDF %>% reframe(Participants = sum(n.participants.hiv.positive + n.participants.hiv.negative, na.rm=TRUE), .by=study.rurality)
bactModelDF %>% mutate(n.participants.hiv.positive = ifelse(study.rurality == "Rural only", 
                                                     n.participants.hiv.positive + 10407, 
                                                    n.participants.hiv.positive), 
                       n.participants.hiv.negative = ifelse(study.rurality == "Rural only", 
                                                     n.participants.hiv.negative + 10443,
                                                     n.participants.hiv.negative)) %>% 
                       
                                                reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=study.rurality)
```

# Model 1: Main effect model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|study.country) + (1|id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectMod)
plot(mainEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- bactModelDF %>%
      select(study.country, LogOdds_HivNeg) %>%
      mutate(logOR_se=1)

mainEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=nd,
                                 re_formula = ~ 1, 
                                 allow_new_levels = TRUE) %>% 
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectModEst %>% ungroup %>% mean_qi(hivRatio)

### output for contribution table
exp(as.numeric(summary(mainEffectMod)[["fixed"]][1]))
exp(as.numeric(summary(mainEffectMod)[["fixed"]][3]))
exp(as.numeric(summary(mainEffectMod)[["fixed"]][4]))
```

Survey-level effects

```{r}
surveyEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=bactModelDF,
                                 re_formula = ~ 1  + (1|study.country) + (1|id),
                                 allow_new_levels = TRUE) %>%
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

surveyEffectModEst %>% mean_qi(hivRatio)

# saveRDS(surveyEffectModEst, file="output/surveyModelEstimates.rds", version=2)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
### Get the empirical values 
empirical_prev <- bactModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        bias = bias,
        symptomScreen = symptomScreen, 
        p_hivneg_empirical = plogis(LogOdds_HivNeg),
        p_hivpos_empirical = plogis(LogOdds_HivPos),
        or_empirical = p_hivpos_empirical / p_hivneg_empirical,
        or_empirical = (p_hivpos_empirical / (1 - p_hivpos_empirical)) / 
            (p_hivneg_empirical / (1 - p_hivneg_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific TB prevalence")
    
```

Bring estimates from regional model together to make a forest plot for the manuscript.

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.width = 20, fig.height = 28}

# Join survey labels
names <- cleanHivDF0 %>%
    mutate(id = row_number()) %>%
    select(id, figure.id.yr)

# Survey-level estimates
survey_sum <- surveyEffectModEst %>%
    group_by(id) %>%
    mean_qi(hivRatio) %>%
    mutate(grouping = "survey") %>%
    left_join(names, by = "id") %>%
    left_join(bactModelDF %>% select(id, WHO.region), by = "id")

# Global summary
global_sum <- mainEffectModEst %>%
    ungroup() %>%
    mean_qi(hivRatio) %>%
    mutate(grouping = "Overall",  
           figure.id.yr = "Overall")

# Combine all
forestDF <- bind_rows(survey_sum, global_sum)

# Fill display label and facet group
forestDF <- forestDF %>%
    mutate(
        display_label_raw = ifelse(is.na(figure.id.yr), grouping, figure.id.yr),
        facet_group = case_when(
            grouping == "survey" ~ WHO.region,
            # grouping == "regional" ~ figure.id.yr,
            grouping == "Overall" ~ "Overall"
        ), 
        type="Posterior mean and 95% credible interval"
    )

# Define facet order
group_order <- c(
    # "Africa region", 
    # "Americas region",
    # "Eastern Mediterranean region", 
    # "South-East Asia region", 
    # "Western Pacific region", 
    "Overall"
)

# Separate surveys and summaries
surveys_only <- forestDF %>%
    filter(grouping == "survey") %>%
    # group_by(facet_group) %>%
    arrange(desc(hivRatio), .by_group = TRUE) %>%
    ungroup()

summaries_only <- forestDF %>%
    filter(grouping != "survey")

# Recombine with surveys first, then summaries (preserves facet-wise order)
forestDF_plot <- bind_rows(surveys_only, summaries_only) %>%
    # mutate(facet_group = factor(facet_group, levels = group_order)) %>%
    mutate(display_label = case_when(
        grouping == "Overall" ~ "Overall",
        # grouping == "regional" ~ paste0("<b>", display_label_raw, "</b>"),
        TRUE ~ display_label_raw
    )) %>%
    mutate(display_label = factor(display_label, levels = rev(unique(display_label))))

# Prepare empirical data
empirical_prev_labels <- left_join(empirical_prev, survey_sum) %>%
    rename(display_label_raw = figure.id.yr) %>%
    select(display_label_raw, or_empirical) %>%
    mutate(
        grouping = "survey",
        display_label = display_label_raw
    ) %>%
    mutate(
           type = "Ratios calculated from reported prevalence")

# Plot
forestRatio <- ggplot(forestDF_plot) +
    geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
    geom_point(data = empirical_prev_labels,
               aes(x = display_label, y = or_empirical, color=type), alpha=0.8, size=8) +
    geom_pointinterval(aes(x = display_label, y = hivRatio, ymin = .lower, ymax = .upper, fill=type),
                       size = 0.25) +
    coord_flip() +
    scale_color_manual(values = myPal[c(2,1)]) +
    scale_y_log10(
        breaks = c(0.25, 0.5, 1, 2, 4),
        labels = c("0.25", "0.5", "1", "2", "4"),
    ) +
    # scale_x_discrete(labels = label_wrap(30)) + 
    labs(
        x = NULL,
        y = "HIV positive-to-HIV negative ratio of bacteriologically-confirmed TB prevalence"
    ) +
    theme_minimal(base_size = 10) +
    theme(
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        # panel.spacing.y = unit(1, "lines"),
        # axis.text.y = ggtext::element_markdown(hjust = 1),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.background = element_rect(fill = "white"),
        legend.position = "bottom", legend.title=element_blank(),
        legend.spacing.y = unit(-.25, 'cm'),
        legend.box.margin=margin(-10,-10,-6,-10),
        panel.spacing = unit(0.18, "cm"),
        plot.margin = unit(c(.1, .7, 0, .1), "cm")
    ) ; forestRatio

# Save
ggsave(plot = forestRatio, "output/hivFigure2.png", height = 12.5, width=8.5, units = "in")
```

# Model 3: Effect over time 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
timeEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z + (1 + year_z | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(timeEffectMod)
plot(timeEffectMod)
```

Prediction

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}
ndTime <- bactModelDF %>%
    select(logOR_se) %>%
    crossing(year_z = seq(min(cleanHivDF$year_z, na.rm = TRUE), max(cleanHivDF$year_z, na.rm = TRUE), 1))

timeEffectModEst <- add_epred_draws(object = timeEffectMod,
                                 newdata=na.omit(ndTime),
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>%
                           mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")
```

Main estimate
```{r}
timeEffectModEst %>%
    group_by(year_z) %>%
    mean_qi(hivRatio) %>%
    mutate(Year = year_z + round(mean(cleanHivDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, hivRatio, .lower, .upper)

### output for contribution table
exp(as.numeric(summary(timeEffectMod)[["fixed"]][2,1]))
exp(as.numeric(summary(timeEffectMod)[["fixed"]][2,3]))
exp(as.numeric(summary(timeEffectMod)[["fixed"]][2,4]))

bayes_R2(timeEffectMod)

timeEffectModEst 

```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    y <- timeEffectModEst  %>%
    group_by(year_z) %>%
    mean_qi(hivRatio) %>%
    mutate(Year = year_z + round(mean(cleanHivDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, hivRatio, .lower, .upper) 

    ggplot(data = y) +
    geom_lineribbon(.width = 0.95, alpha = 0.25, fill="mediumturquoise", aes(x = Year,
               y = hivRatio, ymin = .lower, ymax = .upper)) +
    geom_point(data = empirical_prev %>% rename("Study standard error" = logOR_se),
               aes(x=year_z + round(mean(cleanHivDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = `Study standard error`),
               colour = "black") +
    expand_limits(y=0) +
    labs(y = "HivPos-to-hivneg odds ratio",
         x = "Year") +
    theme_minimal(base_size = 30)  + theme(legend.position = "bottom")
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
#### Annual rate of change 
minYr <- min(timeEffectModEst$year_z)
maxYr <- max(timeEffectModEst$year_z)

startYr <- timeEffectModEst %>% filter(year_z == minYr) %>% select(hivRatio) %>% ungroup()
endYr <- timeEffectModEst %>% filter(year_z == maxYr) %>% select(hivRatio)%>% ungroup()
         data.frame( hivRatio = ((endYr[,"hivRatio"] - startYr[,"hivRatio"])/startYr[,"hivRatio"])/length(minYr:maxYr)*100) %>%
          mean_qi(hivRatio) 
```
# Sub group analysis

## National surveys

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
natEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(natRep=="Yes"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

subNatEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(natRep=="No"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(natEffectMod)
plot(natEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
ndNat <- bactModelDF %>%
         filter(natRep == "Yes") %>%
         select(study.country, LogOdds_HivNeg) %>%
         mutate(logOR_se=1)

natEffectModEst <- add_epred_draws(object = natEffectMod,
                                 newdata=ndNat,
                                 re_formula = NA) %>% 
                           mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

ndSubnat <- bactModelDF %>%
         filter(natRep == "No") %>%
         select(study.country, LogOdds_HivNeg) %>%
         mutate(logOR_se=1)

subnatEffectModEst <- add_epred_draws(object = subNatEffectMod,
                                 newdata=ndSubnat,
                                 re_formula = NA) %>% 
                           mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - National surveys
```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
natEffectModEst %>% ungroup() %>%
    mean_qi(hivRatio) %>% 
    select(hivRatio, .lower, .upper)
```

Main estimate - subnational surveys
```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
subnatEffectModEst %>% ungroup() %>%
    mean_qi(hivRatio) %>% 
    select(hivRatio, .lower, .upper)
```

## Study quality (Low vs. not low)

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}

lowBiasEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(bias == "Low"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

notLowBiasEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(bias == "Not low"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(lowBiasEffectMod)
plot(lowBiasEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
ndLowBias <- bactModelDF %>%
         filter(bias == "Low") %>%
         select(study.country, LogOdds_HivNeg) %>%
         mutate(logOR_se=1)

ndNotLowBias <- bactModelDF %>%
         filter(bias == "Not low") %>%
         select(study.country, LogOdds_HivNeg) %>%
         mutate(logOR_se=1)

lowBiasEffectModEst <- add_epred_draws(object = lowBiasEffectMod,
                                 newdata=ndLowBias,
                                 re_formula = NA) %>% 
                           mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

notLowBiasEffectModEst <- add_epred_draws(object = notLowBiasEffectMod,
                                 newdata=ndNotLowBias,
                                 re_formula = NA) %>% 
                           mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - low risk of bias
```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
lowBiasEffectModEst %>% 
    ungroup() %>%
    mean_qi(hivRatio) %>% 
    select(hivRatio, .lower, .upper)
```

Main estimate - not low risk of bias
```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
notLowBiasEffectModEst %>% 
    ungroup() %>%
    mean_qi(hivRatio) %>% 
    select(hivRatio, .lower, .upper)
```


## Symptom screen used vs. no symptom screen used

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}

sympYesEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(symptomScreen == "Yes"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

sympNoEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(symptomScreen == "No"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(sympYesEffectMod)
plot(sympYesEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
ndSympYes <- bactModelDF %>%
         filter(symptomScreen == "Yes") %>%
         select(study.country, LogOdds_HivNeg) %>%
         mutate(logOR_se=1)

ndSympNo <- bactModelDF %>%
         filter(symptomScreen == "No") %>%
         select(study.country, LogOdds_HivNeg) %>%
         mutate(logOR_se=1)

sympYesEffectModEst <- add_epred_draws(object = sympYesEffectMod,
                                 newdata=ndSympYes,
                                 re_formula = NA) %>% 
                           mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

sympNoEffectModEst <- add_epred_draws(object = sympNoEffectMod,
                                 newdata=ndSympNo,
                                 re_formula = NA) %>% 
                           mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - used symptom screening
```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
sympYesEffectModEst %>% ungroup() %>%
    mean_qi(hivRatio) %>% 
    select(hivRatio, .lower, .upper)
```

Main estimate - did not use symptom screening
```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
sympNoEffectModEst %>% ungroup() %>%
    mean_qi(hivRatio) %>% 
    select(hivRatio, .lower, .upper)
```

## Reported xpert or culture required for bact-tb diagnosis vs. xpert or culture not required

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
xpCltReqYesEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(XpertCultRequired == "Yes"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

xpCltReqNoEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(XpertCultRequired == "No"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(xpCltReqYesEffectMod)
plot(xpCltReqYesEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
ndxpCltReqYes <- bactModelDF %>%
         filter(XpertCultRequired == "Yes") %>%
         select(study.country, LogOdds_HivNeg) %>%
         mutate(logOR_se=1)

ndxpCltReqNo <- bactModelDF %>%
         filter(XpertCultRequired == "No") %>%
         select(study.country, LogOdds_HivNeg) %>%
         mutate(logOR_se=1)

xpCltReqYesEffectModEst <- add_epred_draws(object = xpCltReqYesEffectMod,
                                 newdata=ndxpCltReqYes,
                                 re_formula = NA) %>% 
                           mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

xpCltReqNoEffectModEst <-  add_epred_draws(object = xpCltReqNoEffectMod,
                                 newdata=ndxpCltReqNo,
                                 re_formula = NA) %>% 
                           mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - required xpert or culture
```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
xpCltReqYesEffectModEst %>% ungroup() %>%
    mean_qi(hivRatio) %>% 
    select(hivRatio, .lower, .upper)
```

Main estimate - did not require xpert or culture
```{r, echo=FALSE, warning = FALSE, message=FALSE, cache=TRUE}
xpCltReqNoEffectModEst %>% ungroup() %>%
    mean_qi(hivRatio) %>% 
    select(hivRatio, .lower, .upper)
```

Country-level effects

```{r}
# ndCountry <- newBactModelDF %>%
#              select(study.country, LogOdds_HivNeg, 
#                     id, GDI,
#                     alcohol.use.disorder,
#                     type2.diabetes.mellitus, 
#                     bmi, HIV.AIDS, smoking) %>%
#              mutate(logOR_se=1, 
#                     year_z = 10)
# 
# countryEffectModEst <- add_epred_draws(object = multiMod,
#                                  newdata=ndCountry,
#                                  re_formula = ~ 1  + (1|study.country), 
#                                  allow_new_levels = TRUE) %>% 
#                     mutate(hivRatio = exp(.epred),
#                            type="Posterior mean and 95% credible interval") 
# 
# countryEffectModEst %>% group_by(study.country) %>% mean_qi(hivRatio) 
# # write_csv(countryEffectModEst %>% group_by(study.country) %>% mean_qi(hivRatio), here("output/countryMFRatios.csv"))
# countryEffectModEst %>% group_by(study.country) %>% mean_qi(hivRatio) %>% select(study.country, hivRatio) -> countryRatios
# 
# ### Setup a world map 
# world_map <- map_data("world")
# world_map <- subset(world_map, region != "Antarctica")
# 
# world <- map_data("world")
# world <- world[world$region != "Antarctica", ]
# 
# countryMap <- left_join(countryRatios %>% rename(region=study.country), 
#                          world) %>% rename(study.country = region)
# library(geosphere) 
# centroids <- world_map %>% 
#     group_by(region) %>% 
#     group_modify(~ data.frame(centroid(cbind(.x$long, .x$lat))))
# 
# nCountriesGeo <- left_join(countryMap, centroids %>% rename(study.country = region))
# 
# ggplot(countryMap) +
#     geom_map(
#         dat = world_map, map = world_map, aes(map_id = region),
#         fill = "lightgrey", color = "black", size = 0.25) +
#     geom_map(dat = countryMap, map = world_map, aes(map_id = study.country, fill = hivRatio), color="black") +
#     scale_fill_gradient(low = "pink", high = "darkred", 
#                         name = "HivPos-to-hivneg ratio") +
#     expand_limits(x = world_map$long, y = world_map$lat) + 
#     theme_void() + theme(legend.position = "inside",
#                             legend.position.inside = c(.15,.2))
# 

```

# Smear positive analysis 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
cleanHivSmrDF0 <- smrPosIndicator("hiv") %>% filter(hiv.smr.analysis.indicator !="none")

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanHivSmrDF <- cleanHivSmrDF0 %>% 
                 filter(hiv.smr.analysis.indicator != "none") %>% 
                 mutate("id" = row_number()) %>%
                 ### Scale all prevalence estimates back to decimals 
                 mutate_at(vars(matches('prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>% 
                 select(id,
                       covidence.id, 
                       figure.id.yr, 
                       study.geography, 
                       title.extracted,
                       study.country,
                       WHO.region,
                       study.start.year, 
                       study.end.year, 
                       ends_with("hiv.positive"), ends_with("hiv.negative"), ends_with("hiv.total"),
                       hiv.smr.analysis.indicator) %>%
    
    mutate("tempPrevHivPos" =  case_when(hiv.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.positive,
                                       hiv.smr.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive, 
                                       hiv.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.positive/n.participants.hiv.positive,
                                        hiv.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive),
           "tempPrevHivNeg" = case_when(hiv.smr.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.negative,
                                        hiv.smr.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative, 
                                       hiv.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.negative/n.participants.hiv.negative,
                                       hiv.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative)) %>%
    pivot_longer(cols = c(tempPrevHivNeg,tempPrevHivPos), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivPos" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivNeg" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                       
                                        hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivPos" ~ (prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivNeg" ~ (prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                        hiv.smr.analysis.indicator %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "HivPos" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.positive),
                                        hiv.smr.analysis.indicator  %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "HivNeg" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.negative)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "HivPos" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                              hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "HivNeg" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92,
                                                 hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "HivPos" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                                 hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "HivNeg" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92, 
                                            hiv.smr.analysis.indicator %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "HivPos" ~ 
                                            log(1/(n.participants.hiv.positive*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        hiv.smr.analysis.indicator  %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "HivNeg" ~ 
                                            log(1/(n.participants.hiv.negative*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       # year_z = ifelse(!is.na(study.end.year), 
                                       #                 study.end.year - round(mean(study.end.year, na.rm=TRUE))
                                       #                 (publication.year - meanLag) - round(mean((publication.year - meanLag), na.rm=TRUE))),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))  %>% filter(LogOddsStandardError < Inf)

smrModelDF <- cleanHivSmrDF %>%
    select(id, WHO.region, study.country, Sex, natRep, LogOdds, LogOddsStandardError, 
           n.participants.hiv.positive, n.participants.hiv.negative) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_HivPos - LogOdds_HivNeg,
        logOR_se = sqrt(LogOddsStandardError_HivPos^2 + LogOddsStandardError_HivNeg^2) 
    ) 
```

# Smear positive descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(cleanHivSmrDF0)[1], "unique surveys with sex-stratified smear positive TB"))

# sort(unique(cleanHivDF0$title.extracted))
print(paste(length(unique(cleanHivSmrDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanHivSmrDF0$n.participants.hiv.negative)), "studies report participant numbers totaling", 
            sum(cleanHivSmrDF0$n.participants.hiv.negative, cleanHivSmrDF0$n.participants.hiv.positive, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanHivSmrDF0$n.participants.hiv.positive, na.rm=TRUE)/sum(cleanHivSmrDF0$n.participants.hiv.negative, cleanHivSmrDF0$n.participants.hiv.positive, na.rm = TRUE)*100, 2), "% of participants were hivpos" ))
```
More descriptives 

```{r}
table(smrModelDF$WHO.region)
smrModelDF %>% reframe(Participants = sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE), .by=WHO.region)
smrModelDF %>% reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=WHO.region)
```

# Model X: Smear positive main effect 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | study.country) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectSmrMod)
plot(mainEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- smrModelDF %>%
      select(study.country, LogOdds_HivNeg) %>%
      mutate(logOR_se=1)


mainEffectSmrModEst <- add_epred_draws(object = mainEffectSmrMod,
                                 newdata=nd,
                                 re_formula = ~ 1) %>% 
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectSmrModEst %>% ungroup %>% mean_qi(hivRatio)
```

Survey specific estimate
```{r}
surveyEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=bactModelDF,
                                 re_formula = ~ 1  + (1|study.country) + (1|id),
                                 allow_new_levels = TRUE) %>%
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

surveyEffectModEst %>% mean_qi(hivRatio) %>% 
    select(id, study.country, hivRatio, .lower, .upper)
```
