---
title: "Tuberculosis Prevalence among persons living with HIV in Low- and Middle-income Countries: A Systematic Review and Meta-analysis" 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

About this file:

This script reads in the clean dataset from data folder and then runs fits Bayesian generalized multilevel models on the studies that report TB prevalence for persons living with and without HIV (PLWH and PLWoH). These models aim to estimate the prevalence risk ratio among PLWH compared with PLWoH; the notification rate ratio among PLWH compared with PLWoH; the notification-to-prevalence ratio for PLWH and PLWoH. 


# Setup the code environment 
```{r, echo = FALSE,  warning = FALSE, message=FALSE}
##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
library(tidybayes)
library(wesanderson)
library(ggplot2)
library(maps)
library(patchwork)
### Add in helper function 
expit <- function (x) {exp(x)/(1 + exp(x))}

### Define palette for plots
myPal <- c("black", "mediumturquoise")

### Source in codes from repository
source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))

### Read in country-level HIV prevalence 
### available for calculations below.
artCoverage <- read.csv(here("data/hivAnalysis/unaids-2025-adult-art-coverage.csv")) %>% 
               rename(study.country=spectrum_country, 
                      study.end.year = year) %>% 
               ### We need to replace the HIV prevalence value for 2009 Uganda to 
               ### correspond to the Kampala City, which is 6.8 percent 
               mutate(hivPrev = ifelse(study.country == "Uganda" & study.end.year==2009, 
                                       0.084, # From Jeff Imai-Eaton via Spectrum UNAIDS
                                       hivpop/totpop),
                      totpop = ifelse(study.country == "Uganda" & study.end.year==2009, 
                                       1.5*1e6, # https://www.un.org/en/development/desa/population/publications/pdf/urbanization/urbanization-wallchart2009.pdf
                                       totpop),
                      hivpop = ifelse(study.country == "Uganda" & study.end.year==2009, 
                                       totpop * hivPrev, 
                                       hivpop))

              

```

# Data setup 

## Step 1: Load in clean dataset and filter to those studies that report TB stratified by HIV status. 

```{r, echo=FALSE, warning=FALSE}
##############################################################################|
##### All surveys that we extracted HIV data from
##############################################################################|
allHIV <- readRDS(here("data/fullDataClean.rds")) %>% 
          filter(report.hiv == "Yes") %>% 
               filter(
##### Remove duplicate survey 
                      covidence.id != 26886)

##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING HIV status for 
##### QUANTITATIVE ANALYSIS                                  #################

cleanHivDF0 <- bactPostIndicator("hiv") %>% 
##### Manually move the Uganda study because they don't report an appropriate 
##### denominator 
              mutate(hiv.analysis.indicator = ifelse(covidence.id == 26870, 
                                                     "none", 
                     hiv.analysis.indicator)) %>%
##### Manually move the two South African studies that report "prevalent TB" which 
##### meets our bacteriologically-confirmed TB threshold back into the analysis. 
               mutate(hiv.analysis.indicator = ifelse(covidence.id==26953, 
                                                      "prev100k.ci.prevalent.tb", 
                                                      ifelse(covidence.id == 26911, 
                                                             "n.prevalent.tb", 
                                                             hiv.analysis.indicator)), 
                      study.end.year = as.integer(study.end.year)) %>% 
               filter(
# ##### Remove non-use of HIV test
#                       hiv.status.definition != "Self-report",
##### Remove duplicate survey 
                      covidence.id != 26886) %>%
## Join in the HIV prevalence data
              left_join(artCoverage %>% 
                        select(study.country, study.end.year, hivPrev)) %>% 
              mutate(n.participants.hiv.positive.calc = ifelse(is.na(n.participants.hiv.positive) | covidence.id == 26947, 
                                                          round(n.participants.sex.total * hivPrev), 
                                                          n.participants.hiv.positive),
                     n.participants.hiv.negative.calc = ifelse(is.na(n.participants.hiv.negative) | covidence.id == 26947, 
                                                          n.participants.sex.total-n.participants.hiv.positive.calc, 
                                                          n.participants.hiv.negative), 
                     hiv.analysis.indicator = ifelse(hiv.analysis.indicator == "none" & 
                                                      !is.na(n.bacteriological.tb.hiv.negative) &
                                                      !is.na(n.bacteriological.tb.hiv.positive), 
                                                      "n.bacteriological.tb", 
                                                      hiv.analysis.indicator)
                     )

### Calculate mean publication lag 
meanLag <- round(mean(cleanHivDF0$publication.year - cleanHivDF0$study.end.year, na.rm=TRUE))
```

### Look at the data distribution

Figure of countries of all identified surveys

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 35}
nCountryAll <- allHIV %>% 
            count(study.country) %>% 
            mutate(n=as.factor(n))

world_map <- map_data("world")
world_map <- world_map[world_map$region != "Antarctica", ]

nCountryMap <- left_join(world_map, nCountryAll %>% rename(region=study.country), by="region")

ggplot(nCountryMap, aes(x=long, y=lat)) +
          geom_polygon(aes(group = region, subgroup = group), 
                       fill="lightgrey")  +
      geom_polygon(aes(group = region, subgroup = group,
                       fill = n), colour = "white")  +
    scale_fill_manual(values= c("lightblue", "dodgerblue3", "blue4"),
                      name = "Number of\nprevalence surveys",
                      na.translate = FALSE) +
    theme_void() + theme(legend.position = "inside",
                            legend.position.inside = c(.47,.35))


# ggsave(here("output/HIV/allHivWorld.png"), plot=last_plot(), 
#        width = 15, height = 8.5, units = "in")
```

## Step 2: Create a tidy dataset of relevant variables for modeling. 
```{r, echo = FALSE,  warning = FALSE, message=FALSE}
##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanHivDF <-  cleanHivDF0 %>% 
               mutate("id" = row_number()) %>% 
    ### Scale all prevalence estimates back to decimals 
    mutate_at(vars(matches('prev100k.bacteriological.tb.|prev100k.ci.([a-z]{3})er.bacteriological.tb.|prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.|prev100k.prevalent.tb.|prev100k.ci.([a-z]{3})er.prevalent.tb.')),  ~ (. / 1e5)) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           study.rurality, 
           WHO.region,
           screening.tests,
           hiv.status.definition,
           participants.sputum.collected,
           bacteriological.tb.definition,
           study.start.year, 
           study.end.year, 
           publication.year,
           study.quality.summary,
           n.participants.male, n.participants.female,
           starts_with("n.participants.age.grp"), 
           ends_with("hiv.positive"), ends_with("hiv.negative"), ends_with("hiv.total"),
           n.participants.hiv.positive.calc, n.participants.hiv.negative.calc,
           hiv.analysis.indicator, 
           hivPrev) %>%
    
    mutate("tempPrevHivPos" =  case_when(hiv.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.positive,
                                       hiv.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive, 
                                        hiv.analysis.indicator == 
                                           "prev100k.ci.prevalent.tb" ~ prev100k.prevalent.tb.hiv.positive, 
                                       hiv.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.hiv.positive/n.participants.hiv.positive.calc, 
                                       hiv.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.positive/n.participants.hiv.positive.calc,
                                       hiv.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.hiv.positive/n.participants.hiv.positive.calc,
                                       hiv.analysis.indicator == 
                                           "n.prevalent.tb" ~ n.prevalent.tb.hiv.positive/n.participants.hiv.positive.calc,
                                       hiv.analysis.indicator == 
                                           "adj.prev100k.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.hiv.positive,
                                       hiv.analysis.indicator == 
                                           "prev100k.bacteriological.tb" ~ prev100k.bacteriological.tb.hiv.positive,
                                       hiv.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive),
           "tempPrevHivNeg" = case_when(hiv.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.hiv.negative, 
                                        hiv.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.negative,
                                        hiv.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.hiv.negative, 
                                        hiv.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative, 
                                        hiv.analysis.indicator == 
                                           "prev100k.ci.prevalent.tb" ~ prev100k.prevalent.tb.hiv.negative, 
                                       hiv.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.hiv.negative/n.participants.hiv.negative.calc, 
                                       hiv.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.negative/n.participants.hiv.negative.calc,
                                       hiv.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.hiv.negative/n.participants.hiv.negative.calc,
                                       hiv.analysis.indicator == 
                                           "n.prevalent.tb" ~ n.prevalent.tb.hiv.negative/n.participants.hiv.negative.calc,
                                       hiv.analysis.indicator == 
                                           "adj.prev100k.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.hiv.negative,
                                       hiv.analysis.indicator == 
                                           "prev100k.bacteriological.tb" ~ prev100k.bacteriological.tb.hiv.negative,
                                       hiv.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative)) %>%
    pivot_longer(cols = c(tempPrevHivNeg,tempPrevHivPos), 
                 names_to = "HIVStatus", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivPos" ~ (adj.prev100k.ci.upper.bacteriological.tb.hiv.positive-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivNeg" ~ (adj.prev100k.ci.upper.bacteriological.tb.hiv.negative-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivPos" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivNeg" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivPos" ~ (prev100k.ci.upper.bacteriological.tb.hiv.positive-
                                                                 prev100k.ci.lower.bacteriological.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            HIVStatus == "HivNeg" ~ (prev100k.ci.upper.bacteriological.tb.hiv.negative-
                                                                   prev100k.ci.lower.bacteriological.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivPos" ~ (prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            HIVStatus == "HivNeg" ~ (prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.prevalent.tb" & 
                                            HIVStatus == "HivPos" ~ (prev100k.ci.upper.prevalent.tb.hiv.positive-
                                                                 prev100k.ci.lower.prevalent.tb.hiv.positive)/3.92,
                                        hiv.analysis.indicator == "prev100k.ci.prevalent.tb" & 
                                            HIVStatus == "HivNeg" ~ (prev100k.ci.upper.prevalent.tb.hiv.negative-
                                                                   prev100k.ci.lower.prevalent.tb.hiv.negative)/3.92,
                                        (hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb",
                                                                      "n.culture.positive.tb",  "n.prevalent.tb",
                                                                      "prev100k.bacteriological.tb", "prev100k.smear.positive.tb") & 
                                            HIVStatus == "HivPos"  & !is.na(n.participants.hiv.positive.calc)) ~
                                                 sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.positive.calc),
                                        (hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "n.prevalent.tb",
                                                                      "adj.prev100k.bacteriological.tb", 
                                                                      "prev100k.bacteriological.tb", "prev100k.smear.positive.tb") & 
                                            HIVStatus == "HivNeg" & !is.na(n.participants.hiv.negative.calc)) ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.negative.calc),
                                        (hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb",
                                                                      "n.culture.positive.tb",  "n.prevalent.tb",
                                                                      "adj.prev100k.bacteriological.tb", 
                                                                      "prev100k.bacteriological.tb", "prev100k.smear.positive.tb") & 
                                            HIVStatus == "HivPos"  & !is.na(n.bacteriological.tb.hiv.positive)) ~
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.bacteriological.tb.hiv.positive),
                                        (hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "n.prevalent.tb",
                                                                      "adj.prev100k.bacteriological.tb", 
                                                                      "prev100k.bacteriological.tb", "prev100k.smear.positive.tb") & 
                                            HIVStatus == "HivNeg" & !is.na(n.bacteriological.tb.hiv.negative)) ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.bacteriological.tb.hiv.negative)),
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  HIVStatus == "HivPos" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.hiv.positive/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.hiv.positive)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.hiv.positive/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.hiv.positive))) /3.92,
                                              hiv.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  HIVStatus == "HivNeg" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.hiv.negative/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.hiv.negative)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.hiv.negative /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.hiv.negative))) /3.92,
                                              hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  HIVStatus == "HivPos" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                              hiv.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  HIVStatus == "HivNeg" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92,
                                              hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 HIVStatus == "HivPos" ~ (log(prev100k.ci.upper.bacteriological.tb.hiv.positive/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.hiv.positive)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.hiv.positive/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.hiv.positive))) /3.92,
                                              hiv.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 HIVStatus == "HivNeg" ~ (log(prev100k.ci.upper.bacteriological.tb.hiv.negative/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.hiv.negative)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.hiv.negative /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.hiv.negative))) /3.92,
                                              hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     HIVStatus == "HivPos" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                              hiv.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     HIVStatus == "HivNeg" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92, 
                                              hiv.analysis.indicator == "prev100k.ci.prevalent.tb" &
                                                     HIVStatus == "HivPos" ~ (log(prev100k.ci.upper.prevalent.tb.hiv.positive/
                                                                          (1-prev100k.ci.upper.prevalent.tb.hiv.positive)) -
                                                                      log(prev100k.ci.lower.prevalent.tb.hiv.positive /
                                                                          (1-prev100k.ci.lower.prevalent.tb.hiv.positive)))/3.92,
                                              hiv.analysis.indicator == "prev100k.ci.prevalent.tb" &
                                                     HIVStatus == "HivNeg" ~ (log(prev100k.ci.upper.prevalent.tb.hiv.negative/
                                                                            (1-prev100k.ci.upper.prevalent.tb.hiv.negative)) -
                                                                        log(prev100k.ci.lower.prevalent.tb.hiv.negative /
                                                                            (1-prev100k.ci.lower.prevalent.tb.hiv.negative)))/3.92, 
                                              hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb",
                                                                            "n.culture.positive.tb",  "n.prevalent.tb",
                                                                            "adj.prev100k.bacteriological.tb", 
                                                                            "prev100k.bacteriological.tb", "prev100k.smear.positive.tb") & 
                                                    HIVStatus == "HivPos"  & !is.na(n.participants.hiv.negative.calc) ~ 
                                                                (1/(n.participants.hiv.positive.calc*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                            hiv.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb",
                                                                          "n.culture.positive.tb",  "n.prevalent.tb",
                                                                          "adj.prev100k.bacteriological.tb", 
                                                                          "prev100k.bacteriological.tb","prev100k.smear.positive.tb") & 
                                                    HIVStatus == "HivNeg"  & !is.na(n.participants.hiv.negative.calc) ~ 
                                                                (1/(n.participants.hiv.negative.calc*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                            hiv.analysis.indicator %in% c("prev100k.bacteriological.tb",
                                                                          "prev100k.smear.positive.tb") & 
                                                    HIVStatus == "HivPos"  & is.na(n.participants.hiv.negative.calc) ~ 
                                                                (1/(n.bacteriological.tb.hiv.positive*(1-`Adjusted Prevalence`)))^2,
                                            hiv.analysis.indicator %in% c("prev100k.bacteriological.tb",
                                                                          "prev100k.smear.positive.tb") & 
                                                    HIVStatus == "HivNeg"  & is.na(n.participants.hiv.negative.calc) ~ 
                                                                (1/(n.bacteriological.tb.hiv.negative*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(is.na(study.geography), 
                                                       NA, 
                                                       ifelse(study.geography == "Nationally representative", 
                                                       "Yes", "No")),

                                       year_z = ifelse(!is.na(study.end.year), 
                                                       study.end.year - round(mean(study.end.year, na.rm=TRUE)),
                                                       (publication.year - meanLag) - round(mean((publication.year - meanLag), na.rm=TRUE))),                                       
                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       
                                       symptomScreen = ifelse(is.na(screening.tests), 
                                                                    NA, 
                                                                    ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No")),
                                       symptomsRequired = ifelse(is.na(participants.sputum.collected), 
                                                                       NA, 
                                                                       ifelse(!grepl("x-ray|all participants|unknown|other",
                                                                        participants.sputum.collected,ignore.case=TRUE),
                                                                 "Yes", "No")),
                                       XpertCultRequired = ifelse(is.na(bacteriological.tb.definition),
                                                                  NA, ifelse(!grepl("smear|micros|field|NTP", 
                                                                              bacteriological.tb.definition,
                                                                              ignore.case=TRUE), 
                                                                  "Yes", 
                                                                   "No")),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))

##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("hiv")

bactModelHivDFAll <- cleanHivDF %>%
    select(covidence.id, id, WHO.region, study.country, study.rurality,
           HIVStatus, year_z, hiv.status.definition, study.quality.summary,
           natRep, bias, symptomScreen, symptomsRequired, 
           XpertCultRequired, #`Adjusted Prevalence`, `Standard Error`,
           LogOdds, LogOddsStandardError, hiv.analysis.indicator, 
           n.participants.female, n.participants.male, 
           n.participants.hiv.positive.calc, n.participants.hiv.negative.calc, 
           n.participants.hiv.positive, n.participants.hiv.negative, 
           n.participants.hiv.total,
           starts_with("n.participants.age.grp"),
           publication.year, study.end.year) %>%
    pivot_wider(names_from = HIVStatus,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(logOR = LogOdds_HivPos - LogOdds_HivNeg,
           unadj_logOR_se = sqrt(LogOddsStandardError_HivPos^2 + LogOddsStandardError_HivNeg^2), 
           logOR_se = ifelse(grepl("adj", hiv.analysis.indicator),
                                     unadj_logOR_se,
                                     ifelse(grepl("smear", hiv.analysis.indicator),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Bact Variance Median"]))),
           logOR_se = unadj_logOR_se) 

bactModelHivDF <- bactModelHivDFAll %>%
    
            #### Filter out self-testing 
            filter(!hiv.status.definition %in% c("Self-report","Not specified"),
                   covidence.id != 26856)

```

### Descriptive statistics of model data

```{r, echo=FALSE}
#### Total participants among the main analysis surveys
cleanHivDF0 %>% filter(!hiv.status.definition %in% c("Self-report", "Not specified"),
                   covidence.id != 26856) %>% 
                mutate(totalParticipants = ifelse(is.na(n.participants.sex.total), 
                                                  n.participants.hiv.positive + n.participants.hiv.negative, 
                                                  n.participants.sex.total)) %>%
                select(totalParticipants) %>% 
                sum(na.rm = TRUE)

#### Total participants among the secondary analysis surveys (self-reported HIV status)
cleanHivDF0 %>% filter(hiv.status.definition %in% c("Self-report", "Not specified"),
                   covidence.id != 26856) %>% 
                mutate(totalParticipants = ifelse(is.na(n.participants.sex.total), 
                                                  n.participants.hiv.positive + n.participants.hiv.negative, 
                                                  n.participants.sex.total)) %>%
                select(totalParticipants) %>% 
                sum(na.rm = TRUE)

print(paste0( 
      sum(bactModelHivDF$natRep == "Yes"), " (",
      round(sum(bactModelHivDF$natRep == "Yes") /
      length(unique(bactModelHivDF$covidence.id)) * 100, 2), 
      "%) of the surveys were nationally representative"))
```

### Create a CSV of the data used in quantitative analysis 

```{r, echo=FALSE}

studyDetailsDF <- left_join(bactModelHivDF, cleanHivDF %>% select(covidence.id, 
                                                                  figure.id.yr, 
                                                                  title.extracted,
                                                                  HIVStatus,
                                                                  `Adjusted Prevalence`) %>% 
                                                            pivot_wider(id_cols = c(covidence.id, 
                                                                                    figure.id.yr, 
                                                                                    title.extracted), 
                                                                        names_from = HIVStatus, 
                                                                        values_from = `Adjusted Prevalence`)) %>%
                  mutate(n.participants.hiv.total = ifelse(!is.na(n.participants.hiv.total), 
                                                           n.participants.hiv.total, 
                                                           n.participants.hiv.negative + n.participants.hiv.positive),
                         percHIVpos = round(n.participants.hiv.positive/n.participants.hiv.total*100, 2),
                         HivPos = round(HivPos * 1e5),
                         HivNeg = round(HivNeg * 1e5)) %>%
                  select(figure.id.yr, 
                         title.extracted,
                         hiv.analysis.indicator, 
                         HivPos, HivNeg,
                         n.participants.hiv.total, 
                         n.participants.hiv.positive,
                         study.quality.summary
                         )


studyDetailsDFBroad <-  cleanHivDF %>% select(covidence.id, figure.id.yr, 
                                                                  title.extracted,
                                              hiv.analysis.indicator, 
                                                                                    n.participants.hiv.total,
                                                                                    n.participants.hiv.positive,
                                              n.participants.hiv.negative,
                                                                                    study.quality.summary,
                                                                  HIVStatus,
                                                                  `Adjusted Prevalence`) %>% 
                                                            pivot_wider(id_cols = c(covidence.id, 
                                                                                    figure.id.yr, 
                                                                                    title.extracted, 
                                                                                    hiv.analysis.indicator, 
                                                                                    n.participants.hiv.total,
                                                                                    n.participants.hiv.positive,
                                                                                    n.participants.hiv.negative,
                                                                                    study.quality.summary), 
                                                                        names_from = HIVStatus, 
                                                                        values_from = `Adjusted Prevalence`) %>%
                  mutate(n.participants.hiv.total = ifelse(!is.na(n.participants.hiv.total), 
                                                           n.participants.hiv.total, 
                                                           n.participants.hiv.negative + n.participants.hiv.positive),
                         percHIVpos = round(n.participants.hiv.positive/n.participants.hiv.total*100, 2),
                         HivPos = round(HivPos * 1e5),
                         HivNeg = round(HivNeg * 1e5)) %>%
                  select(figure.id.yr, 
                         title.extracted,
                         hiv.analysis.indicator, 
                         HivPos, HivNeg,
                         n.participants.hiv.total, 
                         n.participants.hiv.positive,
                         study.quality.summary
                         )

# View(studyDetailsDF)
# write.csv(allHIV %>% select(study.country, 
#                   study.setting,
#                   study.years,
#                       title.extracted,
#                       n.participants.hiv.total, 
#                       n.participants.hiv.positive,
#                       n.participants.hiv.negative,
#                       study.quality.summary), here("output/HIV/studyDetailsTableAll.csv"))



# write.csv(studyDetailsDF, here("output/HIV/studyDetailsTable.csv"))

```

### Look at the data distribution

Figure of countries used in quantitative analysis 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 35}
nCountry <- bactModelHivDF %>% 
            count(study.country) %>% 
            mutate(n=as.factor(n))

world_map <- map_data("world", region = c("Morocco", "Algeria", "Tunisia", "Libya",
             "Egypt", "Mauritania", "Mali", "Burkina Faso",
             "Niger", "Chad", "Sudan", "Ethiopia", "Eritrea",
             "Djibouti", "South Sudan", "Somalia", "Senegal",
             "Cape Verde", "Gambia", "Guinea", "Guinea-Bissau",
             "Liberia", "Sierra Leone", "Ivory Coast", "Ghana",
             "Togo", "Benin", "Nigeria", "Cameroon", 
             "Central African Republic", "Gabon", "Equatorial Guinea",
             "Sao Tome and Principe", "Uganda", "Republic of Congo", 
             "Democratic Republic of the Congo", "Rwanda", "Burundi",
             "Kenya", "Tanzania", "Angola", "Mozambique", "Malawi", 
             "Mauritius", "Seychelles", "Comoros", 
             "Botswana", "Zimbabwe", "Zambia", "Namibia", "Swaziland", 
             "South Africa", "Lesotho", "Western Sahara"))

nCountryMap <- left_join(world_map, nCountry %>% rename(region=study.country), by="region")

ggplot(nCountryMap, aes(x=long, y=lat)) +
          geom_polygon(aes(group = region, subgroup = group), 
                       fill="lightgrey")  +
      geom_polygon(aes(group = region, subgroup = group, 
                       fill = n), colour = "white")  +
  coord_map(projection = "mercator",
            xlim = c(-26,64), ylim = c(-35,38)) +
    # # scale_fill_gradient(low = "lightblue", high = "dodgerblue4", 
    # #                     name = "Number of\nprevalence surveys") +
    scale_fill_manual(values= c("lightblue", "dodgerblue3", "darkblue"), 
                      name = "Number of\nprevalence surveys",
                      na.translate = FALSE) +
    theme_void() + theme(legend.position = "none",
                            legend.position.inside = c(.3,.4))


# ggsave(here("output/HIV/hivWorld.png"), plot=last_plot(), width = 15, height = 8.5, units = "in")

```

### Survey end years 

```{r, echo=FALSE, warning = FALSE, message=FALSE}

quantAnalysisIDs <- bactModelHivDF$covidence.id
allHIVtestIDs <- bactModelHivDFAll$covidence.id

studyYearPlotDF <- bactModelHivDFAll %>%
                   mutate("IncExc" = case_when(covidence.id %in% quantAnalysisIDs ~ "Included in primary quantitative analysis",
                                               covidence.id %in% allHIVtestIDs ~ "Excluded due to self-reported HIV status",
                                               .default = "Excluded. Did not provide sufficient data.")) %>%
    transmute(
        id,
        WHO.region = WHO.region,
        IncExc = IncExc,
        study.end.year = study.end.year,
        study.country = study.country,
        natRep = ifelse(natRep == "Yes", "Nationally representative survey", "Subnational survey"),
        bias = bias,
        p_hivneg_empirical = plogis(LogOdds_HivNeg),
        p_hivpos_empirical = plogis(LogOdds_HivPos),
        # or_empirical = p_hivpos_empirical / p_hivneg_empirical,
        or_empirical = (p_hivpos_empirical / (1 - p_hivpos_empirical)) / 
                        (p_hivneg_empirical / (1 - p_hivneg_empirical))  # odds ratio
    ) 

ggplot(data = studyYearPlotDF, 
       aes(x=study.end.year, 
       y=or_empirical)) + 
       geom_point(aes(color=natRep, shape=IncExc), size = 3) +
       geom_text(aes(label=study.country), size = 4, nudge_y = -0.15) +
           geom_hline(aes(yintercept=1), linetype = "dashed") +
           theme_minimal(base_size = 20) + theme(
                                            legend.position = c(.80, .90),
                                            legend.background = element_rect(fill="white"),
                                          legend.title = element_blank(), 
                                          panel.grid.minor = element_blank(),
                                          panel.grid.major = element_blank()) + 
           guides(color=guide_legend(nrow=2), shape=guide_legend(nrow=2))+
           ylab("Empirical HIV+ / HIV- odds ratio") + xlab("Study end year") + 
           scale_color_manual(values = c("darkblue", "darkorange")) + 
           scale_y_continuous(breaks=seq(0,9,1)) + 
           scale_x_continuous(breaks=seq(2005,2022,1)) -> timePlot

ggsave(plot = timePlot, "output/HIV/time-plot.png", height = 8.5, width=13.5, units = "in")


```

# Primary pooled estimate of prevalence risk ratio

## Model 1: Main effect model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|study.country) + (1|id),
                       data = bactModelHivDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 10000,
                       warmup = 2000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectMod)
plot(mainEffectMod)

nd <- bactModelHivDF %>%
      select(study.country, LogOdds_HivNeg) %>%
      mutate(logOR_se=1)

mainEffectModEst <- add_epred_draws(object = mainEffectMod,
                                    newdata = nd,
                                    re_formula = ~ 1, 
                                    allow_new_levels = TRUE) %>% 
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r, echo=FALSE, warning = FALSE, message=FALSE}
mainEffectModEst %>% ungroup() %>% mean_qi(hivRatio)  %>% 
mutate(across(c(hivRatio, .lower, .upper), round, 2))
```

Survey-level effects

```{r, echo=FALSE, warning = FALSE, message=FALSE}
surveyEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=bactModelHivDF,
                                 re_formula = ~ 1  + (1|study.country) + (1|id),
                                 allow_new_levels = TRUE) %>%
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

surveyEffectModEst %>% mean_qi(hivRatio)
```

Country-level estimates

```{r, echo=FALSE, warning = FALSE, message=FALSE}
nd_country <- bactModelHivDF %>% select(study.country) %>% unique() %>% 
              mutate(logOR_se = 1)

countryEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=nd_country,
                                 re_formula = ~ 1  + (1|study.country),
                                 allow_new_levels = TRUE) %>%
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

countryEffectModEst %>% mean_qi(hivRatio) %>% 
mutate(across(c(hivRatio, .lower, .upper), round, 2))
```

Set the empirical values from the extracted data. 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
### Get the empirical values 
empirical_prev <- bactModelHivDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        bias = bias,
        symptomScreen = symptomScreen, 
        p_hivneg_empirical = plogis(LogOdds_HivNeg),
        p_hivpos_empirical = plogis(LogOdds_HivPos),
        or_empirical = p_hivpos_empirical / p_hivneg_empirical,
        or_empirical = (p_hivpos_empirical / (1 - p_hivpos_empirical)) / 
            (p_hivneg_empirical / (1 - p_hivneg_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific TB prevalence")
    
```

Bring estimates from study model together to make a forest plot for the manuscript.

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.width = 20, fig.height = 28}

# Join survey labels
names <- cleanHivDF0 %>%
    mutate(id = row_number()) %>%
    select(id, figure.id.yr)  %>%
    mutate(figure.id.yr = ifelse(grepl(", c)" , figure.id.yr), 
                                 "South Africa (Western Cape Province) 2005", 
                                 figure.id.yr))

# Survey-level estimates
survey_sum <- surveyEffectModEst %>%
    group_by(id) %>%
    mean_qi(hivRatio) %>%
    mutate(grouping = "survey") %>%
    left_join(names, by = "id") %>%
    left_join(bactModelHivDF %>% select(id, WHO.region), by = "id")


# Global summary
global_sum <- mainEffectModEst %>%
    ungroup() %>%
    mean_qi(hivRatio) %>%
    mutate(grouping = "Overall",  
           figure.id.yr = "Overall")

# Combine all
forestDF <- bind_rows(survey_sum, global_sum)

# Fill display label and facet group
forestDF <- forestDF %>%
    mutate(figure.id.yr = ifelse(grepl(", c)" , figure.id.yr), 
                                  "South Africa (Western Cape Province) 2005", 
                                 figure.id.yr),
        display_label_raw = ifelse(is.na(figure.id.yr), grouping, figure.id.yr),
        facet_group = case_when(
            grouping == "regional" ~ figure.id.yr,
            grouping == "Overall" ~ "Overall"
        ), 
        type="Posterior mean and 95% credible interval"
    )

# Define facet order
# group_order <- c(
#     "Africa region", 
#     "Americas region",
#     "Eastern Mediterranean region", 
#     "South-East Asia region", 
#     "Western Pacific region", 
#     "Overall"
# )

# Separate surveys and summaries
surveys_only <- forestDF %>%
    filter(grouping == "survey") %>%
    group_by(facet_group) %>%
    arrange(desc(hivRatio), .by_group = TRUE) %>%
    ungroup()

summaries_only <- forestDF %>%
    filter(grouping != "survey")

# Recombine with surveys first, then summaries (preserves facet-wise order)
forestDF_plot <- bind_rows(surveys_only, summaries_only) %>%
    # mutate(facet_group = factor(facet_group, levels = group_order)) %>%
    mutate(display_label =
               case_when(
                   grouping == "Overall" ~ "<b>Overall</b>",
                   grouping == "regional" ~ paste0("<b>", display_label_raw, "</b>"),
                   TRUE ~ display_label_raw
               )) %>%
    mutate(display_label = factor(display_label, levels = rev(unique(display_label))))

# Prepare empirical data
empirical_prev_labels <- left_join(empirical_prev, survey_sum) %>%
    rename(display_label_raw = figure.id.yr) %>%
    select(display_label_raw, or_empirical) %>%
    mutate(
        grouping = "survey",
        display_label = display_label_raw
    ) %>%
    mutate(type = "Ratios calculated from reported prevalence")

###############################################################################
###############################################################################
forestRatio <- ggplot(forestDF_plot) +
    geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
    geom_point(data = empirical_prev_labels,
               aes(x = display_label, y = or_empirical, color=type), alpha=0.75, size = 3) +
    geom_pointinterval(aes(x = display_label, y = hivRatio, 
                           ymin = .lower, ymax = .upper, fill=type),
                       size = 1) +
    coord_flip() +
    # facet_grid(facet_group ~ ., scales = "free", space = "free") +
    scale_color_manual(values = myPal[c(2,1)]) +
    scale_y_log10(
        breaks = c(0.25, 0.5, 1, 2, 4),
        labels = c("0.25", "0.5", "1", "2", "4"),
    ) +
    # scale_x_discrete(labels = label_wrap(30)) + 
    labs(
        x = NULL,
        y = "Risk ratio of bacteriologically-confirmed TB prevalence among PLWH compared to PLWoH"
    ) +
    theme_minimal(base_size = 30)  + 
    theme(axis.line.y = element_blank(),
          axis.ticks.y= element_blank(),
          axis.text.y= element_blank(),
          axis.title.x= element_text(size=18), 
          axis.text.x = element_text(size=14),
          axis.title.y= element_blank(), 
          strip.text.y = element_blank(),
          strip.background = element_blank(),
          legend.position = "bottom", 
          legend.text = element_text(size=16), 
          legend.title=element_blank()); forestRatio

###############################################################################
###############################################################################
###############################################################################
###############################################################################
# nameRow <- forestDF_plot[1,]; nameRow$mfRatio <- "M:F ratio"
# nameRow[,c(1,3:ncol(nameRow))] <- NA

forestDF_plot1 <- forestDF_plot  %>% 
    mutate(hivRatio = paste0(sprintf("%.2f",round(hivRatio,2)), " (", sprintf("%.2f",round(.lower,2)), ", ", sprintf("%.2f",round(.upper, 2)), ")")) #%>%
# 
# forestDF_plot1 <- forestDF_plot1 %>% 
#     bind_rows(nameRow)
#     rbind(nameRow, forestDF_plot1)

p_right <- forestDF_plot1 %>% 
    ggplot(aes(y = display_label))  + 
    # facet_grid(facet_group ~ ., scales = "free", space = "free") +
    geom_text(size=4.5, 
              aes(x = 1, label = hivRatio),
              hjust = 0, 
              fontface = ifelse(forestDF_plot1$figure.id.yr=="Overall", "bold", "plain")
    ) + theme_void(base_size = 16) +
    theme(
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.y = unit(1, "lines"),
        # axis.text.y = ggtext::element_markdown(hjust = 1),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        # plot.background = element_rect(fill = "white"),
        legend.position = "bottom", legend.title=element_blank(),
        legend.spacing.y = unit(-.25, 'cm'),
        legend.box.margin=margin(-10,-10,-6,-10),
        panel.spacing = unit(0.18, "cm")
        # plot.margin = unit(c(.1, .7, 0, .1), "cm")
    ) 

###############################################################################
###############################################################################
###############################################################################
###############################################################################
tempForParts <- cleanHivDF0 %>% left_join(cleanHivDF %>% select(covidence.id, id))

forestDF_plot2 <- forestDF_plot %>% left_join(tempForParts %>% 
                                                  select(id, n.participants.hiv.negative, n.participants.hiv.positive) , by = "id") 


#### Add overall participants 
forestDF_plot2[which(forestDF_plot2$figure.id.yr == "Overall"), 
               "n.participants.hiv.negative"] <- as.numeric(cleanHivDF0 %>% reframe(n.participants.hiv.negative = sum(n.participants.hiv.negative, na.rm=TRUE)) )

forestDF_plot2[which(forestDF_plot2$figure.id.yr == "Overall"), 
               "n.participants.hiv.positive"] <- as.numeric(cleanHivDF0 %>% reframe(n.participants.hiv.positive = sum(n.participants.hiv.positive, na.rm=TRUE)) )


### Add the male ratio  
forestDF_plot2 <- forestDF_plot2 %>%
    mutate(hivpartRatio = paste0( scales::comma(n.participants.hiv.positive), "/",  scales::comma(n.participants.hiv.positive + n.participants.hiv.negative)), 
           hivpartRatio = ifelse(hivpartRatio== "NA/NA", "Not reported", hivpartRatio))


# nameRow2 <- forestDF_plot2[1,]; nameRow2$maleRatio <- "Males (n/N)"
# forestDF_plot2 <- forestDF_plot2  %>% 
#     bind_rows(as.data.frame(nameRow2)) 

p_left <- forestDF_plot2 %>%
    ggplot(aes(y = display_label)) +
    # facet_grid(facet_group ~ ., scales = "free", space = "free") +
    geom_text(size=4.5,
              aes(x = 0, label = hivpartRatio),
              hjust = 0,
              fontface = ifelse(forestDF_plot2$figure.id.yr == "Overall", "bold", "plain")
    ) +
    theme_minimal(base_size = 16) + expand_limits(x=c(0,1))+
    theme(
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.border = element_blank(),
        panel.spacing.y = unit(1, "lines"),
        # ggtext::element_markdown(hjust = 1),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        # plot.background = element_rect(fill = "white"),
        legend.position = "bottom", legend.title=element_blank(),
        legend.spacing.y = unit(-.25, 'cm'),
        legend.box.margin=margin(-10,-10,-6,-10),
        # panel.spacing = unit(0.18, "cm"),
        plot.margin = unit(c(.1, .7, 0, .1), "cm"), 
        axis.text.y = ggtext::element_markdown(hjust = 1), 
        axis.title.y=element_blank()
    ); p_left


layout <- c(
    area(t = 0, l = 0, b = 30, r = 3), # left plot, starts at the top of the page (0) and goes 30 units down and 3 units to the right
    area(t = 1.2, l = 3, b = 30, r = 9), # middle plot starts a little lower (t=1) because there's no title. starts 1 unit right of the left plot (l=4, whereas left plot is r=3), goes to the bottom of the page (30 units), and 6 units further over from the left plot (r=9 whereas left plot is r=3)
    area(t = 0, l = 9, b = 30, r = 11) # right most plot starts at top of page, begins where middle plot ends (l=9, and middle plot is r=9), goes to bottom of page (b=30), and extends two units wide (r=11)
)
# final plot arrangement
finalForestPlot <- p_left + forestRatio+ p_right + plot_layout(design = layout)

finalForestPlot


# Save
ggsave(plot = finalForestPlot, "output/HIV/hiv-forest-plot.png", height = 8, width=14, units = "in")
```
# Secondary pooled estimate

## Model 2: Main effect model with self-reported HIV status 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffecSlfReptMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|study.country) + (1|id),
                       data = bactModelHivDFAll %>% 
                              filter(!covidence.id %in% bactModelHivDF$covidence.id),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 10000,
                       warmup = 2000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffecSlfReptMod)
plot(mainEffecSlfReptMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- bactModelHivDFAll %>% 
      filter(!covidence.id %in% bactModelHivDF$covidence.id) %>%
      select(study.country, LogOdds_HivNeg) %>%
      mutate(logOR_se=1)

mainEffecSlfReptModEst <- add_epred_draws(object = mainEffecSlfReptMod,
                                 newdata=nd,
                                 re_formula = ~ 1, 
                                 allow_new_levels = TRUE) %>% 
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r, echo=FALSE, warning = FALSE, message=FALSE}
mainEffecSlfReptModEst %>% ungroup %>% mean_qi(hivRatio)
```


# Covariate analysis 

## Model 3: Covariate analysis

Add ART coverage into the dataset 

```{r, warning = FALSE, message=FALSE, cache=TRUE}
bactModelHivCovDF <- left_join(bactModelHivDF, artCoverage %>% 
                                   select(study.country,
                                          study.end.year, 
                                          art_coverage)) %>%
                     mutate(art_coverage = art_coverage * 100)
```

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
artEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                           1 + art_coverage + (1|study.country) + (1|id),
                       data = bactModelHivCovDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(student_t(7, 0, 1.5), class = b),
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 10000,
                       warmup = 2000)
```

Summarize and check

```{r, warning = FALSE, message=FALSE, cache=TRUE}
summary(artEffectMod)
plot(artEffectMod)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
bayes_R2(artEffectMod)

### output for contribution table 
round(exp(as.numeric(summary(artEffectMod)[["fixed"]][2,1])),2)
round(exp(as.numeric(summary(artEffectMod)[["fixed"]][2,3])),2)
round(exp(as.numeric(summary(artEffectMod)[["fixed"]][2,4])),2)

### Compare with using spread_draws instead of model summary
artEffectMod %>% 
spread_draws(b_art_coverage) %>% 
mutate(artCovEff = exp(b_art_coverage)) %>% 
select(artCovEff) %>% mean_qi() 

### Consistent! 

### Calculate the difference in risk ratio across the IQR of ART coverage

### A methodological clarification: 
### Note that brms automatically mean centers your predictors and labels the 
### the intercept on this mean centered scale "b_Intercept" and the real-scaled
### intercept "Intercept" which is shown in the summary(artEffectMod); confusingly
### this naming convention is flipped in spread_draws, so we want to use the 
### "b_Intercept" column, NOT the "Intercept" column. 
### If you're in doubt, run artEffectMod %>% spread_draws(Intercept) %>% mean_qi() [1.29]
### and compare the output to summary(artEffectMod)[["fixed"]][2,1]) [1.58]. 
### Pretty large difference; then confirm by running: 
### artEffectMod %>% spread_draws(b_Intercept) %>% mean_qi() [1.59]

artEffectMod %>% spread_draws(b_Intercept, b_art_coverage) %>% 
mutate(rrArtIqrFstQrt = exp((b_art_coverage*
                             quantile(bactModelHivCovDF$art_coverage, 
                                      probs=c(0.25))) + 
                             b_Intercept), 
       rrArtIqrThrdQrt = exp((b_art_coverage*
                             quantile(bactModelHivCovDF$art_coverage,
                                      probs=c(0.75))) +
                             b_Intercept),
      percChng = (rrArtIqrFstQrt - rrArtIqrThrdQrt) / rrArtIqrFstQrt *100) %>%
select(rrArtIqrFstQrt, rrArtIqrThrdQrt, percChng) %>% 
mean_qi() %>% 
mutate(across(rrArtIqrFstQrt:percChng.upper, 
                                round, 2)) %>% t()

```

# Notification analysis 

Below we use WHO Reported TB Notifications and UNAIDS population estimates to calculate TB notifications rates disaggregated by HIV status. We then estimate the pooled notification rate ratio between PLWH and PLWoH. 

Set up notification data

```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Population with HIV data from UNAIDS
pop <- artCoverage

### Read in TB notification data from WHO
notif0 <- read.csv(here("data/TB_notifications_2025-07-18.csv"))

### Calculate the HIV-status stratified TB notification 
### rate and transform to log odds space
notif0 <- notif0 %>% 
          ## Add 2007 because 2006 data for Zimbabwe does not exist
          filter(year %in% c(unique(bactModelHivDF$study.end.year), 2007),  
                 country %in% unique(bactModelHivDF$study.country)) %>% 
          select(country, g_whoregion, 
                year, 
                newrel_m15plus, 
                newrel_f15plus,
                newrel_hivpos_014,
                new_sp, new_sn, 
                newrel_hivpos, 
                hivtest_pos, 
                hivtest) %>%
         rename(study.country = country, 
                study.end.year = year) %>%
         left_join(pop, by=c("study.country", "study.end.year")) %>%
         mutate(notifRateHivPos = ifelse(study.end.year < 2019, 
                                     hivtest_pos / hivpop, 
                                     ifelse(is.na(newrel_hivpos_014), 
                                            newrel_hivpos / hivpop,
                                           (newrel_hivpos - newrel_hivpos_014)/hivpop)), 
                notifRateHivNeg = ifelse(study.end.year == 2014, 
                                     (newrel_m15plus + newrel_f15plus - hivtest_pos) / (totpop - hivpop), 
                                     ifelse(study.end.year < 2014, 
                                     (new_sp + new_sn - hivtest_pos) / (totpop-hivpop), 
                                     ifelse(is.na(newrel_hivpos_014), 
                                            (newrel_m15plus + newrel_f15plus - newrel_hivpos) / (totpop-hivpop), 
                                            (newrel_m15plus + newrel_f15plus - newrel_hivpos - newrel_hivpos_014) / (totpop-hivpop)))), 
                notifRateHivPos = ifelse(study.country=="Uganda" & study.end.year == 2015,
                                         newrel_hivpos / hivpop,
                                         notifRateHivPos),
                notifRateHivNeg = ifelse(study.country=="Cambodia" & study.end.year == 2013,
                                         (newrel_m15plus + newrel_f15plus - hivtest_pos) / (totpop-hivpop), 
                                         notifRateHivNeg)) %>%
         pivot_longer(cols=starts_with("notifRate"),
                      names_to = "HIVStatus",
                      names_prefix = "notifRate",
                      values_to = "notifRate",
                      values_drop_na = TRUE) %>%
        select(study.country, study.end.year, `HIVStatus`, notifRate, totpop, hivpop) %>%
        mutate(notifRate = ifelse(notifRate==0, 
                                  NA, notifRate), 
               "Standard Error" = case_when(HIVStatus == "HivPos" ~ 
                                                sqrt((notifRate*(1-notifRate))/hivpop), 
                                            HIVStatus == "HivNeg" ~ 
                                                sqrt((notifRate*(1-notifRate))/(totpop-hivpop))),
               "LogOdds_notif" = log(notifRate/(1-notifRate)),
               "LogOddsStandardError_notif" = case_when(HIVStatus == "HivPos" ~ 
                                                  (1/(hivpop*notifRate*(1-notifRate)))^2,
                                                        HIVStatus == "HivNeg" ~ 
                                                  (1/((totpop-hivpop)*notifRate*(1-notifRate)))^2))
       
### Widen and calculate the log odds ratio and associated standard error
notif_ratio  <- notif0 %>%
          select(study.country, study.end.year, HIVStatus, 
                 LogOdds_notif, LogOddsStandardError_notif) %>%
          pivot_wider(names_from = HIVStatus,
                     values_from = c(LogOdds_notif, 
                                     LogOddsStandardError_notif),
                     names_sep = "_") %>%
          mutate(logOR_notif = LogOdds_notif_HivPos - LogOdds_notif_HivNeg,
                 logOR_se_notif = sqrt(LogOddsStandardError_notif_HivPos^2 + LogOddsStandardError_notif_HivNeg^2)) %>%
          right_join(bactModelHivDF %>% 
                     filter(natRep == "Yes") %>%
                     mutate(study.end.year = ifelse(study.end.year == 2006, 
                                                    2007, study.end.year)),
                                
                                # recode(study.end.year, `2006` = 2007)),
                     by=c("study.country", "study.end.year"))
```

## Model 4: TB notifications by country

Model specification

```{r, warning=FALSE}
# notification rate ratio model:
notif_ratio_model <- bf(
  logOR_notif | se(logOR_se_notif, sigma = TRUE) ~ 1  + (1 | p | study.country)
)
 
notif_model_res <- brm(
                      formula = notif_ratio_model,
                      data = notif_ratio,
                      family = gaussian(), 
                      # sample_prior = "only",
                      prior = c(
                        prior(student_t(7, 0, 1.5), class = Intercept),
                        prior(exponential(2), class = sd)
                      ),
                      control = list(adapt_delta = 0.99),
                      chains = 4,
                      cores = 4,
                      iter = 10000,
                      warmup = 2000)

```

Summarize and check

```{r, warnings=FALSE}
summary(notif_model_res)
plot(notif_model_res)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd_n <- notif_ratio %>% 
        unique() %>%
        mutate(logOR_se_notif=1)

mainEffectNotifEst <- add_epred_draws(object = notif_model_res,
                                      newdata=nd_n,
                                      re_formula =   ~ 1 + (1 | study.country),
                                      allow_new_levels = TRUE) %>%
                      mutate(model_notif = .epred,
                             type="Posterior mean and 95% credible interval")


mainEffectNotifEstSummary <- mainEffectNotifEst %>% 
                             ungroup() %>%
                             mean_qi(model_notif, na.rm=T) %>% 
                             mutate(across(c(model_notif, .lower, .upper),
                                           round, 2)); mainEffectNotifEstSummary

mainEffectNotifEstCountrySummary <- mainEffectNotifEst %>%
                                    group_by(study.country) %>%
                                    mean_qi(model_notif, na.rm=T)  %>% 
                                    mutate(across(c(model_notif, .lower, .upper),
                                           round, 2)); mainEffectNotifEstCountrySummary

datPlot <- mainEffectNotifEstCountrySummary %>% 
           select(study.country, model_notif) %>% 
           left_join(notif %>% select(study.country, logOR_notif)) 

ggplot(data=datPlot) + 
    geom_point(aes(x=study.country, y=exp(model_notif)), size=3) + 
    geom_point(aes(x=study.country, y=exp(logOR_notif)), color="orchid2", size=3) + 
    theme_minimal()
```

# Prevalence to notification comparison 

Use a multivariate Bayesian multi-level model to jointly estimate notification and prevalence to account for the uncertainty and potential correlation. Use these estimates to calculate the prevalence to notification ratio for PLWH and PLWoH. 

Setup data 

```{r, echo=FALSE, warning=FALSE}
notif <- rbind(notif_ratio %>% 
                 select(study.country, 
                        LogOdds_HivNeg, 
                        LogOddsStandardError_HivNeg,
                        LogOdds_notif_HivNeg,
                        LogOddsStandardError_notif_HivNeg, 
                        study.end.year) %>%
                  mutate("HIVstatus" = "HivNeg") %>%
                  rename(LogOddsPrev = LogOdds_HivNeg, 
                         LogOddsStandardErrorPrev = LogOddsStandardError_HivNeg, 
                         LogOddsNotif = LogOdds_notif_HivNeg,
                         LogOddsStandardErrorNotif = LogOddsStandardError_notif_HivNeg),
                 notif_ratio %>% 
                 select(study.country, 
                        LogOdds_HivPos, 
                        LogOddsStandardError_HivPos,
                        LogOdds_notif_HivPos,
                        LogOddsStandardError_notif_HivPos, 
                        study.end.year) %>%
                  mutate("HIVstatus" = "HivPos") %>%
                  rename(LogOddsPrev = LogOdds_HivPos, 
                         LogOddsStandardErrorPrev = LogOddsStandardError_HivPos, 
                         LogOddsNotif = LogOdds_notif_HivPos,
                         LogOddsStandardErrorNotif = LogOddsStandardError_notif_HivPos))
```


## Model 5: Multivariate model of TB prevalence risk and notification rate

Model specification

```{r, echo=FALSE, warning = FALSE, message=FALSE}
###############################################################################
### Run a multivariate model
#prevalence model:
prev_model <- bf(
  LogOddsPrev | se(LogOddsStandardErrorPrev, sigma = TRUE) ~ 1 + HIVstatus + (1 | p | study.country)
)
 
#notifications model:
notif_model <- bf(
  LogOddsNotif | se(LogOddsStandardErrorNotif, sigma = TRUE) ~ 1  + HIVstatus + (1 | p | study.country)
)

#fit the model
prev_notif_model <- brm(
                    formula = prev_model + notif_model + set_rescor(FALSE),
                    data = notif,
                    family = gaussian(), 
                    # sample_prior = "only",
                    prior = c(
                          prior(student_t(7, 0, 1.5), class = Intercept, resp = "LogOddsPrev"),
                          prior(student_t(7, 0, 1.5), class = Intercept, resp = "LogOddsNotif"),
                          prior(student_t(7, 0, 1.5), class = b, resp = "LogOddsPrev"),
                          prior(student_t(7, 0, 1.5), class = b, resp = "LogOddsNotif"),
                          prior(exponential(1), class = sd, resp = "LogOddsPrev"),
                          prior(exponential(1), class = sd, resp = "LogOddsNotif")
                    ),
                    control = list(adapt_delta = 0.99),
                    chains = 4,
                    cores = 4,
                    iter = 10000, 
                    warmup = 2000
)
```

Summarize and check

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(prev_notif_model)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
prev_notif_est <-  prev_notif_model %>% 
                   add_epred_draws(newdata = crossing(study.country=unique(notif$study.country), 
                                                      HIVstatus = c("HivNeg", "HivPos"), 
                                                      LogOddsStandardErrorPrev = 1,
                                                      LogOddsStandardErrorNotif = 1),
                                   re_formula =  ~ 1 + (1 | study.country)) %>%
                   ungroup() %>% 
                   dplyr::select(study.country, .category, HIVstatus, .draw, .epred) %>%
                   pivot_wider(names_from = .category, values_from = .epred) %>%
                  # Note we use expit here to recover the probability; 
                  # if we use exp we would recover the odds ratio. 
                   mutate(npRatio = expit(LogOddsNotif - LogOddsPrev)) 

prev_notif_estSummary <- prev_notif_est %>%
                         group_by(HIVstatus) %>%
                         mean_qi(npRatio, na.rm=T) %>% 
                         mutate(across(c(npRatio, .lower, .upper), 
                                round, 2)); prev_notif_estSummary
```

```{r}
prev_notif_estSummary_country <- prev_notif_est %>%
                                 group_by(HIVstatus, study.country) %>%
                                 mean_qi(npRatio, na.rm=T) %>% 
                                 mutate(across(c(npRatio, .lower, .upper), 
                                        round, 2)); prev_notif_estSummary_country

### Overall HP:HN ratio of P:N ratios 
hphn_np_estSummary <- (prev_notif_est %>%
                      filter(HIVstatus == "HivPos") %>%
                      select(npRatio) /
                      prev_notif_est %>%
                      filter(HIVstatus == "HivNeg") %>%
                      select(npRatio)) %>% 
                      rename(ratioOfRatios = npRatio)

### Calculate the ratios of the ratios
hphn_pn_estSummary <- hphn_np_estSummary %>%
                      mean_qi(ratioOfRatios, na.rm=T) %>% 
                      mutate(across(c(ratioOfRatios, .lower, .upper), 
                             round, 2)); hphn_pn_estSummary

```

## Model 6: Covariate analysis with prevalence to notification ratios (no interaction)

Model specification
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Prevalence model with ART coverage
prev_model_covariate <- bf(
  LogOddsPrev | se(LogOddsStandardErrorPrev, sigma = TRUE) ~ 1 + art_coverage + HIVstatus + (1 | p | study.country)
)

# Notification model with ART coverage
notif_model_covariate <- bf(
  LogOddsNotif | se(LogOddsStandardErrorNotif, sigma = TRUE) ~ 1  + art_coverage + HIVstatus + (1 | p | study.country)
)

#fit the model
p_to_n_hiv_covariate_model <- brm(
  formula = prev_model_covariate + notif_model_covariate + set_rescor(FALSE),
  data = notif %>% 
         left_join(bactModelHivCovDF %>% 
         select(study.country, study.end.year, art_coverage)) %>%
         mutate(art_coverage = art_coverage*10),
  family = gaussian(), 
  # sample_prior = "only",
  prior = c(
      prior(student_t(7, 0, 1.5), class = Intercept, resp = "LogOddsPrev"),
      prior(student_t(7, 0, 1.5), class = Intercept, resp = "LogOddsNotif"),
      prior(student_t(7, 0, 1.5), class = b, resp = "LogOddsPrev"),
      prior(student_t(7, 0, 1.5), class = b, resp = "LogOddsNotif"),
      prior(exponential(1), class = sd, resp = "LogOddsPrev"),
      prior(exponential(1), class = sd, resp = "LogOddsNotif")
  ),
  control = list(adapt_delta = 0.99),
  chains = 4,
  cores = 4,
  iter = 10000, 
  warmup = 2000
)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}

nd_pn_cov <- notif %>% 
             left_join(bactModelHivCovDF %>% 
                        select(study.country, 
                               study.end.year, 
                               art_coverage)) %>% 
             mutate(LogOddsStandardErrorPrev = 1,
                    LogOddsStandardErrorNotif =1)
prev_notif_cov_est <-  p_to_n_hiv_covariate_model %>% 
                       add_epred_draws(newdata = nd_pn_cov,
                                   re_formula =  ~ 1 + (1 | study.country)) %>%
                       ungroup() %>% 
                       dplyr::select(study.country, .category, HIVstatus, .draw, .epred) %>%
                       pivot_wider(names_from = .category, values_from = .epred) %>%
                      # Note we use expit here to recover the probability; 
                      # if we use exp we would recover the odds ratio. 
                       mutate(npRatio = expit(LogOddsNotif - LogOddsPrev)) 

prev_notif_cov_estSummary <- prev_notif_cov_est %>%
                         group_by(HIVstatus) %>%
                         mean_qi(npRatio, na.rm=T) %>% 
                         mutate(across(c(npRatio, .lower, .upper), 
                                round, 2)); prev_notif_cov_estSummary
```

## Model 7: Covariate analysis with prevalence to notification ratios (with interaction)

Model specification
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Prevalence model with ART coverage
prev_model_covariate_intx <- bf(
  LogOddsPrev | se(LogOddsStandardErrorPrev, sigma = TRUE) ~ 1 + HIVstatus*art_coverage + (1 | p | study.country)
)

# Notification model with ART coverage
notif_model_covariate_intx <- bf(
  LogOddsNotif | se(LogOddsStandardErrorNotif, sigma = TRUE) ~ 1  + HIVstatus*art_coverage + (1 | p | study.country)
)

#fit the model
p_to_n_hiv_covariate_model_intx <- brm(
  formula = prev_model_covariate_intx + notif_model_covariate_intx + set_rescor(FALSE),
  data = notif %>% 
         left_join(bactModelHivCovDF %>% 
         select(study.country, study.end.year, art_coverage)),
  family = gaussian(), 
  # sample_prior = "only",
  prior = c(
      prior(student_t(7, 0, 1.5), class = Intercept, resp = "LogOddsPrev"),
      prior(student_t(7, 0, 1.5), class = Intercept, resp = "LogOddsNotif"),
      prior(student_t(7, 0, 1.5), class = b, resp = "LogOddsPrev"),
      prior(student_t(7, 0, 1.5), class = b, resp = "LogOddsNotif"),
      prior(exponential(1), class = sd, resp = "LogOddsPrev"),
      prior(exponential(1), class = sd, resp = "LogOddsNotif")
  ),
  control = list(adapt_delta = 0.99),
  chains = 4,
  cores = 4,
  iter = 10000, 
  warmup = 2000
)
```


Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}

prev_notif_cov_intx_est <-  p_to_n_hiv_covariate_model_intx %>% 
                       add_epred_draws(newdata = nd_pn_cov,
                                   re_formula =  ~ 1 + (1 | study.country)) %>%
                       ungroup() %>% 
                       dplyr::select(study.country, .category, HIVstatus, .draw, .epred) %>%
                       pivot_wider(names_from = .category, values_from = .epred) %>%
                      # Note we use expit here to recover the probability; 
                      # if we use exp we would recover the odds ratio. 
                       mutate(npRatio = expit(LogOddsNotif - LogOddsPrev)) 

prev_notif_cov_intx_estSummary <- prev_notif_cov_intx_est %>%
                         group_by(HIVstatus) %>%
                         mean_qi(npRatio, na.rm=T) %>% 
                         mutate(across(c(npRatio, .lower, .upper), 
                                round, 2)); prev_notif_cov_intx_estSummary
```


Impact of ART without interaction model

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(p_to_n_hiv_covariate_model)

meanARTCoverage <- mean(nd_pn_cov$art_coverage)
                   
#### Check that this is approximately the same as above
#### HIV positive notification to prevalence ratio
round(expit(
             #HIV positive notification rate with ART coverage at observed mean 
             (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][2,1]) + 
              as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][6,1]) + 
              (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][5,1])*meanARTCoverage)) -
             #HIV positive prevalence risk with ART coverage at observed mean 
             (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][1,1]) + 
              as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][4,1]) + 
              (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][3,1])*meanARTCoverage))
             ) ,2)


artCovHivPos <- expit(
                 #HIV positive notification rate with ART coverage 
                 (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][2,1]) + 
                  as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][6,1]) + 
                  (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][5,1])*unique(nd_pn_cov$art_coverage))) -
                 #HIV positive prevalence risk with ART coverage
                 (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][1,1]) + 
                  as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][4,1]) + 
                  (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][3,1])*unique(nd_pn_cov$art_coverage)))
             )

artCovHivNeg <- expit(
                 #HIV negative notification rate with ART coverage 
                 (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][2,1]) + 
                  (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][5,1])*unique(nd_pn_cov$art_coverage))) -
                 #HIV negative prevalence risk with ART coverage
                 (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][1,1]) + 
                  (as.numeric(summary(p_to_n_hiv_covariate_model)[["fixed"]][3,1])*unique(nd_pn_cov$art_coverage)))
                 ) 


npCovDF <- data.frame(npCovHivPos = artCovHivPos, 
                      npCovHivNeg = artCovHivNeg, 
                      artCov = unique(nd_pn_cov$art_coverage))

plot(npCovDF$npCovHivPos, npCovDF$artCov)


lm(npCovHivPos ~ artCov, 
   data=npCovDF)

lm(npCovHivNeg ~ artCov, 
   data=npCovDF)

```


Impact of ART with interaction model

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(p_to_n_hiv_covariate_model_intx)

meanARTCoverage <- mean(nd_pn_cov$art_coverage)
                   
#### Check that this is approximately the same as above
#### HIV positive notification to prevalence ratio
round(expit(
             #HIV positive notification rate with ART coverage at observed mean 
             (as.numeric(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][2,1]) + 
              as.numeric(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][6,1]) + 
              sum(as.numeric(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][7:8,1])*meanARTCoverage)) -
             #HIV positive prevalence risk with ART coverage at observed mean 
             (as.numeric(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][1,1]) + 
              as.numeric(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][3,1]) + 
              sum(as.numeric(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][4:5,1])*meanARTCoverage))
             ) ,2)


artCovIntxHivNeg <-expit(
             #HIV positive notification rate with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_Intercept","Estimate"] + 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_art_coverage","Estimate"])*unique(nd_pn_cov$art_coverage)) -
             #HIV positive prevalence risk with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_Intercept","Estimate"] + 
              (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_art_coverage","Estimate"])*unique(nd_pn_cov$art_coverage))
             )

artCovIntxHivPos <-expit(
             #HIV positive notification rate with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_Intercept","Estimate"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_HIVstatusHivPos","Estimate"] + 
              sum(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][7:8,"Estimate"])*unique(nd_pn_cov$art_coverage)) -
             #HIV positive prevalence risk with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_Intercept","Estimate"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_HIVstatusHivPos","Estimate"] + 
              sum(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][4:5,"Estimate"])*unique(nd_pn_cov$art_coverage))
             )

npCovIntxDF <- data.frame(npCovHivPos = artCovIntxHivPos, 
                          npCovHivNeg = artCovIntxHivNeg, 
                          artCov = unique(nd_pn_cov$art_coverage))

plot(npCovIntxDF$npCovHivPos, npCovIntxDF$artCov)

### Upper and lower bounds of the interaction 
npCovIntxDF$npCovHivPosLB <- expit(
             #HIV positive notification rate with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_Intercept","l-95% CI"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_HIVstatusHivPos","l-95% CI"] + 
              sum(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][7:8,"l-95% CI"])*unique(nd_pn_cov$art_coverage)) -
             #HIV positive prevalence risk with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_Intercept","l-95% CI"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_HIVstatusHivPos","l-95% CI"] + 
              sum(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][4:5,"l-95% CI"])*unique(nd_pn_cov$art_coverage))
             )

npCovIntxDF$npCovHivPosUB <- expit(
             #HIV positive notification rate with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_Intercept","u-95% CI"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_HIVstatusHivPos","u-95% CI"] + 
              sum(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][7:8,"u-95% CI"])*unique(nd_pn_cov$art_coverage)) -
             #HIV positive prevalence risk with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_Intercept","u-95% CI"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_HIVstatusHivPos","u-95% CI"] + 
              sum(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][4:5,"u-95% CI"])*unique(nd_pn_cov$art_coverage))
             )


npCovIntxDF$npCovHivNegLB <- expit(
             #HIV positive notification rate with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_Intercept","l-95% CI"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_art_coverage","l-95% CI"]*unique(nd_pn_cov$art_coverage)) -
             #HIV positive prevalence risk with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_Intercept","l-95% CI"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_art_coverage","l-95% CI"]*unique(nd_pn_cov$art_coverage))
             )

npCovIntxDF$npCovHivNegUB <- expit(
             #HIV positive notification rate with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_Intercept","u-95% CI"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_art_coverage","u-95% CI"]*unique(nd_pn_cov$art_coverage)) -
             #HIV positive prevalence risk with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_Intercept","u-95% CI"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_art_coverage","u-95% CI"]*unique(nd_pn_cov$art_coverage))
             )

print(paste0("Multiplicative impact of ART on HIV positive notification to prevalence ratio: ",
            round(lm(npCovHivPos ~ artCov, 
               data=npCovIntxDF)$coefficient["artCov"],4), 
            " (", 
            round(lm(npCovHivPosLB ~ artCov, 
               data=npCovIntxDF)$coefficient["artCov"],4),
            ", ",
            round(lm(npCovHivPosUB ~ artCov, 
               data=npCovIntxDF)$coefficient["artCov"],4),
            ")"))


print(paste0("Multiplicative impact of ART on HIV negative notification to prevalence ratio: ",
            round(lm(npCovHivNeg ~ artCov, 
               data=npCovIntxDF)$coefficient["artCov"],4), 
            " (", 
            round(lm(npCovHivNegLB ~ artCov, 
               data=npCovIntxDF)$coefficient["artCov"],4),
            ", ",
            round(lm(npCovHivNegUB ~ artCov, 
               data=npCovIntxDF)$coefficient["artCov"],4),
            ")"))
```

```{r}
get_variables(p_to_n_hiv_covariate_model_intx)

artCovIntxHivPos <-expit(
             #HIV positive notification rate with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_Intercept","Estimate"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsNotif_HIVstatusHivPos","Estimate"] + 
              sum(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][7:8,"Estimate"])*unique(nd_pn_cov$art_coverage)) -
             #HIV positive prevalence risk with ART coverage 
             (summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_Intercept","Estimate"] + 
              summary(p_to_n_hiv_covariate_model_intx)[["fixed"]]["LogOddsPrev_HIVstatusHivPos","Estimate"] + 
              sum(summary(p_to_n_hiv_covariate_model_intx)[["fixed"]][4:5,"Estimate"])*unique(nd_pn_cov$art_coverage))
             )

artFrstQrt <- quantile(bactModelHivCovDF$art_coverage, 
                                      probs=c(0.25))
artThrdQrt <- quantile(bactModelHivCovDF$art_coverage, 
                                      probs=c(0.75))
p_to_n_hiv_covariate_model_intx %>% 
spread_draws(b_LogOddsPrev_Intercept,
             b_LogOddsNotif_Intercept,
             b_LogOddsPrev_HIVstatusHivPos,
             b_LogOddsNotif_HIVstatusHivPos,
             b_LogOddsPrev_art_coverage,
             b_LogOddsNotif_art_coverage,
             `b_LogOddsPrev_HIVstatusHivPos:art_coverage`,
             `b_LogOddsNotif_HIVstatusHivPos:art_coverage`
             ) %>% 
mutate(rrArtIqrFrstQrtHIVp = expit((b_LogOddsNotif_Intercept + 
                                    b_LogOddsNotif_HIVstatusHivPos +
                                    sum(b_LogOddsNotif_art_coverage*artFrstQrt,
                                        `b_LogOddsNotif_HIVstatusHivPos:art_coverage`*artFrstQrt)) /
                                    (b_LogOddsPrev_Intercept + 
                                    b_LogOddsPrev_HIVstatusHivPos +
                                    sum(b_LogOddsPrev_art_coverage*artFrstQrt,
                                        `b_LogOddsPrev_HIVstatusHivPos:art_coverage`*artFrstQrt))), 
       rrArtIqrThrdQrtHIVp = expit((b_LogOddsNotif_Intercept + 
                                    b_LogOddsNotif_HIVstatusHivPos +
                                    sum(b_LogOddsNotif_art_coverage*artThrdQrt,
                                        `b_LogOddsNotif_HIVstatusHivPos:art_coverage`*artThrdQrt)) /
                                    (b_LogOddsPrev_Intercept + 
                                     b_LogOddsPrev_HIVstatusHivPos +
                                     sum(b_LogOddsPrev_art_coverage*artThrdQrt,
                                    `b_LogOddsPrev_HIVstatusHivPos:art_coverage`*artThrdQrt))), 
       rrArtIqrFrstQrtHIVn = expit((b_LogOddsNotif_Intercept + 
                                    b_LogOddsNotif_art_coverage*artFrstQrt) /
                                    (b_LogOddsPrev_Intercept + 
                                    b_LogOddsPrev_art_coverage*artFrstQrt)), 
       rrArtIqrThrdQrtHIVn = expit((b_LogOddsNotif_Intercept + 
                                    b_LogOddsNotif_art_coverage*artThrdQrt) /
                                    (b_LogOddsPrev_Intercept + 
                                    b_LogOddsPrev_art_coverage*artThrdQrt)),
       
      percChngHIVn = (rrArtIqrFrstQrtHIVn - rrArtIqrThrdQrtHIVn) / rrArtIqrFrstQrtHIVn *100,
      percChngHIVp = (rrArtIqrFrstQrtHIVp - rrArtIqrThrdQrtHIVp) / rrArtIqrFrstQrtHIVp *100) %>%
select(rrArtIqrFrstQrtHIVp, rrArtIqrThrdQrtHIVp, percChngHIVp) %>% 
mean_qi() %>% 
mutate(across(rrArtIqrFrstQrtHIVp:percChngHIVp.upper,
                                round, 5)) %>%
t()
```

# Smear positive analysis 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
cleanHivSmrDF0 <- smrPosIndicator("hiv") %>% 
                  ## Join in the HIV prevalence data
                  left_join(artCoverage %>% 
                            select(study.country, study.end.year, hivPrev)) %>% 
                  mutate(n.participants.hiv.positive = ifelse(covidence.id == 26947, 
                                                              round(n.participants.sex.total * hivPrev), 
                                                              n.participants.hiv.positive),
                         n.participants.hiv.negative = ifelse(covidence.id == 26947, 
                                                              n.participants.sex.total-n.participants.hiv.positive, 
                                                              n.participants.hiv.negative), 
                         hiv.smr.analysis.indicator = ifelse(covidence.id == 26947, 
                                                          "adj.prev100k.smear.positive.tb", 
                                                          hiv.smr.analysis.indicator)
                         ) %>% 
                  filter(hiv.smr.analysis.indicator !="none") 

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanHivSmrDF <- cleanHivSmrDF0 %>% 
                 # filter(hiv.smr.analysis.indicator != "none") %>% 
                 mutate("id" = row_number()) %>%
                 ### Scale all prevalence estimates back to decimals 
                 mutate_at(vars(matches('prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>% 
                 select(id,
                       covidence.id, 
                       figure.id.yr, 
                       study.geography, 
                       title.extracted,
                       study.country,
                       WHO.region,
                       study.start.year, 
                       study.end.year, 
                       n.participants.sex.total,
                       ends_with("hiv.positive"), ends_with("hiv.negative"), ends_with("hiv.total"),
                       hiv.smr.analysis.indicator, 
                       hiv.status.definition) %>%
    
                 mutate("tempPrevHivPos" =  case_when(hiv.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.positive,
                                            hiv.smr.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive, 
                                            hiv.smr.analysis.indicator == 
                                            "adj.prev100k.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.positive,
                                            hiv.smr.analysis.indicator == 
                                            "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.positive/n.participants.hiv.positive,
                                            hiv.smr.analysis.indicator == 
                                            "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.positive),
                        "tempPrevHivNeg" = case_when(hiv.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.negative,
                                           hiv.smr.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative, 
                                           hiv.smr.analysis.indicator == 
                                           "adj.prev100k.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.hiv.negative,
                                           hiv.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.hiv.negative/n.participants.hiv.negative,
                                           hiv.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.hiv.negative)) %>%
    pivot_longer(cols = c(tempPrevHivNeg,tempPrevHivPos), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                                                    Sex == "HivPos" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivNeg" ~ (adj.prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                       
                                        hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivPos" ~ (prev100k.ci.upper.smear.positive.tb.hiv.positive-
                                                                 prev100k.ci.lower.smear.positive.tb.hiv.positive)/3.92,
                                        hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "HivNeg" ~ (prev100k.ci.upper.smear.positive.tb.hiv.negative-
                                                                   prev100k.ci.lower.smear.positive.tb.hiv.negative)/3.92,
                                        hiv.smr.analysis.indicator %in% c("n.smear.positive.tb","prev100k.smear.positive.tb", 
                                                                          "adj.prev100k.smear.positive.tb") & Sex == "HivPos" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.positive),
                                        hiv.smr.analysis.indicator  %in% c("n.smear.positive.tb","prev100k.smear.positive.tb", 
                                                                          "adj.prev100k.smear.positive.tb") & Sex == "HivNeg" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.hiv.negative)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "HivPos" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                              hiv.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "HivNeg" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92,
                                                 hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "HivPos" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.positive/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.hiv.positive)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.hiv.positive /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.hiv.positive)))/3.92,
                                                 hiv.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "HivNeg" ~ (log(prev100k.ci.upper.smear.positive.tb.hiv.negative/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.hiv.negative)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.hiv.negative /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.hiv.negative)))/3.92, 
                                            hiv.smr.analysis.indicator %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb", 
                                                                          "adj.prev100k.smear.positive.tb") & Sex == "HivPos" ~ 
                                            log(1/(n.participants.hiv.positive*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        hiv.smr.analysis.indicator  %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb", 
                                                                          "adj.prev100k.smear.positive.tb") & Sex == "HivNeg" ~ 
                                            log(1/(n.participants.hiv.negative*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       # year_z = ifelse(!is.na(study.end.year), 
                                       #                 study.end.year - round(mean(study.end.year, na.rm=TRUE))
                                       #                 (publication.year - meanLag) - round(mean((publication.year - meanLag), na.rm=TRUE))),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))  #%>% filter(LogOddsStandardError < Inf)

smrModelAllDF <- cleanHivSmrDF %>%
    select(covidence.id, id, WHO.region, study.country, Sex, natRep, LogOdds, LogOddsStandardError, 
           n.participants.sex.total, 
           n.participants.hiv.positive, n.participants.hiv.negative, hiv.status.definition, title.extracted) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_HivPos - LogOdds_HivNeg,
        logOR_se = sqrt(LogOddsStandardError_HivPos^2 + LogOddsStandardError_HivNeg^2) 
    ) 

smrModelDF <- smrModelAllDF %>% 
    filter(hiv.status.definition != "Self-report")
```

## Smear positive descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(smrModelDF)[1], "unique surveys with sex-stratified smear positive TB"))

# sort(unique(cleanHivDF0$title.extracted))
print(paste(length(unique(smrModelDF$title.extracted)), "unique publications"))

print(paste(sum(!is.na(smrModelDF$n.participants.sex.total)), "studies report participant numbers totaling", 
            sum(smrModelDF$n.participants.sex.total, na.rm = TRUE), "participants"))

print(paste(sum(!is.na(smrModelDF$n.participants.hiv.negative)), "studies report number of persons knowing their HIV status totaling", 
            sum(smrModelDF$n.participants.hiv.negative, smrModelDF$n.participants.hiv.positive, na.rm = TRUE), "participants"))

print(paste0(round(sum(smrModelDF$n.participants.hiv.positive, na.rm=TRUE)/sum(smrModelDF$n.participants.hiv.negative, smrModelDF$n.participants.hiv.positive, na.rm = TRUE)*100, 2), "% of participants were hivpos" ))
```

## More descriptives 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
table(smrModelDF$WHO.region)
smrModelDF %>% reframe(Participants = sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE), .by=WHO.region)
smrModelDF %>% reframe(PercHivPos = sum(n.participants.hiv.positive, na.rm=TRUE)/sum(n.participants.hiv.negative + n.participants.hiv.positive, na.rm=TRUE)*100, .by=WHO.region)
```

# Secondary pooled estimate of smear-positive prevalence risk ratio

## Model 8: Smear positive main effect model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                                        1  + (1 | study.country) + (1|id),
                              data = smrModelDF,
                              prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                                      # prior(normal(0, 10), class = b) +
                                      prior(exponential(2), class = sd),
                              family = gaussian(),
                              control = list(adapt_delta = 0.99),
                              chains = 4,
                              cores = 4,
                              iter = 10000,
                              warmup = 2000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectSmrMod)
plot(mainEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
nd <- smrModelDF %>%
      select(study.country, LogOdds_HivNeg) %>% 
      mutate(logOR_se=1)


mainEffectSmrModEst <- add_epred_draws(object = mainEffectSmrMod,
                                 newdata=nd,
                                 re_formula = ~ 1) %>% 
                    mutate(hivRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate

```{r, echo=FALSE, warning = FALSE, message=FALSE}
mainEffectSmrModEst %>% ungroup %>% mean_qi(hivRatio)
```

Survey specific estimates

```{r, echo=FALSE, warning = FALSE, message=FALSE}
surveyEffectModEst <- add_epred_draws(object = mainEffectSmrMod,
                                      newdata=smrModelDF,
                                      re_formula = ~ 1  + (1|study.country) + (1|id),
                                      allow_new_levels = TRUE) %>%
                      mutate(hivRatio = exp(.epred),
                             type="Posterior mean and 95% credible interval")

surveyEffectModEst %>% 
mean_qi(hivRatio) %>% 
select(id, study.country, hivRatio, .lower, .upper)
```

## Comparison of bacteriologically confirmed estimates from studies above
## Model 9: Bacteriologically positive prevalence risk ratio main effect model fit to only surveys that report smear-positive TB prevalence. 

Model specification

```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectBCSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | study.country) + (1|id),
                       data = bactModelHivDF %>% 
                              filter(covidence.id %in% smrModelDF$covidence.id),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 10000,
                       warmup = 2000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(mainEffectBCSmrMod)
plot(mainEffectBCSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
nd <- smrModelDF %>% 
      filter(covidence.id %in% smrModelDF$covidence.id) %>%
      select(study.country, LogOdds_HivNeg) %>%
      mutate(logOR_se=1) 


mainEffectBCSmrModEst <- add_epred_draws(object = mainEffectBCSmrMod,
                                         newdata=nd,
                                         re_formula = ~ 1) %>% 
                         mutate(hivRatio = exp(.epred),
                                type="Posterior mean and 95% credible interval") 
```

Main estimate

``````{r, echo=FALSE, warning = FALSE, message=FALSE}
mainEffectBCSmrModEst %>% ungroup %>% mean_qi(hivRatio)
```

# Risk of bias analysis

```{r, echo=FALSE, warning = FALSE, message=FALSE}
##############################################################################|
##### RISK OF BIAS SUMMARY ####################################################
##############################################################################|
### Standardize the study quality/risk of bias data 
riskData <- cleanHivDF0 %>% 
    filter(covidence.id %in% bactModelHivDF$covidence.id) %>%
    select(contains("study.quality"), study.geography) %>%
    select(! "study.quality.comments") %>%
    mutate("study.quality.representative" = ifelse(grepl("Unknown", study.quality.representative),
                                                   "Unknown",  as.character(study.quality.representative))) %>%
    mutate(bias.risk = case_when(grepl("High", study.quality.summary) ~ "High risk of bias",
                                 grepl("Low", study.quality.summary) ~ "Low risk of bias",
                                 grepl("Moderate", study.quality.summary) | grepl("Medium", study.quality.summary) ~ "Moderate risk of bias",
                                 grepl("N/A", study.quality.summary) ~ "Unknown"))

### Setup palette
myPal <- c( "#44AA99", "#882255", "#332288", "#117733", "#6699CC", "#CC6677",
            "#AA4499", "#999933", "#A41034","#88CCEE", "#DDCC77", "#888888")

### Open PDF
filename2 <- here(paste0("output/descriptive/riskOfBiasHIV.pdf"))
pdf(file = filename2, width = 8.5, height = 11.5)

risk1 <- riskData %>%
        mutate(study.quality.representative = fct_relevel(study.quality.representative, c("Unknown", "No (high risk)", "Yes (low risk)"))) %>%
        ggplot() +
        geom_bar(aes(study.quality.representative, fill = bias.risk), position = "stack") +
        theme_minimal() + theme(legend.position = "bottom",
                                panel.grid.minor = element_blank(),
                                panel.grid.major = element_blank()) +
        scale_y_continuous(breaks=seq(0,12,1))+
        coord_flip() +
        labs(y="", x="") +
        # geom_text(stat='count', aes(x=study.quality.representative, label=after_stat(count)), 
        #           position = position_stack(vjust = 1.05)) +
        scale_fill_manual(values = myPal[c(9,4,11,12)], name = "Overall bias assessment") +
        # geom_text(aes(y = 1, x=study.quality.representative, label = bias.risk), size = 3, position = position_stack(vjust = 0.5)) +
        ggtitle("Was the study’s sample population a true or close representation of the target population?")

risk2 <- riskData %>%
    mutate(study.quality.random.selection = fct_relevel(study.quality.random.selection, c("Unknown", "No (high risk)", "Yes (low risk)"))) %>%
    ggplot() +
    geom_bar(aes(study.quality.random.selection, fill = bias.risk), position = "stack") +
    theme_minimal() + theme(legend.position = "bottom",
                            panel.grid.minor = element_blank(),
                            panel.grid.major = element_blank()) + coord_flip() +
    labs(y="", x="") +
            scale_y_continuous(breaks=seq(0,12,1))+

    # geom_text(stat='count', aes(x=study.quality.representative, label=after_stat(count)), 
    #           position = position_stack(vjust = 1.05)) +
    scale_fill_manual(values = myPal[c(9,4,11,12)], name = "Overall bias assessment") +
    # geom_text(aes(y = 1, x=study.quality.representative, label = bias.risk), size = 3, position = position_stack(vjust = 0.5)) +
    ggtitle("Was some form of random selection used to select the sample or was a census undertaken?")

risk3 <- riskData %>%
    mutate(study.quality.nonresponse = fct_relevel(study.quality.nonresponse, c("Unknown", "No (high risk)", "Yes (low risk)"))) %>%
    ggplot() +
    geom_bar(aes(study.quality.nonresponse, fill = bias.risk), position = "stack") +
    theme_minimal() + theme(legend.position = "bottom",
                            panel.grid.minor = element_blank(),
                            panel.grid.major = element_blank()) + coord_flip() +
    labs(y="", x="") +
            scale_y_continuous(breaks=seq(0,12,1))+

    # geom_text(stat='count', aes(x=study.quality.representative, label=after_stat(count)), 
    #           position = position_stack(vjust = 1.05)) +
    scale_fill_manual(values = myPal[c(9,4,11,12)], name = "Overall bias assessment") +
    # geom_text(aes(y = 1, x=study.quality.representative, label = bias.risk), size = 3, position = position_stack(vjust = 0.5)) +
    ggtitle("Was the likelihood of non-response bias minimal?")

risk4 <- riskData %>%
    mutate(study.quality.direct.data.collect = fct_relevel(study.quality.direct.data.collect, c("Unknown", "No (high risk)", "Yes (low risk)"))) %>%
    ggplot() +
    geom_bar(aes(study.quality.direct.data.collect, fill = bias.risk), position = "stack") +
            scale_y_continuous(breaks=seq(0,12,1))+

    theme_minimal() + theme(legend.position = "bottom",
                            panel.grid.minor = element_blank(),
                            panel.grid.major = element_blank()) + coord_flip() +
    labs(y="", x="") +
    # geom_text(stat='count', aes(x=study.quality.representative, label=after_stat(count)), 
    #           position = position_stack(vjust = 1.05)) +
    scale_fill_manual(values = myPal[c(9,4,11,12)], name = "Overall bias assessment") +
    # geom_text(aes(y = 1, x=study.quality.representative, label = bias.risk), size = 3, position = position_stack(vjust = 0.5)) +
    ggtitle("Were data collected directly from subjects (as opposed to a proxy)?")

risk5 <- riskData %>%
        mutate(study.quality.case.definition = fct_relevel(study.quality.case.definition, c("Unknown", "No (high risk)", "Yes (low risk)"))) %>%
        ggplot() +
        geom_bar(aes(study.quality.case.definition, fill = bias.risk), position = "stack") +
            scale_y_continuous(breaks=seq(0,12,1))+

        theme_minimal() + theme(legend.position = "bottom",
                                panel.grid.minor = element_blank(),
                                panel.grid.major = element_blank()) + coord_flip() +
        labs(y="", x="") +
        # geom_text(stat='count', aes(x=study.quality.representative, label=after_stat(count)), 
        #           position = position_stack(vjust = 1.05)) +
        scale_fill_manual(values = myPal[c(9,4,11,12)], name = "Overall bias assessment") +
        # geom_text(aes(y = 1, x=study.quality.representative, label = bias.risk), size = 3, position = position_stack(vjust = 0.5)) +
        ggtitle("Was an acceptable case definition used in the study?")

risk6 <- riskData %>%
        mutate(study.quality.valid.instrument = fct_relevel(study.quality.valid.instrument, c("Unknown", "No (high risk)", "Yes (low risk)"))) %>%
        ggplot() +
        geom_bar(aes(study.quality.valid.instrument, fill = bias.risk), position = "stack") +
            scale_y_continuous(breaks=seq(0,12,1))+

        theme_minimal() + theme(legend.position = "bottom",
                                panel.grid.minor = element_blank(),
                                panel.grid.major = element_blank()) + coord_flip() +
        labs(y="", x="") +
        # geom_text(stat='count', aes(x=study.quality.representative, label=after_stat(count)), 
        #           position = position_stack(vjust = 1.05)) +
        scale_fill_manual(values = myPal[c(9,4,11,12)], name = "Overall bias assessment") +
        # geom_text(aes(y = 1, x=study.quality.representative, label = bias.risk), size = 3, position = position_stack(vjust = 0.5)) +
        ggtitle("Was the study instrument that measured the parameter of interest shown to have\nreliability and validity?")

risk7 <- riskData %>%
        mutate(study.quality.same.data.collect = fct_relevel(study.quality.same.data.collect, c("Unknown", "No (high risk)", "Yes (low risk)"))) %>%
        ggplot() +
        geom_bar(aes(study.quality.same.data.collect, fill = bias.risk), position = "stack") +
            scale_y_continuous(breaks=seq(0,12,1))+

        theme_minimal() + theme(legend.position = "bottom",
                                panel.grid.minor = element_blank(),
                                panel.grid.major = element_blank()) + coord_flip() +
        labs(y="", x="") +
        # geom_text(stat='count', aes(x=study.quality.representative, label=after_stat(count)), 
        #           position = position_stack(vjust = 1.05)) +
        scale_fill_manual(values = myPal[c(9,4,11,12)], name = "Overall bias assessment") +
        # geom_text(aes(y = 1, x=study.quality.representative, label = bias.risk), size = 3, position = position_stack(vjust = 0.5)) +
        ggtitle("Was the same mode of data collection used for all subjects?")

risk8 <- riskData %>%
        mutate(study.quality.same.data.collect = fct_relevel(study.quality.same.data.collect, c("Unknown", "No (high risk)", "Yes (low risk)"))) %>%
        ggplot(aes(x = study.quality.same.data.collect)) +
        geom_bar(aes(fill = bias.risk), position = "stack") +
        theme_minimal() + theme(legend.position = "bottom",
                                panel.grid.minor = element_blank(),
                                panel.grid.major = element_blank()) + coord_flip() +
            scale_y_continuous(breaks=seq(0,12,1))+

        labs(y="", x="") +
        # geom_text(stat='count', aes(x=study.quality.representative, label=after_stat(count)), 
        #           position = position_stack(vjust = 1.05)) +
        scale_fill_manual(values = myPal[c(9,4,11,12)], name = "Overall bias assessment") +
        # geom_text(aes(label = after_stat(count)), position = "stack", stat = "count", hjust = 1.2, colour = "white") +
        ggtitle("Were the numerator and denominator for the parameter of interest appropriate?")

library(ggpubr)
print(
    ggarrange(risk1, risk2, risk3, risk4,
          risk5, risk6, risk7, risk8,
          ncol=1, nrow=8, common.legend = TRUE, legend="bottom")
)


dev.off()
```


```{r}
### Primary analysis table
hivTbl0 <- cleanHivDF0 %>% 
            filter(covidence.id %in% bactModelHivDF$covidence.id) %>% unique() %>%
            dplyr::select(starts_with("n.") & ends_with(c("hiv.positive", "hiv.negative", "hiv.total"))) %>% 
            colSums(na.rm=TRUE)

hivTbl <- matrix(hivTbl0,3,length(hivTbl0)/3, byrow = FALSE)[-3,]

rownames(hivTbl) <- c("HIV positive (N)", "HIV negative (N)")

library(gt)

hivTbl_1 <- hivTbl %>% t() %>% as.data.frame() %>% 
    mutate("Description" = unique(sapply(names(hivTbl0), function(x) sub("[^.]+\\.([^.]+)\\..*", "\\1", x))),
           "Total (N)" = `HIV positive (N)` + `HIV negative (N)`, 
           "HIV negative (%)" = round(`HIV negative (N)`/`Total (N)`*100,1), 
           "HIV positive (%)" = round(`HIV positive (N)`/`Total (N)`*100,1)) %>% 
    dplyr::select(Description, `HIV negative (N)`, `HIV negative (%)`, `HIV positive (N)`,`HIV positive (%)`, `Total (N)`) %>%
    na.omit() %>%
    gt() %>% 
    tab_header(
        title = paste("HIV status distribution totals across", 
                      length(bactModelHivDF$covidence.id),  
                      "TB prevalence surveys")) %>%
    fmt_number(drop_trailing_zeros = TRUE); hivTbl_1


### Secondary analysis table
hivTbl0_b <- cleanHivDF0 %>% 
          filter(!covidence.id %in% bactModelHivDF$covidence.id) %>% unique() %>%
          dplyr::select(starts_with("n.") & ends_with(c("hiv.positive", "hiv.negative", "hiv.total"))) %>% 
          colSums(na.rm=TRUE)

hivTbl_b <- matrix(hivTbl0_b,3,length(hivTbl0_b)/3, byrow = FALSE)[-3,]

rownames(hivTbl_b) <- c("HIV positive (N)", "HIV negative (N)")

library(gt)

hivTbl_1_b <- hivTbl_b %>% t() %>% as.data.frame() %>% 
    mutate("Description" = unique(sapply(names(hivTbl0_b), function(x) sub("[^.]+\\.([^.]+)\\..*", "\\1", x))),
           "Total (N)" = `HIV positive (N)` + `HIV negative (N)`, 
           "HIV negative (%)" = round(`HIV negative (N)`/`Total (N)`*100,1), 
           "HIV positive (%)" = round(`HIV positive (N)`/`Total (N)`*100,1)) %>% 
    dplyr::select(Description, `HIV negative (N)`, `HIV negative (%)`, `HIV positive (N)`,`HIV positive (%)`, `Total (N)`) %>%
    na.omit() %>%
    gt() %>% 
    tab_header(
        title = paste("HIV status distribution totals across", 
                      length(bactModelHivDF$covidence.id),  
                      "TB prevalence surveys")) %>%
    fmt_number(drop_trailing_zeros = TRUE); hivTbl_1_b

```

