---
title: "Relationship between sex-disaggregated TB Prevalence and Notification Rates in Low- and Middle- Income Countries " 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
# classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

```{r Setup, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
# library(posterior)
library(tidybayes)
library(wesanderson)
library(ggplot2)

expit <- function (x) {exp(x)/(1 + exp(x))}

myPal <- c("black", "violet")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))

##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################
cleanSexNatDF0 <- bactPostIndicator("sex") %>% 
                  filter(sex.analysis.indicator != "none", 
                  study.geography == "Nationally representative") 

### Check if any publication years are missing
sum(is.na(cleanSexNatDF0$study.end.year))

### Distribution of samples across reporting data type 
table(cleanSexNatDF0$sex.analysis.indicator)

```

```{r Prevalence Estimates, echo = FALSE,  warning = FALSE, message=FALSE}

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexNatDF <- cleanSexNatDF0 %>% 
               filter(sex.analysis.indicator != "none") %>% 
               mutate("id" = row_number()) %>% 
    ### Scale all prevalence estimates back to decimals 
    mutate_at(vars(matches('prev100k.bacteriological.tb.|prev100k.ci.([a-z]{3})er.bacteriological.tb.|prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           study.rurality, 
           WHO.region,
           screening.tests,
           participants.sputum.collected,
           bacteriological.tb.definition,
           xpert.used,
           xpert.samples,
           other.xpert.samples,
           culture.used,
           culture.samples,
           other.culture.samples,
           study.start.year, 
           study.end.year, 
           publication.year,
           study.quality.summary,
           ends_with("male"), ends_with("sex"),
           sex.analysis.indicator) %>%
    
    mutate("tbPrev.male" =  case_when(sex.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.male/n.participants.male, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.male/n.participants.male,
                                        sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tbPrev.female" = case_when(sex.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.female/n.participants.female, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female),
           logTbPrev.male = log(tbPrev.male), 
           logTbPrev.female = log(tbPrev.female)) %>%
    #### Create a confidence interval for those without 
    #### these are using the wilson method with continuity correction
    #### cannot seem to get case_when to work so this is going to be verbose bc deadline
    rowwise() %>%
    mutate(calc.ci.lower.male = ifelse(sex.analysis.indicator == "n.bacteriological.tb",  
                                              (prop.test(n.bacteriological.tb.male, n.participants.male))$conf.int[1],
                                       ifelse(sex.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.male, n.participants.male))$conf.int[1],
                                          ifelse(sex.analysis.indicator == "n.culture.positive.tb", 
                                              (prop.test(n.culture.positive.tb.male, n.participants.male))$conf.int[1],
                                                ifelse(sex.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.male*n.participants.male, n.participants.male))$conf.int[1], 
                                              NA)))),
            calc.ci.upper.male = ifelse(sex.analysis.indicator == "n.bacteriological.tb",  
                                              (prop.test(n.bacteriological.tb.male, n.participants.male))$conf.int[2],
                                       ifelse(sex.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.male, n.participants.male))$conf.int[2],
                                          ifelse(sex.analysis.indicator == "n.culture.positive.tb", 
                                              (prop.test(n.culture.positive.tb.male, n.participants.male))$conf.int[2],
                                                ifelse(sex.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.male*n.participants.male, n.participants.male))$conf.int[2], 
                                              NA)))),
           calc.ci.lower.female = ifelse(sex.analysis.indicator == "n.bacteriological.tb",  
                                              (prop.test(n.bacteriological.tb.female, n.participants.female))$conf.int[1],
                                       ifelse(sex.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.female, n.participants.female))$conf.int[1],
                                          ifelse(sex.analysis.indicator == "n.culture.positive.tb", 
                                              (prop.test(n.culture.positive.tb.female, n.participants.female))$conf.int[1],
                                                ifelse(sex.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.female*n.participants.female, n.participants.female))$conf.int[1], 
                                              NA)))),
            calc.ci.upper.female = ifelse(sex.analysis.indicator == "n.bacteriological.tb",  
                                              (prop.test(n.bacteriological.tb.female, n.participants.female))$conf.int[2],
                                       ifelse(sex.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.female, n.participants.female))$conf.int[2],
                                          ifelse(sex.analysis.indicator == "n.culture.positive.tb", 
                                              (prop.test(n.culture.positive.tb.female, n.participants.female))$conf.int[2],
                                                ifelse(sex.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.female*n.participants.female, n.participants.female))$conf.int[2], 
                                              NA))))) %>%
        ungroup() %>%
        mutate("logTbPrevSE.male" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb"  ~ 
                                         log(adj.prev100k.ci.upper.bacteriological.tb.male /
                                                                 adj.prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         log(adj.prev100k.ci.upper.smear.positive.tb.male /
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" ~ 
                                            log(prev100k.ci.upper.bacteriological.tb.male /
                                                                 prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            log(prev100k.ci.upper.smear.positive.tb.male /
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb")  ~ 
                                            log(calc.ci.upper.male / calc.ci.lower.male)/3.92),
           "logTbPrevSE.female" =  case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb"  ~ 
                                         log(adj.prev100k.ci.upper.bacteriological.tb.female  /
                                                                 adj.prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         log(adj.prev100k.ci.upper.smear.positive.tb.female /
                                                                 adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" ~ 
                                            log(prev100k.ci.upper.bacteriological.tb.female /
                                                                 prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            log(prev100k.ci.upper.smear.positive.tb.female / 
                                                                 prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb")  ~ 
                                                        log(calc.ci.upper.female / calc.ci.lower.female)/3.92)) %>% 
    
        pivot_longer(cols = c(tbPrev.male, tbPrev.female,
                              logTbPrev.male, logTbPrev.female, 
                              logTbPrevSE.male, logTbPrevSE.female), 
                 names_to = c("Measure","Sex"), 
                 names_pattern = "(.*)\\.(.*)$", 
                 ) %>% 
        pivot_wider(names_from = Measure, values_from = value) %>%
    mutate(natRep = ifelse(is.na(study.geography), 
                                                       NA, 
                                                       ifelse(study.geography == "Nationally representative", 
                                                       "Yes", "No")),
                                       ## preserve the extracted study end year
                                       study.end.year.ext = study.end.year, 
                                       ## update missing study end year to be publication year - mean lag
                                       study.end.year = ifelse(!is.na(study.end.year), 
                                                               study.end.year.ext, 
                                                               publication.year - meanLag),
                                       ## Mean center year for temporal analysis
                                       year_z = study.end.year - round(mean(study.end.year, na.rm=TRUE)),

                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       
                                       symptomScreen = ifelse(is.na(screening.tests), 
                                                                    NA, 
                                                                    ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No")),
                                       symptomsRequired = ifelse(is.na(participants.sputum.collected) | grepl("unknown", 
                                                                                                              participants.sputum.collected,ignore.case=TRUE), 
                                                                       NA, 
                                                                       ifelse(!grepl("x-ray|all participants|other",
                                                                        participants.sputum.collected,ignore.case=TRUE),
                                                                 "Yes", "No")),

                                       XpertCultRequired = ifelse(is.na(bacteriological.tb.definition),
                                                                  NA, ifelse(!grepl("smear|micros|field|NTP", 
                                                                              bacteriological.tb.definition,
                                                                              ignore.case=TRUE), 
                                                                  "Yes", 
                                                                   "No")),
                                       ### Fix those that were missed with the above attempt 
                                       XpertCultRequired = ifelse(covidence.id %in% c(26989,26928,26920,26918,26892,26861,26857), 
                                                                  "Yes", XpertCultRequired), 
                                       XpertCultRequired = ifelse(covidence.id %in% c(28002), 
                                                                  "No", XpertCultRequired),
                                       #### if we used the culture positive count, then xpert culture was requried
                                       XpertCultRequired = ifelse(grepl("culture", sex.analysis.indicator), 
                                                                  "Yes", XpertCultRequired),
                                       #### if we used the smear positive prevs, then xpert culture was not requried
                                       XpertCultRequired = ifelse(grepl("smear", sex.analysis.indicator), 
                                                                  "No", XpertCultRequired),
                                       XpertCultUsed = ifelse((grepl("Yes", xpert.used) & 
                                                                 (grepl("All samples", xpert.samples) |
                                                                  !grepl("mear|NAAT", other.xpert.samples))) |
                                                                  (grepl("Yes", culture.used) & 
                                                                  grepl("All samples|Subset", culture.samples) & 
                                                                  !grepl("HIV", other.culture.samples)), 
                                                                 "Yes", "No"),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))


##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("sex")


bactModelNatDF1 <- cleanSexNatDF %>%  
    select(id, WHO.region, study.country, study.end.year, 
           sex.analysis.indicator,
           Sex, tbPrev, 
           logTbPrev, logTbPrevSE,
           XpertCultRequired, XpertCultUsed,
           symptomsRequired,
           study.rurality, bias, 
           n.participants.male, n.participants.female, 
           bias, publication.year, year_z) 

bactModelNatDF <- cleanSexNatDF %>%  
    select(id, WHO.region, study.country, study.end.year, 
           sex.analysis.indicator,
           Sex, tbPrev,
           logTbPrev, logTbPrevSE,
           XpertCultRequired, XpertCultUsed,
           symptomsRequired,
           study.rurality, bias, 
           n.participants.male, n.participants.female, 
           bias, publication.year, year_z) %>%
    pivot_wider(names_from = Sex,
                values_from = c(logTbPrev, logTbPrevSE),
                names_sep = "_") %>%
    mutate(logPR = logTbPrev_male - logTbPrev_female,
           unadj_logPR_se = sqrt(logTbPrevSE_male^2 + logTbPrevSE_female^2), 
           log_PR_se = ifelse(grepl("adj", sex.analysis.indicator), 
                                     unadj_logPR_se,
                                     ifelse(grepl("smear", sex.analysis.indicator),
                                         sqrt(`unadj_logPR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logPR_se`^2 + varAdj["Bact Variance Median"]))))

```

```{r Notification Data, echo = FALSE,  warning = FALSE, message=FALSE, width=11.5, height=8.5}

##### NOTIFICATION DATA ##############################################
notif00 <- read.csv(here("data/covariates/TB_notifications_2025-07-18.csv"))
sexPop0 <- read.csv(here("data/covariates/populationData.csv"))

sexPop <- sexPop0 %>% melt(id.vars = c("Country.Name", "Series.Name")) %>%
          mutate(year = as.integer(gsub("X", "", variable))) %>% 
          select(! variable) %>%
          pivot_wider(id_cols = c(Country.Name, year), 
                      names_from = Series.Name, values_from = value) %>% 
          rename(country = Country.Name, 
                 pop_male = male, 
                 pop_female = female)

sexPop[grep("Viet", sexPop$country), "country"] <- 
    notif00[grep("Viet", notif00$country), "country"] <- "Vietnam"
sexPop[grep("Lao", sexPop$country), "country"] <- 
    notif00[grep("Lao", notif00$country), "country"] <- "Laos"
sexPop[grep("People's", sexPop$country), "country"] <- 
    notif00[grep("People's", notif00$country), "country"] <- "North Korea"
sexPop[grep("Gambia", sexPop$country), "country"] <- 
    notif00[grep("Gambia", notif00$country), "country"] <- "Gambia"

#### Convert to male-to-female ratios 
notif0 <- notif00 %>% 
         filter(year > 1993, 
                country %in% unique(bactModelNatDF$study.country)) %>%
         select(country, g_whoregion, 
                year, 
                newrel_m15plus,
                newrel_f15plus, 
                new_sn_m15plus,
                new_sp_m1524, new_sp_m2534, 
                new_sp_m3544, new_sp_m4554,
                new_sp_m5564, new_sp_m65, new_sp_mu,
                new_sn_f15plus,
                new_sp_f1524, new_sp_f2534, 
                new_sp_f3544, new_sp_f4554,
                new_sp_f5564, new_sp_f65, new_sp_fu) %>% 
         left_join(sexPop) %>%
         rename(cases_male = newrel_m15plus,
                cases_female = newrel_f15plus)

### for some reason if else wouldn't work so let' just loop 
for (i in 1:nrow(notif0)){
    if (is.na(notif0$cases_male[i]) & notif0$year[i] < 2013){
        notif0$cases_female[i] <- sum((notif0 %>% select(contains(c("sn_f", "sp_f"))))[i,], na.rm = T)
        notif0$cases_male[i]   <- sum((notif0 %>% select(contains(c("sn_m", "sp_m"))))[i,], na.rm = T)
    }
}


notifStrat <- notif0 %>% 
              pivot_longer(cols=c(pop_male, pop_female, 
                                 cases_male, cases_female), 
                          names_to = c("Measure", "Sex"), 
                          # names_prefix = "notifRate",
                          names_sep = "_",
                          values_to = "values",
                          values_drop_na = TRUE) %>%
              pivot_wider(names_from = Measure, 
                          values_from = values) %>% 
              mutate(### Poisson uncertainty
                     casesSE = sqrt(cases / pop),
                     ### Notification rate 
                     notifRate = cases/pop, 
                     ### Notification rate uncertainty
                     notifRateSE = casesSE/pop, 
                     ### Log notification rate 
                     logNotifRate = log(notifRate), 
                     ### Uncertainty in log notification rate
                     logNotifRateSE = log((cases + casesSE) / (cases-casesSE))/pop)  %>%
              select(country, year, Sex, pop, cases, casesSE, 
                     notifRate, notifRateSE, logNotifRate, logNotifRateSE)


notifRatio <- notif0 %>% 
              mutate(notifRateMale = cases_male/pop_male, 
                     notifRateFemale = cases_female/pop_female, 
                     logNotifRateMale = log(notifRateMale),
                     logNotifRateFemale = log(notifRateFemale),
                     mfRatioNotif = notifRateMale/notifRateFemale,
                     logMfRatioNotif = log(mfRatioNotif)) 


ggplot(data=notifRatio) +
    geom_line(aes(x=year, y=mfRatioNotif, group=country, color=country)) +
    theme_minimal() + theme(legend.position = "none",
                            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    facet_grid() + ggtitle("Male to female ratio of TB notifications 1994-2021") +
    scale_x_continuous(breaks = seq(1994, 2022, 4)) +
    facet_wrap(vars(country))
```

```{r Single Dataset Creation, echo=FALSE, warning=FALSE, message=FALSE}
notifStrat <- bactModelNatDF1  %>%
             select(id,
                    Sex,
                    logTbPrev,
                    logTbPrevSE,
                    study.country, study.end.year,
                    # `Adjusted Prevalence`,
                    # sex.analysis.indicator,
                    # XpertCultRequired,
                    year_z) %>%
             rename(country = study.country,
                    year = study.end.year) %>%
             left_join(notifStrat, by=c("country", "year", "Sex")) %>%
             rename(study.country = country,
                    study.end.year = year) %>%
             mutate(notifRate = ifelse(notifRate==0, NA, notifRate))


npRatioDF <- left_join(bactModelNatDF1  %>%
                         select(id, Sex,
                                tbPrev,
                                logTbPrev, 
                                logTbPrevSE,
                                study.country, study.end.year, 
                                year_z), 
                       notifStrat) 

npRatioDF %>% mutate(empiricalNP = notifRate / tbPrev) %>% 
    select(study.country, Sex, notifRate, tbPrev, empiricalNP) -> y; View(y)


# tinyDF <- notifPrev %>% mutate(prevMale = expit(LogOdds_Male), 
#                                prevFemale = expit(LogOdds_Female)) %>% 
#                         select(study.country, 
#                                study.end.year, 
#                                sex.analysis.indicator, 
#                                XpertCultRequired, 
#                                cases_male, 
#                                cases_female, 
#                                prevMale, 
#                                prevFemale) 

# write.csv(tinyDF, here("output/pnRatioSummaryTable.csv"))

# notifPrevModel <- notifPrev %>% ungroup() %>%
#                   mutate(LogOdds_NotifMale = log(notifRateMale/(1-notifRateMale)), 
#                          LogOdds_NotifFemale = log(notifRateFemale/(1-notifRateFemale))) %>%
#                   select(logOR, logOR_se, mfRatioNotif, 
#                          cases_male, cases_female,
#                          notifRateMale, notifRateFemale,
#                          LogOdds_Male, LogOdds_Female, 
#                          LogOddsStandardError_Male, LogOddsStandardError_Female,
#                          LogOdds_NotifMale, LogOdds_NotifFemale, 
#                          study.country, id, year_z, study.end.year)
```

### Lets fit the prevalence model 
```{r Sex-stratified Prevalence Model, warning=FALSE, message=FALSE}

tbPrevMod <- brms::brm(formula = logTbPrev | se(logTbPrevSE, sigma=TRUE) ~ 
                           1 + Sex + year_z + (1 + Sex | study.country) + (1 | study.end.year),
                       data = notifStrat,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 10000, warmup = 2500, thin=3)

summary(tbPrevMod)

tb_prev_mod_est <-  tbPrevMod %>% 
    add_epred_draws(newdata = crossing(study.country=unique(notifStrat$study.country), 
                                       year_z = unique(notifStrat$year_z),
                                       Sex = c("female", "male"), 
                                       logTbPrevSE = 0,
                                       logNotifRateSE = 0),
                    re_formula =  ~ 1 + Sex + year_z + (1 + Sex | study.country), 
                    allow_new_levels=TRUE) %>%
    ungroup() %>% 
    dplyr::select(study.country, year_z, Sex, .draw, .epred) %>%
    group_by(Sex, year_z, study.country) %>%  
    mutate(tbPrev = exp(.epred) * 1e5, 
           study.end.year = year_z + round(mean(cleanSexNatDF$study.end.year, na.rm=TRUE)))


tb_prev_mod_est_summary <- 
    tb_prev_mod_est %>%
    group_by(Sex, study.country, study.end.year) %>%
    median_qi(tbPrev, na.rm=T) %>% 
    mutate(across(c(tbPrev, .lower, .upper), 
                  round, 2)); tb_prev_mod_est_summary

ggplot(data=tb_prev_mod_est_summary) +
    geom_line(aes(x=study.end.year, y=tbPrev, group=Sex, color=Sex)) +
    theme_minimal() + theme(legend.position = "none",
                            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    facet_grid() + ggtitle("Sex-stratified TB prevalence by country, 1994-2021") +
    scale_x_continuous(breaks = seq(1994, 2022, 4)) +
    facet_wrap(vars(study.country))

tb_prev_mod_est_summary2 <- 
    tb_prev_mod_est %>%
    group_by(Sex, study.end.year) %>%
    median_qi(tbPrev, na.rm=T) %>% 
    mutate(across(c(tbPrev, .lower, .upper), 
                  round, 2)); tb_prev_mod_est_summary2

ggplot(data=tb_prev_mod_est_summary2) +
    geom_line(aes(x=study.end.year, y=tbPrev, group=Sex, color=Sex)) +
    theme_minimal() + theme(legend.position = "none",
                            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    ggtitle("Sex-stratified TB prevalence by country, 1994-2021") +
    scale_x_continuous(breaks = seq(1994, 2022, 4))
```



### Sex-stratified prevalence to notification ratios 

```{r Sex-stratified Specification, warning=FALSE, message=FALSE}
### Run a multivariate model
#prevalence model:
prev_model <- bf(
    logTbPrev | se(logTbPrevSE, sigma = TRUE) ~ 1 + Sex + (1 + Sex | p | study.country) + (1 + Sex | q | year_z)
)

#notifications model:
notif_model <- bf(
    logNotifRate | se(logNotifRateSE, sigma = TRUE) ~ 1 + Sex + (1 + Sex | p | study.country) + (1 + Sex | q | year_z)
)

#fit the model
sex_prev_notif_model <- brm(
    formula = prev_model + notif_model + set_rescor(FALSE),
    data = notifStrat,
    family = gaussian(), 
    # sample_prior = "only",
    prior = c(
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logTbPrev"),
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logNotifRate"),
        prior(normal(0,1), class = b, resp = "logTbPrev"),
        prior(normal(0,1), class = b, resp = "logNotifRate"),
        prior(exponential(2), class = sd, resp = "logTbPrev"),
        prior(exponential(2), class = sd, resp = "logNotifRate")
    ),
    control = list(adapt_delta = 0.99),
    chains = 4,
    cores = 4,
    iter = 10000, 
    warmup = 2500
)

summary(sex_prev_notif_model)
            
sex_prev_notif_est <-  sex_prev_notif_model %>% 
    add_epred_draws(newdata = crossing(study.country=unique(notifStrat$study.country), 
                                       year_z = unique(notifStrat$year_z),
                                       Sex = c("female", "male"), 
                                       logTbPrevSE = 1,
                                       logNotifRateSE = 0),
                    re_formula =  ~ 1 + Sex + (1 + Sex | study.country) + (1 + Sex | year_z), 
                    allow_new_levels=TRUE) %>%
    ungroup() %>% 
    dplyr::select(study.country, year_z, .category, Sex, .draw, .epred) %>%
    pivot_wider(names_from = .category, values_from = .epred) %>%
    mutate(npRatio = exp(logNotifRate - logTbPrev), 
           notifRateMod = exp(logNotifRate)) 

sex_prev_notif_estSummary <- 
    sex_prev_notif_est %>%
    group_by(Sex, study.country) %>%
    mean_qi(npRatio, na.rm=T) %>% 
    mutate(across(c(npRatio, .lower, .upper), 
                  round, 2)); sex_prev_notif_estSummary

sex_prev_notif_estSummaryTot <- 
    sex_prev_notif_est %>%
    group_by(Sex) %>%
    mean_qi(npRatio, na.rm=T) %>% 
    mutate(across(c(npRatio, .lower, .upper), 
                  round, 2)); sex_prev_notif_estSummaryTot

### output for contribution table
intcpt.effect <- sex_prev_notif_model %>% 
              spread_draws(b_logTbPrev_Intercept, b_logNotifRate_Intercept, 
                          b_logTbPrev_Sexmale, b_logNotifRate_Sexmale) %>%
            mutate(year.female.prev = exp(b_logTbPrev_Intercept), 
                   year.female.notif = exp(b_logNotifRate_Intercept), 
                   year.male.prev = exp(b_logTbPrev_Intercept + b_logTbPrev_Sexmale), 
                   year.male.notif = exp(b_logNotifRate_Intercept + b_logNotifRate_Sexmale), 
                   year.female.npRatio = year.female.notif/year.female.prev,
                   year.male.npRatio = year.male.notif/year.male.prev) %>% 
            median_qi(year.female.npRatio, year.male.npRatio) %>% 
            select(! c(.width, .point, .interval)) %>% round(digits = 2) %>%
           matrix(nrow=2, byrow = TRUE) ;

intcpt.effect
```

```{r check notifications modeled vs. reported}

notifComp <- sex_prev_notif_est %>% 
            select(study.country, year_z, Sex, notifRateMod) %>% 
            right_join(notifStrat %>% 
                           select(study.country, year_z, Sex, notifRate, notifRateSE)) %>% 
            group_by(study.country, year_z, Sex) %>% 
            median_qi(notifRateMod) %>% 
            select(! c(.width, .point, .interval)) %>% 
            right_join(notifStrat %>% 
                           select(study.country, year_z, study.end.year, Sex, notifRate, notifRateSE)) 

# ggplot(data=notifComp %>% mutate(ID = paste(study.country, study.end.year))) + 
#         geom_point(aes(x=reorder(ID, notifRateMod), y=notifRate*1e5, group=Sex, fill = Sex), size=3, alpha=0.5, position=position_dodge(width = .6), shape=21, color = "black")  + 
#         geom_errorbar(aes(x = reorder(ID, notifRateMod), ymin=(notifRate - 1.96*notifRateSE)*1e5, ymax=(notifRate + 1.96*notifRateSE)*1e5, color=Sex), position=position_dodge(width = .6), size=1, alpha=.5) +
#     theme_minimal(base_size = 20) + coord_flip() + xlab("") + ylab("Notification Rate (per 100K)")  +
#     theme(legend.position = "bottom")
    
ggplot(data=notifComp %>% na.omit() %>% mutate(ID = paste(study.country, study.end.year))) + 
    geom_point( 
               aes(x=reorder(ID, notifRateMod), y=notifRateMod*1e5, group=Sex, color = Sex),
               position=position_dodge(width = .6), size=2) + 
    geom_errorbar(aes(x = reorder(ID, notifRateMod), ymin=.lower*1e5, ymax=.upper*1e5, color=Sex), position=position_dodge(width = .6), size=1) +
    geom_point(aes(x=reorder(ID, notifRateMod), y=notifRate*1e5, group=Sex, fill = Sex), size=3, alpha=0.5, position=position_dodge(width = .6), shape=21, color = "black")  + 
    theme_minimal(base_size = 15) + coord_flip() + xlab("") + ylab("Notification Rate (per 100K)")  +
    theme(legend.position = "bottom")

ggsave(plot=get_last_plot(), here("output/sexNP/notificationComp.pdf"), 
       height=13, width = 8.5, units = "in") 

notifCompDiff <- notifComp %>% 
                 left_join(notifStrat %>% select(study.country, study.end.year, Sex, logTbPrevSE)) %>% 
                 mutate(diff = round(abs(notifRate-notifRateMod) * 1e5), 
                        tbPrevSE = exp(logTbPrevSE), 
                        combinedSE = tbPrevSE + notifRateSE) 

    ggplot(data = notifCompDiff %>% na.omit(), aes(x=tbPrevSE, y=diff)) + 
        geom_point() + theme_minimal() + coord_flip() + 
        xlab("Standard error of the TB prevalence risk") + 
        ylab("Absolute difference in model estimated and observed notification rates (per 100K)") + 
        expand_limits(x=1)
    
    ggsave(plot=get_last_plot(), "~/Desktop/prevSE-notifDiff.pdf", 
       height=8.5, width = 8.5, units = "in") 
    
        ggplot(data = notifCompDiff %>% na.omit(), aes(x=notifRateSE, y=diff)) + 
        geom_point() + theme_minimal() + coord_flip() + 
        xlab("Standard error of the TB notification rates") + 
        ylab("Absolute difference in model estimated and observed notification rates (per 100K)")
    
                 
    ggsave(plot=get_last_plot(), "~/Desktop/notifSE-notifDiff.pdf", 
       height=8.5, width = 8.5, units = "in") 
    # ggplot(data = notifCompDiff %>% na.omit(), aes(x=combinedSE, y=diff)) + 
    #     geom_point() + theme_minimal() + coord_flip() + 
    #     xlab("Standard error of the TB prevalence risk") + 
    #     ylab("Absolute difference in model estimated and observed notification rates (per 100K)") + 
    #     expand_limits(x=0)
   

```

```{r Sex-specific PN Ratio Plot, warning=FALSE, message=FALSE, width=8.5, length=11.5}
ggplot() + 
    geom_point(data=sex_prev_notif_estSummary, aes(x=reorder(study.country, npRatio), y=npRatio, group=Sex, color = Sex), position=position_dodge(width = .6), size=2) + 
    geom_errorbar(data=sex_prev_notif_estSummary, aes(x = reorder(study.country, npRatio), ymin=.lower, ymax=.upper, color=Sex), position=position_dodge(width = .6), size=1) +
    # geom_point(data=notifPrevStrat2, aes(x=reorder(study.country, npRatio), y=empiricalRatio), color="black") + 
    theme_minimal() + 
    scale_color_manual(values = c("darkred", "dodgerblue4"))+
    coord_flip() + ylab("Notification to prevalence ratio") + xlab("") +
    theme( legend.position=c(.9,.2), legend.background = element_rect(fill="white")) +
    expand_limits(y=1) + geom_hline(yintercept = 1, linetype = 2) + 
    scale_y_continuous(breaks=seq(0,1.25, .25)) +
    theme(legend.position = "bottom")
    
ggsave(here("output/sexSpecificPNratios.png"), plot = last_plot(),
       width = 8.5, height=11, units = "in")

# notifPrevStrat2 <- notifPrevStrat %>% mutate(empiricalRatio = `Adjusted Prevalence`/notifRate) %>% 
#     left_join(notifModel_est_summary %>% select(study.country, pnRatio))
```



### Time-effect Model Specification 
```{r Time Model Specification, warning=FALSE, message=FALSE}

sex_prev_time_model <- bf(
    logTbPrev | se(logTbPrevSE, sigma = TRUE) ~ 1 + Sex*year_z + (1 + Sex*year_z | p | study.country) 
)

#notifications model:
sex_notif_time_model <- bf(
    logNotifRate | se(logNotifRateSE, sigma = TRUE) ~ 1  + Sex*year_z + (1 + Sex*year_z | p | study.country) 
)

#fit the model
sex_prev_notif_time_model <- brm(
    formula = sex_prev_time_model + sex_notif_time_model + set_rescor(FALSE),
    data = notifStrat,
    family = gaussian(), 
    # sample_prior = "only",
    prior = c(
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logTbPrev"),
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logNotifRate"),
        prior(normal(0,1), class = b, resp = "logTbPrev"),
        prior(normal(0,1), class = b, resp = "logNotifRate"),
        prior(exponential(2), class = sd, resp = "logTbPrev"),
        prior(exponential(2), class = sd, resp = "logNotifRate")
    ),
    control = list(adapt_delta = 0.99),
    chains = 4,
    cores = 4,
    iter = 10000, 
    warmup = 2500
)

```

### Time-effect Model Estimation 
```{r Time Model Estimation, echo=FALSE, warning=FALSE, message=FALSE}
sex_prev_notif_time_est <-  sex_prev_notif_time_model %>% 
    add_epred_draws(newdata = crossing(study.country=unique(notifStrat$study.country), 
                                       Sex = c("female", "male"), 
                                       year_z = min(notifStrat$year_z):max(notifStrat$year_z), 
                                       logTbPrevSE = 1,
                                       logNotifRateSE = 1),
                    re_formula =  ~ 1 + Sex*year_z + (1 + | study.country)) %>%
    ungroup() %>% 
    dplyr::select(study.country, .category, Sex, year_z,.draw, .epred) %>%
    pivot_wider(names_from = .category, values_from = .epred) %>%
    mutate(npRatio = exp(logNotifRate - logTbPrev)) 

# sex_prev_notif_estSummary <- 
#     sex_prev_notif_est %>%
#     group_by(Sex, study.country) %>%
#     mean_qi(npRatio, na.rm=T) %>% 
#     mutate(across(c(npRatio, .lower, .upper), 
#                   round, 2)); sex_prev_notif_estSummary


### output for contribution table
year.effect <- sex_prev_notif_time_model %>% 
              spread_draws(b_logTbPrev_Intercept, b_logNotifRate_Intercept, 
                          b_logTbPrev_Sexmale, b_logNotifRate_Sexmale,
                          b_logTbPrev_year_z, b_logNotifRate_year_z, 
                          `b_logTbPrev_Sexmale:year_z`,`b_logNotifRate_Sexmale:year_z`) %>%
            mutate(year.female.prev = exp(b_logTbPrev_year_z), 
                   year.female.notif = exp(b_logNotifRate_year_z), 
                   year.male.prev = exp(b_logTbPrev_year_z + `b_logTbPrev_Sexmale:year_z`), 
                   year.male.notif = exp(b_logNotifRate_year_z + `b_logNotifRate_Sexmale:year_z`), 
                   year.female.npRatio = year.female.notif/year.female.prev,
                   year.male.npRatio = year.male.notif/year.male.prev) %>% 
            median_qi(year.female.npRatio, year.male.npRatio) %>% 
            select(! c(.width, .point, .interval)) %>% round(digits = 2) %>%
           matrix(nrow=2, byrow = TRUE) 

```

### Time-effect Model Visualization 
```{r}
myPal <- c("darkred", "dodgerblue4")
y <- sex_prev_notif_time_est  %>%
     group_by(year_z, Sex) %>%
      mean_qi(npRatio) %>%
      mutate(Year = year_z + round(mean(cleanSexNatDF$study.end.year, na.rm=TRUE))) %>%
     select(Year, Sex, npRatio, .lower, .upper) 


ggplot(data = y) +
        geom_line(aes(x=Year, y=npRatio, group=Sex, color = Sex),  size=2) + 
    geom_ribbon(aes(x = Year, ymin=.lower, ymax=.upper, fill=Sex), alpha=.5, size=1) +
    
    expand_limits(y=0) +
    scale_fill_manual(values = myPal) + scale_color_manual(values = myPal) + 
    labs(y = "TB notification to prevalence ratio",
         x = "Year") +
    theme_minimal(base_size = 20)  + theme(legend.position = "bottom")


ggsave(here("output/sexSpecificPNratioTime.png"), plot = last_plot(),
       width = 11, height=8.5, units = "in")
```

Covariates 
```{r Setup covariate dataset, echo=FALSE}
covariates <- read.csv(file=here("data/covariates/covariates_1994-2023.csv")) %>% 
                  dplyr::rename(study.country = location_name) %>%
              filter(sex_name != "Both", 
                     measure_name == "Prevalence", 
                     metric_name == "Rate") %>% 
    select(study.country, year, sex_name, cause_name, val) %>% 
    pivot_wider(names_from = cause_name, 
                values_from = val) %>% 
    group_by(sex_name) %>% 
    mutate(alcohol.use.disorder.unscld = `Alcohol use disorders`,
           # alcohol.use.disorder = scale(`Alcohol use disorders`)[,1], 
           type2.diabetes.mellitus.unscld = `Diabetes mellitus type 2`, 
           # type2.diabetes.mellitus = scale(`Diabetes mellitus type 2`)[,1], 
           HIV.AIDS.unscld = `HIV/AIDS`, 
           # HIV.AIDS = scale(`HIV/AIDS`)[,1]
           ) %>% 
    rename(study.end.year = year, 
           sex = sex_name) %>% 
    select(study.country, 
           study.end.year, 
           sex,
           alcohol.use.disorder.unscld, 
           type2.diabetes.mellitus.unscld,
           HIV.AIDS.unscld)

### BMI data
bmi <- read.csv(file=here("data/covariates/IHME_GBD_2023_COV_1994_2023_ADULT_UNDERWEIGHT_33 Countries.csv")) %>% 
    rename(study.country = location_name) %>%
    select(study.country, year_id, sex, covariate_name_short, mean_value) %>% 
    group_by(year_id, study.country, sex) %>%
    summarize(val = sum(mean_value)) %>%
    group_by(sex) %>%
    mutate(bmi.unscld = val, 
           # bmi = scale(val)
           ) %>% 
    select(study.country, 
           year_id, 
           bmi.unscld) %>% 
    rename(study.end.year = year_id) %>% 
    ungroup()

### Smoking data 
smoking.prevalence <- read.csv(file=here("data/covariates/IHME_GBD_2021_COV_1980_2021_SMOKING_PREV_33 Countries.csv")) %>% 
    rename(study.country = location_name) %>%
    select(study.country, year_id, sex, covariate_name_short, mean_value) %>% 
    group_by(year_id, study.country, sex) %>%
    summarize(val = sum(mean_value)) %>%
    group_by(sex) %>%
    mutate(smoking.unscld = val, 
           # smoking = scale(val)
           ) %>% 
    select(study.country, 
           year_id, 
           smoking.unscld) %>%
    rename(study.end.year = year_id) %>% 
    ungroup()

# smoking.prevalence$smoking <- scale(smoking.prevalence$smoking)

### Gender development index 
GDI <- read_csv(here("data/covariates/GenderDevelopmentIndex.csv"))[,-c(2,5)] %>%
    mutate(GDI.unscld = as.numeric(`Gender Development Index`)) %>% 
    select(!`Gender Development Index`) %>%
    rename(study.country = Entity, 
           study.end.year = Year)
### Rename several countries 
GDI[which(GDI$study.country == "Eswatini (Kingdom of)"), "study.country"] <- "Eswatini"
GDI[which(GDI$study.country == "Tanzania (United Republic of)"), "study.country"] <- "Tanzania"

### Join all covariates 
covariates <- full_join(covariates, smoking.prevalence)
covariates <- full_join(covariates, bmi)
covariates[which(covariates$study.country == "United Republic of Tanzania"), "study.country"] <- "Tanzania"

covariates <- full_join(covariates, GDI) %>% 
               mutate(Sex = tolower(sex))

notifPrevRatio2 <- left_join(notifStrat, covariates, by=c("study.country", "study.end.year", "Sex"))

### Now scale covariates 
notifPrevRatio2 <- notifPrevRatio2 %>% 
    mutate(alcohol.use.disorder = scale(alcohol.use.disorder.unscld)[,1], 
           type2.diabetes.mellitus = scale(type2.diabetes.mellitus.unscld)[,1], 
           HIV.AIDS = scale(HIV.AIDS.unscld)[,1],
           smoking = scale(smoking.unscld)[,1], 
           GDI = scale(GDI.unscld)[,1], 
           bmi = scale(bmi.unscld)[,1]
) %>% na.omit()
```

# Univariate analysis

Model specifications 
```{r GDI univariate model specification}

sex_prev_gdi_model <- bf(
    logTbPrev | se(logTbPrevSE, sigma = TRUE) ~ 1 + Sex*GDI + (1 | p | study.country)
)

#notifications model:
sex_notif_gdi_model <- bf(
    logNotifRate | se(logNotifRateSE, sigma = TRUE) ~ 1  + Sex*GDI + (1 | p | study.country)
)

#fit the model
sex_prev_notif_gdi_model <- brm(
    formula = sex_prev_gdi_model + sex_notif_gdi_model + set_rescor(FALSE),
    data = notifPrevRatio2,
    family = gaussian(), 
    # sample_prior = "only",
    prior = c(
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logTbPrev"),
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logNotifRate"),
        prior(normal(0,1), class = b, resp = "logTbPrev"),
        prior(normal(0,1), class = b, resp = "logNotifRate"),
        prior(exponential(2), class = sd, resp = "logTbPrev"),
        prior(exponential(2), class = sd, resp = "logNotifRate")
    ),
    control = list(adapt_delta = 0.99),
    chains = 4,
    cores = 4,
    iter = 10000, 
    warmup = 2500
)


### output for contribution table
gdi.effect <- sex_prev_notif_gdi_model %>% 
             spread_draws(b_logTbPrev_Intercept, b_logNotifRate_Intercept, 
                          b_logTbPrev_Sexmale, b_logNotifRate_Sexmale,
                          b_logTbPrev_GDI, b_logNotifRate_GDI, 
                          `b_logTbPrev_Sexmale:GDI`,`b_logNotifRate_Sexmale:GDI`) %>%
            mutate(gdi.female.prev = exp(b_logTbPrev_GDI), 
                   gdi.female.notif = exp(b_logNotifRate_GDI), 
                   gdi.male.prev = exp(b_logTbPrev_GDI + `b_logTbPrev_Sexmale:GDI`), 
                   gdi.male.notif = exp(b_logNotifRate_GDI + `b_logNotifRate_Sexmale:GDI`), 
                   gdi.female.npRatio = gdi.female.notif/gdi.female.prev,
                   gdi.male.npRatio = gdi.male.notif/gdi.male.prev) %>% 
            median_qi(gdi.female.npRatio, gdi.male.npRatio ) %>% 
            select(! c(.width, .point, .interval)) %>% round(digits = 2) %>%
           matrix(nrow=2, byrow = TRUE) 

gdi.effect

gdi.loo <- loo(sex_prev_notif_gdi_model)
```

```{r alcohol univariate model specification}

sex_prev_alc_model <- bf(
    logTbPrev | se(logTbPrevSE, sigma = TRUE) ~ 1 + Sex*alcohol.use.disorder + (1 | p | study.country)
)

#notifications model:
sex_notif_alc_model <- bf(
    logNotifRate | se(logNotifRateSE, sigma = TRUE) ~ 1  + Sex*alcohol.use.disorder + (1 | p | study.country)
)

#fit the model
sex_prev_notif_alc_model <- brm(
    formula = sex_prev_alc_model + sex_notif_alc_model + set_rescor(FALSE),
    data = notifPrevRatio2,
    family = gaussian(), 
    # sample_prior = "only",
    prior = c(
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logTbPrev"),
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logNotifRate"),
        prior(normal(0,1), class = b, resp = "logTbPrev"),
        prior(normal(0,1), class = b, resp = "logNotifRate"),
        prior(exponential(2), class = sd, resp = "logTbPrev"),
        prior(exponential(2), class = sd, resp = "logNotifRate")
    ),
    control = list(adapt_delta = 0.99),
    chains = 4,
    cores = 4,
    iter = 10000, 
    warmup = 2500
)


### output for contribution table
alc.effect <- sex_prev_notif_alc_model %>% 
             spread_draws(b_logTbPrev_Intercept, b_logNotifRate_Intercept, 
                          b_logTbPrev_Sexmale, b_logNotifRate_Sexmale,
                          b_logTbPrev_alcohol.use.disorder, b_logNotifRate_alcohol.use.disorder, 
                          `b_logTbPrev_Sexmale:alcohol.use.disorder`,`b_logNotifRate_Sexmale:alcohol.use.disorder`) %>%
            mutate(alc.female.prev = exp(b_logTbPrev_alcohol.use.disorder), 
                   alc.female.notif = exp(b_logNotifRate_alcohol.use.disorder), 
                   alc.male.prev = exp(b_logTbPrev_alcohol.use.disorder + `b_logTbPrev_Sexmale:alcohol.use.disorder`), 
                   alc.male.notif = exp(b_logNotifRate_alcohol.use.disorder + `b_logNotifRate_Sexmale:alcohol.use.disorder`), 
                   alc.female.npRatio = alc.female.notif/alc.female.prev,
                   alc.male.npRatio = alc.male.notif/alc.male.prev) %>% 
            median_qi(alc.female.npRatio, alc.male.npRatio ) %>% 
            select(! c(.width, .point, .interval)) %>% round(digits = 2) %>%
           matrix(nrow=2, byrow = TRUE) 

alc.effect

alc.loo <- loo(sex_prev_notif_alc_model)

```

```{r smoking univariate model specification}

sex_prev_smoke_model <- bf(
    logTbPrev | se(logTbPrevSE, sigma = TRUE) ~ 1 + Sex*smoking + (1 | p | study.country)
)

#notifications model:
sex_notif_smoke_model <- bf(
    logNotifRate | se(logNotifRateSE, sigma = TRUE) ~ 1  + Sex*smoking + (1 | p | study.country)
)

#fit the model
sex_prev_notif_smoke_model <- brm(
    formula = sex_prev_smoke_model + sex_notif_smoke_model + set_rescor(FALSE),
    data = notifPrevRatio2,
    family = gaussian(), 
    # sample_prior = "only",
    prior = c(
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logTbPrev"),
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logNotifRate"),
        prior(normal(0,1), class = b, resp = "logTbPrev"),
        prior(normal(0,1), class = b, resp = "logNotifRate"),
        prior(exponential(2), class = sd, resp = "logTbPrev"),
        prior(exponential(2), class = sd, resp = "logNotifRate")
    ),
    control = list(adapt_delta = 0.99),
    chains = 4,
    cores = 4,
    iter = 10000, 
    warmup = 2500
)


### output for contribution table
smoke.effect <- sex_prev_notif_smoke_model %>% 
             spread_draws(b_logTbPrev_Intercept, b_logNotifRate_Intercept, 
                          b_logTbPrev_Sexmale, b_logNotifRate_Sexmale,
                          b_logTbPrev_smoking, b_logNotifRate_smoking, 
                          `b_logTbPrev_Sexmale:smoking`,`b_logNotifRate_Sexmale:smoking`) %>%
            mutate(smoke.female.prev = exp(b_logTbPrev_smoking), 
                   smoke.female.notif = exp(b_logNotifRate_smoking), 
                   smoke.male.prev = exp(b_logTbPrev_smoking + `b_logTbPrev_Sexmale:smoking`), 
                   smoke.male.notif = exp(b_logNotifRate_smoking + `b_logNotifRate_Sexmale:smoking`), 
                   smoke.female.npRatio = smoke.female.notif/smoke.female.prev,
                   smoke.male.npRatio = smoke.male.notif/smoke.male.prev) %>% 
            median_qi(smoke.female.npRatio, smoke.male.npRatio) %>% 
            select(! c(.width, .point, .interval)) %>% round(digits = 2) %>%
           matrix(nrow=2, byrow = TRUE) 

smoke.effect
```

```{r hiv aids univariate model specification}

sex_prev_hiv_model <- bf(
    logTbPrev | se(logTbPrevSE, sigma = TRUE) ~ 1 + Sex*HIV.AIDS + (1 | p | study.country)
)

#notifications model:
sex_notif_hiv_model <- bf(
    logNotifRate | se(logNotifRateSE, sigma = TRUE) ~ 1  + Sex*HIV.AIDS + (1 | p | study.country)
)

#fit the model
sex_prev_notif_hiv_model <- brm(
    formula = sex_prev_hiv_model + sex_notif_hiv_model + set_rescor(FALSE),
    data = notifPrevRatio2,
    family = gaussian(), 
    # sample_prior = "only",
    prior = c(
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logTbPrev"),
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logNotifRate"),
        prior(normal(0,1), class = b, resp = "logTbPrev"),
        prior(normal(0,1), class = b, resp = "logNotifRate"),
        prior(exponential(2), class = sd, resp = "logTbPrev"),
        prior(exponential(2), class = sd, resp = "logNotifRate")
    ),
    control = list(adapt_delta = 0.99),
    chains = 4,
    cores = 4,
    iter = 10000, 
    warmup = 2500
)


### output for contribution table
hiv.effect <- sex_prev_notif_hiv_model %>% 
             spread_draws(b_logTbPrev_Intercept, b_logNotifRate_Intercept, 
                          b_logTbPrev_Sexmale, b_logNotifRate_Sexmale,
                          b_logTbPrev_HIV.AIDS, b_logNotifRate_HIV.AIDS, 
                          `b_logTbPrev_Sexmale:HIV.AIDS`,`b_logNotifRate_Sexmale:HIV.AIDS`) %>%
            mutate(hiv.female.prev = exp(b_logTbPrev_HIV.AIDS), 
                   hiv.female.notif = exp(b_logNotifRate_HIV.AIDS), 
                   hiv.male.prev = exp(b_logTbPrev_HIV.AIDS + `b_logTbPrev_Sexmale:HIV.AIDS`), 
                   hiv.male.notif = exp(b_logNotifRate_HIV.AIDS + `b_logNotifRate_Sexmale:HIV.AIDS`), 
                   hiv.female.npRatio = hiv.female.notif/hiv.female.prev,
                   hiv.male.npRatio = hiv.male.notif/hiv.male.prev) %>% 
            median_qi(hiv.female.npRatio, hiv.male.npRatio ) %>% 
            select(! c(.width, .point, .interval)) %>% round(digits = 2) %>%
           matrix(nrow=2, byrow = TRUE) 

hiv.effect
```

```{r diabetes univariate model specification}

sex_prev_diab_model <- bf(
    logTbPrev | se(logTbPrevSE, sigma = TRUE) ~ 1 + Sex*type2.diabetes.mellitus + (1 | p | study.country)
)

#notifications model:
sex_notif_diab_model <- bf(
    logNotifRate | se(logNotifRateSE, sigma = TRUE) ~ 1  + Sex*type2.diabetes.mellitus + (1 | p | study.country)
)

#fit the model
sex_prev_notif_diab_model <- brm(
    formula = sex_prev_diab_model + sex_notif_diab_model + set_rescor(FALSE),
    data = notifPrevRatio2,
    family = gaussian(), 
    # sample_prior = "only",
    prior = c(
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logTbPrev"),
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logNotifRate"),
        prior(normal(0,1), class = b, resp = "logTbPrev"),
        prior(normal(0,1), class = b, resp = "logNotifRate"),
        prior(exponential(2), class = sd, resp = "logTbPrev"),
        prior(exponential(2), class = sd, resp = "logNotifRate")
    ),
    control = list(adapt_delta = 0.99),
    chains = 4,
    cores = 4,
    iter = 10000, 
    warmup = 2500
)


### output for contribution table
diab.effect <- sex_prev_notif_diab_model %>% 
             spread_draws(b_logTbPrev_Intercept, b_logNotifRate_Intercept, 
                          b_logTbPrev_Sexmale, b_logNotifRate_Sexmale,
                          b_logTbPrev_type2.diabetes.mellitus, b_logNotifRate_type2.diabetes.mellitus, 
                          `b_logTbPrev_Sexmale:type2.diabetes.mellitus`,`b_logNotifRate_Sexmale:type2.diabetes.mellitus`) %>%
            mutate(diab.female.prev = exp(b_logTbPrev_type2.diabetes.mellitus), 
                   diab.female.notif = exp(b_logNotifRate_type2.diabetes.mellitus), 
                   diab.male.prev = exp(b_logTbPrev_type2.diabetes.mellitus + `b_logTbPrev_Sexmale:type2.diabetes.mellitus`), 
                   diab.male.notif = exp(b_logNotifRate_type2.diabetes.mellitus + `b_logNotifRate_Sexmale:type2.diabetes.mellitus`), 
                   diab.female.npRatio = diab.female.notif/diab.female.prev,
                   diab.male.npRatio = diab.male.notif/diab.male.prev) %>% 
            median_qi(diab.female.npRatio, diab.male.npRatio ) %>% 
            select(! c(.width, .point, .interval)) %>% round(digits = 2) %>%
           matrix(nrow=2, byrow = TRUE) 

diab.effect

```

```{r bmi univariate model specification}

sex_prev_bmi_model <- bf(
    logTbPrev | se(logTbPrevSE, sigma = TRUE) ~ 1 + Sex*bmi + (1 | p | study.country)
)

#notifications model:
sex_notif_bmi_model <- bf(
    logNotifRate | se(logNotifRateSE, sigma = TRUE) ~ 1  + Sex*bmi + (1 | p | study.country)
)

#fit the model
sex_prev_notif_bmi_model <- brm(
    formula = sex_prev_bmi_model + sex_notif_bmi_model + set_rescor(FALSE),
    data = notifPrevRatio2,
    family = gaussian(), 
    # sample_prior = "only",
    prior = c(
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logTbPrev"),
        prior(student_t(7, 0, 1.5), class = Intercept, resp = "logNotifRate"),
        prior(normal(0,1), class = b, resp = "logTbPrev"),
        prior(normal(0,1), class = b, resp = "logNotifRate"),
        prior(exponential(2), class = sd, resp = "logTbPrev"),
        prior(exponential(2), class = sd, resp = "logNotifRate")
    ),
    control = list(adapt_delta = 0.99),
    chains = 4,
    cores = 4,
    iter = 10000, 
    warmup = 2500
)


### output for contribution table
bmi.effect <- sex_prev_notif_bmi_model %>% 
             spread_draws(b_logTbPrev_Intercept, b_logNotifRate_Intercept, 
                          b_logTbPrev_Sexmale, b_logNotifRate_Sexmale,
                          b_logTbPrev_bmi, b_logNotifRate_bmi, 
                          `b_logTbPrev_Sexmale:bmi`,`b_logNotifRate_Sexmale:bmi`) %>%
            mutate(bmi.female.prev = exp(b_logTbPrev_bmi), 
                   bmi.female.notif = exp(b_logNotifRate_bmi), 
                   bmi.male.prev = exp(b_logTbPrev_bmi + `b_logTbPrev_Sexmale:bmi`), 
                   bmi.male.notif = exp(b_logNotifRate_bmi + `b_logNotifRate_Sexmale:bmi`), 
                   bmi.female.npRatio = bmi.female.notif/bmi.female.prev,
                   bmi.male.npRatio = bmi.male.notif/bmi.male.prev) %>% 
            median_qi(bmi.female.npRatio, bmi.male.npRatio) %>% 
            select(! c(.width, .point, .interval)) %>% round(digits = 2) %>%
           matrix(nrow=2, byrow = TRUE) 

bmi.effect
```

```{r combined univariate}
rbind(cbind(unlist(year.effect[1,]), unlist(year.effect[2,])),
      cbind(unlist(gdi.effect[1,]), unlist(gdi.effect[2,])),
      cbind(unlist(alc.effect[1,]), unlist(alc.effect[2,])),
      cbind(unlist(diab.effect[1,]), unlist(diab.effect[2,])),
      cbind(unlist(hiv.effect[1,]), unlist(hiv.effect[2,])),
      cbind(unlist(smoke.effect[1,]), unlist(smoke.effect[2,])),
      cbind(unlist(bmi.effect[1,]), unlist(bmi.effect[2,]))
)



```

# Multivariate analysis

Model specifications 
```{r Multivariate model specifications}
sex_prev_multivariate_model <- bf(
    logTbPrev | se(logTbPrevSE, sigma = TRUE) ~ 1 + Sex + 
                                                    GDI + Sex:GDI + 
                                                    year_z + Sex:year_z +
                                                    alcohol.use.disorder + Sex:alcohol.use.disorder + 
                                                    smoking + Sex:smoking +
                                                    HIV.AIDS + Sex:HIV.AIDS + 
                                                    type2.diabetes.mellitus + Sex:type2.diabetes.mellitus +
                                                    bmi + Sex:bmi +
                                                (1 | p | study.country)
)

#notifications model:
sex_notif_multivariate_model <- bf(
    logNotifRate | se(logNotifRateSE, sigma = TRUE) ~ 1 + Sex + 
                                                          year_z + Sex:year_z +
                                                          GDI + Sex:GDI + 
                                                          alcohol.use.disorder + Sex:alcohol.use.disorder + 
                                                          smoking + Sex:smoking +
                                                          HIV.AIDS + Sex:HIV.AIDS + 
                                                          type2.diabetes.mellitus + Sex:type2.diabetes.mellitus +
                                                          bmi + Sex:bmi +
                                                        (1 | p | study.country)
)

sex_prev_notif_multi_model <- brm(
                formula = sex_prev_multivariate_model + sex_notif_multivariate_model + set_rescor(FALSE),
                data = notifPrevRatio2,
                family = gaussian(), 
                # sample_prior = "only",
                prior = c(
                    prior(student_t(7, 0, 1.5), class = Intercept, resp = "logTbPrev"),
                    prior(student_t(7, 0, 1.5), class = Intercept, resp = "logNotifRate"),
                    prior(normal(0,1), class = b, resp = "logTbPrev"),
                    prior(normal(0,1), class = b, resp = "logNotifRate"),
                    prior(exponential(2), class = sd, resp = "logTbPrev"),
                    prior(exponential(2), class = sd, resp = "logNotifRate")
                ),
                control = list(adapt_delta = 0.99),
                chains = 4,
                cores = 4,
                iter = 10000, 
                warmup = 2500
            )

### output for contribution table
multi.effect <- sex_prev_notif_multi_model %>% 
             spread_draws(b_logTbPrev_Intercept, b_logNotifRate_Intercept, 
                          b_logTbPrev_Sexmale, b_logNotifRate_Sexmale,
                           b_logTbPrev_year_z, b_logNotifRate_year_z, 
                          `b_logTbPrev_Sexmale:year_z`,`b_logNotifRate_Sexmale:year_z`, 
                           b_logTbPrev_GDI, b_logNotifRate_GDI, 
                          `b_logTbPrev_Sexmale:GDI`,`b_logNotifRate_Sexmale:GDI`, 
                           b_logTbPrev_alcohol.use.disorder, b_logNotifRate_alcohol.use.disorder, 
                          `b_logTbPrev_Sexmale:alcohol.use.disorder`,`b_logNotifRate_Sexmale:alcohol.use.disorder`, 
                           b_logTbPrev_smoking, b_logNotifRate_smoking, 
                          `b_logTbPrev_Sexmale:smoking`,`b_logNotifRate_Sexmale:smoking`, 
                           b_logTbPrev_HIV.AIDS, b_logNotifRate_HIV.AIDS, 
                          `b_logTbPrev_Sexmale:HIV.AIDS`,`b_logNotifRate_Sexmale:HIV.AIDS`, 
                           b_logTbPrev_type2.diabetes.mellitus, b_logNotifRate_type2.diabetes.mellitus, 
                          `b_logTbPrev_Sexmale:type2.diabetes.mellitus`,`b_logNotifRate_Sexmale:type2.diabetes.mellitus`,                               b_logTbPrev_bmi, b_logNotifRate_bmi, 
                          `b_logTbPrev_Sexmale:bmi`,`b_logNotifRate_Sexmale:bmi`, 
                          ) %>%
            mutate(intcpt.female.prev = exp(b_logTbPrev_Intercept), 
                   intcpt.female.notif = exp(b_logNotifRate_Intercept), 
                   intcpt.male.prev = exp(b_logTbPrev_Intercept + b_logTbPrev_year_z), 
                   intcpt.male.notif = exp(b_logNotifRate_Intercept + b_logNotifRate_Sexmale), 
                   intcpt.female.npRatio = intcpt.female.notif/intcpt.female.prev,
                   intcpt.male.npRatio = intcpt.male.notif/intcpt.male.prev,
                
                   year.female.prev = exp(b_logTbPrev_year_z), 
                   year.female.notif = exp(b_logNotifRate_year_z), 
                   year.male.prev = exp(b_logTbPrev_year_z + `b_logTbPrev_Sexmale:year_z`), 
                   year.male.notif = exp(b_logNotifRate_year_z + `b_logNotifRate_Sexmale:year_z`), 
                   year.female.npRatio = year.female.notif/year.female.prev,
                   year.male.npRatio = year.male.notif/year.male.prev,
                
                
                   gdi.female.prev = exp(b_logTbPrev_GDI), 
                   gdi.female.notif = exp(b_logNotifRate_GDI), 
                   gdi.male.prev = exp(b_logTbPrev_GDI + `b_logTbPrev_Sexmale:GDI`), 
                   gdi.male.notif = exp(b_logNotifRate_GDI + `b_logNotifRate_Sexmale:GDI`), 
                   gdi.female.npRatio = gdi.female.notif/gdi.female.prev,
                   gdi.male.npRatio = gdi.male.notif/gdi.male.prev,
                   
                   alc.female.prev = exp(b_logTbPrev_alcohol.use.disorder), 
                   alc.female.notif = exp(b_logNotifRate_alcohol.use.disorder), 
                   alc.male.prev = exp(b_logTbPrev_alcohol.use.disorder + 
                                           `b_logTbPrev_Sexmale:alcohol.use.disorder`), 
                   alc.male.notif = exp(b_logNotifRate_alcohol.use.disorder +
                                            `b_logNotifRate_Sexmale:alcohol.use.disorder`), 
                   alc.female.npRatio = alc.female.notif/alc.female.prev,
                   alc.male.npRatio = alc.male.notif/alc.male.prev, 
                
                   smoke.female.prev = exp(b_logTbPrev_smoking), 
                   smoke.female.notif = exp(b_logNotifRate_smoking), 
                   smoke.male.prev = exp(b_logTbPrev_smoking + `b_logTbPrev_Sexmale:smoking`), 
                   smoke.male.notif = exp(b_logNotifRate_smoking + `b_logNotifRate_Sexmale:smoking`), 
                   smoke.female.npRatio = smoke.female.notif/smoke.female.prev,
                   smoke.male.npRatio = smoke.male.notif/smoke.male.prev,
                
                   hiv.female.prev = exp(b_logTbPrev_HIV.AIDS), 
                   hiv.female.notif = exp(b_logNotifRate_HIV.AIDS), 
                   hiv.male.prev = exp(b_logTbPrev_HIV.AIDS + `b_logTbPrev_Sexmale:HIV.AIDS`), 
                   hiv.male.notif = exp(b_logNotifRate_HIV.AIDS + `b_logNotifRate_Sexmale:HIV.AIDS`), 
                   hiv.female.npRatio = hiv.female.notif/hiv.female.prev,
                   hiv.male.npRatio = hiv.male.notif/hiv.male.prev,
                   
                   diab.female.prev = exp(b_logTbPrev_type2.diabetes.mellitus), 
                   diab.female.notif = exp(b_logNotifRate_type2.diabetes.mellitus), 
                   diab.male.prev = exp(b_logTbPrev_type2.diabetes.mellitus +
                                            `b_logTbPrev_Sexmale:type2.diabetes.mellitus`), 
                   diab.male.notif = exp(b_logNotifRate_type2.diabetes.mellitus +
                                             `b_logNotifRate_Sexmale:type2.diabetes.mellitus`), 
                   diab.female.npRatio = diab.female.notif/diab.female.prev,
                   diab.male.npRatio = diab.male.notif/diab.male.prev, 
                   
                   bmi.female.prev = exp(b_logTbPrev_bmi), 
                   bmi.female.notif = exp(b_logNotifRate_bmi), 
                   bmi.male.prev = exp(b_logTbPrev_bmi +
                                            `b_logTbPrev_Sexmale:bmi`), 
                   bmi.male.notif = exp(b_logNotifRate_bmi +
                                             `b_logNotifRate_Sexmale:bmi`), 
                   bmi.female.npRatio = bmi.female.notif/bmi.female.prev,
                   bmi.male.npRatio = bmi.male.notif/bmi.male.prev) %>% 
    
            median_qi(intcpt.female.npRatio, intcpt.male.npRatio, 
                      year.female.npRatio, year.male.npRatio,
                      gdi.female.npRatio, gdi.male.npRatio,
                      alc.female.npRatio, alc.male.npRatio, 
                      smoke.female.npRatio, smoke.male.npRatio, 
                      hiv.female.npRatio, hiv.male.npRatio, 
                      diab.female.npRatio, diab.male.npRatio, 
                      bmi.female.npRatio, bmi.male.npRatio
                      ) %>% 
            select(! c(.width, .point, .interval)) %>% round(digits = 2) %>%
           matrix(nrow=8, byrow = TRUE) 

cbind(c("intercept", "year", "GDI", "alcohol", "smoking", "hiv", "diabetes", "bmi"),  multi.effect)

loo_saturated_og <- loo(sex_prev_notif_multi_model)

```
Country-level effects

```{r}
ndCountry <- notifPrevModel2 %>%
             select(study.country, logOR,
                    mfRatioNotif, 
                    id, year_z, #GDI,
                    alcohol.use.disorder,
                    type2.diabetes.mellitus,
                    # bmi, smoking, 
                    HIV.AIDS) %>%
             mutate(logOR_se=1)

countryEffectModEst <- add_epred_draws(object = multiNotifMod,
                                 newdata=ndCountry,
                                 re_formula = ~ 1  + (1|study.country),
                                 allow_new_levels = TRUE) %>%
                    mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

countryEffectModEst %>% group_by(study.country) %>% mean_qi(mfRatio)
#write_csv(countryEffectModEst %>% group_by(study.country) %>% mean_qi(mfRatio), here("output/countryMFRatios.csv"))
countryEffectModEst %>% group_by(study.country) %>% mean_qi(mfRatio) %>% select(study.country, mfRatio) -> countryRatios

### Setup a world map
world_map <- map_data("world")
world_map <- subset(world_map, region != "Antarctica")

world <- map_data("world")
world <- world[world$region != "Antarctica", ]

countryMap <- left_join(countryRatios %>% rename(region=study.country),
                         world) %>% rename(study.country = region)
library(geosphere)
centroids <- world_map %>%
    group_by(region) %>%
    group_modify(~ data.frame(centroid(cbind(.x$long, .x$lat))))

nCountriesGeo <- left_join(countryMap, centroids %>% rename(study.country = region))

ggplot(countryMap) +
    geom_map(
        dat = world_map, map = world_map, aes(map_id = region),
        fill = "lightgrey", color = "black", size = 0.25) +
    geom_map(dat = countryMap, map = world_map, aes(map_id = study.country, fill = mfRatio), color="black") +
    scale_fill_gradient(low = "pink", high = "darkred",
                        name = "Male-to-female ratio") +
    expand_limits(x = world_map$long, y = world_map$lat) +
    theme_void() + theme(legend.position = "inside",
                            legend.position.inside = c(.15,.2))

```

### Main Model Specification 
```{r Main Model Specification, warning=FALSE, message=FALSE}

notifMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                                 1  + offset(log(mfRatioNotif)) +
                                 (1 | study.country) + (1 | id),
                             data = notifPrevRatio,
                             prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                                 prior(exponential(2), class = sd),
                             family = gaussian(),
                             control = list(adapt_delta = 0.99),
                             chains = 4,
                             cores = 4,
                             iter = 4000)

```


### Offset approach
### Main Model Examination
```{r Main Model Examination, echo=FALSE, warning=FALSE, message=FALSE}
summary(notifMod)
```

### Main Model Estimation
```{r Main Model Estimation, echo=FALSE, warning=FALSE, message=FALSE}
nd <- notifPrevRatio %>%
      select(study.country, logOR, mfRatioNotif) %>%
      mutate(logOR_se=1)

mainEffectModEst <- add_epred_draws(object = notifMod,
                                 newdata=nd,
                                 re_formula = ~ 1 + offset(log(mfRatioNotif)), 
                                 allow_new_levels = TRUE) %>% 
                    mutate(pnRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

mainEffectModEst %>% ungroup() %>% mean_qi(pnRatio, na.rm=T)

### output for contribution table
exp(as.numeric(summary(notifMod)[["fixed"]][1]))
exp(as.numeric(summary(notifMod)[["fixed"]][3]))
exp(as.numeric(summary(notifMod)[["fixed"]][4]))
```

### Country-level Estimation
```{r Country-level Estimation, echo=FALSE, warning=FALSE, message=FALSE}
ndCountry<- notifPrevRatio %>%
    select(study.country, logOR) %>%
    mutate(logOR_se=1, 
           mfRatioNotif =1)

countryEffectModEst <- add_epred_draws(object = notifMod,
                                      newdata=ndCountry,
                                      re_formula = ~ 1 + offset(log(mfRatioNotif)) +
                                                     (1 | study.country),
                                      allow_new_levels = TRUE) %>% 
    mutate(pnRatio = exp(.epred),
           type="Posterior mean and 95% credible interval") 

ggplot(countryEffectModEst) + 
    geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
    stat_pointinterval(aes(x=reorder(study.country, pnRatio), y=pnRatio), .width = 0.95, size=5) + 
    theme_minimal() + 
    # scale_x_discrete(limits=rev) + 
    scale_y_log10(
        breaks = c(0.25, 0.5, 1, 2, 4, 6),
        labels = c("0.25", "0.5", "1", "2", "4", "6"),
    ) + coord_flip() + 
    labs(
        x = NULL,
        y = "Prevalence to notification ratio of TB prevalence"
    ) 

countryEffectModEst %>% group_by(study.country) %>% mean_qi(pnRatio)

ggsave(here("output/mfPNratios.png"), plot = last_plot(), 
       width = 8.5, height=10, units = "in")
```
