---
title: "Sex Stratified Relationships of Tuberculosis Prevalence and Notification in Low- and Middle-income Countries: A Systematic Review and Meta-analysis" 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

ABOUT THIS SCRIPT

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multivariate multilevel model on the studies that report sex stratified prevalence results. This model aims to estimate sex stratified prevalence adjusted for geographic factors. 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(posterior)
library(tidybayes)

library(ggplot2)

expit <- function (x) {exp(x)/(1 + exp(x))}
##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################
source(here("code/bacterialPositiveIndicator.R"))
cleanSexDF0 <- bactPostIndicator("sex") %>% filter(sex.analysis.indicator !="none")

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexDF <- cleanSexDF0 %>% 
    filter(sex.analysis.indicator %in% c("adj.prev100k.ci.bacteriological.tb",
                                         "adj.prev100k.ci.smear.positive.tb",
                                         "prev100k.ci.bacteriological.tb",
                                         "prev100k.ci.smear.positive.tb")) %>% 
    mutate("id" = row_number(),
           adj.prev100k.bacteriological.tb.male = adj.prev100k.bacteriological.tb.male/1e5,
           adj.prev100k.ci.upper.bacteriological.tb.male = adj.prev100k.ci.upper.bacteriological.tb.male/1e5,
           adj.prev100k.ci.lower.bacteriological.tb.male = adj.prev100k.ci.lower.bacteriological.tb.male/1e5,
           
           adj.prev100k.bacteriological.tb.female = adj.prev100k.bacteriological.tb.female/1e5,
           adj.prev100k.ci.upper.bacteriological.tb.female = adj.prev100k.ci.upper.bacteriological.tb.female/1e5,
           adj.prev100k.ci.lower.bacteriological.tb.female = adj.prev100k.ci.lower.bacteriological.tb.female/1e5,
           
           adj.prev100k.smear.positive.tb.male = adj.prev100k.smear.positive.tb.male/1e5,
           adj.prev100k.ci.upper.smear.positive.tb.male = adj.prev100k.ci.upper.smear.positive.tb.male/1e5,
           adj.prev100k.ci.lower.smear.positive.tb.male = adj.prev100k.ci.lower.smear.positive.tb.male/1e5,
           
           adj.prev100k.smear.positive.tb.female = adj.prev100k.smear.positive.tb.female/1e5,
           adj.prev100k.ci.upper.smear.positive.tb.female = adj.prev100k.ci.upper.smear.positive.tb.female/1e5,
           adj.prev100k.ci.lower.smear.positive.tb.female = adj.prev100k.ci.lower.smear.positive.tb.female/1e5,
           
           prev100k.bacteriological.tb.male = prev100k.bacteriological.tb.male/1e5,
           prev100k.ci.upper.bacteriological.tb.male = prev100k.ci.upper.bacteriological.tb.male/1e5,
           prev100k.ci.lower.bacteriological.tb.male = prev100k.ci.lower.bacteriological.tb.male/1e5,
           
           prev100k.bacteriological.tb.female = prev100k.bacteriological.tb.female/1e5,
           prev100k.ci.upper.bacteriological.tb.female = prev100k.ci.upper.bacteriological.tb.female/1e5,
           prev100k.ci.lower.bacteriological.tb.female = prev100k.ci.lower.bacteriological.tb.female/1e5,
           
           prev100k.smear.positive.tb.male = prev100k.smear.positive.tb.male/1e5,
           prev100k.ci.upper.smear.positive.tb.male = prev100k.ci.upper.smear.positive.tb.male/1e5,
           prev100k.ci.lower.smear.positive.tb.male = prev100k.ci.lower.smear.positive.tb.male/1e5,
           
           prev100k.smear.positive.tb.female = prev100k.smear.positive.tb.female/1e5,
           prev100k.ci.upper.smear.positive.tb.female = prev100k.ci.upper.smear.positive.tb.female/1e5,
           prev100k.ci.lower.smear.positive.tb.female = prev100k.ci.lower.smear.positive.tb.female/1e5) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           WHO.region,
           study.start.year, 
           sex.analysis.indicator,
           ends_with("male"), ends_with("sex")) %>%
    
    mutate("tempPrevMale" =  case_when(sex.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tempPrevFemale" = case_when(sex.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female)) %>%
    pivot_longer(cols = c(tempPrevMale,
                          tempPrevFemale), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.bacteriological.tb.male-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.bacteriological.tb.female-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.smear.positive.tb.male-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.smear.positive.tb.female-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.bacteriological.tb.male-
                                                                 prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.bacteriological.tb.female-
                                                                   prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.smear.positive.tb.male-
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.smear.positive.tb.female-
                                                                   prev100k.ci.lower.smear.positive.tb.female)/3.92), 
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.male/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.male/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.female/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.male/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.male /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.female/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.female)))/3.92,
                                              sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Male" ~ (log(prev100k.ci.upper.bacteriological.tb.male/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.male)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.male/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Female" ~ (log(prev100k.ci.upper.bacteriological.tb.female/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.female)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.female /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Male" ~ (log(prev100k.ci.upper.smear.positive.tb.male/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.male)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.male /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Female" ~ (log(prev100k.ci.upper.smear.positive.tb.female/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.female)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.female /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.female)))/3.92
           ))

```


Model building 

## Model 1: Only survey level random effects

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
model1 <- brms::brm(formula = `LogOdds`|se(`LogOddsStandardError`, 
                                           sigma=TRUE) ~ 1 + (1|id) + Sex,
                    data = cleanSexDF,
                    prior = prior(cauchy(231/1e5, 90/1e5), class = Intercept),
                    family = "gaussian",
                    control = list(adapt_delta = 0.99),
                    cores = 4,
                    chains = 4,
                    iter = 4000)
```

Model diagnostics  

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(model1)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
plot(model1)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
pairs(model1)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
pp_check(model1, ndraws = 100)
```

Model results

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
model1_est0 <- model1 %>% 
               as_draws_df() %>%
               select(b_Intercept, b_SexMale) %>%
               transmute(Female = expit(b_Intercept)*1e5, 
                         Male = expit(b_Intercept + b_SexMale)*1e5) %>%
               mutate(mfRatio = Male/Female, 
                      model = "Only random effects")

model1_est <- melt(model1_est0)

model1_est_summary <- model1_est %>% reframe(meanPrev = mean(value), .by = c(variable)); model1_est_summary

ggplot(data = model1_est) +
    stat_halfeye(aes(x=value, fill=model), normalize="panels", alpha=0.5, size =20) +
    facet_grid(model~fct_relevel(variable,
                                 "Male", "Female", "mfRatio"), 
               scales = "free_x") +
    labs(x="", y="") +
    theme_minimal(base_size = 30) + theme(legend.position = "bottom")
```

## Model 2: survey level random effects + WHO region level fixed effects

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
model2 <- brms::brm(formula = `LogOdds`|se(`LogOddsStandardError`, 
                                           sigma=TRUE) ~ 1 + Sex + WHO.region + (1|id),
                    data = cleanSexDF,
                    prior = prior(cauchy(231/1e5, 90/1e5), class = Intercept),
                    family = "gaussian",
                    control = list(adapt_delta = 0.90),
                    cores = 4,
                    chains = 4,
                    iter = 4000)
```

Model diagnostics  
```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(model2)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
plot(model2)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
pairs(model2)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
pp_check(model2, ndraws = 100)
```

Model results

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
model2_est0 <- model2 %>% 
    add_epred_draws(newdata = crossing(WHO.region = unique(cleanSexDF$WHO.region), 
                                       Sex = c("Male", "Female"),
                                       LogOddsStandardError = 1),
                    re_formula = NA) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    dplyr::select(WHO.region, Sex, .draw, value) %>%
    pivot_wider(names_from = Sex, values_from = value) %>%
    mutate(mfRatio = Male/Female) %>%
    select(-.draw) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>%
    mutate(model = "Country level fixed effects") %>%
    rename(Sex = name)

model2_est_summary <- model2_est0 %>% reframe(meanPrev = mean(value), .by = c(Sex, WHO.region)); model2_est_summary

# ggplot(data = model2_est0) + geom_density(aes(x=value, fill = WHO.region, group=WHO.region))

ggplot(data = model2_est0) +
    stat_pointinterval(aes(x=value, colour=WHO.region), position = position_dodge(width=0.5), size =20) +
    facet_grid(model~fct_relevel(Sex,
                                 "Male prev", "Female prev", "mfRatio"),
               scales = "free_x") +
    labs(x="", y="") + 
    theme_minimal( base_size = 30) + theme(legend.position = "bottom")
```

## Model 3: survey level random effects + study level effects (multi-level model)

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
model3 <- brms::brm(formula = LogOdds | se(LogOddsStandardError, 
                                           sigma=TRUE) ~ 1 + Sex + (1|id/title.extracted),
                    data = cleanSexDF,
                    prior = prior(cauchy(231/1e5, 90/1e5), class = Intercept),
                    family = "gaussian",
                    control = list(adapt_delta = 0.99),
                    cores = 4,
                    chains = 4,
                    iter = 4000)
```

Model diagnostics  

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
summary(model3)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
plot(model3)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pairs(model3)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pp_check(model3, ndraws =100)
```

Model results 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
model3_est0 <- model3 %>% 
               as_draws_df() %>%
               select(b_Intercept, b_SexMale) %>%
               transmute(Female = expit(b_Intercept)*1e5, 
                         Male = expit(b_Intercept + b_SexMale)*1e5) %>%
               mutate(mfRatio = Male/Female, 
                      model = "survey level random effects + study level effects (multi-level model)")

model3_est <- melt(model3_est0)
model3_est_summary <- model3_est %>% reframe(meanPrev = mean(value), .by = c(variable)); model3_est_summary


ggplot(data = model3_est) +
    stat_halfeye(aes(x=value, fill=model), normalize="panels", alpha=0.5, size =20) +
    facet_grid(model~fct_relevel(variable,
                                 "Male", "Female", "mfRatio"), 
               scales = "free_x") +
    labs(x="", y="") +
    theme_minimal(base_size = 30) + theme(legend.position = "bottom")

```

## Model 4: random intercepts for countries nested within regions, and surveys nested within studies

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
model4 <- brms::brm(formula = LogOdds | se(LogOddsStandardError, 
                                             sigma=TRUE) ~ 1 + Sex + (1 + Sex |WHO.region/study.country) + (1 + Sex|id/title.extracted),
                      data = cleanSexDF,
                      prior = prior(cauchy(231/1e5, 90/1e5), class = Intercept),
                      family = "gaussian",
                      control = list(adapt_delta = 0.99),
                      cores = 4,
                      chains = 4,
                      iter = 4000)
```

Model diagnostics  

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
summary(model4)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
plot(model4)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pairs(model4)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pp_check(model4, ndraws =100)
```

Model results 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
model4_est <- model4 %>%
    add_epred_draws(newdata = cleanSexDF, re_formula = NULL) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    select(study.country, WHO.region, Sex, value, .draw) %>%
    pivot_wider(
        names_from = Sex,
        values_from = value
    ) %>%
    unnest_longer(col = c(Male, Female)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>% 
    rename(Sex = name)

# model3_est <- melt(model3_est0)
# model3_est_summary <- model3_est %>% reframe(meanPrev = mean(value), .by = c(variable)); model3_est_summary


ggplot(data = model4_est) +
        stat_pointinterval(aes(x=value, y=study.country, color = WHO.region), .width = 0.95, size=20) +
        facet_grid(.~fct_relevel(Sex,
                                     "Male", "Female", "mfRatio"), scale = "free_x") + 
        labs(x="", y="") +
        theme_minimal(base_size = 30) + theme(legend.position = "bottom")

```