---
title: "Sex Stratified Relationships of Tuberculosis Prevalence and Notification in Low- and Middle-income Countries: A Systematic Review and Meta-analysis" 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

ABOUT THIS SCRIPT

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multivariate multilevel model on the studies that report sex stratified prevalence results. This model aims to estimate sex stratified prevalence adjusted for geographic factors. 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
library(posterior)
library(tidybayes)

library(ggplot2)

expit <- function (x) {exp(x)/(1 + exp(x))}

myPal <- c("black", "violet")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))

##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################


cleanSexDF0 <- bactPostIndicator("sex") %>% filter(sex.analysis.indicator !="none")
```

# Descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(cleanSexDF0)[1], "unique surveys with sex-stratified bacteriological positive TB"))

# sort(unique(cleanSexDF0$title.extracted))
print(paste(length(unique(cleanSexDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanSexDF0$n.participants.female)), "studies report participant numbers totaling", 
            sum(cleanSexDF0$n.participants.female, cleanSexDF0$n.participants.male, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanSexDF0$n.participants.male, na.rm=TRUE)/sum(cleanSexDF0$n.participants.female, cleanSexDF0$n.participants.male, na.rm = TRUE)*100, 2), "% of participants were male" ))
```

# Risk of bias statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE}
table(cleanSexDF0$study.quality.summary)
```

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexDF <-  cleanSexDF0 %>% 
               filter(sex.analysis.indicator != "none") %>% 
               mutate("id" = row_number()) %>% 
    ### Scale all prevalence estimates back to decimals 
    mutate_at(vars(matches('prev100k.bacteriological.tb.|prev100k.ci.([a-z]{3})er.bacteriological.tb.|prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           WHO.region,
           screening.tests,
           study.start.year, 
           study.end.year, 
           study.quality.summary,
           ends_with("male"), ends_with("sex"),
           sex.analysis.indicator) %>%
    
    mutate("tempPrevMale" =  case_when(sex.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.male/n.participants.male, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.male/n.participants.male,
                                        sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tempPrevFemale" = case_when(sex.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.female/n.participants.female, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female)) %>%
    pivot_longer(cols = c(tempPrevFemale,tempPrevMale), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.bacteriological.tb.male-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.bacteriological.tb.female-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.smear.positive.tb.male-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.smear.positive.tb.female-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.bacteriological.tb.male-
                                                                 prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.bacteriological.tb.female-
                                                                   prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.smear.positive.tb.male-
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.smear.positive.tb.female-
                                                                   prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.male),
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.female)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.male/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.male/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.female/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.male/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.male /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.female/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.female)))/3.92,
                                              sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Male" ~ (log(prev100k.ci.upper.bacteriological.tb.male/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.male)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.male/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Female" ~ (log(prev100k.ci.upper.bacteriological.tb.female/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.female)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.female /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Male" ~ (log(prev100k.ci.upper.smear.positive.tb.male/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.male)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.male /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Female" ~ (log(prev100k.ci.upper.smear.positive.tb.female/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.female)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.female /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.female)))/3.92, 
                                            sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            log(1/(n.participants.male*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            log(1/(n.participants.female*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       year_z = study.start.year - round(mean(study.start.year, na.rm=TRUE)),
                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       symptomScreen = ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No"),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))

##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("sex")

bactModelDF <- cleanSexDF %>%
    select(id, WHO.region, study.country, Sex, year_z, natRep, bias, symptomScreen, LogOdds, LogOddsStandardError, sex.analysis.indicator) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(logOR = LogOdds_Male - LogOdds_Female,
           unadj_logOR_se = sqrt(LogOddsStandardError_Male^2 + LogOddsStandardError_Female^2), 
           logOR_se = ifelse(grepl("adj", sex.analysis.indicator), 
                                     unadj_logOR_se,
                                     ifelse(grepl("smear", sex.analysis.indicator),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Bact Variance Median"]))))

```

# Model 1: Main effect model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|study.country) + (1|id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectMod)
plot(mainEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- bactModelDF %>%
      select(study.country, LogOdds_Female) %>%
      mutate(logOR_se=1)

mainEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=nd,
                                 re_formula = ~ 1, 
                                 allow_new_levels = TRUE) %>% 
                    mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectModEst %>% ungroup %>% mean_qi(mfRatio)
```

# Model 2: Effect by world region model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regionEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(regionEffectMod)
plot(regionEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndRegion <- bactModelDF %>%
      select(WHO.region, LogOdds_Female) %>%
      mutate(logOR_se=1)

regionEffectModEst <- add_epred_draws(object = regionEffectMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region),
                                 allow_new_levels = TRUE) %>% 
                       mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectModEst %>% group_by(WHO.region) %>% mean_qi(mfRatio)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
### Get the empirical values 
empirical_prev <- bactModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        bias = bias,
        symptomScreen = symptomScreen, 
        p_female_empirical = plogis(LogOdds_Female),
        p_male_empirical = plogis(LogOdds_Male),
        or_empirical = p_male_empirical / p_female_empirical,
        or_empirical = (p_male_empirical / (1 - p_male_empirical)) / 
            (p_female_empirical / (1 - p_female_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific TB prevalence")

    ggplot() +
    # geom_hline(aes(yintercept = 1), linetype=2) +
    geom_point(data = empirical_prev, aes(y=WHO.region, x=or_empirical, color = type2),alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectModEst, aes(x=mfRatio, y=WHO.region, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio",
         y = "World region") +
    theme_bw(base_size = 30) + theme(legend.title = element_blank(), legend.position = c(.83,.1), 
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
    
```

# Model 3: Effect over time 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
timeEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z + (1 + year_z | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(timeEffectMod)
plot(timeEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndTime <- bactModelDF %>%
    select(logOR_se) %>%
    crossing(year_z = seq(min(cleanSexDF$year_z, na.rm = TRUE), max(cleanSexDF$year_z, na.rm = TRUE), 1))

timeEffectModEst <- add_epred_draws(object = timeEffectMod,
                                 newdata=ndTime,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")
```

Main estimate
```{r}
timeEffectModEst %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper)

```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    timeEffectModEst %>%
    ggplot(aes(x = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)),
               y = mfRatio)) +
    stat_lineribbon(width = 0.95, alpha = 0.25) +
    geom_point(data = empirical_prev,
               aes(x=year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = logOR_se),
               colour = "black") +
    geom_hline(yintercept = 1, linetype = 2) +
    labs(y = "Male-to-female odds ratio",
         x = "Year") +
    theme_minimal(base_size = 30)  + theme(legend.position = "bottom")
```


# Model 4: Region effect over time 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
timeEffectRegionMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                              1 + year_z * WHO.region + (1 + year_z | study.country) + (1 | id), 
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(timeEffectRegionMod)
plot(timeEffectRegionMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndTimeReg <- bactModelDF %>%
             select(logOR_se, WHO.region) %>%
             distinct() %>%
             crossing(year_z = seq(min(cleanSexDF$year_z, na.rm = TRUE),
                                   max(cleanSexDF$year_z, na.rm = TRUE), 1))

timeEffectRegionModEst <- add_epred_draws(object = timeEffectRegionMod,
                                 newdata=ndTimeReg,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")
```

Main estimate
```{r}
timeEffectRegionModEst %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper)

```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    timeEffectRegionModEst %>%
    filter(WHO.region !="Americas region",
           WHO.region !="Eastern Mediterranean region") %>%
    ggplot(aes(x = year_z + round(mean(cleanSexDF$study.end.year, na.rm = TRUE)),
               y = mfRatio,
               fill = WHO.region,
               color = WHO.region)) +
    stat_lineribbon(aes(color = WHO.region, fill = WHO.region), .width = 0.95, alpha = 0.25) +
    geom_point(data = empirical_prev %>%
                   filter(WHO.region !="Americas region",
                          WHO.region !="Eastern Mediterranean region"),
               aes(x=year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = logOR_se),
               colour = "black") +
    geom_hline(yintercept = 1, linetype = 2) +
    facet_wrap(~ WHO.region) +
    labs(y = "Male-to-female odds ratio!",
         x = "Year",
         fill = "WHO region",
         color = "WHO region") +
    theme_minimal(base_size = 30)  + theme(legend.position = "bottom")
```

# Univariate analysis of other covariates 

## Model 6: National versus subnational 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
natEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + natRep + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(natEffectMod)
plot(natEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndNat <- bactModelDF %>%
    select(logOR_se, natRep)

natEffectModEst <- add_epred_draws(object = natEffectMod,
                                 newdata=ndNat,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
natEffectModEst %>% 
    group_by(natRep) %>% 
    mean_qi(mfRatio) %>% 
    select(natRep, mfRatio, .lower, .upper)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    ggplot() +
    # geom_hline(aes(yintercept = 1), linetype=2) +
    geom_point(data = empirical_prev, aes(y=natRep, x=or_empirical, color = type2, size=logOR_se), alpha = 0.5) +
    stat_pointinterval(data=natEffectModEst, aes(x=mfRatio, y=natRep, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio", color = "",
         y = "Survey representativeness") +
    scale_size_continuous(range = c(8,16)) +
    theme_bw(base_size = 30) + theme(
                                     legend.position = c(.83,.2), legend.box.just = "right",
                                     # legend.direction = "horizontal",
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
```

## Model 6: Study quality (High vs. not high)

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}

biasEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + bias + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(biasEffectMod)
plot(biasEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndBias <- bactModelDF %>%
    select(logOR_se, bias)

biasEffectModEst <- add_epred_draws(object = biasEffectMod,
                                 newdata=ndBias,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
biasEffectModEst %>% 
    group_by(bias) %>% 
    mean_qi(mfRatio) %>% 
    select(bias, mfRatio, .lower, .upper)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    ggplot() +
    geom_point(data = empirical_prev, aes(y=bias, x=or_empirical, color = type2, size=logOR_se), alpha = 0.5) +
    stat_pointinterval(data=biasEffectModEst, aes(x=mfRatio, y=bias, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio", color = "",
         y = "Low probability of survey bias vs. not low probability of survey bias") +
    scale_size_continuous(range = c(8,16)) +
    theme_bw(base_size = 30) + theme(
                                     legend.position = c(.83,.2), legend.box.just = "right",
                                     # legend.direction = "horizontal",
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
```

## Model 7: Symptom screen vs. no symptom screen

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}

sympEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + symptomScreen + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(sympEffectMod)
plot(sympEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndSymp <- bactModelDF %>%
    select(logOR_se, symptomScreen)

sympEffectModEst <- add_epred_draws(object = sympEffectMod,
                                 newdata=ndSymp,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
sympEffectModEst %>% 
    group_by(symptomScreen) %>% 
    mean_qi(mfRatio) %>% 
    select(symptomScreen, mfRatio, .lower, .upper)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    ggplot() +
    geom_point(data = empirical_prev, aes(y=symptomScreen, x=or_empirical, color = type2, size=logOR_se), alpha = 0.5) +
    stat_pointinterval(data=sympEffectModEst, aes(x=mfRatio, y=symptomScreen, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio", color = "",
         y = "Symptom screening required") +
    scale_size_continuous(range = c(8,16)) +
    theme_bw(base_size = 30) + theme(
                                     legend.position = c(.83,.2), legend.box.just = "right",
                                     # legend.direction = "horizontal",
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
```

# Smear positive analysis 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
cleanSexSmrDF0 <- smrPosIndicator("sex") %>% filter(sex.smr.analysis.indicator !="none")

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexSmrDF <- cleanSexSmrDF0 %>% 
                 filter(sex.smr.analysis.indicator != "none") %>% 
                 mutate("id" = row_number()) %>%
                 ### Scale all prevalence estimates back to decimals 
                 mutate_at(vars(matches('prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>% 
                 select(id,
                       covidence.id, 
                       figure.id.yr, 
                       study.geography, 
                       title.extracted,
                       study.country,
                       WHO.region,
                       study.start.year, 
                       study.end.year, 
                       ends_with("male"), ends_with("sex"),
                       sex.smr.analysis.indicator) %>%
    
    mutate("tempPrevMale" =  case_when(sex.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.smr.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                        sex.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tempPrevFemale" = case_when(sex.smr.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.smr.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female)) %>%
    pivot_longer(cols = c(tempPrevFemale,tempPrevMale), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.smear.positive.tb.male-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.smear.positive.tb.female-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                       
                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.smear.positive.tb.male-
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.smear.positive.tb.female-
                                                                   prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.smr.analysis.indicator %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.male),
                                        sex.smr.analysis.indicator  %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.female)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.male/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.male /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                              sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.female/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.female)))/3.92,
                                                 sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Male" ~ (log(prev100k.ci.upper.smear.positive.tb.male/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.male)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.male /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                                 sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Female" ~ (log(prev100k.ci.upper.smear.positive.tb.female/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.female)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.female /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.female)))/3.92, 
                                            sex.smr.analysis.indicator %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            log(1/(n.participants.male*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        sex.smr.analysis.indicator  %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            log(1/(n.participants.female*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       year_z = study.start.year - round(mean(study.start.year, na.rm=TRUE)),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))  %>% filter(LogOddsStandardError < Inf)

smrModelDF <- cleanSexSmrDF %>%
    select(id, WHO.region, study.country, Sex, year_z, natRep, LogOdds, LogOddsStandardError) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_Male - LogOdds_Female,
        logOR_se = sqrt(LogOddsStandardError_Male^2 + LogOddsStandardError_Female^2) 
    ) 
```

# Smear positive descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(cleanSexSmrDF0)[1], "unique surveys with sex-stratified smear positive TB"))

# sort(unique(cleanSexDF0$title.extracted))
print(paste(length(unique(cleanSexSmrDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanSexSmrDF0$n.participants.female)), "studies report participant numbers totaling", 
            sum(cleanSexSmrDF0$n.participants.female, cleanSexSmrDF0$n.participants.male, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanSexSmrDF0$n.participants.male, na.rm=TRUE)/sum(cleanSexSmrDF0$n.participants.female, cleanSexSmrDF0$n.participants.male, na.rm = TRUE)*100, 2), "% of participants were male" ))
```

# Model X: Smear positive main effect 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | study.country) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectSmrMod)
plot(mainEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- smrModelDF %>%
      select(study.country, LogOdds_Female) %>%
      mutate(logOR_se=1)

mainEffectSmrModEst <- add_epred_draws(object = mainEffectSmrMod,
                                 newdata=nd,
                                 re_formula = ~ 1) %>% 
                    mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectSmrModEst %>% ungroup %>% mean_qi(mfRatio)
```

# Model 2: Smear positive effect by world region model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regionEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(regionEffectSmrMod)
plot(regionEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndRegion <- smrModelDF %>%
      select(WHO.region, LogOdds_Female) %>%
      mutate(logOR_se=1)

regionEffectSmrModEst <- add_epred_draws(object = regionEffectSmrMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region)) %>% 
                       mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectSmrModEst %>% group_by(WHO.region) %>% mean_qi(mfRatio)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
### Get the empirical values 
empirical_smr_prev <- smrModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        p_female_empirical = plogis(LogOdds_Female),
        p_male_empirical = plogis(LogOdds_Male),
        or_empirical = p_male_empirical / p_female_empirical,
        or_empirical = (p_male_empirical / (1 - p_male_empirical)) / 
            (p_female_empirical / (1 - p_female_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific smear positive TB prevalence")

    ggplot() +
    # geom_hline(aes(yintercept = 1), linetype=2) +
    geom_point(data = empirical_smr_prev, aes(y=WHO.region, x=or_empirical, color = type2),alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectSmrModEst, aes(x=mfRatio, y=WHO.region, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio",
         y = "World region") +
    theme_bw(base_size = 30) + theme(legend.title = element_blank(), legend.position = c(.83,.07), 
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
    
```