---
title: "Sex Stratified Relationships of Tuberculosis Prevalence and Notification in Low- and Middle-income Countries: A Systematic Review and Meta-analysis" 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

ABOUT THIS SCRIPT

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multivariate multilevel model on the studies that report sex stratified prevalence results. This model aims to estimate sex stratified prevalence adjusted for geographic factors. 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(posterior)
library(tidybayes)

library(ggplot2)

expit <- function (x) {exp(x)/(1 + exp(x))}
##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################
source(here("code/bacterialPositiveIndicator.R"))
cleanSexDF0 <- bactPostIndicator("sex") %>% filter(sex.analysis.indicator !="none")

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexDF <- cleanSexDF0 %>% 
    # filter(sex.analysis.indicator %in% c("adj.prev100k.ci.bacteriological.tb",
    #                                      "adj.prev100k.ci.smear.positive.tb",
    #                                      "prev100k.ci.bacteriological.tb",
    #                                      "prev100k.ci.smear.positive.tb")) %>% 
        filter(sex.analysis.indicator != "none") %>% 
    mutate("id" = row_number(),
           adj.prev100k.bacteriological.tb.male = adj.prev100k.bacteriological.tb.male/1e5,
           adj.prev100k.ci.upper.bacteriological.tb.male = adj.prev100k.ci.upper.bacteriological.tb.male/1e5,
           adj.prev100k.ci.lower.bacteriological.tb.male = adj.prev100k.ci.lower.bacteriological.tb.male/1e5,
           
           adj.prev100k.bacteriological.tb.female = adj.prev100k.bacteriological.tb.female/1e5,
           adj.prev100k.ci.upper.bacteriological.tb.female = adj.prev100k.ci.upper.bacteriological.tb.female/1e5,
           adj.prev100k.ci.lower.bacteriological.tb.female = adj.prev100k.ci.lower.bacteriological.tb.female/1e5,
           
           adj.prev100k.smear.positive.tb.male = adj.prev100k.smear.positive.tb.male/1e5,
           adj.prev100k.ci.upper.smear.positive.tb.male = adj.prev100k.ci.upper.smear.positive.tb.male/1e5,
           adj.prev100k.ci.lower.smear.positive.tb.male = adj.prev100k.ci.lower.smear.positive.tb.male/1e5,
           
           adj.prev100k.smear.positive.tb.female = adj.prev100k.smear.positive.tb.female/1e5,
           adj.prev100k.ci.upper.smear.positive.tb.female = adj.prev100k.ci.upper.smear.positive.tb.female/1e5,
           adj.prev100k.ci.lower.smear.positive.tb.female = adj.prev100k.ci.lower.smear.positive.tb.female/1e5,
           
           prev100k.bacteriological.tb.male = prev100k.bacteriological.tb.male/1e5,
           prev100k.ci.upper.bacteriological.tb.male = prev100k.ci.upper.bacteriological.tb.male/1e5,
           prev100k.ci.lower.bacteriological.tb.male = prev100k.ci.lower.bacteriological.tb.male/1e5,
           
           prev100k.bacteriological.tb.female = prev100k.bacteriological.tb.female/1e5,
           prev100k.ci.upper.bacteriological.tb.female = prev100k.ci.upper.bacteriological.tb.female/1e5,
           prev100k.ci.lower.bacteriological.tb.female = prev100k.ci.lower.bacteriological.tb.female/1e5,
           
           prev100k.smear.positive.tb.male = prev100k.smear.positive.tb.male/1e5,
           prev100k.ci.upper.smear.positive.tb.male = prev100k.ci.upper.smear.positive.tb.male/1e5,
           prev100k.ci.lower.smear.positive.tb.male = prev100k.ci.lower.smear.positive.tb.male/1e5,
           
           prev100k.smear.positive.tb.female = prev100k.smear.positive.tb.female/1e5,
           prev100k.ci.upper.smear.positive.tb.female = prev100k.ci.upper.smear.positive.tb.female/1e5,
           prev100k.ci.lower.smear.positive.tb.female = prev100k.ci.lower.smear.positive.tb.female/1e5) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           WHO.region,
           study.start.year, 
           study.end.year, 
           ends_with("male"), ends_with("sex"),
           sex.analysis.indicator) %>%
    
    mutate("tempPrevMale" =  case_when(sex.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.male/n.participants.male, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.male/n.participants.male,
                                        sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tempPrevFemale" = case_when(sex.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.female/n.participants.female, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female)) %>%
    pivot_longer(cols = c(tempPrevFemale,tempPrevMale), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.bacteriological.tb.male-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.bacteriological.tb.female-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.smear.positive.tb.male-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.smear.positive.tb.female-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.bacteriological.tb.male-
                                                                 prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.bacteriological.tb.female-
                                                                   prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.smear.positive.tb.male-
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.smear.positive.tb.female-
                                                                   prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.male),
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.female)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.male/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.male/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.female/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.male/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.male /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.female/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.female)))/3.92,
                                              sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Male" ~ (log(prev100k.ci.upper.bacteriological.tb.male/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.male)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.male/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Female" ~ (log(prev100k.ci.upper.bacteriological.tb.female/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.female)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.female /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Male" ~ (log(prev100k.ci.upper.smear.positive.tb.male/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.male)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.male /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Female" ~ (log(prev100k.ci.upper.smear.positive.tb.female/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.female)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.female /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.female)))/3.92, 
                                            sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            log(1/(n.participants.male*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            log(1/(n.participants.female*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2))
```


Model building 

## Model 1: Only survey level random effects

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
model1 <- brms::brm(formula = `LogOdds`|se(`LogOddsStandardError`, 
                                           sigma=TRUE) ~ 1 + (1|id) + Sex,
                    data = cleanSexDF,
                    prior = prior(normal(0,10), class=Intercept) +
                            prior(normal(0, 10), class = b) +
                            prior(exponential(1), class=sd), 
                    family = "gaussian",
                    control = list(adapt_delta = 0.99),
                    cores = 4,
                    chains = 4,
                    iter = 4000)

model1rev <- brms::brm(formula = `LogOdds`|se(`revisedLogOddsStandardErr`, 
                                           sigma=TRUE) ~ 1 + (1|id) + Sex,
                    data = cleanSexDF,
                    prior = prior(normal(0,10), class=Intercept) +
                            prior(normal(0, 10), class = b) +
                            prior(exponential(1), class=sd), 
                    family = "gaussian",
                    control = list(adapt_delta = 0.99),
                    cores = 4,
                    chains = 4,
                    iter = 4000)
```

Model diagnostics  

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(model1)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
plot(model1)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
pairs(model1)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
pp_check(model1, ndraws = 100)
```

Model results

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
model1_est0 <- model1 %>% 
               as_draws_df() %>%
               select(b_Intercept, b_SexMale) %>%
               transmute(Female = expit(b_Intercept)*1e5, 
                         Male = expit(b_Intercept + b_SexMale)*1e5) %>%
               mutate(mfRatio = Male/Female, 
                      model = "Only random effects")

model1_est <- melt(model1_est0)

model1_est_summary <- model1_est %>% reframe(meanPrev = mean(value), .by = c(variable)); model1_est_summary

ggplot(data = model1_est) +
    stat_halfeye(aes(x=value, fill=model), normalize="panels", alpha=0.5, size =20) +
    facet_grid(model~fct_relevel(variable,
                                 "Male", "Female", "mfRatio"), 
               scales = "free_x") +
    labs(x="", y="") +
    theme_minimal(base_size = 30) + theme(legend.position = "bottom")
```

## Model 2: survey level random effects + WHO region level fixed effects

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
model2 <- brms::brm(formula = `LogOdds`|se(`LogOddsStandardError`, 
                                           sigma=TRUE) ~ 1 + Sex + WHO.region + (1|id),
                    data = cleanSexDF,
                    prior = prior(normal(0,10), class=Intercept) +
                            prior(normal(0, 10), class = b) +
                            prior(exponential(1), class=sd),
                    family = "gaussian",
                    control = list(adapt_delta = 0.90),
                    cores = 4,
                    chains = 4,
                    iter = 4000)
```

Model diagnostics  
```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(model2)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
plot(model2)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
pairs(model2)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
pp_check(model2, ndraws = 100)
```

Model results

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
model2_est0 <- model2 %>% 
    add_epred_draws(newdata = crossing(WHO.region = unique(cleanSexDF$WHO.region), 
                                       Sex = c("Male", "Female"),
                                       LogOddsStandardError = 1),
                    re_formula = NA) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    dplyr::select(WHO.region, Sex, .draw, value) %>%
    pivot_wider(names_from = Sex, values_from = value) %>%
    mutate(mfRatio = Male/Female) %>%
    select(-.draw) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>%
    mutate(model = "Country level fixed effects") %>%
    rename(Sex = name)

model2_est_summary <- model2_est0 %>% reframe(meanPrev = mean(value), .by = c(Sex, WHO.region)); model2_est_summary

# ggplot(data = model2_est0) + geom_density(aes(x=value, fill = WHO.region, group=WHO.region))

ggplot(data = model2_est0) +
    stat_pointinterval(aes(x=value, colour=WHO.region), position = position_dodge(width=0.5), size =20) +
    facet_grid(model~fct_relevel(Sex,
                                 "Male prev", "Female prev", "mfRatio"),
               scales = "free_x") +
    labs(x="", y="") + 
    theme_minimal( base_size = 30) + theme(legend.position = "bottom")
```

## Model 3: survey level random effects + study level effects (multi-level model)

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
model3 <- brms::brm(formula = LogOdds | se(LogOddsStandardError, 
                                           sigma=TRUE) ~ 1 + Sex + (1|id/title.extracted),
                    data = cleanSexDF,
                    prior = prior(normal(0,10), class=Intercept) +
                            prior(normal(0, 10), class = b) +
                            prior(exponential(1), class=sd),
                    family = "gaussian",
                    control = list(adapt_delta = 0.99),
                    cores = 4,
                    chains = 4,
                    iter = 4000)
```

Model diagnostics  

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
summary(model3)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
plot(model3)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pairs(model3)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pp_check(model3, ndraws =100)
```

Model results 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
model3_est0 <- model3 %>% 
               as_draws_df() %>%
               select(b_Intercept, b_SexMale) %>%
               transmute(Female = expit(b_Intercept)*1e5, 
                         Male = expit(b_Intercept + b_SexMale)*1e5) %>%
               mutate(mfRatio = Male/Female, 
                      model = "survey level random effects + study level effects (multi-level model)")

model3_est <- melt(model3_est0)
model3_est_summary <- model3_est %>% reframe(meanPrev = mean(value), .by = c(variable)); model3_est_summary


ggplot(data = model3_est) +
    stat_halfeye(aes(x=value, fill=model), normalize="panels", alpha=0.5, size =20) +
    facet_grid(model~fct_relevel(variable,
                                 "Male", "Female", "mfRatio"), 
               scales = "free_x") +
    labs(x="", y="") +
    theme_minimal(base_size = 30) + theme(legend.position = "bottom")

```

## Model 4: random intercepts for countries nested within regions, and surveys nested within studies

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
model4 <- brms::brm(formula = LogOdds | se(LogOddsStandardError, 
                                             sigma=TRUE) ~ 1 + Sex + (1 + Sex |WHO.region/study.country) +
                    (1 + Sex|id/title.extracted),
                      data = cleanSexDF,
                      prior = prior(normal(0,10), class=Intercept) +
                          prior(normal(0, 10), class = b) +
                          prior(exponential(1), class=sd) +
                          prior(lkj(4), class=cor),
                      family = "gaussian",
                      control = list(adapt_delta = 0.99),
                      cores = 4,
                      chains = 4,
                      iter = 4000)

### What happens to precision when we remove the WHO.region

model4b <- brms::brm(formula = LogOdds | se(LogOddsStandardError, 
                                             sigma=TRUE) ~ 1 + Sex + (1 + Sex |study.country) + (1 + Sex|id/title.extracted),
                      data = cleanSexDF,
                      prior = prior(normal(0,10), class=Intercept) +
                          prior(normal(0, 10), class = b) +
                          prior(exponential(1), class=sd) +
                          prior(lkj(4), class=cor),
                      family = "gaussian",
                      control = list(adapt_delta = 0.99),
                      cores = 4,
                      chains = 4,
                      iter = 4000)

model4c <- brms::brm(formula = LogOdds | se(LogOddsStandardError, 
                                             sigma=TRUE) ~ 1 + Sex + (1 + Sex |study.country) + (1 + Sex|id),
                      data = cleanSexDF,
                      prior = prior(normal(0,10), class=Intercept) +
                          prior(normal(0, 10), class = b) +
                          prior(exponential(1), class=sd) +
                          prior(lkj(4), class=cor),
                      family = "gaussian",
                      control = list(adapt_delta = 0.99),
                      cores = 4,
                      chains = 4,
                      iter = 4000)

model4d <- brms::brm(formula = LogOdds | se(LogOddsStandardError, 
                                             sigma=TRUE) ~ 1 + Sex + (1 + Sex |WHO.region/study.country) + (1 + Sex|id),
                      data = cleanSexDF,
                      prior = prior(normal(0,10), class=Intercept) +
                          prior(normal(0, 10), class = b) +
                          prior(exponential(1), class=sd) +
                          prior(lkj(4), class=cor),
                      family = "gaussian",
                      control = list(adapt_delta = 0.99),
                      cores = 4,
                      chains = 4,
                      iter = 4000)
```

Model diagnostics  

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
summary(model4)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
plot(model4)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pairs(model4)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pp_check(model4, ndraws =100)
```

Model results 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- cleanSexDF %>%
    select(WHO.region, study.country, Sex) %>%
    distinct() %>%
    mutate(LogOddsStandardError=1)
           
model4_est <- model4 %>%
    add_epred_draws(newdata = nd, re_formula = ~(1 + Sex |WHO.region/study.country)) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    select(study.country, WHO.region, Sex, value, .draw) %>%
    pivot_wider(
        names_from = Sex,
        values_from = value
    ) %>%
    unnest_longer(col = c(Male, Female)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>% 
    rename(Sex = name)

model4b_est <- model4 %>%
    add_epred_draws(newdata = nd, re_formula = ~(1 + Sex | study.country)) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    select(study.country, WHO.region, Sex, value, .draw) %>%
    pivot_wider(
        names_from = Sex,
        values_from = value
    ) %>%
    unnest_longer(col = c(Male, Female)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>% 
    rename(Sex = name)


model4_Mainest <- model4 %>%
    add_epred_draws(newdata = nd, re_formula = ~(1 + Sex)) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    select(study.country, WHO.region, Sex, value, .draw) %>%
    pivot_wider(
        names_from = Sex,
        values_from = value
    ) %>%
    unnest_longer(col = c(Male, Female)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>% 
    rename(Sex = name)

model4b_Mainest <- model4b %>%
    add_epred_draws(newdata = nd, re_formula = ~(1 + Sex)) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    select(study.country, WHO.region, Sex, value, .draw) %>%
    pivot_wider(
        names_from = Sex,
        values_from = value
    ) %>%
    unnest_longer(col = c(Male, Female)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>% 
    rename(Sex = name)


ggplot(data = model4_est) +
        stat_pointinterval(aes(x=value, y=study.country, color = WHO.region), .width = 0.95, size=20) +
        facet_grid(.~fct_relevel(Sex,
                                     "Male", "Female", "mfRatio"), scale = "free_x") + 
        labs(x="", y="") + expand_limits(x=0) + 
        theme_minimal(base_size = 30) + theme(legend.position = "bottom")

model4_estb <- model4 %>%
    add_epred_draws(newdata = nd, re_formula = ~(1 + Sex |WHO.region)) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    select(study.country, WHO.region, Sex, value, .draw) %>%
    pivot_wider(
        names_from = Sex,
        values_from = value
    ) %>%
    unnest_longer(col = c(Male, Female)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>% 
    rename(Sex = name)


ggplot(data = model4_estb) +
        stat_pointinterval(aes(x=value, y=WHO.region), .width = 0.95, size=20) +
        facet_grid(.~fct_relevel(Sex,
                                     "Male", "Female", "mfRatio"), scale = "free_x") + 
        labs(x="", y="") + #expand_limits(x=0) + 
        theme_minimal(base_size = 30) + theme(legend.position = "bottom")

```

## Model 5: random intercepts for countries nested within regions, and surveys nested within studies; add in survey end years

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
model5 <- brms::brm(formula = LogOdds | se(LogOddsStandardError, 
                                             sigma=TRUE) ~ 1 + Sex + (1 + Sex |WHO.region/study.country) + (1 + Sex|id/title.extracted) + (1 + Sex | study.end.year),
                      data = cleanSexDF,
                      prior = prior(normal(0,10), class=Intercept) +
                          prior(normal(0, 10), class = b) +
                          prior(exponential(1), class=sd) +
                          prior(lkj(4), class=cor),
                      family = "gaussian",
                      control = list(adapt_delta = 0.99),
                      cores = 4,
                      chains = 4,
                      iter = 4000)

```
Model diagnostics  

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
summary(model5)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
plot(model5)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pairs(model5)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE,fig.height = 20, fig.width = 28}
pp_check(model5, ndraws =100)
```

Model results 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- cleanSexDF %>%
    filter(!is.na(study.end.year)) %>%
    select(Sex, study.country, id, WHO.region, study.end.year, title.extracted) %>%
    distinct() %>%
    mutate(LogOddsStandardError=1)
           
model5_est <- model5 %>%
    add_epred_draws(newdata = nd, re_formula = ~ (1 + Sex | study.end.year)) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    select(Sex, study.end.year,  value, .draw) %>%
    pivot_wider(
        names_from = Sex,
        values_from = value
    ) %>%
    unnest_longer(col = c(Male, Female)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>% 
    rename(Sex = name) %>%
    mutate(study.end.year = as.factor(study.end.year))


ggplot(data = model5_est) +
        stat_pointinterval(aes(x=value, y=study.end.year), .width = 0.95, size=20) +
        facet_grid(.~fct_relevel(Sex,
                                     "Male", "Female", "mfRatio"), scale = "free_x") + 
        labs(x="", y="") + expand_limits(x=0) + 
        theme_minimal(base_size = 30) + theme(legend.position = "bottom")

model5_estb <- model5 %>%
    add_epred_draws(newdata = nd, re_formula = ~ (1 + Sex | study.end.year) + (1 + Sex | WHO.region)) %>%
    ungroup() %>%
    mutate(value = expit(.epred)*1e5) %>%
    select(Sex, study.end.year, WHO.region, value, .draw) %>%
    pivot_wider(
        names_from = Sex,
        values_from = value
    ) %>%
    unnest_longer(col = c(Male, Female)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Female, Male, mfRatio)) %>% 
    rename(Sex = name) %>%
    mutate(study.end.year = as.factor(study.end.year))


ggplot(data = model5_estb %>% filter(Sex != "mfRatio")) +
        stat_pointinterval(aes(x=value, y=study.end.year), .width = 0.95, size=20) +
        facet_grid(.~fct_relevel(Sex,
                                     "Male", "Female", "mfRatio") ~ WHO.region, scale = "free_x") + 
        labs(x="", y="") + expand_limits(x=0) + 
        theme_minimal(base_size = 30) + theme(legend.position = "bottom")

ggplot(data = model5_estb %>% filter(Sex == "mfRatio")) +
        stat_pointinterval(aes(x=value, y=study.end.year), .width = 0.95, size=20) +
        facet_grid(.~fct_relevel(Sex,
                                     "Male", "Female", "mfRatio") ~ WHO.region, scale = "free_x") + 
        labs(x="", y="") + expand_limits(x=0) + 
        theme_minimal(base_size = 30) + theme(legend.position = "bottom")

```


Peter MacP: 7th April 2025

Modify Model 5 above to 
 a) Look at priors
 b) model sigma
 c) Look at year effect
 
Need to centre year 

```{r}

cleanSexDF <- cleanSexDF %>%
    mutate(year_z =study.end.year - round(mean(study.end.year, na.rm=TRUE)))

cleanSexDF %>%
    ggplot(aes(x=year_z)) +
    geom_histogram()


```




```{r}

set.seed(123)

data.frame(theta = plogis(ggdist::rstudent_t(n=20000, df=7, mu=0, sigma=1.5))) |>
  bayesplot::mcmc_hist() +
  xlim(c(0,1)) +
  labs(title='student_t(7, 0, 1.5) for Intercept')

```


Fit a model, sampling from the prior distributions only (i.e. not including data)


```{r}

model6_priors <- brms::brm(formula = LogOdds | se(LogOddsStandardError, sigma=TRUE) ~ 
                               1 + Sex + year_z +  (1 + Sex |WHO.region/study.country),
                               sigma ~ Sex + year_z,
                      data = cleanSexDF,
                      prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                          prior(normal(0, 10), class = b) +
                          prior(exponential(2), class = sd) +
                          prior(lkj(2), class=cor),
                      family = "gaussian",
                      control = list(adapt_delta = 0.99),
                      sample_prior = "only",
                      cores = 4,
                      chains = 4,
                      iter = 4000)


summary(model6_priors)
plot(model6_priors)
conditional_effects(model6_priors)

```
Now fit the model with the data
 
```{r}

model6 <- brms::brm(formula = LogOdds | se(LogOddsStandardError, sigma=TRUE) ~ 
                               1 + Sex*year_z +  (1 + Sex*year_z |WHO.region/study.country),
                               sigma ~ Sex*year_z,
                      data = cleanSexDF,
                      prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                          prior(normal(0, 10), class = b) +
                          prior(exponential(2), class = sd) +
                          prior(lkj(4), class=cor),
                      family = "gaussian",
                      control = list(adapt_delta = 0.99),
                      cores = 4,
                      chains = 4,
                      iter = 4000)


summary(model6)
plot(model6)
conditional_effects(model6)

```

Overall sex effect, ignoring random effects.
 
```{r}

as_draws_df(model6) %>%
    select(b_Intercept, b_SexMale) %>%
    transmute(Female = expit(b_Intercept)*1e5, 
              Male = expit(b_Intercept + b_SexMale)*1e5) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(everything()) %>%
    group_by(name) %>%
    mean_qi()



```
 
Random effect for country

```{r}

nd_country <-  cleanSexDF %>%
    select(id, WHO.region, study.country, Sex, year_z) %>%
    distinct() %>%
    mutate(LogOddsStandardError = 1)


model6_draws_country <- 
    add_epred_draws(object = model6,
                    newdata = nd_country,
                    re_formula = ~(1 +  Sex*year_z | WHO_region/study.country)) %>%
    ungroup() %>%
    group_by(id, study.country) %>%
    mutate(value = expit(.epred) * 1e5) %>%
    select(id, study.country, Sex, value, .draw) %>%
    ungroup()

model6_draws_country <- model6_draws_country %>%
    pivot_wider(names_from = Sex, values_from = value) %>%
    mutate(mfRatio = Male / Female) %>%
    pivot_longer(cols = c(Male, Female, mfRatio)) 


model6_draws_country %>%
    group_by(study.country, name) %>%
    mean_qi(value)


#EMPERICAL VALUES
m6_empirical <- cleanSexDF %>%
  select(id, WHO.region, study.country, Sex,LogOdds) %>%
  mutate(value = expit(LogOdds)*1e5) %>%
    pivot_wider(names_from=Sex,
                values_from = value,
                id_cols = c(id, study.country, WHO.region)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Male, Female, mfRatio))

model6_draws_country %>%
  ggplot() +
    geom_vline(data = m6_empirical %>% filter(name=="mfRatio"),
               aes(xintercept = 1), colour="grey50", linetype=2) +
  stat_pointinterval(aes(x = value, y = study.country), .width = 0.95) +
  geom_point(data = m6_empirical, aes(y = study.country, x = value),
             colour = "mediumseagreen", shape = 1) +
  scale_y_discrete(limits=rev) +
  facet_grid(. ~ fct_relevel(name, "Male", "Female"), scale = "free_x") +
  theme_tidybayes() +
  theme(panel.background = element_rect(fill = NA, colour = "grey78")) +
  labs(x="", y="")


```


Random effect for country

```{r}

nd_region <-  cleanSexDF %>%
    select(id, WHO.region, Sex, year_z) %>%
    distinct() %>%
    mutate(LogOddsStandardError = 1)


model6_draws_region <- 
    add_epred_draws(object = model6,
                    newdata = nd_region,
                    re_formula = ~(1 +  Sex*year_z | WHO_region)) %>%
    ungroup() %>%
    group_by(id, WHO.region) %>%
    mutate(value = expit(.epred) * 1e5) %>%
    select(id, WHO.region, Sex, value, .draw) %>%
    ungroup()

model6_draws_region <- model6_draws_region %>%
    pivot_wider(names_from = Sex, values_from = value) %>%
    mutate(mfRatio = Male / Female) %>%
    pivot_longer(cols = c(Male, Female, mfRatio)) 


model6_draws_region %>%
    group_by(WHO.region, name) %>%
    mean_qi(value, na.rm=TRUE)


#EMPERICAL VALUES
m6_empirical <- cleanSexDF %>%
  select(id, WHO.region, study.country, Sex,LogOdds) %>%
  mutate(value = expit(LogOdds)*1e5) %>%
    pivot_wider(names_from=Sex,
                values_from = value,
                id_cols = c(id, study.country, WHO.region)) %>%
    mutate(mfRatio = Male/Female) %>%
    pivot_longer(cols = c(Male, Female, mfRatio))

model6_draws_region %>%
  ggplot() +
    geom_vline(data = m6_empirical %>% filter(name=="mfRatio"),
               aes(xintercept = 1), colour="grey50", linetype=2) +
  stat_pointinterval(aes(x = value, y = WHO.region), .width = 0.95) +
  geom_point(data = m6_empirical, aes(y = WHO.region, x = value),
             colour = "mediumseagreen", shape = 1) +
  scale_y_discrete(limits=rev) +
  facet_grid(. ~ fct_relevel(name, "Male", "Female"), scale = "free_x") +
  theme_tidybayes() +
  theme(panel.background = element_rect(fill = NA, colour = "grey78")) +
  labs(x="", y="")


```

Experiment a bit more with taking draws and summarising.

```{r}

#updated function to also convert to prev per 100k scale
expit_p100k <- function (x) {(exp(x)/(1 + exp(x)))*1e5}

#get posteriors in rvars format
post_model6_country <- cleanSexDF %>% tidybayes::add_epred_rvars(model6, value = "logodds_epred",
                                                                 re_formula = ~(1 +  Sex*year_z | WHO_region/study.country))

#mutate a new column to get prev scale
post_model6_rvars <- post_model6_country %>%
    select(Sex, WHO.region, study.country, logodds_epred) %>%
    mutate(prev_epred = rvar_apply(.x = logodds_epred, 1, expit_p100k))

#now get the m-to-f ratio
post_model6_rvars_ratio <- post_model6_rvars %>%
  group_by(WHO.region, study.country) %>%
  reframe(
    Male = prev_epred[Sex == "Male"],
    Female = prev_epred[Sex == "Female"],
    mfRatio = Male / Female
  ) %>%
    mutate(id = 1:n())

#pivot and unnesay to dataframe for summary and plotting
post_model6_rvars_long <- post_model6_rvars_ratio %>%
    ungroup() %>%
    pivot_longer(cols = c(Male, Female, mfRatio)) %>%
    unnest_rvars() %>%
    ungroup()

#Country summaries
post_model6_rvars_long %>%
    group_by(study.country, name) %>%
    mean_qi(value, na.rm=TRUE)

#Now plot
post_model6_rvars_long %>%
    ggplot() +
    geom_vline(data = m6_empirical %>% filter(name=="mfRatio"),
               aes(xintercept = 1), colour="grey50", linetype=2) +
    stat_pointinterval(aes(x = value, y = study.country), .width = 0.95) +  
    geom_point(data = m6_empirical, aes(y = study.country, x = value),
             colour = "mediumseagreen", shape = 1) +
    facet_wrap(~name, scale="free_x") +
    scale_y_discrete(limits=rev)


```

```{r}


#get posteriors in rvars format
post_model6_region <- cleanSexDF %>% tidybayes::add_epred_rvars(model6, value = "logodds_epred",
                                                                 re_formula = ~(1 +  Sex*year_z | WHO_region))

#mutate a new column to get prev scale
post_model6_rvars_region <- post_model6_region %>%
    select(Sex, WHO.region, logodds_epred) %>%
    mutate(prev_epred = rvar_apply(.x = logodds_epred, 1, expit_p100k))

#now get the m-to-f ratio
post_model6_rvars_ratio_region <- post_model6_rvars_region %>%
  group_by(WHO.region) %>%
  reframe(
    Male = prev_epred[Sex == "Male"],
    Female = prev_epred[Sex == "Female"],
    mfRatio = Male / Female
  ) %>%
    mutate(id = 1:n())

#pivot and unnesay to dataframe for summary and plotting
post_model6_rvars_long_region <- post_model6_rvars_ratio_region %>%
    ungroup() %>%
    pivot_longer(cols = c(Male, Female, mfRatio)) %>%
    unnest_rvars() %>%
    ungroup()

#Country summaries
post_model6_rvars_long_region %>%
    group_by(WHO.region, name) %>%
    mean_qi(value, na.rm=TRUE)

#Now plot
post_model6_rvars_long_region %>%
    ggplot() +
    geom_vline(data = m6_empirical %>% filter(name=="mfRatio"),
               aes(xintercept = 1), colour="grey50", linetype=2) +
    stat_pointinterval(aes(x = value, y = WHO.region), .width = 0.95) +  
    geom_point(data = m6_empirical, aes(y = WHO.region, x = value),
             colour = "mediumseagreen", shape = 1) +
    facet_wrap(~name, scale="free_x") +
    scale_y_discrete(limits=rev)

```

Year effects by country and sex

```{r}

nd <- cleanSexDF %>%
    select(id, study.country, WHO.region, LogOddsStandardError) %>%
    crossing(year_z = seq(min(cleanSexDF$year_z, na.rm = TRUE), max(cleanSexDF$year_z, na.rm = TRUE), 1),
             Sex = c("Male", "Female"))

model6_year_draws <- add_epred_draws(object = model6,
                                     newdata=nd,
                                     re_formula = ~(1 + Sex*year_z |WHO.region/study.country))

model6_year_sum <- model6_year_draws %>%
    mutate(prev = expit_p100k(.epred)) %>%
    group_by(year_z, study.country, Sex) %>%
    mean_qi(prev)

model6_year_sum %>%
    mutate(year_z = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    ggplot() +
    geom_ribbon(aes(x=year_z, ymin=.lower, ymax=.upper, fill=Sex), alpha=0.5) +
    geom_line(aes(x=year_z, y=prev, colour=Sex), alpha=0.5) +
    geom_point(data = cleanSexDF, aes(
                       x=study.end.year,
                       y=`Adjusted Prevalence`*1e5,
               colour=Sex, shape=Sex)) +
    facet_wrap(~study.country, scales="free_y") +
    scale_colour_manual(values = c("darkorange", "mediumseagreen")) +
    scale_fill_manual(values = c("darkorange", "mediumseagreen")) +
    theme_ggdist()

model6_year_draws2 <- add_epred_draws(object = model6,
                                     newdata=nd,
                                     re_formula = ~(1 + Sex*year_z |WHO.region))

model6_year_sum2 <- model6_year_draws2 %>%
    mutate(prev = expit_p100k(.epred)) %>%
    group_by(year_z, WHO.region, Sex) %>%
    mean_qi(prev)

model6_year_sum2 %>%
    mutate(year_z = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    ggplot() +
    geom_ribbon(aes(x=year_z, ymin=.lower, ymax=.upper, fill=Sex), alpha=0.5) +
    geom_line(aes(x=year_z, y=prev, colour=Sex), alpha=0.5) +
    geom_point(data = cleanSexDF, aes(
                       x=study.end.year,
                       y=`Adjusted Prevalence`*1e5,
               colour=Sex, shape=Sex)) +
    facet_wrap(~WHO.region, scales="free_y") +
    scale_colour_manual(values = c("darkorange", "mediumseagreen")) +
    scale_fill_manual(values = c("darkorange", "mediumseagreen")) +
    theme_ggdist()


```



