---
title: "Sex Stratified Relationships of Tuberculosis Prevalence and Notification in Low- and Middle-income Countries: A Systematic Review and Meta-analysis" 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

ABOUT THIS SCRIPT

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multivariate multilevel model on the studies that report sex stratified prevalence results. This model aims to estimate sex stratified prevalence adjusted for geographic factors. 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
library(posterior)
library(tidybayes)

library(ggplot2)

expit <- function (x) {exp(x)/(1 + exp(x))}

myPal <- c("black", "violet")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))

##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################


cleanSexDF0 <- bactPostIndicator("sex") %>% filter(sex.analysis.indicator !="none")
```

# Descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(cleanSexDF0)[1], "unique surveys with sex-stratified bacteriological positive TB"))

# sort(unique(cleanSexDF0$title.extracted))
print(paste(length(unique(cleanSexDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanSexDF0$n.participants.female)), "studies report participant numbers totaling", 
            sum(cleanSexDF0$n.participants.female, cleanSexDF0$n.participants.male, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanSexDF0$n.participants.male, na.rm=TRUE)/sum(cleanSexDF0$n.participants.female, cleanSexDF0$n.participants.male, na.rm = TRUE)*100, 2), "% of participants were male" ))
```

# Risk of bias statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE}
table(cleanSexDF0$study.quality.summary)
```

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexDF <-  cleanSexDF0 %>% 
               filter(sex.analysis.indicator != "none") %>% 
               mutate("id" = row_number()) %>% 
    ### Scale all prevalence estimates back to decimals 
    mutate_at(vars(matches('prev100k.bacteriological.tb.|prev100k.ci.([a-z]{3})er.bacteriological.tb.|prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           WHO.region,
           screening.tests,
           study.start.year, 
           study.end.year, 
           publication.year,
           study.quality.summary,
           ends_with("male"), ends_with("sex"),
           sex.analysis.indicator) %>%
    
    mutate("tempPrevMale" =  case_when(sex.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.male/n.participants.male, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.male/n.participants.male,
                                        sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tempPrevFemale" = case_when(sex.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.female/n.participants.female, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female)) %>%
    pivot_longer(cols = c(tempPrevFemale,tempPrevMale), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.bacteriological.tb.male-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.bacteriological.tb.female-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.smear.positive.tb.male-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.smear.positive.tb.female-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.bacteriological.tb.male-
                                                                 prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.bacteriological.tb.female-
                                                                   prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.smear.positive.tb.male-
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.smear.positive.tb.female-
                                                                   prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.male),
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.female)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.male/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.male/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.female/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.male/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.male /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.female/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.female)))/3.92,
                                              sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Male" ~ (log(prev100k.ci.upper.bacteriological.tb.male/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.male)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.male/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Female" ~ (log(prev100k.ci.upper.bacteriological.tb.female/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.female)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.female /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Male" ~ (log(prev100k.ci.upper.smear.positive.tb.male/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.male)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.male /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Female" ~ (log(prev100k.ci.upper.smear.positive.tb.female/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.female)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.female /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.female)))/3.92, 
                                            sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            log(1/(n.participants.male*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            log(1/(n.participants.female*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       year_z = study.start.year - round(mean(study.start.year, na.rm=TRUE)),
                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       symptomScreen = ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No"),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))

##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("sex")

bactModelDF <- cleanSexDF %>%
    select(id, WHO.region, study.country, Sex, year_z, natRep, bias, symptomScreen, LogOdds, LogOddsStandardError, sex.analysis.indicator, n.participants.male, n.participants.female, publication.year) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(logOR = LogOdds_Male - LogOdds_Female,
           unadj_logOR_se = sqrt(LogOddsStandardError_Male^2 + LogOddsStandardError_Female^2), 
           logOR_se = ifelse(grepl("adj", sex.analysis.indicator), 
                                     unadj_logOR_se,
                                     ifelse(grepl("smear", sex.analysis.indicator),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Bact Variance Median"]))))

```
More descriptives 

```{r}
table(bactModelDF$WHO.region)
bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=WHO.region)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=WHO.region)
table(bactModelDF$natRep)
bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=natRep)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=natRep)
table(bactModelDF$bias)
bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=bias)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=bias)
table(bactModelDF$symptomScreen)
bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=symptomScreen)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=symptomScreen)
```

# Model 1: Main effect model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|study.country) + (1|id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectMod)
plot(mainEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- bactModelDF %>%
      select(study.country, LogOdds_Female) %>%
      mutate(logOR_se=1)

mainEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=nd,
                                 re_formula = ~ 1, 
                                 allow_new_levels = TRUE) %>% 
                    mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectModEst %>% ungroup %>% mean_qi(mfRatio)

### output for contribution table
exp(as.numeric(summary(mainEffectMod)[["fixed"]][1]))
exp(as.numeric(summary(mainEffectMod)[["fixed"]][3]))
exp(as.numeric(summary(mainEffectMod)[["fixed"]][4]))
```

Survey-level effects

```{r}
surveyEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=bactModelDF,
                                 re_formula = ~ 1  + (1|study.country) + (1|id), 
                                 allow_new_levels = TRUE) %>% 
                    mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

# Model 2: Effect by world region model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regionEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(regionEffectMod)
plot(regionEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndRegion <- bactModelDF %>%
      select(WHO.region, LogOdds_Female) %>%
      mutate(logOR_se=1)

regionEffectModEst <- add_epred_draws(object = regionEffectMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region),
                                 allow_new_levels = TRUE) %>% 
                       mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectModEst %>% group_by(WHO.region) %>% mean_qi(mfRatio)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
### Get the empirical values 
empirical_prev <- bactModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        bias = bias,
        symptomScreen = symptomScreen, 
        p_female_empirical = plogis(LogOdds_Female),
        p_male_empirical = plogis(LogOdds_Male),
        or_empirical = p_male_empirical / p_female_empirical,
        or_empirical = (p_male_empirical / (1 - p_male_empirical)) / 
            (p_female_empirical / (1 - p_female_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific TB prevalence")

    ggplot() +
    # geom_hline(aes(yintercept = 1), linetype=2) +
    geom_point(data = empirical_prev, aes(y=WHO.region, x=or_empirical, color = type2),alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectModEst, aes(x=mfRatio, y=WHO.region, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio",
         y = "World region") +
    theme_bw(base_size = 30) + theme(legend.title = element_blank(), legend.position = c(.83,.1), 
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
    
```

Bring estimates from regional model together to make a forest plot for the manuscript.

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.width = 20, fig.height = 28}

# Join survey labels
names <- cleanSexDF0 %>%
    mutate(id = row_number()) %>%
    select(id, figure.id.yr)

# Survey-level estimates
survey_sum <- surveyEffectModEst %>%
    group_by(id) %>%
    mean_qi(mfRatio) %>%
    mutate(grouping = "survey") %>%
    left_join(names, by = "id") %>%
    left_join(bactModelDF %>% select(id, WHO.region), by = "id")

# Regional summaries
region_sum <- regionEffectModEst %>%
    group_by(WHO.region) %>%
    mean_qi(mfRatio) %>%
    mutate(grouping = "regional") %>%
    rename(figure.id.yr = WHO.region)

# Global summary
global_sum <- mainEffectModEst %>%
    ungroup() %>%
    mean_qi(mfRatio) %>%
    mutate(grouping = "Overall",  
           figure.id.yr = "Overall")

# Combine all
forestDF <- bind_rows(survey_sum, region_sum, global_sum)

# Fill display label and facet group
forestDF <- forestDF %>%
    mutate(
        display_label_raw = ifelse(is.na(figure.id.yr), grouping, figure.id.yr),
        facet_group = case_when(
            grouping == "survey" ~ WHO.region,
            grouping == "regional" ~ figure.id.yr,
            grouping == "Overall" ~ "Overall"
        ), 
        type="Posterior mean and 95% credible interval"
    )

# Define facet order
group_order <- c(
    "Africa region", 
    "Americas region",
    "Eastern Mediterranean region", 
    "South-East Asia region", 
    "Western Pacific region", 
    "Overall"
)

# Separate surveys and summaries
surveys_only <- forestDF %>%
    filter(grouping == "survey") %>%
    group_by(facet_group) %>%
    arrange(desc(mfRatio), .by_group = TRUE) %>%
    ungroup()

summaries_only <- forestDF %>%
    filter(grouping != "survey")

# Recombine with surveys first, then summaries (preserves facet-wise order)
forestDF_plot <- bind_rows(surveys_only, summaries_only) %>%
    mutate(facet_group = factor(facet_group, levels = group_order)) %>%
    mutate(display_label = case_when(
        grouping == "Overall" ~ "<b>Overall</b>",
        grouping == "regional" ~ paste0("<b>", display_label_raw, "</b>"),
        TRUE ~ display_label_raw
    )) %>%
    mutate(display_label = factor(display_label, levels = rev(unique(display_label))))

# Prepare empirical data
empirical_prev_labels <- left_join(empirical_prev, survey_sum) %>%
    rename(display_label_raw = figure.id.yr) %>%
    select(display_label_raw, WHO.region, or_empirical) %>%
    mutate(
        grouping = "survey",
        facet_group = WHO.region,
        display_label = display_label_raw
    ) %>%
    mutate(facet_group = factor(facet_group, levels = group_order),
           type = "Ratios calculated from reported prevalence")

# Plot
forestRatio <- ggplot(forestDF_plot) +
    geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
    geom_point(data = empirical_prev_labels,
               aes(x = display_label, y = or_empirical, color=type), alpha=0.5) +
    geom_pointinterval(aes(x = display_label, y = mfRatio, ymin = .lower, ymax = .upper, color=type),
                       size = 0.25) +
    coord_flip() +
    facet_grid(facet_group ~ ., scales = "free", space = "free") +
    scale_color_manual(values = myPal) +
    scale_y_log10(
        breaks = c(0.25, 0.5, 1, 2, 4),
        labels = c("0.25", "0.5", "1", "2", "4"),
    ) +
    # scale_x_discrete(labels = label_wrap(30)) + 
    labs(
        x = NULL,
        y = "Male-to-female ratio of bacteriologically-confirmed TB prevalence"
    ) +
    theme_minimal(base_size = 10) +
    theme(
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.spacing.y = unit(1, "lines"),
        axis.text.y = ggtext::element_markdown(hjust = 1),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.background = element_rect(fill = "white"),
        legend.position = "bottom", legend.title=element_blank(),
        legend.spacing.y = unit(-.25, 'cm'),
        legend.box.margin=margin(-10,-10,-6,-10),
        panel.spacing = unit(0.18, "cm"),
        plot.margin = unit(c(.1, .7, 0, .1), "cm")
    )

# Save
ggsave(plot = forestRatio, "output/mfFigure2.png", height = 12.5, width=8.5, units = "in")
```

# Model 3: Effect over time 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
timeEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z + (1 + year_z | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

### Also run this model with only those studies in the horton review
timeEffectHortMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z + (1 + year_z | study.country) + (1 | id),
                       data = bactModelDF %>% filter(year_z <= 5),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(timeEffectMod)
plot(timeEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndTime <- bactModelDF %>%
    select(logOR_se) %>%
    crossing(year_z = seq(min(cleanSexDF$year_z, na.rm = TRUE), max(cleanSexDF$year_z, na.rm = TRUE), 1))

timeEffectModEst <- add_epred_draws(object = timeEffectMod,
                                 newdata=ndTime,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

timeEffectHortModEst <- add_epred_draws(object = timeEffectHortMod,
                                 newdata=ndTime,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

```

Main estimate
```{r}
timeEffectModEst %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper)

### output for contribution table
exp(as.numeric(summary(timeEffectMod)[["fixed"]][2,1]))
exp(as.numeric(summary(timeEffectMod)[["fixed"]][2,3]))
exp(as.numeric(summary(timeEffectMod)[["fixed"]][2,4]))

timeEffectHortModEst %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper)

bayes_R2(timeEffectMod)

timeEffectModEst 

```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    y <- timeEffectModEst  %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper) 

    ggplot(data = y) +
    geom_lineribbon(.width = 0.95, alpha = 0.25, fill="violet", aes(x = Year,
               y = mfRatio, ymin = .lower, ymax = .upper)) +
    geom_point(data = empirical_prev %>% rename("Study standard error" = logOR_se),
               aes(x=year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = `Study standard error`),
               colour = "black") +
    expand_limits(y=0) +
    labs(y = "Male-to-female odds ratio",
         x = "Year") +
    theme_minimal(base_size = 30)  + theme(legend.position = "bottom")
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    yHort <- timeEffectHortModEst  %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper)

yFull <- rbind(
    mutate(y, review = "Current review"), 
    mutate(yHort, review = "Horton et al. review")
)

    ggplot(data = yFull) +
    geom_ribbon(.width = 0.95, alpha = 0.4, aes(x = Year,
               y = mfRatio, ymin = .lower, ymax = .upper, fill=review)) +
    geom_line(aes(x = Year, y = mfRatio, color=review), size=1.5) +
    geom_point(data = empirical_prev %>% rename("Study standard error" = logOR_se),
               aes(x=year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = `Study standard error`),
               colour = "black") +
    expand_limits(y=0) +
    scale_fill_manual("", values = c("violet", "seagreen2")) + 
    scale_color_manual("", values = c("darkviolet", "seagreen")) + 
    labs(y = "Male-to-female odds ratio",
         x = "Year") +
    theme_minimal(base_size = 30)  + theme(legend.position = "bottom")
```


# Model 4: Region effect over time 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
timeEffectRegionMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                              1 + year_z * WHO.region + (1 + year_z | study.country) + (1 | id), 
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(timeEffectRegionMod)
plot(timeEffectRegionMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndTimeReg <- bactModelDF %>%
             select(logOR_se, WHO.region) %>%
             distinct() %>%
             crossing(year_z = seq(min(cleanSexDF$year_z, na.rm = TRUE),
                                   max(cleanSexDF$year_z, na.rm = TRUE), 1))

timeEffectRegionModEst <- add_epred_draws(object = timeEffectRegionMod,
                                 newdata=ndTimeReg,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")
```

Main estimate
```{r}
timeEffectRegionModEst %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper)

```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    timeEffectRegionModEst %>%
    filter(WHO.region !="Americas region",
           WHO.region !="Eastern Mediterranean region") %>%
    ggplot(aes(x = year_z + round(mean(cleanSexDF$study.end.year, na.rm = TRUE)),
               y = mfRatio,
               fill = WHO.region,
               color = WHO.region)) +
    stat_lineribbon(aes(color = WHO.region, fill = WHO.region), .width = 0.95, alpha = 0.4, size=1.5) +
    geom_point(data = empirical_prev %>% rename("Study standard error" = logOR_se) %>% 
                   filter(WHO.region !="Americas region",
                          WHO.region !="Eastern Mediterranean region"),
               aes(x=year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = `Study standard error`),
               colour = "black") +
    geom_hline(yintercept = 1, linetype = 2) +
    facet_wrap(~ WHO.region) +
    labs(y = "Male-to-female odds ratio",
         x = "Year",
         fill = "WHO region",
         color = "WHO region") +
    scale_color_manual("", values = wes_palette("Cavalcanti1")) +
    scale_fill_manual("", values = wes_palette("Cavalcanti1")) +
    theme_minimal(base_size = 30)  + theme(legend.position = "bottom")
```

# Sub group analysis

## World regions 
Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regSubGrpDF <- data.frame(region = NULL, 
                          mfRatio = NULL,
                          .lower = NULL, .upper = NULL)

for (i in 1:length(unique(bactModelDF$WHO.region))){
    region <- unique(bactModelDF$WHO.region)[i]
    print(region)
    
    tmpRegionEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                               1 + (1 | study.country) + (1 | id),
                           data = bactModelDF %>% filter(WHO.region == region),
                           prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                                   prior(exponential(2), class = sd),
                           family = gaussian(),
                           control = list(adapt_delta = 0.99),
                           chains = 4,
                           cores = 4,
                           iter = 4000)
    
    ndTmpRegion <- bactModelDF %>%
         filter(WHO.region == region) %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

    tmpRegionEffectModEst <- add_epred_draws(object = tmpRegionEffectMod,
                                 newdata=ndTmpRegion,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")   
    

    regSubGrpDF[i,1] <- region
    regSubGrpDF[i,2:4] <- round(tmpRegionEffectModEst %>% ungroup() %>%
                          mean_qi(mfRatio) %>% 
                          select(mfRatio, .lower, .upper),2)
}

regSubGrpDF
```


## National surveys

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
natEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(natRep=="Yes"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

subNatEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(natRep=="No"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(natEffectMod)
plot(natEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndNat <- bactModelDF %>%
         filter(natRep == "Yes") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

natEffectModEst <- add_epred_draws(object = natEffectMod,
                                 newdata=ndNat,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

ndSubnat <- bactModelDF %>%
         filter(natRep == "No") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

subnatEffectModEst <- add_epred_draws(object = subNatEffectMod,
                                 newdata=ndSubnat,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
natEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper)

subnatEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper)
```

## Study quality (Low vs. not low)

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}

lowBiasEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(bias == "Low"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

notLowBiasEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(bias == "Not low"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(biasEffectMod)
plot(biasEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndLowBias <- bactModelDF %>%
         filter(bias == "Low") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

ndNotLowBias <- bactModelDF %>%
         filter(bias == "Not low") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

lowBiasEffectModEst <- add_epred_draws(object = lowBiasEffectMod,
                                 newdata=ndLowBias,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

notLowBiasEffectModEst <- add_epred_draws(object = notLowBiasEffectMod,
                                 newdata=ndNotLowBias,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
lowBiasEffectModEst %>% 
    ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper)

notLowBiasEffectModEst %>% 
    ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper)
```


## Symptom screen vs. no symptom screen

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}

sympYesEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(symptomScreen == "Yes"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

sympNoEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(symptomScreen == "No"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(sympEffectMod)
plot(sympEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndSympYes <- bactModelDF %>%
         filter(symptomScreen == "Yes") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

ndSympNo <- bactModelDF %>%
         filter(symptomScreen == "No") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

sympYesEffectModEst <- add_epred_draws(object = sympYesEffectMod,
                                 newdata=ndSympYes,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

sympNoEffectModEst <- add_epred_draws(object = sympNoEffectMod,
                                 newdata=ndSympNo,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
sympYesEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper)

sympNoEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    ggplot() +
    geom_point(data = empirical_prev, aes(y=symptomScreen, x=or_empirical, color = type2, size=logOR_se), alpha = 0.5) +
    stat_pointinterval(data=sympEffectModEst, aes(x=mfRatio, y=symptomScreen, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio", color = "",
         y = "Symptom screening required") +
    scale_size_continuous(range = c(8,16)) +
    theme_bw(base_size = 30) + theme(
                                     legend.position = c(.83,.2), legend.box.just = "right",
                                     # legend.direction = "horizontal",
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
```

## Multivariate analysis of male-to-female prevalence ratio

Setup and read in external data sources 
```{r, echo=FALSE, warning = FALSE, message=FALSE}
tbBurdenEst <- read_csv("data/WHOEstimates.csv") %>% 
             dplyr::select(country, e_inc_100k, e_inc_num, 
                           e_tbhiv_prct) %>% 
             mutate(e_inc_100k = e_inc_100k/1e5, 
                    e_tbhiv_prct = e_tbhiv_prct/100) %>%
             rename("study.country" = country,
                    "tb.inc.100k" = e_inc_100k, 
                    "tb.cases" = e_inc_num, 
                    "tb.hiv.perc" = e_tbhiv_prct) %>% 
            mutate("tb.inc.100k_z" = tb.inc.100k - mean(tb.inc.100k, na.rm=TRUE),
                   "tb.cases_z" = tb.cases - mean(tb.cases, na.rm=TRUE), 
                   "tb.hiv.perc_z" = tb.hiv.perc - mean(tb.hiv.perc, na.rm=TRUE))

### Gender development index 
GDI <- read_csv("data/GenderDevelopmentIndex.csv")[,1:2] %>% 
       mutate(GDI = as.numeric(Value), 
              GDI_z = GDI - mean(GDI, na.rm=TRUE)) %>% select(!Value) %>%
       rename(study.country = Country)

### Join 
covariates <- full_join(tbBurdenEst, GDI,  by=c("study.country"))

### Join these to the modeling dataset 
bactModelDF <- full_join(bactModelDF, covariates, by=c("study.country"))
```

Model specification 

```{r}
tbIncMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + tb.inc.100k_z + 
                               (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

summary(tbIncMod)

bayes_R2(tbIncMod)

ndTbInc <- bactModelDF %>%
             select(logOR_se,
                    tb.inc.100k_z) %>%
             distinct()

tbIncModEst <- add_epred_draws(object = tbIncMod,
                                 newdata=ndTbInc %>% na.omit(),
                                 re_formula = NA) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

tbIncModEst %>% ungroup() %>%
    mean_qi(mfRatio) 

### output for contribution table
round(exp(as.numeric(summary(tbIncMod)[["fixed"]][2,1])),2)
round(exp(as.numeric(summary(tbIncMod)[["fixed"]][2,3])),2)
round(exp(as.numeric(summary(tbIncMod)[["fixed"]][2,4])),2)
```


Model specification 

```{r}
tbHivMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + tb.hiv.perc_z + 
                               (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

bayes_R2(tbHivMod)

ndtbHiv <- bactModelDF %>%
             select(logOR_se,
                    tb.hiv.perc_z) %>%
             distinct()

tbHivModEst <- add_epred_draws(object = tbHivMod,
                                 newdata=ndtbHiv %>% na.omit(),
                                 re_formula = NA) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

tbHivModEst %>% ungroup() %>%
    mean_qi(mfRatio)

### output for contribution table
exp(as.numeric(summary(tbHivMod)[["fixed"]][2,1]))
exp(as.numeric(summary(tbHivMod)[["fixed"]][2,3]))
exp(as.numeric(summary(tbHivMod)[["fixed"]][2,4]))

```

Model specification 

```{r}
gdiMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + GDI_z + 
                               (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

summary(gdiMod)

bayes_R2(gdiMod)

ndGDI <- bactModelDF %>%
             select(logOR_se,
                    GDI_z) %>%
             distinct()

gdiModEst <- add_epred_draws(object = gdiMod,
                                 newdata=ndGDI %>% na.omit(),
                                 re_formula = NA) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

gdiModEst %>% ungroup() %>%
    mean_qi(mfRatio) 

### output for contribution table
round(exp(as.numeric(summary(gdiMod)[["fixed"]][2,1])),2)
round(exp(as.numeric(summary(gdiMod)[["fixed"]][2,3])),2)
round(exp(as.numeric(summary(gdiMod)[["fixed"]][2,4])),2)
```

Model specification 

```{r}
multiMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                                1 + year_z + 
                                tb.inc.100k_z + tb.hiv.perc_z + 
                                GDI_z + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(multiMod)

### output for contribution table
x <- data.frame(parameter = rownames(summary(multiMod)[["fixed"]]),
                PE = round(exp(as.numeric(summary(multiMod)[["fixed"]][,1])),2), 
                CrI = paste(round(exp(as.numeric(summary(multiMod)[["fixed"]][,3])),2), 
                          ",",
                    round(exp(as.numeric(summary(multiMod)[["fixed"]][,4])),2)))

print(x)

bayes_R2(multiMod)
plot(multiMod)

stanplot(multiMod)
```

# Replicate the covariate analysis by world region

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
regCoVList <- list()

for (i in 2:length(unique(bactModelDF$WHO.region))){
    region <- unique(bactModelDF$WHO.region)[i]
    print(region)
    
    regionMainMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                                1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(WHO.region == region),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
    
    regionTBIncMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                                1 + tb.inc.100k_z + 
                                (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(WHO.region == region),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
    
    regionTBHivIncMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                            1 + tb.hiv.perc_z + 
                            (1 | study.country) + (1 | id),
                   data = bactModelDF %>% filter(WHO.region == region),
                   prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                       prior(normal(0, 10), class = b) +
                       prior(exponential(2), class = sd),
                   family = gaussian(),
                   control = list(adapt_delta = 0.99),
                   chains = 4,
                   cores = 4,
                   iter = 4000)
    
    regionTBYearMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                            1 + year_z + 
                            (1 | study.country) + (1 | id),
                   data = bactModelDF %>% filter(WHO.region == region),
                   prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                       prior(normal(0, 10), class = b) +
                       prior(exponential(2), class = sd),
                   family = gaussian(),
                   control = list(adapt_delta = 0.99),
                   chains = 4,
                   cores = 4,
                   iter = 4000)
    
    regionTBGdiMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                            1 + GDI_z + 
                            (1 | study.country) + (1 | id),
                   data = bactModelDF %>% filter(WHO.region == region),
                   prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                       prior(normal(0, 10), class = b) +
                       prior(exponential(2), class = sd),
                   family = gaussian(),
                   control = list(adapt_delta = 0.99),
                   chains = 4,
                   cores = 4,
                   iter = 4000)
    
    tmp <- data.frame(parameter = c(rownames(summary(regionMainMod)[["fixed"]]), 
                                  rownames(summary(regionTBGdiMod)[["fixed"]])[2], 
                                  rownames(summary(regionTBYearMod)[["fixed"]])[2],
                                  rownames(summary(regionTBIncMod)[["fixed"]])[2],
                                  rownames(summary(regionTBHivIncMod)[["fixed"]])[2]), 
                    PE = c(round(exp(as.numeric(summary(regionMainMod)[["fixed"]][1,1])),2), 
                           round(exp(as.numeric(summary(regionTBGdiMod)[["fixed"]][2,1])),2), 
                           round(exp(as.numeric(summary(regionTBYearMod)[["fixed"]][2,1])),2), 
                           round(exp(as.numeric(summary(regionTBIncMod)[["fixed"]][2,1])),2), 
                           round(exp(as.numeric(summary(regionTBHivIncMod)[["fixed"]][2,1])),2)), 
                    CrI = c(paste(round(exp(as.numeric(summary(regionMainMod)[["fixed"]][1,3])),2), ",",
                                round(exp(as.numeric(summary(regionMainMod)[["fixed"]][1,4])),2)),
                        paste(round(exp(as.numeric(summary(regionTBGdiMod)[["fixed"]][2,3])),2), ",",
                                    round(exp(as.numeric(summary(regionTBGdiMod)[["fixed"]][2,4])),2)),
                        paste(round(exp(as.numeric(summary(regionTBYearMod)[["fixed"]][2,3])),2), ",",
                                    round(exp(as.numeric(summary(regionTBYearMod)[["fixed"]][2,4])),2)), 
                        paste(round(exp(as.numeric(summary(regionTBIncMod)[["fixed"]][2,3])),2), ",",
                                    round(exp(as.numeric(summary(regionTBIncMod)[["fixed"]][2,4])),2)), 
                        paste(round(exp(as.numeric(summary(regionTBHivIncMod)[["fixed"]][2,3])),2), ",",
                                    round(exp(as.numeric(summary(regionTBHivIncMod)[["fixed"]][2,4])),2))))
    
    regCoVList[[region]] <- tmp
    rm(tmp)
}
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
regMultiList <- list()

for (i in 1:length(unique(bactModelDF$WHO.region))){
    region <- unique(bactModelDF$WHO.region)[i]
    print(region)
    
    regionMultiMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                                1 + year_z + 
                                tb.inc.100k_z + tb.hiv.perc_z + 
                                GDI_z + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(WHO.region == region),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
    
    x <- data.frame(parameter = rownames(summary(regionMultiMod)[["fixed"]]),
                PE = round(exp(as.numeric(summary(regionMultiMod)[["fixed"]][,1])),2), 
                CrI = paste(round(exp(as.numeric(summary(regionMultiMod)[["fixed"]][,3])),2), 
                          ",",
                    round(exp(as.numeric(summary(regionMultiMod)[["fixed"]][,4])),2)))
    
    regMultiList[[region]] <- x 
    rm(x)
}
```



# Smear positive analysis 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
cleanSexSmrDF0 <- smrPosIndicator("sex") %>% filter(sex.smr.analysis.indicator !="none")

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexSmrDF <- cleanSexSmrDF0 %>% 
                 filter(sex.smr.analysis.indicator != "none") %>% 
                 mutate("id" = row_number()) %>%
                 ### Scale all prevalence estimates back to decimals 
                 mutate_at(vars(matches('prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>% 
                 select(id,
                       covidence.id, 
                       figure.id.yr, 
                       study.geography, 
                       title.extracted,
                       study.country,
                       WHO.region,
                       study.start.year, 
                       study.end.year, 
                       ends_with("male"), ends_with("sex"),
                       sex.smr.analysis.indicator) %>%
    
    mutate("tempPrevMale" =  case_when(sex.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.smr.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                        sex.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tempPrevFemale" = case_when(sex.smr.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.smr.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female)) %>%
    pivot_longer(cols = c(tempPrevFemale,tempPrevMale), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.smear.positive.tb.male-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.smear.positive.tb.female-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                       
                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.smear.positive.tb.male-
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.smear.positive.tb.female-
                                                                   prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.smr.analysis.indicator %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.male),
                                        sex.smr.analysis.indicator  %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.female)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.male/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.male /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                              sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.female/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.female)))/3.92,
                                                 sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Male" ~ (log(prev100k.ci.upper.smear.positive.tb.male/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.male)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.male /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                                 sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Female" ~ (log(prev100k.ci.upper.smear.positive.tb.female/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.female)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.female /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.female)))/3.92, 
                                            sex.smr.analysis.indicator %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            log(1/(n.participants.male*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        sex.smr.analysis.indicator  %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            log(1/(n.participants.female*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       year_z = study.start.year - round(mean(study.start.year, na.rm=TRUE)),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))  %>% filter(LogOddsStandardError < Inf)

smrModelDF <- cleanSexSmrDF %>%
    select(id, WHO.region, study.country, Sex, year_z, natRep, LogOdds, LogOddsStandardError, 
           n.participants.male, n.participants.female) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_Male - LogOdds_Female,
        logOR_se = sqrt(LogOddsStandardError_Male^2 + LogOddsStandardError_Female^2) 
    ) 
```

# Smear positive descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(cleanSexSmrDF0)[1], "unique surveys with sex-stratified smear positive TB"))

# sort(unique(cleanSexDF0$title.extracted))
print(paste(length(unique(cleanSexSmrDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanSexSmrDF0$n.participants.female)), "studies report participant numbers totaling", 
            sum(cleanSexSmrDF0$n.participants.female, cleanSexSmrDF0$n.participants.male, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanSexSmrDF0$n.participants.male, na.rm=TRUE)/sum(cleanSexSmrDF0$n.participants.female, cleanSexSmrDF0$n.participants.male, na.rm = TRUE)*100, 2), "% of participants were male" ))
```
More descriptives 

```{r}
table(smrModelDF$WHO.region)
smrModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=WHO.region)
smrModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=WHO.region)
```

# Model X: Smear positive main effect 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | study.country) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectSmrMod)
plot(mainEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- smrModelDF %>%
      select(study.country, LogOdds_Female) %>%
      mutate(logOR_se=1)

mainEffectSmrModEst <- add_epred_draws(object = mainEffectSmrMod,
                                 newdata=nd,
                                 re_formula = ~ 1) %>% 
                    mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectSmrModEst %>% ungroup %>% mean_qi(mfRatio)
```

# Model 2: Smear positive effect by world region model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regionEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(regionEffectSmrMod)
plot(regionEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndRegion <- smrModelDF %>%
      select(WHO.region, LogOdds_Female) %>%
      mutate(logOR_se=1)

regionEffectSmrModEst <- add_epred_draws(object = regionEffectSmrMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region)) %>% 
                       mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectSmrModEst %>% group_by(WHO.region) %>% mean_qi(mfRatio)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
### Get the empirical values 
empirical_smr_prev <- smrModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        p_female_empirical = plogis(LogOdds_Female),
        p_male_empirical = plogis(LogOdds_Male),
        or_empirical = p_male_empirical / p_female_empirical,
        or_empirical = (p_male_empirical / (1 - p_male_empirical)) / 
            (p_female_empirical / (1 - p_female_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific smear positive TB prevalence")

    ggplot() +
    # geom_hline(aes(yintercept = 1), linetype=2) +
    geom_point(data = empirical_smr_prev, aes(y=WHO.region, x=or_empirical, color = type2),alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectSmrModEst, aes(x=mfRatio, y=WHO.region, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio",
         y = "World region") +
    theme_bw(base_size = 30) + theme(legend.title = element_blank(), legend.position = c(.83,.07), 
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
    
```

# Subgroup analysis

## World regions 
Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regSubGrpSmrDF <- data.frame(region = NULL, 
                          mfRatio = NULL,
                          .lower = NULL, .upper = NULL)

for (i in 1:length(unique(smrModelDF$WHO.region))){
    region <- unique(smrModelDF$WHO.region)[i]
    print(region)
    
    tmpRegionEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                               1 + (1 | study.country) + (1 | id),
                           data = smrModelDF %>% filter(WHO.region == region),
                           prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                                   prior(exponential(2), class = sd),
                           family = gaussian(),
                           control = list(adapt_delta = 0.99),
                           chains = 4,
                           cores = 4,
                           iter = 4000)
    
    ndTmpRegion <- smrModelDF %>%
         filter(WHO.region == region) %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

    tmpRegionEffectModEst <- add_epred_draws(object = tmpRegionEffectMod,
                                 newdata=ndTmpRegion,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")   
    

    regSubGrpSmrDF[i,1] <- region
    regSubGrpSmrDF[i,2:4] <- round(tmpRegionEffectModEst %>% ungroup() %>%
                             mean_qi(mfRatio) %>% 
                             select(mfRatio, .lower, .upper),2)
}

regSubGrpSmrDF
```