---
title: "Sex Stratified Relationships of Tuberculosis Prevalence and Notification in Low- and Middle-income Countries: A Systematic Review and Meta-analysis" 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

# About this script 

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multivariate multilevel model on the studies that report sex stratified prevalence results. This model aims to estimate sex stratified prevalence adjusted for geographic factors. 

# Setup

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
# library(posterior)
library(tidybayes)
library(wesanderson)
library(ggplot2)
library(patchwork)

expit <- function (x) {exp(x)/(1 + exp(x))}

myPal <- c("black", "violet")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))

##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################
cleanSexDF0 <- readRDS("data/sex-stratified-clean-data.rds")
### Calculate mean publication lag 
meanLag <- round(mean(cleanSexDF0$publication.year - cleanSexDF0$study.end.year, na.rm=TRUE))
```

# Descriptive statistics

```{r, warning = FALSE, message=FALSE, echo=FALSE}

print(paste(dim(cleanSexDF0)[1], "unique surveys with sex-stratified bacteriological positive TB"))

# sort(unique(cleanSexDF0$title.extracted))
print(paste(length(unique(cleanSexDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanSexDF0$n.participants.sex.total)), "studies report participant numbers totaling", 
            sum(cleanSexDF0$n.participants.sex.total, na.rm = TRUE), "participants"))

print(paste(sum(!is.na(cleanSexDF0$n.participants.female)), "studies report sex-stratified participant numbers totaling", sum(cleanSexDF0$n.participants.female, cleanSexDF0$n.participants.male, na.rm=TRUE), " and ", round(sum(cleanSexDF0$n.participants.male, na.rm=TRUE)/sum(cleanSexDF0$n.participants.female, cleanSexDF0$n.participants.male, na.rm = TRUE)*100, 2), "% of participants were male" ))
```

# Risk of bias statistics

```{r, warning = FALSE, message=FALSE}
table(cleanSexDF0$study.quality.summary)
```
# Bacteriologically-confirmed TB analysis 
## Format data for modeling 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexDF <-  cleanSexDF0 %>% 
               filter(sex.analysis.indicator != "none") %>% 
               mutate("id" = row_number()) %>% 
    ### Scale all prevalence estimates back to decimals 
    mutate_at(vars(matches('prev100k.bacteriological.tb.|prev100k.ci.([a-z]{3})er.bacteriological.tb.|prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           study.rurality, 
           WHO.region,
           screening.tests,
           participants.sputum.collected,
           bacteriological.tb.definition,
           xpert.used,
           xpert.samples,
           other.xpert.samples,
           culture.used,
           culture.samples,
           other.culture.samples,
           study.start.year, 
           study.end.year, 
           publication.year,
           study.quality.summary,
           ends_with("male"), ends_with("sex"),
           sex.analysis.indicator) %>%
    
    mutate("tempPrevMale" =  case_when(sex.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.male/n.participants.male, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.male/n.participants.male,
                                        sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tempPrevFemale" = case_when(sex.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.female/n.participants.female, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female)) %>%
    pivot_longer(cols = c(tempPrevFemale,tempPrevMale), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.bacteriological.tb.male-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.bacteriological.tb.female-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.smear.positive.tb.male-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.smear.positive.tb.female-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.bacteriological.tb.male-
                                                                 prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.bacteriological.tb.female-
                                                                   prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.smear.positive.tb.male-
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.smear.positive.tb.female-
                                                                   prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.male),
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.female)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.male/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.male/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.female/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.male/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.male /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.female/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.female)))/3.92,
                                              sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Male" ~ (log(prev100k.ci.upper.bacteriological.tb.male/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.male)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.male/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Female" ~ (log(prev100k.ci.upper.bacteriological.tb.female/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.female)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.female /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Male" ~ (log(prev100k.ci.upper.smear.positive.tb.male/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.male)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.male /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Female" ~ (log(prev100k.ci.upper.smear.positive.tb.female/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.female)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.female /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.female)))/3.92, 
                                            sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            log(1/(n.participants.male*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            log(1/(n.participants.female*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(is.na(study.geography), 
                                                       NA, 
                                                       ifelse(study.geography == "Nationally representative", 
                                                       "Yes", "No")),

                                       year_z = ifelse(!is.na(study.end.year), 
                                                       study.end.year - round(mean(study.end.year, na.rm=TRUE)),
                                                       (publication.year - meanLag) - round(mean((publication.year - meanLag), na.rm=TRUE))),                                       
                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       
                                       symptomScreen = ifelse(is.na(screening.tests), 
                                                                    NA, 
                                                                    ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No")),
                                       symptomsRequired = ifelse(is.na(participants.sputum.collected) | grepl("unknown", 
                                                                                                              participants.sputum.collected,ignore.case=TRUE), 
                                                                       NA, 
                                                                       ifelse(!grepl("x-ray|all participants|other",
                                                                        participants.sputum.collected,ignore.case=TRUE),
                                                                 "Yes", "No")),

                                       XpertCultRequired = ifelse(is.na(bacteriological.tb.definition),
                                                                  NA, ifelse(!grepl("smear|micros|field|NTP", 
                                                                              bacteriological.tb.definition,
                                                                              ignore.case=TRUE), 
                                                                  "Yes", 
                                                                   "No")),
                                       ### Fix those that were missed with the above attempt 
                                       XpertCultRequired = ifelse(covidence.id %in% c(26989,26928,26920,26918,26892,26861,26857), 
                                                                  "Yes", XpertCultRequired), 
                                       XpertCultRequired = ifelse(covidence.id %in% c(28002), 
                                                                  "No", XpertCultRequired),
                                       #### if we used the culture positive count, then xpert culture was requried
                                       XpertCultRequired = ifelse(grepl("culture", sex.analysis.indicator), 
                                                                  "Yes", XpertCultRequired),
                                       #### if we used the smear positive prevs, then xpert culture was not requried
                                       XpertCultRequired = ifelse(grepl("smear", sex.analysis.indicator), 
                                                                  "No", XpertCultRequired),
                                       XpertCultUsed = ifelse((grepl("Yes", xpert.used) & 
                                                                 (grepl("All samples", xpert.samples) |
                                                                  !grepl("mear|NAAT", other.xpert.samples))) |
                                                                  (grepl("Yes", culture.used) & 
                                                                  grepl("All samples|Subset", culture.samples) & 
                                                                  !grepl("HIV", other.culture.samples)), 
                                                                 "Yes", "No"),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))

##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("sex")

bactModelDF <- cleanSexDF %>%
    select(id, covidence.id, title.extracted, WHO.region, study.country, study.rurality,
           Sex, year_z, 
           natRep, bias, symptomScreen, symptomsRequired, 
           XpertCultRequired,  XpertCultUsed,
           LogOdds, LogOddsStandardError, sex.analysis.indicator, 
           n.participants.male, n.participants.female, 
           publication.year, study.end.year) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(logOR = LogOdds_Male - LogOdds_Female,
           unadj_logOR_se = sqrt(LogOddsStandardError_Male^2 + LogOddsStandardError_Female^2), 
           logOR_se = ifelse(grepl("adj", sex.analysis.indicator), 
                                     unadj_logOR_se,
                                     ifelse(grepl("smear", sex.analysis.indicator),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Bact Variance Median"]))))

```

## Additional descriptive output

```{r}
table(bactModelDF$WHO.region)


### How many persons diagnosed with bacteriologically-confirmed TB 
### Case 1: we used the reported count so simply sum 
### 38 total surveys; no missing 
cleanSexDF %>% filter(grepl("n.", sex.analysis.indicator, fixed = TRUE)) %>% 
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = case_when(sex.analysis.indicator == 
                                          "n.bacteriological.tb" ~ n.bacteriological.tb.male + n.bacteriological.tb.female, 
                                      sex.analysis.indicator == 
                                          "n.culture.positive.tb" ~ n.culture.positive.tb.male + n.culture.positive.tb.female, 
                                      sex.analysis.indicator == 
                                          "n.smear.positive.tb" ~ n.smear.positive.tb.male + n.smear.positive.tb.female)) %>% 
    select(diagnosedCases) %>% sum() -> cases1

cleanSexDF %>% filter(grepl("n.", sex.analysis.indicator, fixed = TRUE)) %>% 
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = case_when(sex.analysis.indicator == 
                                          "n.bacteriological.tb" ~ n.bacteriological.tb.male , 
                                      sex.analysis.indicator == 
                                          "n.culture.positive.tb" ~ n.culture.positive.tb.male , 
                                      sex.analysis.indicator == 
                                          "n.smear.positive.tb" ~ n.smear.positive.tb.male )) %>% 
    select(diagnosedCases) %>% sum() -> maleCases1

### Case 2: we used bacteriological confirmed prevalence risk so pull their reported 
### counts (if available) or we need participant counts. 
### 55 total surveys; 36 reported case counts; 15 estimated; 4 missing 
cleanSexDF %>% filter(grepl("prev100k.ci.bacteriological", sex.analysis.indicator, fixed = TRUE)) %>%
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = case_when(!is.na(n.bacteriological.tb.male) ~ 
                                           n.bacteriological.tb.female + n.bacteriological.tb.male, 
                                       !is.na(n.participants.male)  & !is.na(adj.prev100k.bacteriological.tb.male) ~
                                           adj.prev100k.bacteriological.tb.male * n.participants.male +
                                           adj.prev100k.bacteriological.tb.female * n.participants.female, 
                                       !is.na(n.participants.male)  & !is.na(prev100k.bacteriological.tb.male) ~
                                           prev100k.bacteriological.tb.male * n.participants.male +
                                           prev100k.bacteriological.tb.female * n.participants.female
    )) %>% 
    # select(sex.analysis.indicator, 
    #        n.bacteriological.tb.female, 
    #        n.bacteriological.tb.male, 
    #        adj.prev100k.bacteriological.tb.female,
    #        adj.prev100k.bacteriological.tb.male,
    #        prev100k.bacteriological.tb.female, 
    #        prev100k.bacteriological.tb.male,
    #        n.participants.female, 
    #        n.participants.male, 
    #        diagnosedCases) %>% View() 
    select(diagnosedCases) %>% round() %>% sum(na.rm = TRUE) -> cases2


cleanSexDF %>% filter(grepl("prev100k.ci.bacteriological", sex.analysis.indicator, fixed = TRUE)) %>%
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = case_when(!is.na(n.bacteriological.tb.male) ~ 
                                           n.bacteriological.tb.male, 
                                       !is.na(n.participants.male)  & !is.na(adj.prev100k.bacteriological.tb.male) ~
                                           adj.prev100k.bacteriological.tb.male * n.participants.male, 
                                       !is.na(n.participants.male)  & !is.na(prev100k.bacteriological.tb.male) ~
                                           prev100k.bacteriological.tb.male * n.participants.male 
                                           
    )) %>% 
    select(diagnosedCases) %>% round() %>% sum(na.rm = TRUE) -> maleCases2

### Case 2: we used smear positive  prevalence risk so pull their reported counts  
### (if available) or we need participant counts. 
### 9 total surveys; 8 reported case counts; 1 estimated; no missings
cleanSexDF %>% filter(sex.analysis.indicator %in% c("adj.prev100k.ci.smear.positive.tb", 
                                                    "prev100k.ci.smear.positive.tb",
                                                    "prev100k.smear.positive.tb")) %>%
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = case_when(!is.na(n.smear.positive.tb.male) ~ 
                                           n.smear.positive.tb.female + n.smear.positive.tb.male, 
                                       !is.na(n.participants.male)  & !is.na(adj.prev100k.smear.positive.tb.male) ~
                                           adj.prev100k.smear.positive.tb.male * n.participants.male +
                                           adj.prev100k.smear.positive.tb.female * n.participants.female, 
                                       !is.na(n.participants.male)  & !is.na(prev100k.smear.positive.tb.male) ~
                                           prev100k.smear.positive.tb.male * n.participants.male +
                                           prev100k.smear.positive.tb.female * n.participants.female
    )) %>% 
    # select(sex.analysis.indicator,
    #        n.smear.positive.tb.female,
    #        n.smear.positive.tb.male,
    #        adj.prev100k.smear.positive.tb.female,
    #        adj.prev100k.smear.positive.tb.male,
    #        prev100k.smear.positive.tb.female,
    #        prev100k.smear.positive.tb.male,
    #        n.participants.female,
    #        n.participants.male,
    #        diagnosedCases) %>% View()
    select(diagnosedCases) %>% round() %>% sum(na.rm = TRUE) -> cases3


cleanSexDF %>% filter(sex.analysis.indicator %in% c("adj.prev100k.ci.smear.positive.tb", 
                                                    "prev100k.ci.smear.positive.tb",
                                                    "prev100k.smear.positive.tb")) %>%
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = case_when(!is.na(n.smear.positive.tb.male) ~ 
                                           n.smear.positive.tb.male, 
                                       !is.na(n.participants.male)  & !is.na(adj.prev100k.smear.positive.tb.male) ~
                                           adj.prev100k.smear.positive.tb.male * n.participants.male,
                                       !is.na(n.participants.male)  & !is.na(prev100k.smear.positive.tb.male) ~
                                           prev100k.smear.positive.tb.male * n.participants.male,
    )) %>% 
    select(diagnosedCases) %>% round() %>% sum(na.rm = TRUE) -> maleCases3

### Total TB cases diagnosed (approximately)
sum(cases1, cases2, cases3)

### Total TB cases diagnosed among men(approximately)
sum(maleCases1, maleCases2, maleCases3)

### Percentage male 
sum(maleCases1, maleCases2, maleCases3) / sum(cases1, cases2, cases3) *100

bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=WHO.region)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=WHO.region)
table(bactModelDF$natRep)
bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=natRep)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=natRep)
table(bactModelDF$bias)
bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=bias)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=bias)
table(bactModelDF$symptomScreen)
bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=symptomScreen)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=symptomScreen)

bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=symptomsRequired)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=symptomsRequired)

bactModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=XpertCultRequired)
bactModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=XpertCultRequired)

```

## Model 1: Main effect model

Model specification
```{r, warning = FALSE, message=FALSE}
mainEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|study.country) + (1|id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept),
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

bayes_R2(mainEffectMod)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectMod)
plot(mainEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- bactModelDF %>%
      select(study.country) %>%
      unique() %>% 
      mutate(logOR_se=1)

mainEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=nd,
                                 re_formula = ~ 1, 
                                 allow_new_levels = TRUE) %>% 
                    mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
### Posterior mean with 95% credible interval
mainEffectModEst %>% ungroup() %>% mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### Posterior median with 95% credible interval
mainEffectModEst %>% ungroup() %>% median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### output for contribution table
exp(as.numeric(summary(mainEffectMod)[["fixed"]][1]))
exp(as.numeric(summary(mainEffectMod)[["fixed"]][3]))
exp(as.numeric(summary(mainEffectMod)[["fixed"]][4]))
```

Survey-level effects

```{r}
surveyEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=bactModelDF,
                                 re_formula = ~ 1  + (1|study.country) + (1|id), 
                                 allow_new_levels = TRUE) %>% 
                      mutate(mfRatio = exp(.epred),
                             type="Posterior mean and 95% credible interval") 
```

## Model 2: Effect by world region model

Model specification
```{r, warning = FALSE, message=FALSE}
regionEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(regionEffectMod)
plot(regionEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndRegion <- bactModelDF %>%
            select(WHO.region) %>%
            unique() %>%
            mutate(logOR_se=1)

regionEffectModEst <- add_epred_draws(object = regionEffectMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region),
                                 allow_new_levels = TRUE) %>% 
                       mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectModEst %>% group_by(WHO.region) %>% mean_qi(mfRatio) 
```


```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 24, fig.width = 20}
### Get the empirical values 
empirical_prev <- bactModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        bias = bias,
        symptomScreen = symptomScreen, 
        p_female_empirical = plogis(LogOdds_Female),
        p_male_empirical = plogis(LogOdds_Male),
        or_empirical = p_male_empirical / p_female_empirical,
        or_empirical = (p_male_empirical / (1 - p_male_empirical)) / 
            (p_female_empirical / (1 - p_female_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific TB prevalence")
myPalNew <- c("black", "deepskyblue")
    ggplot() +
    # geom_hline(aes(yintercept = 1), linetype=2) +
    geom_point(data = empirical_prev, aes(y=WHO.region, x=or_empirical, color = type2), alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectModEst, aes(x=mfRatio, y=WHO.region, fill=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPalNew[c(2,1)]) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio",
         y = "") +
    theme_bw(base_size = 20) + 
    geom_vline(xintercept=1, linetype = 2, size =2) +
        theme(legend.title = element_blank(), 
              legend.position = "bottom",
              # legend.position = c(.76,.06), 
              legend.background = element_blank(), 
          panel.grid.minor=element_blank(), 
          plot.title = element_text(size=30)) -> plot1
    
    ggsave(plot1, filename=here("output/SEX/UnionPosterPlot.png"), width = 20, height=24, units = "in", dpi=600)
    
```
## Manuscript figure 3: Forest plot
Bring estimates from regional model together to make a forest plot for the manuscript.

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.width = 30, fig.height = 50}

# Join survey labels
names <- cleanSexDF0 %>%
    mutate(id = row_number()) %>%
    select(id, figure.id.yr)

# Survey-level estimates
survey_sum <- surveyEffectModEst %>%
    group_by(id) %>%
    mean_qi(mfRatio) %>%
    mutate(grouping = "survey") %>%
    left_join(names, by = "id") %>%
    left_join(bactModelDF %>% select(id, WHO.region), by = "id")

# Regional summaries
region_sum <- regionEffectModEst %>%
    group_by(WHO.region) %>%
    mean_qi(mfRatio) %>%
    mutate(grouping = "regional") %>%
    rename(figure.id.yr = WHO.region)

# Global summary
global_sum <- mainEffectModEst %>%
    ungroup() %>%
    mean_qi(mfRatio) %>%
    mutate(grouping = "Overall",  
           figure.id.yr = "Overall")

# Combine all
forestDF <- bind_rows(survey_sum, region_sum, global_sum)

# Fill display label and facet group
forestDF <- forestDF %>%
    mutate(
        display_label_raw = ifelse(is.na(figure.id.yr), grouping, figure.id.yr),
        facet_group = case_when(
            grouping == "survey" ~ WHO.region,
            grouping == "regional" ~ figure.id.yr,
            grouping == "Overall" ~ "Overall"
        ), 
        type="Posterior mean and 95% credible interval"
    )

# Define facet order
group_order <- c(
    "Africa region", 
    "Americas region",
    "Eastern Mediterranean region", 
    "South-East Asia region", 
    "Western Pacific region", 
    "Overall"
)

# Separate surveys and summaries
surveys_only <- forestDF %>%
    filter(grouping == "survey") %>%
    group_by(facet_group) %>%
    arrange(desc(mfRatio), .by_group = TRUE) %>%
    ungroup()

summaries_only <- forestDF %>%
    filter(grouping != "survey")

# Recombine with surveys first, then summaries (preserves facet-wise order)
forestDF_plot <- bind_rows(surveys_only, summaries_only) %>%
    mutate(facet_group = factor(facet_group, levels = group_order)) %>%
    mutate(display_label =
           case_when(
        grouping == "Overall" ~ "<b>Overall</b>",
        grouping == "regional" ~ paste0("<b>", display_label_raw, "</b>"),
        TRUE ~ display_label_raw
    )) %>%
    mutate(display_label = factor(display_label, levels = rev(unique(display_label))))

# Prepare empirical data
empirical_prev_labels <- left_join(empirical_prev, survey_sum) %>%
    rename(display_label_raw = figure.id.yr) %>%
    select(display_label_raw, WHO.region, or_empirical) %>%
    mutate(
        grouping = "survey",
        facet_group = WHO.region,
        display_label = display_label_raw
    ) %>%
    mutate(facet_group = factor(facet_group, levels = group_order),
           type = "Ratios calculated from reported prevalence")

###############################################################################
###############################################################################
forestRatio <- ggplot(forestDF_plot) +
    geom_hline(yintercept = 1, linetype = "dashed", colour = "black") +
    geom_point(data = empirical_prev_labels,
               aes(x = display_label, y = or_empirical, color=type), alpha=0.5, size = 3) +
    geom_pointinterval(aes(x = display_label, y = mfRatio, 
                           ymin = .lower, ymax = .upper, fill=type),
                       size = 1) +
    coord_flip() +
    facet_grid(facet_group ~ ., scales = "free", space = "free") +
    scale_color_manual(values = myPal[c(2,1)]) +
    scale_y_log10(
        breaks = c(0.25, 0.5, 1, 2, 4),
        labels = c("0.25", "0.5", "1", "2", "4"),
    ) +
    # scale_x_discrete(labels = label_wrap(30)) + 
    labs(
        x = NULL,
        y = "Male-to-female ratio of bacteriologically-confirmed TB prevalence"
    ) +
    theme_minimal(base_size = 30)  + 
    theme(axis.line.y = element_blank(),
          axis.ticks.y= element_blank(),
          axis.text.y= element_blank(),
          axis.title.x= element_text(size=18), 
          axis.text.x = element_text(size=14),
          axis.title.y= element_blank(), 
              strip.text.y = element_blank(),
              strip.background = element_blank(),
      legend.position = "bottom", 
        legend.text = element_text(size=16), 
      legend.title=element_blank()); forestRatio

###############################################################################
###############################################################################
###############################################################################
###############################################################################
# nameRow <- forestDF_plot[1,]; nameRow$mfRatio <- "M:F ratio"
# nameRow[,c(1,3:ncol(nameRow))] <- NA

forestDF_plot1 <- forestDF_plot  %>% 
    mutate(mfRatio = paste0(sprintf("%.2f",round(mfRatio,2)), " (", sprintf("%.2f",round(.lower,2)), ", ", sprintf("%.2f",round(.upper, 2)), ")")) #%>%
# 
# forestDF_plot1 <- forestDF_plot1 %>% 
#     bind_rows(nameRow)
#     rbind(nameRow, forestDF_plot1)

p_right <- forestDF_plot1 %>% 
    ggplot(aes(y = display_label))  + 
    facet_grid(facet_group ~ ., scales = "free", space = "free") +
    geom_text(size=4.5, 
        aes(x = 1, label = mfRatio),
        hjust = 0, 
        fontface = ifelse(forestDF_plot1$figure.id.yr %in% c(unique(cleanSexDF$WHO.region), "Overall"), "bold", "plain")
    ) + theme_void(base_size = 16) +
    theme(
    strip.text.y = element_blank(),
    strip.background = element_blank(),
    panel.spacing.y = unit(1, "lines"),
    # axis.text.y = ggtext::element_markdown(hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    # plot.background = element_rect(fill = "white"),
    legend.position = "bottom", legend.title=element_blank(),
    legend.spacing.y = unit(-.25, 'cm'),
    legend.box.margin=margin(-10,-10,-6,-10),
    panel.spacing = unit(0.18, "cm")
    # plot.margin = unit(c(.1, .7, 0, .1), "cm")
    ) 

###############################################################################
###############################################################################
###############################################################################
###############################################################################
tempForParts <- cleanSexDF0 %>% left_join(cleanSexDF %>% select(covidence.id, id))

forestDF_plot2 <- forestDF_plot %>% left_join(tempForParts %>% 
                                                  select(id, n.participants.female, n.participants.male) , by = "id") 
#### Add regional participants 
regionMalePart <- cleanSexDF0 %>% reframe(n.participants.male = sum(n.participants.male, na.rm=TRUE), .by = WHO.region) %>% 
    arrange(WHO.region)
regionFemalePart <- cleanSexDF0 %>% reframe(n.participants.female = sum(n.participants.female, na.rm=TRUE), .by = WHO.region) %>%
    arrange(WHO.region)

forestDF_plot2[which(forestDF_plot2$figure.id.yr %in% unique(cleanSexDF$WHO.region)), 
               "n.participants.female"] <- regionFemalePart$n.participants.female

forestDF_plot2[which(forestDF_plot2$figure.id.yr %in% unique(cleanSexDF$WHO.region)), 
               "n.participants.male"] <- regionMalePart$n.participants.male


#### Add overall participants 
forestDF_plot2[which(forestDF_plot2$figure.id.yr == "Overall"), 
               "n.participants.female"] <- as.numeric(cleanSexDF0 %>% reframe(n.participants.female = sum(n.participants.female, na.rm=TRUE)) )

forestDF_plot2[which(forestDF_plot2$figure.id.yr == "Overall"), 
               "n.participants.male"] <- as.numeric(cleanSexDF0 %>% reframe(n.participants.male = sum(n.participants.male, na.rm=TRUE)) )


### Add the male ratio  
forestDF_plot2 <- forestDF_plot2 %>%
    mutate(maleRatio = paste0( scales::comma(n.participants.male), "/",  scales::comma(n.participants.male + n.participants.female)), 
           maleRatio = ifelse(maleRatio== "NA/NA", "Not reported", maleRatio))


# nameRow2 <- forestDF_plot2[1,]; nameRow2$maleRatio <- "Males (n/N)"
# forestDF_plot2 <- forestDF_plot2  %>% 
#     bind_rows(as.data.frame(nameRow2)) 

p_left <- forestDF_plot2 %>%
    ggplot(aes(y = display_label)) +
    facet_grid(facet_group ~ ., scales = "free", space = "free") +
    geom_text(size=4.5,
              aes(x = 0, label = maleRatio),
              hjust = 0,
              fontface = ifelse(forestDF_plot2$figure.id.yr %in% c(unique(cleanSexDF$WHO.region), "Overall"), "bold", "plain")
    ) +
    theme_minimal(base_size = 16) + expand_limits(x=c(0,1))+
    theme(
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        panel.border = element_blank(),
        panel.spacing.y = unit(1, "lines"),
        # ggtext::element_markdown(hjust = 1),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        # plot.background = element_rect(fill = "white"),
        legend.position = "bottom", legend.title=element_blank(),
        legend.spacing.y = unit(-.25, 'cm'),
        legend.box.margin=margin(-10,-10,-6,-10),
        # panel.spacing = unit(0.18, "cm"),
        plot.margin = unit(c(.1, .7, 0, .1), "cm"), 
        axis.text.y = ggtext::element_markdown(hjust = 1), 
        axis.title.y=element_blank()
    ); p_left


layout <- c(
    area(t = 0, l = 0, b = 30, r = 3), # left plot, starts at the top of the page (0) and goes 30 units down and 3 units to the right
    area(t = 1.2, l = 3, b = 30, r = 9), # middle plot starts a little lower (t=1) because there's no title. starts 1 unit right of the left plot (l=4, whereas left plot is r=3), goes to the bottom of the page (30 units), and 6 units further over from the left plot (r=9 whereas left plot is r=3)
    area(t = 0, l = 9, b = 30, r = 11) # right most plot starts at top of page, begins where middle plot ends (l=9, and middle plot is r=9), goes to bottom of page (b=30), and extends two units wide (r=11)
)
# final plot arrangement
finalForestPlot <- p_left + forestRatio+ p_right + plot_layout(design = layout)
# finalForestPlot
# Save
# ggsave(plot = finalForestPlot, "output/SEX/mfFigure2.png", height = 24, width=16, units = "in")
```

## Model 3: Effect over time 

Model specification
```{r, warning = FALSE, message=FALSE}
timeEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z + (1 + year_z | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

### Also run this model with only those studies in the horton review
timeEffectHortMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z  + (1 + year_z | study.country) + (1 | id),
                       data = bactModelDF %>% filter(publication.year < 2017,
                                                     id != 17, id!=72),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(timeEffectMod)
plot(timeEffectMod)
```

Prediction

```{r, warning = FALSE, message=FALSE, echo=FALSE}
ndTime <- bactModelDF %>%
          select(logOR_se) %>%
            crossing(year_z = seq(min(cleanSexDF$year_z, na.rm = TRUE), max(cleanSexDF$year_z, na.rm = TRUE), 1))

timeEffectModEst <- add_epred_draws(object = timeEffectMod,
                                 newdata=na.omit(ndTime),
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

timeEffectHortModEst <- add_epred_draws(object = timeEffectHortMod,
                                 newdata=na.omit(ndTime),
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

```

Main estimate
```{r}
timeEffectModEst %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper) %>% round(digits = 2)

### output for contribution table
exp(as.numeric(summary(timeEffectMod)[["fixed"]][2,1]))
exp(as.numeric(summary(timeEffectMod)[["fixed"]][2,3]))
exp(as.numeric(summary(timeEffectMod)[["fixed"]][2,4]))

timeEffectHortModEst %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper)

bayes_R2(timeEffectMod)
bayes_R2(timeEffectHortMod)


timeEffectModEst 

```
### Manuscript figure 4A: M:F ratio over time

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    y <- timeEffectModEst  %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper) 

   # fig4A <-
       ggplot(data = y) +
    geom_lineribbon(.width = 0.95, alpha = 0.25, fill="chocolate4", aes(x = Year,
               y = mfRatio, ymin = .lower, ymax = .upper)) +
    geom_point(data = empirical_prev %>% rename("Study standard error" = logOR_se),
               aes(x=year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = 1/`Study standard error`),
               colour = "black") +
    expand_limits(y=0) +
    labs(y = "Male-to-female odds ratio",
         x = "Year") +
    theme_minimal(base_size = 20)  + 
    scale_size_continuous(range = c(1,10),  guide="none")
```

### Alternative plot for Union 2025 poster
```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 30, fig.width = 60}
    y <- timeEffectModEst  %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper) 

    ggplot(data = y) +
    geom_lineribbon(.width = 0.95, alpha = 0.6, fill="navy", aes(x = Year,
               y = mfRatio, ymin = .lower, ymax = .upper), size = 6) +
    # geom_point(data = empirical_prev %>% rename("Study standard error" = logOR_se),
    #            aes(x=year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = `Study standard error`),
    #            colour = "black") +
    geom_hline(yintercept = 1, linetype = 2, size =3) +
    # expand_limits(y=0) +
    labs(y = "Male-to-female odds ratio",
         x = "Year") +
    theme_minimal(base_size = 60)  + theme(legend.position = "none") +  
    scale_size_continuous(range = c(3,11))
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
minYr <- min(timeEffectModEst$year_z)
maxYr <- max(timeEffectModEst$year_z)

startYr <- timeEffectModEst %>% filter(year_z == minYr) %>% select(mfRatio) %>% ungroup()
endYr <- timeEffectModEst %>% filter(year_z == maxYr) %>% select(mfRatio)%>% ungroup()

data.frame( mfRatio = ((endYr[,"mfRatio"] - startYr[,"mfRatio"])/startYr[,"mfRatio"])/length(minYr:maxYr)*100) %>%
          mean_qi(mfRatio) 
```

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    yHort <- timeEffectHortModEst  %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper)

yFull <- rbind(
    mutate(y, review = "Current review"), 
    mutate(yHort, review = "Horton et al. review")
)

    ggplot(data = yFull) +
    geom_ribbon(.width = 0.95, alpha = 0.4, aes(x = Year,
               y = mfRatio, ymin = .lower, ymax = .upper, fill=review)) +
    geom_line(aes(x = Year, y = mfRatio, color=review), size=1.5) +
    geom_point(data = empirical_prev %>% rename("Study standard error" = logOR_se),
               aes(x=year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = `Study standard error`),
               colour = "black") +
    expand_limits(y=0) +
    scale_fill_manual("", values = c("violet", "seagreen2")) + 
    scale_color_manual("", values = c("darkviolet", "seagreen")) + 
    labs(y = "Male-to-female odds ratio",
         x = "Year") +
    theme_minimal(base_size = 30)  + theme(legend.position = "bottom")
```


## Model 4: Region effect over time 

Model specification
```{r, warning = FALSE, message=FALSE}
timeEffectRegionMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                              1 + year_z * WHO.region + (1 + year_z | study.country) + (1 | id), 
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(timeEffectRegionMod)
plot(timeEffectRegionMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndTimeReg <- bactModelDF %>%
             select(logOR_se, WHO.region) %>%
             distinct() %>%
             crossing(year_z = seq(min(cleanSexDF$year_z, na.rm = TRUE),
                                   max(cleanSexDF$year_z, na.rm = TRUE), 1))

timeEffectRegionModEst <- add_epred_draws(object = timeEffectRegionMod,
                                 newdata=ndTimeReg,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")
```

Main estimate (Time and region model)
```{r, echo=FALSE, warning = FALSE, message=FALSE}
timeEffectRegionModEst %>%
    group_by(year_z) %>%
    mean_qi(mfRatio) %>%
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, mfRatio, .lower, .upper)


#### Annual rate of change 
minYr <- min(timeEffectRegionModEst$year_z)
maxYr <- max(timeEffectRegionModEst$year_z)

startYr <- timeEffectRegionModEst %>% filter(year_z == minYr) %>% select(mfRatio) %>% ungroup()
endYr <- timeEffectRegionModEst %>% filter(year_z == maxYr) %>% select(mfRatio)%>% ungroup()
data.frame(WHO.region = startYr$WHO.region, 
          mfRatio = ((endYr[,"mfRatio"] - startYr[,"mfRatio"])/startYr[,"mfRatio"])/length(minYr:maxYr)*100) %>%
          group_by(WHO.region) %>% 
          mean_qi(mfRatio) 
    
```

### Manuscript figure 4B: M:F ratio over time by region

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
    
# fig4B <- 
    timeEffectRegionModEst %>%
    filter(WHO.region !="Americas region",
           WHO.region !="Eastern Mediterranean region") %>%
    ggplot(aes(x = year_z + round(mean(cleanSexDF$study.end.year, na.rm = TRUE)),
               y = mfRatio,
               fill = WHO.region,
               color = WHO.region)) +
    stat_lineribbon(aes(color = WHO.region, fill = WHO.region), .width = 0.95, alpha = 0.4, size=1.5) +
    geom_point(data = empirical_prev %>% rename("Study standard error" = logOR_se) %>% 
                   filter(WHO.region !="Americas region",
                          WHO.region !="Eastern Mediterranean region"),
               aes(x=year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = 1/`Study standard error`),
               colour = "black") +
    geom_hline(yintercept = 1, linetype = 2) +
    facet_wrap(~ WHO.region) +
    labs(y = "Male-to-female odds ratio",
         x = "Year",
         fill = "WHO region",
         color = "WHO region") +
    scale_color_manual("", values = wes_palette("Cavalcanti1")) +
    scale_fill_manual("", values = wes_palette("Cavalcanti1")) +
    theme_minimal(base_size = 20)  + 
    scale_size_continuous(range = c(1,10),  guide="none") +
    theme(legend.position = "bottom") 

### Combine both panels and save for publication submission
# fig4 <- (fig4A + fig4B) + plot_layout(nrow=2)
# ggsave(fig4, file=here("output/SEX/sex-figure4.png"), height=15, width=10, dpi = 300)


```

## Model 5: Alternative time model 
Check the time models without the five surveys that did not report survey end year. 

Model specification
```{r, warning = FALSE, message=FALSE}
timeEffectRegionModAlt <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                              1 + year_z * WHO.region + (1 + year_z | study.country) + (1 | id), 
                       data = bactModelDF %>% 
                              filter(!covidence.id %in% 
                                    as.vector(cleanSexDF0 %>% 
                                    filter(is.na(study.end.year)) %>% 
                                    select(covidence.id))),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd) +
                           prior(lkj(4), class=cor),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

summary(timeEffectRegionMod)
summary(timeEffectRegionModAlt)
```

## Sub group analysis

### Model 6-7: National and subnational surveys

Model specification
```{r, warning = FALSE, message=FALSE}
natEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(natRep=="Yes"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

subNatEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(natRep=="No"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(natEffectMod)
plot(natEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndNat <- data.frame("natRep" == "Yes") %>% 
                    crossing(unique(bactModelDF$study.country)) %>% 
         mutate(logOR_se=1)

colnames(ndNat)[1:2] <- c("natRep", "study.country")

natEffectModEst <- add_epred_draws(object = natEffectMod,
                                 newdata=ndNat,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

ndNat <- data.frame("natRep" == "No") %>% 
                    crossing(unique(bactModelDF$study.country)) %>% 
         mutate(logOR_se=1)

colnames(ndNat)[1:2] <- c("natRep", "study.country")

subnatEffectModEst <- add_epred_draws(object = subNatEffectMod,
                                 newdata=ndSubnat,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - National surveys
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
natEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### Posterior median with 95% credible interval
natEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

Main estimate - subnational surveys
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
subnatEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper)  %>% round(digits=2)

### Posterior median with 95% credible interval
subnatEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper)  %>% round(digits=2)
```

###Model 8-9: Study quality (low vs. not low)

Model specification
```{r, warning = FALSE, message=FALSE}

lowBiasEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(bias == "Low"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

notLowBiasEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(bias == "Not low"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(lowBiasEffectMod)
plot(lowBiasEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndLowBias <- data.frame("bias" == "Low") %>% 
                        crossing(unique(bactModelDF$study.country)) %>% 
            mutate(logOR_se=1)

colnames(ndLowBias)[1:2] <- c("bias", "study.country")

ndNotLowBias <- data.frame("bias" == "Not low") %>% 
                        crossing(unique(bactModelDF$study.country)) %>% 
            mutate(logOR_se=1)

colnames(ndNotLowBias)[1:2] <- c("bias", "study.country")

ndNotLowBias <- bactModelDF %>%
         filter(bias == "Not low") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

lowBiasEffectModEst <- add_epred_draws(object = lowBiasEffectMod,
                                 newdata=ndLowBias,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

notLowBiasEffectModEst <- add_epred_draws(object = notLowBiasEffectMod,
                                 newdata=ndNotLowBias,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - low risk of bias
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
lowBiasEffectModEst %>% 
    ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### Posterior median with 95% credible interval
lowBiasEffectModEst %>% 
    ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

Main estimate - not low risk of bias
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
notLowBiasEffectModEst %>% 
    ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### Posterior median with 95% credible interval
notLowBiasEffectModEst %>% 
    ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```


### Model 10-11: Symptom screen used vs. not used

Model specification
```{r, warning = FALSE, message=FALSE}

sympYesEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(symptomScreen == "Yes"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

sympNoEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(symptomScreen == "No"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(sympYesEffectMod)
plot(sympYesEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndSympYes <- bactModelDF %>%
         filter(symptomScreen == "Yes") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

ndSympNo <- bactModelDF %>%
         filter(symptomScreen == "No") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

sympYesEffectModEst <- add_epred_draws(object = sympYesEffectMod,
                                 newdata=ndSympYes,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

sympNoEffectModEst <- add_epred_draws(object = sympNoEffectMod,
                                 newdata=ndSympNo,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - used symptom screening
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
sympYesEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### Posterior median with 95% credible interval
sympYesEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

Main estimate - did not use symptom screening
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
sympNoEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### Posterior median with 95% credible interval
sympNoEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

### Model 12-13: Symptoms required vs. not required for sputum collection

Model specification
```{r, warning = FALSE, message=FALSE}
sympReqYesEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(symptomsRequired == "Yes"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

sympReqNoEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(symptomsRequired == "No"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(sympReqYesEffectMod)
plot(sympReqYesEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndSympReqYes <- data.frame("symptomsRequired" == "Yes") %>% 
                        crossing(unique(bactModelDF$study.country)) %>% 
                mutate(logOR_se=1)

colnames(ndSympReqYes)[1:2] <- c("symptomsRequired", "study.country")

ndSympReqNo <- data.frame("symptomsRequired" == "No") %>% 
                        crossing(unique(bactModelDF$study.country)) %>% 
                mutate(logOR_se=1)

colnames(ndSympReqNo)[1:2] <- c("symptomsRequired", "study.country")

ndSympReqNo <- bactModelDF %>%
         filter(symptomsRequired == "No") %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

sympReqYesEffectModEst <- add_epred_draws(object = sympReqYesEffectMod,
                                 newdata=ndSympReqYes,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

sympReqNoEffectModEst <-  add_epred_draws(object = sympReqNoEffectMod,
                                 newdata=ndSympReqNo,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - required symptoms reported
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Posterior mean with 95% credible interval
sympReqYesEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

# Posterior median with 95% credible interval
sympReqYesEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

Main estimate - did not require symptoms reported
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Posterior mean with 95% credible interval
sympReqNoEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

# Posterior median with 95% credible interval
sympReqNoEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

### Model 14-15: Required Xpert or culture vs. not
Reported xpert or culture required for bact-tb diagnosis vs. xpert or culture not required

Model specification
```{r, warning = FALSE, message=FALSE}
xpCltReqYesEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(XpertCultRequired == "Yes"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

xpCltReqNoEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(XpertCultRequired == "No"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(xpCltReqYesEffectMod)
plot(xpCltReqYesEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndxpCltReqYes <- data.frame("XpertCultRequired" == "Yes") %>% 
                        crossing(unique(bactModelDF$study.country)) %>% 
                  mutate(logOR_se=1)

colnames(ndxpCltReqYes)[1:2] <- c("XpertCultRequired", "study.country")

ndxpCltReqNo <- data.frame("XpertCultRequired" == "No") %>% 
                        crossing(unique(bactModelDF$study.country)) %>% 
                  mutate(logOR_se=1)

colnames(ndxpCltReqNo)[1:2] <- c("XpertCultRequired", "study.country")

xpCltReqYesEffectModEst <- add_epred_draws(object = xpCltReqYesEffectMod,
                                 newdata=ndxpCltReqYes,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

xpCltReqNoEffectModEst <-  add_epred_draws(object = xpCltReqNoEffectMod,
                                 newdata=ndxpCltReqNo,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - required xpert or culture
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Posterior mean with 95% credible interval
xpCltReqYesEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

# Posterior median with 95% credible interval
xpCltReqYesEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

Main estimate - did not require xpert or culture
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Posterior mean with 95% credible interval
xpCltReqNoEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

# Posterior median with 95% credible interval
xpCltReqNoEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

### Model 16-20: Country specific estimates

Model specification
```{r, warning = FALSE, message=FALSE}

BangladeshEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | id),
                       data = bactModelDF %>% filter(study.country == "Bangladesh"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

ChinaEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | id),
                       data = bactModelDF %>% filter(study.country == "China"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

VietnamEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | id),
                       data = bactModelDF %>% filter(study.country == "Vietnam"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

EthiopiaEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | id),
                       data = bactModelDF %>% filter(study.country == "Ethiopia"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

IndiaEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | id),
                       data = bactModelDF %>% filter(study.country == "India"),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(BangladeshEffectMod)
summary(ChinaEffectMod)
summary(VietnamEffectMod)
summary(EthiopiaEffectMod)
summary(IndiaEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
bangladeshNd <-  bactModelDF %>%
                 filter(study.country == "Bangladesh") %>%
                 select(study.country) %>%
                 mutate(logOR_se=1)

chinaNd <- bactModelDF %>%
                 filter(study.country == "China") %>%
                 select(study.country) %>%
                 mutate(logOR_se=1)

indiaNd <- bactModelDF %>%
                 filter(study.country == "India") %>%
                 select(study.country) %>%
                 mutate(logOR_se=1)

vietnamNd <- bactModelDF %>%
                 filter(study.country == "Vietnam") %>%
                 select(study.country) %>%
                 mutate(logOR_se=1)

ethiopiaNd <- bactModelDF %>%
                 filter(study.country == "Ethiopia") %>%
                 select(study.country) %>%
                 mutate(logOR_se=1)

BangladeshEffectModEst <- add_epred_draws(object = BangladeshEffectMod,
                                 newdata=bangladeshNd,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

ChinaEffectModEst <- add_epred_draws(object = ChinaEffectMod,
                                 newdata=chinaNd,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

VietnamEffectModEst <- add_epred_draws(object = VietnamEffectMod,
                                 newdata=vietnamNd,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

EthiopiaEffectModEst <- add_epred_draws(object = EthiopiaEffectMod,
                                 newdata=ethiopiaNd,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

IndiaEffectModEst <- add_epred_draws(object = IndiaEffectMod,
                                 newdata=indiaNd,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - Bangladesh
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Posterior mean with 95% credible interval 
BangladeshEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

# Posterior median with 95% credible interval 
BangladeshEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

Main estimate - China
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Posterior mean with 95% credible interval 
ChinaEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

# Posterior median with 95% credible interval 
ChinaEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

Main estimate - Vietnam
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Posterior mean with 95% credible interval 
VietnamEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

# Posterior medin with 95% credible interval 
VietnamEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

Main estimate - Ethiopia
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Posterior mean with 95% credible interval 
EthiopiaEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

# Posterior median with 95% credible interval 
EthiopiaEffectModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

Main estimate - India
```{r, echo=FALSE, warning = FALSE, message=FALSE}
# Posterior mean with 95% credible interval 
IndiaEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

# Posterior median with 95% credible interval 
IndiaEffectModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

```{r}
bactModelDF %>% filter(study.country %in% c("Bangladesh", "China", "Ethiopia", 
                                            "India", "Vietnam")) %>% 
                group_by(study.country) %>%
                summarise(n.total.participants = sum(n.participants.male,
                                                     n.participants.female, 
                                                     na.rm = TRUE), 
                          malePerc=sum(n.participants.male, na.rm=TRUE)/n.total.participants*100)
```

### Model 21-25: Five study-end-year bands 

```{r, warning = FALSE, message=FALSE}

year00_04Mod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(study.end.year < 2005, 
                                                     study.end.year > 1999),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

year05_09Mod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(study.end.year < 2010,
                                                     study.end.year > 2004),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

year10_14Mod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(study.end.year < 2015, 
                                                     study.end.year > 2009),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

year15_19Mod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(study.end.year < 2020,
                                                     study.end.year > 2014),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

year20_24Mod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(study.end.year > 2019),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
nd00_04 <- bactModelDF %>% filter(study.end.year > 1999, 
                                  study.end.year < 2005) %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)

nd05_09 <- bactModelDF %>% filter(study.end.year < 2010,
                                  study.end.year > 2004) %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)
nd10_14 <- bactModelDF %>% filter(study.end.year < 2015, 
                                  study.end.year > 2009)%>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)
nd15_19 <- bactModelDF %>% filter(study.end.year < 2020, 
                                  study.end.year > 2014)%>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)
nd20_24 <- bactModelDF %>% filter(study.end.year > 2019)%>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)


year00_04ModEst <- add_epred_draws(object = year00_04Mod,
                                 newdata=nd00_04,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

year05_09ModEst <- add_epred_draws(object = year05_09Mod,
                                 newdata=nd05_09,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

year10_14ModEst <- add_epred_draws(object = year10_14Mod,
                                 newdata=nd10_14,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

year15_19ModEst <- add_epred_draws(object = year15_19Mod,
                                 newdata=nd15_19,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

year20_24ModEst <- add_epred_draws(object = year20_24Mod,
                                 newdata=nd20_24,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

```

Main estimate - 2000-2004
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
year00_04ModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### Posterior median with 95% credible interval
year00_04ModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### 
dim(bactModelDF %>% filter(study.end.year < 2005,
                           study.end.year > 1999))[1]
bactModelDF %>% filter(study.end.year < 2005,
                           study.end.year > 1999) %>% summarize(participants = sum(n.participants.male, n.participants.female, na.rm=TRUE))
bactModelDF %>% filter(study.end.year < 2005,
                           study.end.year > 1999) %>% summarize(percMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.male, n.participants.female, na.rm=TRUE))*100
```

Main estimate - 2005-2009
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
year05_09ModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### 
dim(bactModelDF %>% filter(study.end.year < 2010,
                                  study.end.year > 2004))[1]
bactModelDF %>% filter(study.end.year < 2010,
                                  study.end.year > 2004) %>% 
summarize(participants = sum(n.participants.male, n.participants.female, na.rm=TRUE))
bactModelDF %>% filter(study.end.year < 2010,
                                  study.end.year > 2004) %>% summarize(percMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.male, n.participants.female, na.rm=TRUE))*100
```

Main estimate - 2010-2014
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
year10_14ModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### 
dim(bactModelDF %>% filter(study.end.year < 2015, 
                                  study.end.year > 2009))[1]
bactModelDF %>% filter(study.end.year < 2015, 
                                  study.end.year > 2009) %>% 
summarize(participants = sum(n.participants.male, n.participants.female, na.rm=TRUE))
bactModelDF %>% filter(study.end.year < 2015, 
                                  study.end.year > 2009) %>% summarize(percMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.male, n.participants.female, na.rm=TRUE))*100
```

Main estimate - 2015-2019
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
year15_19ModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### 
dim(bactModelDF %>% filter(study.end.year < 2020, 
                                  study.end.year > 2014))[1]
bactModelDF %>% filter(study.end.year < 2020, 
                                  study.end.year > 2014) %>% 
summarize(participants = sum(n.participants.male, n.participants.female, na.rm=TRUE))
bactModelDF %>% filter(study.end.year < 2020, 
                                  study.end.year > 2014) %>% summarize(percMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.male, n.participants.female, na.rm=TRUE))*100
```

Main estimate - 2020-2024
```{r, echo=FALSE, warning = FALSE, message=FALSE}
### Posterior mean with 95% credible interval
year20_24ModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### 
dim(bactModelDF %>% filter(study.end.year > 2019))[1]
bactModelDF %>% filter(study.end.year > 2019) %>% 
summarize(participants = sum(n.participants.male, n.participants.female, na.rm=TRUE))
bactModelDF %>% filter(study.end.year > 2019) %>% summarize(percMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.male, n.participants.female, na.rm=TRUE))*100
```

### Model 26-27: Included in Horton review vs. not
```{r, warning = FALSE, message=FALSE}
hortonMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(publication.year < 2017,
+                            id != 20, id!=75),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

postHortonMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + (1 | study.country) + (1 | id),
                       data = bactModelDF %>% filter(id ==20 | id ==75 | 
                                                    publication.year > 2016 | 
                                                    is.na(publication.year)),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndHorton <- bactModelDF %>% filter(publication.year < 2017,
                            id != 17, id!=72) %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)
ndPostHorton <- bactModelDF %>% filter(id ==17 | id ==72 | 
                                                    publication.year > 2016 | 
                                                    is.na(publication.year)) %>%
         select(study.country, LogOdds_Female) %>%
         mutate(logOR_se=1)


hortonModEst <- add_epred_draws(object = hortonMod,
                                 newdata=ndHorton,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

postHortonModEst <- add_epred_draws(object = postHortonMod,
                                 newdata=ndPostHorton,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate - Horton
```{r, echo=FALSE, warning = FALSE, message=FALSE}
hortonModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### 
dim(bactModelDF %>% filter(publication.year < 2017,
                            id != 20, id!=75))[1]
bactModelDF %>% filter(publication.year < 2017,
                            id != 20, id!=75) %>% 
summarize(participants = sum(n.participants.male, n.participants.female, na.rm=TRUE))
bactModelDF %>% filter(publication.year < 2017,
                            id != 20, id!=75) %>% summarize(percMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.male, n.participants.female, na.rm=TRUE))*100
```

Main estimate - post Horton
```{r, echo=FALSE, warning = FALSE, message=FALSE}
postHortonModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

### 
dim(bactModelDF %>% filter(id ==20 | id ==75 | 
                            publication.year > 2016 | 
                            is.na(publication.year)))[1]
bactModelDF %>% filter(id ==20 | id ==75 | 
                            publication.year > 2016 | 
                            is.na(publication.year)) %>% 
summarize(participants = sum(n.participants.male, n.participants.female, na.rm=TRUE))
bactModelDF %>% filter(id ==20 | id ==75 | 
                            publication.year > 2016 | 
                            is.na(publication.year)) %>% summarize(percMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.male, n.participants.female, na.rm=TRUE))*100
```

## Covariate analysis of male-to-female prevalence ratio
### Setup and read in external data covariate data  
```{r, echo=FALSE, warning = FALSE, message=FALSE}
##############################################################################|
#### External covariates  #####################################################
##############################################################################|

covariates <- read.csv(file=here("data/covariates_1990-2021.csv")) %>% 
    dplyr::rename(study.country = location_name) %>%
    filter(
           sex_name != "Both", 
           metric_name == "Rate") %>% 
    select(study.country, year, sex_name, cause_name, val) %>% 
    pivot_wider(names_from = c(sex_name, cause_name), 
                values_from = val) %>% 
    mutate(alcohol.use.disorder = 
               scale(`Male_Alcohol use disorders` - `Female_Alcohol use disorders`), 
           diabetes.mellitus = 
               scale(`Male_Diabetes mellitus` - `Female_Diabetes mellitus`), 
           type1.diabetes.mellitus = 
               scale(`Male_Diabetes mellitus type 1` - `Female_Diabetes mellitus type 1`), 
           type2.diabetes.mellitus = 
               scale(`Male_Diabetes mellitus type 2` - `Female_Diabetes mellitus type 2`), 
           nutritional.deficiencies = 
               scale(`Male_Nutritional deficiencies` - `Female_Nutritional deficiencies`), 
           HIV.AIDS = 
               scale(`Male_HIV/AIDS` - `Female_HIV/AIDS`), 
           TB.incidence = 
               scale(Male_Tuberculosis - Female_Tuberculosis)
           ) %>% 
    select(study.country, 
           year, 
           alcohol.use.disorder, 
           diabetes.mellitus, 
           type1.diabetes.mellitus, 
           type2.diabetes.mellitus,
           nutritional.deficiencies, 
           HIV.AIDS, 
           TB.incidence) %>% 
    rename(study.end.year = year)

### BMI data
bmi <- read.csv(file=here("data/IHME_GBD_2021_COV_1980_2021_ADULT_UNDERWEIGHT_33 Countries.csv")) %>% 
    rename(study.country = location_name) %>%
    select(study.country, year_id, sex, covariate_name_short, mean_value) %>% 
    group_by(year_id, study.country, sex) %>%
    summarize(val = sum(mean_value)) %>%
    pivot_wider(names_from = c(sex), 
                values_from = val) %>% 
    mutate(bmi = as.numeric(unlist(Male)) - as.numeric(unlist(Female))) %>% 
    select(study.country, 
           year_id, 
           bmi) %>%
    rename(study.end.year = year_id)

bmi$bmi <- scale(bmi$bmi)


### Smoking data 
smoking.prevalence <- read.csv(file=here("data/IHME_GBD_2021_COV_1980_2021_SMOKING_PREV_33 Countries.csv")) %>% 
    rename(study.country = location_name) %>%
    select(study.country, year_id, sex, covariate_name_short, mean_value) %>% 
    group_by(year_id, study.country, sex) %>%
    summarize(val = sum(mean_value)) %>%
    pivot_wider(names_from = c(sex), 
                values_from = val) %>% 
    mutate(smoking = as.numeric(unlist(Male)) - as.numeric(unlist(Female))) %>% 
    select(study.country, 
           year_id, 
           smoking) %>%
    rename(study.end.year = year_id)

smoking.prevalence$smoking <- scale(smoking.prevalence$smoking)

### Gender development index 
GDI <- read_csv(here("data/GenderDevelopmentIndex.csv"))[,1:2] %>% 
    mutate(GDI = scale(as.numeric(Value)), 
           GDI.unscaled = as.numeric(Value)) %>% select(!Value) %>%
    rename(study.country = Country)
### Rename several countries 
GDI[which(GDI$study.country == "Eswatini (Kingdom of)"), "study.country"] <- "Eswatini"
GDI[which(GDI$study.country == "Tanzania (United Republic of)"), "study.country"] <- "Tanzania"

### Covariates updated 
covariates <- full_join(covariates, smoking.prevalence)
covariates <- full_join(covariates, bmi)
covariates[which(covariates$study.country == "United Republic of Tanzania"), "study.country"] <- "Tanzania"

covariates <- full_join(covariates, GDI)
    
### Rename several countries 
covariates[which(covariates$study.country == "Viet Nam"), "study.country"] <- "Vietnam"
covariates[which(covariates$study.country == "Lao People's Democratic Republic"), "study.country"] <- "Laos"
covariates[which(covariates$study.country == "Democratic People's Republic of Korea"), "study.country"] <- "North Korea"

newBactModelDF <- left_join(bactModelDF, covariates)

### manually adjust some rows that have missing data 
newBactModelDF[which(newBactModelDF$id == 21), 23:33] <- newBactModelDF[which(newBactModelDF$id == 20), 23:33]

newBactModelDF[which(newBactModelDF$id == 58), 23:33] <- newBactModelDF[which(newBactModelDF$id == 57), 23:33]

newBactModelDF[which(newBactModelDF$id == 51), 23:33] <- newBactModelDF[which(newBactModelDF$id == 34), 23:33]

newBactModelDF[which(newBactModelDF$id == 97), 23:33] <- newBactModelDF[which(newBactModelDF$id == 98), 23:33]

newBactModelDF <- as.data.frame(newBactModelDF)
newBactModelDF[which(newBactModelDF$id == 38), 23:33] <- as.vector(unlist((covariates %>% filter(study.end.year==2011, study.country == "Kenya"))[-c(1,2)]))
```

### Covariate correlation assessment
Estimate the correlation between the covariate data using a BRMS multivariate model with fixed intercepts and sigmas. Approach inspired by this post 
https://solomonkurz.netlify.app/blog/2019-02-16-bayesian-correlations-let-s-talk-options/

```{r, warning = FALSE, message=FALSE}
### Rename the columns so they present well in plot
### Use "Z" as a space delimiter as model run will strip "_" or "." 
formatBactModelDF <- newBactModelDF %>% 
                     rename("StudyZendZyear" = year_z, 
                            "AlcoholZuseZdisorder" = alcohol.use.disorder, 
                            "TypeZ2ZDiabetesZMellitus" = type2.diabetes.mellitus, 
                            "LowZBMI" = bmi, 
                            "HIVZAIDS" = HIV.AIDS, 
                            "GenderZdevelopmentZindex" = GDI, 
                            "Smoking" = smoking)

covCorrMod <-  brms::brm(bf(mvbind(`AlcoholZuseZdisorder`,
                                   `GenderZdevelopmentZindex`,
                                   `HIVZAIDS`, 
                                   `LowZBMI`, 
                                   `Smoking`, 
                                   `TypeZ2ZDiabetesZMellitus`, 
                                   `StudyZendZyear`) ~ 0,
                         sigma ~ 0) + 
                         set_rescor(TRUE), 
                         data    = formatBactModelDF,
                         prior   = prior(lkj(2), class = rescor),
                         family  = gaussian(),
                         control = list(adapt_delta = 0.99),
                         chains  = 4,
                         cores   = 4,
                         iter    = 4000)

# summary(covCorrMod)

    as_draws_df(covCorrMod) %>% 
    select(starts_with("rescor")) %>% 
    gather(label, rho) %>% 
    select(rho, label) %>% 
    mutate(label = gsub("Z", " ", gsub("__", " with ", gsub("rescor__", "", label)))) %>% 
    ggplot(aes(x = rho, y = label)) +
    stat_halfeye(.width = .95, size = 5/4) + theme_minimal() + 
    ylab("") + 
    ggtitle("Estimated correlation of covariates") + 
    expand_limits(x=c(-1,1))

    as_draws_df(covCorrMod) %>% 
    select(starts_with("rescor")) %>% 
    gather(label, rho) %>% 
    select(rho, label) %>% 
    mutate(label = gsub("Z", " ", gsub("__", " with ", gsub("rescor__", "", label)))) %>% 
    group_by(label) %>% 
    mean_qi(rho)

```

### Model 28: HIV prevalence univariate model

Model specification 

```{r, warning = FALSE, message=FALSE}
hivPrevMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                          1 + HIV.AIDS +
                          (1 | study.country) + (1 | id),
                      data = newBactModelDF,
                      prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                              prior(normal(0, 10), class = b) +
                              prior(exponential(2), class = sd),
                      family = gaussian(),
                      control = list(adapt_delta = 0.99),
                      chains = 4,
                      cores = 4,
                      iter = 4000)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
bayes_R2(hivPrevMod)

### output for contribution table 
round(exp(as.numeric(summary(hivPrevMod)[["fixed"]][2,1])),2)
round(exp(as.numeric(summary(hivPrevMod)[["fixed"]][2,3])),2)
round(exp(as.numeric(summary(hivPrevMod)[["fixed"]][2,4])),2)
```

### Model 29: Gender disparity index univariate model 

Model specification 

```{r, warning = FALSE, message=FALSE}
gdiMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + GDI + (1 | study.country) + (1 | id),
                       data = newBactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(gdiMod)

bayes_R2(gdiMod)

### output for contribution table
round(exp(as.numeric(summary(gdiMod)[["fixed"]][2,1])),2)
round(exp(as.numeric(summary(gdiMod)[["fixed"]][2,3])),2)
round(exp(as.numeric(summary(gdiMod)[["fixed"]][2,4])),2)
```

### Model 30: Diabetes univariate model

Model specification 

```{r, warning = FALSE, message=FALSE}
diabetesPrevMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                                  1 +  type2.diabetes.mellitus +
                                  (1 | study.country) + (1 | id),
                              data = newBactModelDF,
                              prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                                  prior(normal(0, 10), class = b) +
                                  prior(exponential(2), class = sd),
                              family = gaussian(),
                              control = list(adapt_delta = 0.99),
                              chains = 4,
                              cores = 4,
                              iter = 4000)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(diabetesPrevMod)

bayes_R2(diabetesPrevMod)

### output for contribution table
round(exp(as.numeric(summary(diabetesPrevMod)[["fixed"]][2,1])),2)
round(exp(as.numeric(summary(diabetesPrevMod)[["fixed"]][2,3])),2)
round(exp(as.numeric(summary(diabetesPrevMod)[["fixed"]][2,4])),2)
```

### Model 31: Low BMI univariate model

Model specification 

```{r, warning = FALSE, message=FALSE}
lowBMIMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                                  1 +  bmi +
                                  (1 | study.country) + (1 | id),
                              data = newBactModelDF,
                              prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                                  prior(normal(0, 10), class = b) +
                                  prior(exponential(2), class = sd),
                              family = gaussian(),
                              control = list(adapt_delta = 0.99),
                              chains = 4,
                              cores = 4,
                              iter = 4000)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(lowBMIMod)

bayes_R2(lowBMIMod)

### output for contribution table
round(exp(as.numeric(summary(lowBMIMod)[["fixed"]][2,1])),2)
round(exp(as.numeric(summary(lowBMIMod)[["fixed"]][2,3])),2)
round(exp(as.numeric(summary(lowBMIMod)[["fixed"]][2,4])),2)
```

### Model 32: Smoking univariate model

Model specification 

```{r, warning = FALSE, message=FALSE}
smokingPrevMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                                   1 +  smoking +
                                   (1 | study.country) + (1 | id),
                               data = newBactModelDF,
                               prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                                   prior(normal(0, 10), class = b) +
                                   prior(exponential(2), class = sd),
                               family = gaussian(),
                               control = list(adapt_delta = 0.99),
                               chains = 4,
                               cores = 4,
                               iter = 4000)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(smokingPrevMod)

bayes_R2(smokingPrevMod)

### output for contribution table
round(exp(as.numeric(summary(smokingPrevMod)[["fixed"]][2,1])),2)
round(exp(as.numeric(summary(smokingPrevMod)[["fixed"]][2,3])),2)
round(exp(as.numeric(summary(smokingPrevMod)[["fixed"]][2,4])),2)
```

### Model 33: Alcohol use disorder univariate model

Model specification 

```{r, warning = FALSE, message=FALSE}
alcoholUsePrevMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                            1 +  alcohol.use.disorder +
                            (1 | study.country) + (1 | id),
                        data = newBactModelDF,
                        prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                            prior(normal(0, 10), class = b) +
                            prior(exponential(2), class = sd),
                        family = gaussian(),
                        control = list(adapt_delta = 0.99),
                        chains = 4,
                        cores = 4,
                        iter = 4000)
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(alcoholUsePrevMod)

bayes_R2(alcoholUsePrevMod)

### output for contribution table
round(exp(as.numeric(summary(alcoholUsePrevMod)[["fixed"]][2,1])),2)
round(exp(as.numeric(summary(alcoholUsePrevMod)[["fixed"]][2,3])),2)
round(exp(as.numeric(summary(alcoholUsePrevMod)[["fixed"]][2,4])),2)
```

### Model 34: Multi-variable covariate analysis

Model specification 

```{r, warning = FALSE, message=FALSE}

multiMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                          1 + year_z + GDI +
                          alcohol.use.disorder + type2.diabetes.mellitus + 
                          bmi + HIV.AIDS + smoking + 
                          (1 | study.country) + (1 | id),
                      data = newBactModelDF,
                      prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                          prior(normal(0, 10), class = b) +
                          prior(exponential(2), class = sd),
                      family = gaussian(),
                      control = list(adapt_delta = 0.99),
                      chains = 4,
                      cores = 4,
                      iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(multiMod)

### output for contribution table
x <- data.frame(parameter = rownames(summary(multiMod)[["fixed"]]),
                PE = round(exp(as.numeric(summary(multiMod)[["fixed"]][,1])),2), 
                CrI = paste(round(exp(as.numeric(summary(multiMod)[["fixed"]][,3])),2), 
                          ",",
                    round(exp(as.numeric(summary(multiMod)[["fixed"]][,4])),2)))

print(x)

bayes_R2(multiMod)
plot(multiMod)

stanplot(multiMod)
```

## Covariate analysis by world region

### Model 35-75: Univariable covariate analysis by world region

```{r, warning = FALSE, message=FALSE}
regCoVList <- list()
model.regions <- unique(na.omit(bactModelDF$WHO.region))[c(-1,-5)]

for (i in 1:length(model.regions)){
    region <- model.regions[i]
    print(region)
    
    regionMainMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                                1 + (1 | study.country) + (1 | id),
                       data = newBactModelDF %>% filter(WHO.region == region),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
    
    regionHivMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                            1 + HIV.AIDS + 
                            (1 | study.country) + (1 | id),
                   data = newBactModelDF %>% filter(WHO.region == region),
                   prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                       prior(normal(0, 10), class = b) +
                       prior(exponential(2), class = sd),
                   family = gaussian(),
                   control = list(adapt_delta = 0.99),
                   chains = 4,
                   cores = 4,
                   iter = 4000)
    
    regionYearMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                            1 +  year_z + 
                            (1 | study.country) + (1 | id),
                   data = newBactModelDF %>% filter(WHO.region == region),
                   prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                       prior(normal(0, 10), class = b) +
                       prior(exponential(2), class = sd),
                   family = gaussian(),
                   control = list(adapt_delta = 0.99),
                   chains = 4,
                   cores = 4,
                   iter = 4000)
    
    regionGdiMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                            1 + GDI + 
                            (1 | study.country) + (1 | id),
                   data = newBactModelDF %>% filter(WHO.region == region),
                   prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                       prior(normal(0, 10), class = b) +
                       prior(exponential(2), class = sd),
                   family = gaussian(),
                   control = list(adapt_delta = 0.99),
                   chains = 4,
                   cores = 4,
                   iter = 4000)
    
    regionDiabetesMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                        1 + diabetes.mellitus + 
                        (1 | study.country) + (1 | id),
               data = newBactModelDF %>% filter(WHO.region == region),
               prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                   prior(normal(0, 10), class = b) +
                   prior(exponential(2), class = sd),
               family = gaussian(),
               control = list(adapt_delta = 0.99),
               chains = 4,
               cores = 4,
               iter = 4000)
    
    regionBMIMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                    1 + bmi + 
                    (1 | study.country) + (1 | id),
           data = newBactModelDF %>% filter(WHO.region == region),
           prior = prior(student_t(7, 0, 1.5), class = Intercept) +
               prior(normal(0, 10), class = b) +
               prior(exponential(2), class = sd),
           family = gaussian(),
           control = list(adapt_delta = 0.99),
           chains = 4,
           cores = 4,
           iter = 4000)
    
    regionAlcoholMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                1 + alcohol.use.disorder + 
                (1 | study.country) + (1 | id),
       data = newBactModelDF %>% filter(WHO.region == region),
       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
           prior(normal(0, 10), class = b) +
           prior(exponential(2), class = sd),
       family = gaussian(),
       control = list(adapt_delta = 0.99),
       chains = 4,
       cores = 4,
       iter = 4000)
    
    regionSmokingMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                1 + smoking + 
                (1 | study.country) + (1 | id),
       data = newBactModelDF %>% filter(WHO.region == region),
       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
           prior(normal(0, 10), class = b) +
           prior(exponential(2), class = sd),
       family = gaussian(),
       control = list(adapt_delta = 0.99),
       chains = 4,
       cores = 4,
       iter = 4000)
    
    tmp <- data.frame(parameter = c(rownames(summary(regionMainMod)[["fixed"]]), 
                                  rownames(summary(regionGdiMod)[["fixed"]])[2], 
                                  rownames(summary(regionYearMod)[["fixed"]])[2],
                                  rownames(summary(regionAlcoholMod)[["fixed"]])[2],
                                  rownames(summary(regionDiabetesMod)[["fixed"]])[2],
                                  rownames(summary(regionHivMod)[["fixed"]])[2], 
                                  rownames(summary(regionSmokingMod)[["fixed"]])[2],
                                  rownames(summary(regionBMIMod)[["fixed"]])[2]), 
                      
                    PE = c(round(exp(as.numeric(summary(regionMainMod)[["fixed"]][1,1])),4), 
                           round(exp(as.numeric(summary(regionGdiMod)[["fixed"]][2,1])),4), 
                           round(exp(as.numeric(summary(regionYearMod)[["fixed"]][2,1])),4), 
                           round(exp(as.numeric(summary(regionAlcoholMod)[["fixed"]][2,1])),4), 
                           round(exp(as.numeric(summary(regionDiabetesMod)[["fixed"]][2,1])),4),
                           round(exp(as.numeric(summary(regionHivMod)[["fixed"]][1,1])),4), 
                           round(exp(as.numeric(summary(regionSmokingMod)[["fixed"]][2,1])),4), 
                           # round(exp(as.numeric(summary(regionTBIncMod)[["fixed"]][2,1])),4),
                           round(exp(as.numeric(summary(regionBMIMod)[["fixed"]][2,1])),4)), 
                    
                    CrI = c(paste(round(exp(as.numeric(summary(regionMainMod)[["fixed"]][1,3])),4), ",",
                                round(exp(as.numeric(summary(regionMainMod)[["fixed"]][1,4])),4)),
                            
                        paste(round(exp(as.numeric(summary(regionGdiMod)[["fixed"]][2,3])),4), ",",
                                    round(exp(as.numeric(summary(regionGdiMod)[["fixed"]][2,4])),4)),
                        
                        paste(round(exp(as.numeric(summary(regionYearMod)[["fixed"]][2,3])),4), ",",
                              round(exp(as.numeric(summary(regionYearMod)[["fixed"]][2,4])),4)), 
                        
                        paste(round(exp(as.numeric(summary(regionAlcoholMod)[["fixed"]][2,3])),4), ",",
                              round(exp(as.numeric(summary(regionAlcoholMod)[["fixed"]][2,4])),4)), 
                        
                        paste(round(exp(as.numeric(summary(regionDiabetesMod)[["fixed"]][2,3])),4), ",",
                              round(exp(as.numeric(summary(regionDiabetesMod)[["fixed"]][2,4])),4)), 
                        
                        paste(round(exp(as.numeric(summary(regionHivMod)[["fixed"]][2,3])),4), ",",
                              round(exp(as.numeric(summary(regionHivMod)[["fixed"]][2,4])),4)), 
                        
                        paste(round(exp(as.numeric(summary(regionSmokingMod)[["fixed"]][2,3])),4), ",",
                                    round(exp(as.numeric(summary(regionSmokingMod)[["fixed"]][2,4])),4)), 
                        
                        # paste(round(exp(as.numeric(summary(regionTBIncMod)[["fixed"]][2,3])),4), ",",
                        #         round(exp(as.numeric(summary(regionTBIncMod)[["fixed"]][2,4])),4)),
                    
                        paste(round(exp(as.numeric(summary(regionBMIMod)[["fixed"]][2,3])),4), ",",
                              round(exp(as.numeric(summary(regionBMIMod)[["fixed"]][2,4])),4))))
    
    regCoVList[[region]] <- tmp
}
```

#### Africa region (Univariable)
```{r, echo=FALSE, warning = FALSE, message=FALSE}
regCoVList[["Africa region"]]
```

####South-East Asia region (Univariable)
```{r, echo=FALSE, warning = FALSE, message=FALSE}
regCoVList[["South-East Asia region"]]
```

#### Western Pacific region (Univariable)
```{r, echo=FALSE, warning = FALSE, message=FALSE}
regCoVList[["Western Pacific region"]]
```

### Models 76-80: Multivariable covariate analysis by world region

```{r, echo=FALSE, warning = FALSE, message=FALSE}
regMultiList <- list()

for (i in 1:length(model.regions)){
    region <- unique(model.regions)[i]
    print(region)
    
    regionMultiMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                                1 + year_z + GDI +
                                alcohol.use.disorder + diabetes.mellitus + 
                                bmi + HIV.AIDS + 
                                smoking + 
                                  (1 | study.country) + (1 | id),
                       data = newBactModelDF %>% filter(WHO.region == region),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
    
    x <- data.frame(parameter = rownames(summary(regionMultiMod)[["fixed"]]),
                PE = round(exp(as.numeric(summary(regionMultiMod)[["fixed"]][,1])),4), 
                CrI = paste(round(exp(as.numeric(summary(regionMultiMod)[["fixed"]][,3])),4), 
                          ",",
                    round(exp(as.numeric(summary(regionMultiMod)[["fixed"]][,4])),4)))
    
    regMultiList[[region]] <- x 
    
}
```

#### Africa region (multivariable)
```{r, echo=FALSE, warning = FALSE, message=FALSE}
regMultiList[["Africa region"]]
```

#### South-East Asia region (multivariable)
```{r, echo=FALSE, warning = FALSE, message=FALSE}
regMultiList[["South-East Asia region"]]
```

#### Western Pacific region (multivariable)
```{r, echo=FALSE, warning = FALSE, message=FALSE}
regMultiList[["Western Pacific region"]]
```


# Smear positive analysis 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
cleanSexSmrDF0 <- smrPosIndicator("sex") %>% filter(sex.smr.analysis.indicator !="none")

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexSmrDF <- cleanSexSmrDF0 %>% 
                 filter(sex.smr.analysis.indicator != "none") %>% 
                 mutate("id" = row_number()) %>%
                 ### Scale all prevalence estimates back to decimals 
                 mutate_at(vars(matches('prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>% 
                 select(id,
                       covidence.id, 
                       figure.id.yr, 
                       study.geography, 
                       title.extracted,
                       study.country,
                       WHO.region,
                       study.start.year, 
                       study.end.year, 
                       publication.year,
                       study.quality.summary,
                       screening.tests,
                       bacteriological.tb.definition, 
                       xpert.samples, xpert.used, other.xpert.samples,
                       culture.samples, culture.used, other.culture.samples,
                       participants.sputum.collected,
                       ends_with("male"), ends_with("sex"),
                       sex.smr.analysis.indicator) %>%
    
    mutate("tempPrevMale" =  case_when(sex.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.smr.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                        sex.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tempPrevFemale" = case_when(sex.smr.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.smr.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female)) %>%
    pivot_longer(cols = c(tempPrevFemale,tempPrevMale), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.smear.positive.tb.male-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.smear.positive.tb.female-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                       
                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.smear.positive.tb.male-
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.smear.positive.tb.female-
                                                                   prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.smr.analysis.indicator %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.male),
                                        sex.smr.analysis.indicator  %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.female)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.male/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.male /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                              sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.female/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.female)))/3.92,
                                                 sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Male" ~ (log(prev100k.ci.upper.smear.positive.tb.male/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.male)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.male /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                                 sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Female" ~ (log(prev100k.ci.upper.smear.positive.tb.female/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.female)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.female /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.female)))/3.92, 
                                            sex.smr.analysis.indicator %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            log(1/(n.participants.male*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        sex.smr.analysis.indicator  %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            log(1/(n.participants.female*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(is.na(study.geography), 
                                                       NA, 
                                                       ifelse(study.geography == "Nationally representative", 
                                                       "Yes", "No")),

                                       year_z = ifelse(!is.na(study.end.year), 
                                                       study.end.year - round(mean(study.end.year, na.rm=TRUE)),
                                                       (publication.year - meanLag) - round(mean((publication.year - meanLag), na.rm=TRUE))),                                       
                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       
                                       symptomScreen = ifelse(is.na(screening.tests), 
                                                                    NA, 
                                                                    ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No")),
                                       symptomsRequired = ifelse(is.na(participants.sputum.collected) | grepl("unknown", 
                                                                                                              participants.sputum.collected,ignore.case=TRUE), 
                                                                       NA, 
                                                                       ifelse(!grepl("x-ray|all participants|other",
                                                                        participants.sputum.collected,ignore.case=TRUE),
                                                                 "Yes", "No")),

                                       XpertCultRequired = ifelse(is.na(bacteriological.tb.definition),
                                                                  NA, ifelse(!grepl("smear|micros|field|NTP", 
                                                                              bacteriological.tb.definition,
                                                                              ignore.case=TRUE), 
                                                                  "Yes", 
                                                                   "No")),
                                       ### Fix those that were missed with the above attempt 
                                       XpertCultRequired = ifelse(covidence.id %in% c(26989,26928,26920,26918,26892,26861,26857), 
                                                                  "Yes", XpertCultRequired), 
                                       XpertCultRequired = ifelse(covidence.id %in% c(28002), 
                                                                  "No", XpertCultRequired),
                                       #### if we used the culture positive count, then xpert culture was requried
                                       XpertCultRequired = ifelse(grepl("culture", sex.smr.analysis.indicator), 
                                                                  "Yes", XpertCultRequired),
                                       #### if we used the smear positive prevs, then xpert culture was not requried
                                       XpertCultRequired = ifelse(grepl("smear", sex.smr.analysis.indicator), 
                                                                  "No", XpertCultRequired),
                                       XpertCultUsed = ifelse((grepl("Yes", xpert.used) & 
                                                                 (grepl("All samples", xpert.samples) |
                                                                  !grepl("mear|NAAT", other.xpert.samples))) |
                                                                  (grepl("Yes", culture.used) & 
                                                                  grepl("All samples|Subset", culture.samples) & 
                                                                  !grepl("HIV", other.culture.samples)), 
                                                                 "Yes", "No"),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))  %>% filter(LogOddsStandardError < Inf)

smrModelDF <- cleanSexSmrDF %>%
    select(id,covidence.id, WHO.region, study.country, Sex, year_z, natRep, 
           symptomsRequired, XpertCultRequired, XpertCultUsed, 
           LogOdds, LogOddsStandardError, 
           n.participants.male, n.participants.female) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_Male - LogOdds_Female,
        logOR_se = sqrt(LogOddsStandardError_Male^2 + LogOddsStandardError_Female^2) 
    ) 
```

## Smear positive descriptive statistics

```{r, warning = FALSE, message=FALSE, echo=FALSE}

print(paste(dim(cleanSexSmrDF0)[1], "unique surveys with sex-stratified smear positive TB"))

# sort(unique(cleanSexDF0$title.extracted))
print(paste(length(unique(cleanSexSmrDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanSexSmrDF0$n.participants.female)), "studies report participant numbers totaling", 
            sum(cleanSexSmrDF0$n.participants.female, cleanSexSmrDF0$n.participants.male, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanSexSmrDF0$n.participants.male, na.rm=TRUE)/sum(cleanSexSmrDF0$n.participants.female, cleanSexSmrDF0$n.participants.male, na.rm = TRUE)*100, 2), "% of participants were male" ))

### How many persons diagnosed with bacteriologically-confirmed TB 
### Case 1: we used the reported count so simply sum 
### 27 total surveys; no missing 
cleanSexSmrDF %>% filter(sex.smr.analysis.indicator == "n.smear.positive.tb") %>% 
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = n.smear.positive.tb.male + n.smear.positive.tb.female) %>% 
    select(diagnosedCases) %>% sum() -> smrCases1

## Among males only
cleanSexSmrDF %>% filter(sex.smr.analysis.indicator == "n.smear.positive.tb") %>% 
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = n.smear.positive.tb.male) %>% 
    select(diagnosedCases) %>% sum() -> smrMaleCases1

### Case 2: we used smear positive prevalence risk so pull their reported 
### counts (if available) or we need participant counts. 
### 37 total surveys; 27 reported case counts; 10 estimated; 0 missing 
cleanSexSmrDF %>% filter(sex.smr.analysis.indicator != "n.smear.positive.tb") %>%
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = case_when(!is.na(n.smear.positive.tb.male) ~
                                           n.smear.positive.tb.male + n.smear.positive.tb.female,
                                       !is.na(n.participants.male)  & !is.na(adj.prev100k.smear.positive.tb.male) ~
                                           adj.prev100k.smear.positive.tb.male * n.participants.male +
                                           adj.prev100k.smear.positive.tb.female * n.participants.female,
                                       !is.na(n.participants.male)  & !is.na(prev100k.smear.positive.tb.male) ~
                                           prev100k.smear.positive.tb.male * n.participants.male +
                                           prev100k.smear.positive.tb.female * n.participants.female
    )) %>%
#     select(sex.smr.analysis.indicator,
#            n.smear.positive.tb.female,
#            n.smear.positive.tb.male,
#            adj.prev100k.smear.positive.tb.female,
#            adj.prev100k.smear.positive.tb.male,
#            prev100k.ci.smear.positive.tb.female,
#            prev100k.ci.smear.positive.tb.male,
#            prev100k.smear.positive.tb.female,
#            prev100k.smear.positive.tb.male,
#            n.participants.female,
#            n.participants.male,
#            diagnosedCases) %>% View()
    select(diagnosedCases) %>% round() %>% sum(na.rm = TRUE) -> smrCases2

## Among males only
cleanSexSmrDF %>% filter(sex.smr.analysis.indicator != "n.smear.positive.tb") %>%
    select(!Sex:LogOddsStandardError) %>% distinct() %>%
    mutate(diagnosedCases = case_when(!is.na(n.smear.positive.tb.male) ~ 
                                           n.smear.positive.tb.male, 
                                       !is.na(n.participants.male)  & !is.na(adj.prev100k.smear.positive.tb.male) ~
                                           adj.prev100k.smear.positive.tb.male * n.participants.male,
                                       !is.na(n.participants.male)  & !is.na(prev100k.smear.positive.tb.male) ~
                                           prev100k.smear.positive.tb.male * n.participants.male
                                           
    )) %>% 
    select(diagnosedCases) %>% round() %>% sum(na.rm = TRUE) -> smrMaleCases2

### Total TB cases diagnosed (approximately)
sum(cases1, cases2)

### Total TB cases diagnosed among men(approximately)
sum(maleCases1, maleCases2)

### Percentage male 
sum(maleCases1, maleCases2) / sum(cases1, cases2) *100


### How many with culture positive TB 
cleanSexDF %>% filter(!is.na(n.culture.positive.tb.male)) %>%
    select(!Sex:LogOddsStandardError) %>% distinct() %>% 
    # select(n.culture.positive.tb.male, n.culture.positive.tb.female) %>% 
    # View()
    nrow()
```

## More descriptives 

```{r}
table(smrModelDF$WHO.region)
smrModelDF %>% reframe(Participants = sum(n.participants.female + n.participants.male, na.rm=TRUE), .by=WHO.region)
smrModelDF %>% reframe(PercMale = sum(n.participants.male, na.rm=TRUE)/sum(n.participants.female + n.participants.male, na.rm=TRUE)*100, .by=WHO.region)
```

## Model 81: Smear positive main effect 

Model specification
```{r, warning = FALSE, message=FALSE}
mainEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | study.country) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(mainEffectSmrMod)
plot(mainEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
nd <- smrModelDF %>%
      select(study.country, LogOdds_Female) %>%
      mutate(logOR_se=1)

mainEffectSmrModEst <- add_epred_draws(object = mainEffectSmrMod,
                                 newdata=nd,
                                 re_formula = ~ 1) %>% 
                    mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectSmrModEst %>% ungroup %>% mean_qi(mfRatio) %>% select(mfRatio, .lower, .upper) %>% round(digits=2)
```

## Model 82: Smear positive effect by world region model

Model specification
```{r, warning = FALSE, message=FALSE}
regionEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(regionEffectSmrMod)
plot(regionEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
ndRegion <- smrModelDF %>%
            select(WHO.region, LogOdds_Female) %>%
            mutate(logOR_se=1)

regionEffectSmrModEst <- add_epred_draws(object = regionEffectSmrMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region)) %>% 
                       mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectSmrModEst %>% group_by(WHO.region) %>% mean_qi(mfRatio)
```
## Forest plot 
```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
### Get the empirical values 
empirical_smr_prev <- smrModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        p_female_empirical = plogis(LogOdds_Female),
        p_male_empirical = plogis(LogOdds_Male),
        or_empirical = p_male_empirical / p_female_empirical,
        or_empirical = (p_male_empirical / (1 - p_male_empirical)) / 
            (p_female_empirical / (1 - p_female_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific smear positive TB prevalence")


    ggplot() +
    # geom_hline(aes(yintercept = 1), linetype=2) +
    geom_point(data = empirical_smr_prev, aes(y=WHO.region, x=or_empirical, color = type2),alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectSmrModEst, aes(x=mfRatio, y=WHO.region, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio",
         y = "World region") +
    theme_bw(base_size = 30) + theme(legend.title = element_blank(), legend.position = c(.83,.07), 
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
    
```

# Some comparisons between smear-positive and bacteriologically-confirmed TB 

```{r, warning = FALSE, message=FALSE}
### Reviewer noted that while some of the WHO regions had extremely similar MF ratios for smr vs. bc, others had larger differences. Is there any reason we can see in the data to explain this? 

### Small difference: South East Asia Region
### Relatively large difference: Africa Region

### Is there a difference in the percentage of smear positive used as bacteriologically positive between the two regions? 
### South East Asia Region
nrow(bactModelDF %>% filter(WHO.region == "South-East Asia region", grepl("smear", sex.analysis.indicator))) / 
    nrow(bactModelDF %>% filter(WHO.region == "South-East Asia region")) * 100

### Africa Region
nrow(bactModelDF %>% filter(WHO.region == "Africa region", grepl("smear", sex.analysis.indicator))) / 
    nrow(bactModelDF %>% filter(WHO.region == "South-East Asia region")) * 100

### Consider effect sizes in a moment 

### Do we see differences in symptoms required? 
## South East Asia Region
nrow(bactModelDF %>% filter(WHO.region == "South-East Asia region", symptomsRequired == "Yes")) / 
    nrow(bactModelDF %>% filter(WHO.region == "South-East Asia region")) * 100

nrow(smrModelDF %>% filter(WHO.region == "South-East Asia region", symptomsRequired == "Yes")) / 
    nrow(smrModelDF %>% filter(WHO.region == "South-East Asia region")) * 100

### Africa Region
nrow(bactModelDF %>% filter(WHO.region == "Africa region",  symptomsRequired == "Yes")) / 
    nrow(bactModelDF %>% filter(WHO.region == "South-East Asia region")) * 100

nrow(smrModelDF %>% filter(WHO.region == "Africa region",  symptomsRequired == "Yes")) / 
    nrow(smrModelDF %>% filter(WHO.region == "South-East Asia region")) * 100

### Do we see differences in symptoms required? 
## South East Asia Region
nrow(bactModelDF %>% filter(WHO.region == "South-East Asia region", XpertCultRequired == "Yes")) / 
    nrow(bactModelDF %>% filter(WHO.region == "South-East Asia region")) * 100

### Africa Region
nrow(bactModelDF %>% filter(WHO.region == "Africa region",  XpertCultRequired == "Yes")) / 
    nrow(bactModelDF %>% filter(WHO.region == "South-East Asia region")) * 100

### Some difference -- what about effect sizes?
bactModelDF %>% 
    filter(WHO.region == "South-East Asia region", XpertCultRequired == "Yes") %>% 
    mutate(MFRatio = exp(logOR)) %>% 
    select(MFRatio) %>% mean()

bactModelDF %>% 
    filter(WHO.region == "South-East Asia region", XpertCultRequired == "No") %>% 
    mutate(MFRatio = exp(logOR)) %>% 
    select(MFRatio) %>% mean()

bactModelDF %>% 
    filter(WHO.region == "Africa region", XpertCultRequired == "Yes") %>% 
    mutate(MFRatio = exp(logOR)) %>% 
    select(MFRatio) %>% mean()

bactModelDF %>% 
    filter(WHO.region == "Africa region", XpertCultRequired == "No") %>% 
    mutate(MFRatio = exp(logOR)) %>% 
    select(MFRatio) %>% mean()

```

## Model 83: Bacteriologically-confirmed by region using only surveys that also report smear positive
```{r, warning = FALSE, message=FALSE}
### What if we compare only among the surveys that report both the smear and bacteriologically confirmed TB 
reportBothModelDF <- bactModelDF %>% 
                     filter(covidence.id %in% smrModelDF$covidence.id)

regionEffectRptBothMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = reportBothModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

ndBothRegion <- reportBothModelDF %>%
                select(WHO.region) %>%
                unique() %>%
                mutate(logOR_se=1)

regionBothEffectModEst <- add_epred_draws(object = regionEffectRptBothMod,
                                          newdata=ndBothRegion,
                                          re_formula = ~ 1 + (1|WHO.region),
                                          allow_new_levels = FALSE) %>% 
                          mutate(mfRatio = exp(.epred),
                                 type="Posterior mean and 95% credible interval") 

regionBothEffectModEst %>% group_by(WHO.region) %>% mean_qi(mfRatio) 

```

# Bacteriologically positive TB sensitivity analyses

## Model 84: Symptoms required included as covariate in main model

```{r, warning = FALSE, message=FALSE}
sympReqMainMod<- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + symptomsRequired + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept),
                               prior(student_t(7, 0, 1.5), class = b) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)


ndsympReq <- crossing(unique(bactModelDF$study.country), 
                      unique(na.omit(bactModelDF$symptomsRequired))) %>%
    # 
    # bactModelDF %>%
    #             select(study.country, LogOdds_Female, symptomsRequired) %>%
                mutate(logOR_se=1)

colnames(ndsympReq) <- c("study.country", "symptomsRequired", "logOR_se")

sympReqModEst <- add_epred_draws(object = sympReqMainMod,
                                 newdata=ndsympReq,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

sympReqModEst %>% ungroup() %>%
    # filter(symptomsRequired == "No") %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

sympReqModEst %>% ungroup() %>%
    # filter(symptomsRequired == "No") %>%
    median_qi(mfRatio) %>% 
    select(mfRatio) %>% round(digits=2)

bayes_R2(sympReqMainMod)

```

## Model 85: Xpert or culture required for diagnosis 

```{r, warning = FALSE, message=FALSE}
xpCltReqMainMod<- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                           1 + XpertCultRequired + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)


ndXperCult <- bactModelDF %>%
                select(study.country, LogOdds_Female, XpertCultRequired) %>%
                mutate(logOR_se=1)

ndXperCultModEst <- add_epred_draws(object = xpCltReqMainMod,
                                 newdata=ndXperCult,
                                 re_formula = NA) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

ndXperCultModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>%
    select(mfRatio) %>% round(digits=2)

ndXperCultModEst %>% ungroup() %>%
    median_qi(mfRatio) %>%
    select(mfRatio, .lower, .upper) %>% round(digits=2)
```

## Model 86: Xpert or culture used for diagnosis 

```{r, warning = FALSE, message=FALSE}
xpCltUseMainMod<- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~
                           1 + XpertCultUsed + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)


ndXperCultUse <- bactModelDF %>%
                 select(study.country, LogOdds_Female, XpertCultUsed) %>%
                 mutate(logOR_se=1)

ndXperCultUseModEst <- add_epred_draws(object = xpCltUseMainMod,
                                 newdata=ndXperCultUse,
                                 re_formula = NA) %>%
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval")

ndXperCultUseModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>%
    select(mfRatio, .lower, .upper) %>% round(digits=2)

ndXperCultUseModEst %>% ungroup() %>%
    median_qi(mfRatio) %>%
    select(mfRatio) %>% round(digits=2)
```

## Model 87: Country and regional nested random effects 

```{r, warning = FALSE, message=FALSE}
cntryRegMainMod<- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 +  (1 | WHO.region/study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept),
                               prior(student_t(7, 0, 1.5), class = b) +
                               prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)


ndcntryRegReq <- bactModelDF %>% 
                 select(study.country, WHO.region) %>%
                 mutate(logOR_se=1)


cntryRegMainModEst <- add_epred_draws(object = cntryRegMainMod,
                                 newdata=ndcntryRegReq,
                                 re_formula = NA) %>% 
                           mutate(mfRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

cntryRegMainModEst %>% ungroup() %>%
    mean_qi(mfRatio) %>% 
    select(mfRatio, .lower, .upper) %>% round(digits=2)

cntryRegMainModEst %>% ungroup() %>%
    median_qi(mfRatio) %>% 
    select(mfRatio) %>% round(digits=2)

bayes_R2(cntryRegMainMod)
```

