---
title: "Relationship between sex-disaggregated TB Prevalence and Notification Rates in Low- and Middle- Income Countries " 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

```{r Setup, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(brms)
library(tidybayes)
library(scales)
library(wesanderson)
library(ggplot2)

myPal <- c("black", "violet")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))

##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################
cleanSexNatDF0 <- bactPostIndicator("sex") %>% 
                  filter(sex.analysis.indicator != "none", 
                  study.geography == "Nationally representative") 

### Check if any publication years are missing
sum(is.na(cleanSexNatDF0$study.end.year))

### Distribution of samples across reporting data type 
table(cleanSexNatDF0$sex.analysis.indicator)

```
## Format prevalence data

```{r Format Prevalence Data, echo = FALSE,  warning = FALSE, message=FALSE}

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexNatDF <- cleanSexNatDF0 %>% 
                 filter(sex.analysis.indicator != "none") %>% 
                 mutate("id" = row_number()) %>% 
                 ### Scale all prevalence estimates back to decimals 
                 mutate_at(vars(matches('prev100k.bacteriological.tb.|prev100k.ci.([a-z]{3})er.bacteriological.tb.|prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),
                           ~ (. / 1e5)) %>%
                select(id,
                       covidence.id, 
                       figure.id.yr, 
                       study.geography, 
                       title.extracted,
                       study.country,
                       study.rurality, 
                       WHO.region,
                       screening.tests,
                       participants.sputum.collected,
                       bacteriological.tb.definition,
                       xpert.used,
                       xpert.samples,
                       other.xpert.samples,
                       culture.used,
                       culture.samples,
                       other.culture.samples,
                       study.start.year, 
                       study.end.year, 
                       publication.year,
                       study.quality.summary,
                       ends_with("male"), ends_with("sex"),
                       sex.analysis.indicator) %>%
                
    mutate("tbPrev.male" =  case_when(sex.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.male/n.participants.male, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.male/n.participants.male,
                                        sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tbPrev.female" = case_when(sex.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.female/n.participants.female, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female),
           logTbPrev.male = log(tbPrev.male), 
           logTbPrev.female = log(tbPrev.female)) %>%
    #### Create a confidence interval for those without 
    #### these are using the wilson method with continuity correction
    #### cannot seem to get case_when to work so this is going to be verbose bc deadline
    rowwise() %>%
    mutate(calc.ci.lower.male = ifelse(sex.analysis.indicator == "n.bacteriological.tb",  
                                              (prop.test(n.bacteriological.tb.male, n.participants.male))$conf.int[1],
                                       ifelse(sex.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.male, n.participants.male))$conf.int[1],
                                          ifelse(sex.analysis.indicator == "n.culture.positive.tb", 
                                              (prop.test(n.culture.positive.tb.male, n.participants.male))$conf.int[1],
                                                ifelse(sex.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.male*n.participants.male, n.participants.male))$conf.int[1], 
                                              NA)))),
            calc.ci.upper.male = ifelse(sex.analysis.indicator == "n.bacteriological.tb",  
                                              (prop.test(n.bacteriological.tb.male, n.participants.male))$conf.int[2],
                                       ifelse(sex.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.male, n.participants.male))$conf.int[2],
                                          ifelse(sex.analysis.indicator == "n.culture.positive.tb", 
                                              (prop.test(n.culture.positive.tb.male, n.participants.male))$conf.int[2],
                                                ifelse(sex.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.male*n.participants.male, n.participants.male))$conf.int[2], 
                                              NA)))),
           calc.ci.lower.female = ifelse(sex.analysis.indicator == "n.bacteriological.tb",  
                                              (prop.test(n.bacteriological.tb.female, n.participants.female))$conf.int[1],
                                       ifelse(sex.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.female, n.participants.female))$conf.int[1],
                                          ifelse(sex.analysis.indicator == "n.culture.positive.tb", 
                                              (prop.test(n.culture.positive.tb.female, n.participants.female))$conf.int[1],
                                                ifelse(sex.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.female*n.participants.female, n.participants.female))$conf.int[1], 
                                              NA)))),
            calc.ci.upper.female = ifelse(sex.analysis.indicator == "n.bacteriological.tb",  
                                              (prop.test(n.bacteriological.tb.female, n.participants.female))$conf.int[2],
                                       ifelse(sex.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.female, n.participants.female))$conf.int[2],
                                          ifelse(sex.analysis.indicator == "n.culture.positive.tb", 
                                              (prop.test(n.culture.positive.tb.female, n.participants.female))$conf.int[2],
                                                ifelse(sex.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.female*n.participants.female, n.participants.female))$conf.int[2], 
                                              NA))))) %>%
        ungroup() %>%
        mutate("tbPrevSE.male" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb"  ~ 
                                         (adj.prev100k.ci.upper.bacteriological.tb.male -
                                                                 adj.prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         (adj.prev100k.ci.upper.smear.positive.tb.male -
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" ~ 
                                            (prev100k.ci.upper.bacteriological.tb.male -
                                                                 prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            (prev100k.ci.upper.smear.positive.tb.male -
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb")  ~ 
                                            (calc.ci.upper.male - calc.ci.lower.male)/3.92),
           "tbPrevSE.female" =  case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb"  ~ 
                                         (adj.prev100k.ci.upper.bacteriological.tb.female  -
                                                                 adj.prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         (adj.prev100k.ci.upper.smear.positive.tb.female -
                                                                 adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" ~ 
                                            (prev100k.ci.upper.bacteriological.tb.female -
                                                                 prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            (prev100k.ci.upper.smear.positive.tb.female - 
                                                                 prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb")  ~ 
                                                        (calc.ci.upper.female - calc.ci.lower.female)/3.92),
             "logTbPrevSE.male" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb"  ~ 
                                         log(adj.prev100k.ci.upper.bacteriological.tb.male /
                                                                 adj.prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         log(adj.prev100k.ci.upper.smear.positive.tb.male /
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" ~ 
                                            log(prev100k.ci.upper.bacteriological.tb.male /
                                                                 prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            log(prev100k.ci.upper.smear.positive.tb.male /
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb")  ~ 
                                            log(calc.ci.upper.male / calc.ci.lower.male)/3.92),
           "logTbPrevSE.female" =  case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb"  ~ 
                                         log(adj.prev100k.ci.upper.bacteriological.tb.female  /
                                                                 adj.prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         log(adj.prev100k.ci.upper.smear.positive.tb.female /
                                                                 adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" ~ 
                                            log(prev100k.ci.upper.bacteriological.tb.female /
                                                                 prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            log(prev100k.ci.upper.smear.positive.tb.female / 
                                                                 prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb")  ~ 
                                                        log(calc.ci.upper.female / calc.ci.lower.female)/3.92)) %>% 
    
        pivot_longer(cols = c(tbPrev.male, tbPrev.female,
                              logTbPrev.male, logTbPrev.female, 
                              logTbPrevSE.male, logTbPrevSE.female), 
                 names_to = c("Measure","Sex"), 
                 names_pattern = "(.*)\\.(.*)$", 
                 ) %>% 
        pivot_wider(names_from = Measure, values_from = value) %>%
    mutate(natRep = ifelse(is.na(study.geography), 
                                                       NA, 
                                                       ifelse(study.geography == "Nationally representative", 
                                                       "Yes", "No")),
                                       ## preserve the extracted study end year
                                       study.end.year.ext = study.end.year, 
                                       ## update missing study end year to be publication year - mean lag
                                       study.end.year = ifelse(!is.na(study.end.year), 
                                                               study.end.year.ext, 
                                                               publication.year - meanLag),
                                       ## Mean center year for temporal analysis
                                       year_z = study.end.year - round(mean(study.end.year, na.rm=TRUE)),

                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       
                                       symptomScreen = ifelse(is.na(screening.tests), 
                                                                    NA, 
                                                                    ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No")),
                                       symptomsRequired = ifelse(is.na(participants.sputum.collected) | grepl("unknown", 
                                                                                                              participants.sputum.collected,ignore.case=TRUE), 
                                                                       NA, 
                                                                       ifelse(!grepl("x-ray|all participants|other",
                                                                        participants.sputum.collected,ignore.case=TRUE),
                                                                 "Yes", "No")),

                                       XpertCultRequired = ifelse(is.na(bacteriological.tb.definition),
                                                                  NA, ifelse(!grepl("smear|micros|field|NTP", 
                                                                              bacteriological.tb.definition,
                                                                              ignore.case=TRUE), 
                                                                  "Yes", 
                                                                   "No")),
                                       ### Fix those that were missed with the above attempt 
                                       XpertCultRequired = ifelse(covidence.id %in% c(26989,26928,26920,26918,26892,26861,26857), 
                                                                  "Yes", XpertCultRequired), 
                                       XpertCultRequired = ifelse(covidence.id %in% c(28002), 
                                                                  "No", XpertCultRequired),
                                       #### if we used the culture positive count, then xpert culture was requried
                                       XpertCultRequired = ifelse(grepl("culture", sex.analysis.indicator), 
                                                                  "Yes", XpertCultRequired),
                                       #### if we used the smear positive prevs, then xpert culture was not requried
                                       XpertCultRequired = ifelse(grepl("smear", sex.analysis.indicator), 
                                                                  "No", XpertCultRequired),
                                       XpertCultUsed = ifelse((grepl("Yes", xpert.used) & 
                                                                 (grepl("All samples", xpert.samples) |
                                                                  !grepl("mear|NAAT", other.xpert.samples))) |
                                                                  (grepl("Yes", culture.used) & 
                                                                  grepl("All samples|Subset", culture.samples) & 
                                                                  !grepl("HIV", other.culture.samples)), 
                                                                 "Yes", "No"),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))


##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("sex")


bactModelNatDF1 <- cleanSexNatDF %>%  
    select(id, WHO.region, study.country, study.end.year, 
           sex.analysis.indicator,
           Sex, tbPrev, 
           logTbPrev, logTbPrevSE,
           XpertCultRequired, XpertCultUsed,
           symptomsRequired,
           study.rurality, bias, 
           n.participants.male, n.participants.female, 
           bias, publication.year, year_z) 

bactModelNatDF <- cleanSexNatDF %>%  
    select(id, WHO.region, study.country, study.end.year, 
           sex.analysis.indicator,
           Sex, tbPrev,
           logTbPrev, logTbPrevSE,
           XpertCultRequired, XpertCultUsed,
           symptomsRequired,
           study.rurality, bias, 
           n.participants.male, n.participants.female, 
           bias, publication.year, year_z) %>%
    pivot_wider(id_cols = c(id, WHO.region, study.country, study.end.year, 
           sex.analysis.indicator,
           XpertCultRequired, XpertCultUsed,
           symptomsRequired,
           study.rurality, bias, 
           n.participants.male, n.participants.female, 
           bias, publication.year, year_z), 
                names_from = Sex,
                values_from = c(tbPrev, logTbPrev, logTbPrevSE),
                names_sep = "_") %>%
    mutate(logPR = logTbPrev_male - logTbPrev_female,
           unadj_logPR_se = sqrt(logTbPrevSE_male^2 + logTbPrevSE_female^2), 
           log_PR_se = ifelse(grepl("adj", sex.analysis.indicator), 
                                     unadj_logPR_se,
                                     ifelse(grepl("smear", sex.analysis.indicator),
                                         sqrt(`unadj_logPR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logPR_se`^2 + varAdj["Bact Variance Median"]))))

```
### Plot TB prevalence 

```{r, height = 8, width =11}

ggplot() +
    geom_point(data=bactModelNatDF1 %>% 
                   mutate(label = paste(study.country, study.end.year)), 
              aes(x=reorder(label, tbPrev), y=tbPrev*1e5, group=Sex, color=Sex), size=2) +

    theme_minimal(base_size=14) + theme(legend.position = "none",
                            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    facet_wrap(vars(Sex)) + ggtitle("Reported sex-stratified TB prevalence by survey") +
    coord_flip() + 
    scale_color_manual(values = c("darkred", "dodgerblue4"))+
    ylab("Reported bacteriologically confirmed TB prevalence (per 100,000)") + 
    xlab("") 

```

## Format notification data

```{r Notification Data, echo = FALSE,  warning = FALSE, message=FALSE, width=11.5, height=8.5}
##### NOTIFICATION DATA ##############################################
# notif00 <- read.csv(here("data/covariates/TB_notifications_2025-07-18.csv"))
notif00 <- read.csv("~/Downloads/TB_notifications_2026-04-23.csv")
sexPop0 <- read.csv(here("data/covariates/populationData.csv"))

sexPop <- sexPop0 %>% melt(id.vars = c("Country.Name", "Series.Name")) %>%
          mutate(year = as.integer(gsub("X", "", variable))) %>% 
          select(! variable) %>%
          pivot_wider(id_cols = c(Country.Name, year), 
                      names_from = Series.Name, values_from = value) %>% 
          rename(country = Country.Name, 
                 pop_male = male, 
                 pop_female = female)

### Need cambodia 2024 population


sexPop[grep("Viet", sexPop$country), "country"] <- 
    notif00[grep("Viet", notif00$country), "country"] <- "Vietnam"
sexPop[grep("Lao", sexPop$country), "country"] <- 
    notif00[grep("Lao", notif00$country), "country"] <- "Laos"
sexPop[grep("People's", sexPop$country), "country"] <- 
    notif00[grep("People's", notif00$country), "country"] <- "North Korea"
sexPop[grep("Gambia", sexPop$country), "country"] <- 
    notif00[grep("Gambia", notif00$country), "country"] <- "Gambia"
sexPop[grep("Tanzania", sexPop$country), "country"] <- 
    notif00[grep("Tanzania", notif00$country), "country"] <- "Tanzania"

#### Convert to male-to-female ratios 
notif0 <- notif00 %>% 
         filter(year > 1993, 
                country %in% unique(bactModelNatDF$study.country)) %>%
         select(country, g_whoregion, 
                year, 
                newrel_m15plus,
                newrel_f15plus, 
                new_sn_m15plus, new_ep_m15plus, 
                new_sp_m1524, new_sp_m2534, 
                new_sp_m3544, new_sp_m4554,
                new_sp_m5564, new_sp_m65, new_sp_mu,
                new_sn_f15plus, new_ep_f15plus, 
                new_sp_f1524, new_sp_f2534, 
                new_sp_f3544, new_sp_f4554,
                new_sp_f5564, new_sp_f65, new_sp_fu) %>% 
         left_join(sexPop) %>%
         mutate(cases_male = newrel_m15plus,
                allCases_male = newrel_m15plus,
                cases_female = newrel_f15plus,
                allCases_female = newrel_f15plus,
                smrCases_male = new_sp_m1524 + new_sp_m2534 + new_sp_m3544 + new_sp_m4554 + new_sp_m5564 + new_sp_m65, 
                smrCases_female = new_sp_f1524 + new_sp_f2534 + new_sp_f3544 + new_sp_f4554 + new_sp_f5564 + new_sp_f65)

### for some reason if else wouldn't work so let' just loop 
for (i in 1:nrow(notif0)){
    if (is.na(notif0$cases_male[i]) & notif0$year[i] < 2013){
        notif0$cases_female[i] <- sum((notif0 %>% select(contains(c("sn_f", "sp_f"))))[i,], na.rm = T)
        notif0$cases_male[i]   <- sum((notif0 %>% select(contains(c("sn_m", "sp_m"))))[i,], na.rm = T)
        
        notif0$allCases_female[i] <- sum((notif0 %>% select(contains(c("sn_f", 
                                                                        "sp_f", "ep_f"))))[i,], na.rm = T)
        notif0$allCases_male[i]   <- sum((notif0 %>% select(contains(c("sn_m", 
                                                                        "sp_m", "ep_m"))))[i,], na.rm = T)
    }
}

### Need cases bangladesh 2009 and Myanmar 1994


notifStrat <- notif0 %>% 
              pivot_longer(cols=c(pop_male, pop_female, 
                                 cases_male, cases_female,
                                 allCases_male, allCases_female,
                                 smrCases_male, smrCases_female), 
                          names_to = c("Measure", "Sex"), 
                          # names_prefix = "notifRate",
                          names_sep = "_",
                          values_to = "values",
                          values_drop_na = TRUE) %>%
              pivot_wider(names_from = Measure, 
                          values_from = values) %>% 
              mutate(### Poisson uncertainty
                     casesSE = sqrt(cases / pop),
                     ### Notification rate 
                     notifRate = cases/pop, 
                     allnotifRate = allCases / pop, 
                     smrNotifRate = smrCases / pop,
                     ### Notification rate uncertainty
                     notifRateSE = casesSE/pop, 
                     ### Log notification rate 
                     logNotifRate = log(notifRate), 
                     logSmrNotifRate = log(smrNotifRate), 
                     ### Uncertainty in log notification rate
                     logNotifRateSE = log((cases + casesSE) / (cases-casesSE))/pop)  %>%
              select(country, year, Sex, pop, cases, casesSE, allCases, 
                     notifRate, notifRateSE, allnotifRate, 
                     logNotifRate, logNotifRateSE, 
                     smrNotifRate, logSmrNotifRate)


notifRatio <- notif0 %>% 
              mutate(notifRateMale = cases_male/pop_male, 
                     notifRateFemale = cases_female/pop_female, 
                     # notifRateBoth = (cases_female + cases_male) / (pop_female + pop_male), 
                     allNotifRateMale = allCases_male/pop_male, 
                     allNotifRateFemale = allCases_female/pop_female,
                     logNotifRateMale = log(notifRateMale),
                     logNotifRateFemale = log(notifRateFemale),
                     mfRatioNotif = notifRateMale/notifRateFemale,
                     logMfRatioNotif = log(mfRatioNotif)) %>% 
              rename(study.country = country, 
                     study.end.year = year) 

notifPrevRatio <- left_join(bactModelNatDF, notifRatio) %>% 
              mutate(notifRateMale = ifelse(notifRateMale==0, NA, notifRateMale),
                     allNotifRateMale = ifelse(allNotifRateMale==0, NA, allNotifRateMale),
                     notifRateFemale = ifelse(notifRateFemale==0, NA, notifRateFemale),
                     allNotifRateFemale = ifelse(allNotifRateFemale==0, NA, allNotifRateFemale),
                     
                     empPNMale = tbPrev_male / notifRateMale, 
                     empAllPNMale = tbPrev_male / allNotifRateMale, 
                     empNPMale = notifRateMale / tbPrev_male, 
                     empPNFemale = tbPrev_female / notifRateFemale, 
                     empAllPNFemale = tbPrev_female / allNotifRateFemale, 
                     empNPFemale = notifRateFemale / tbPrev_female, 
                     label = paste(study.country, study.end.year))

bactNotifStrat <- bactModelNatDF1  %>%
             select(id,
                    Sex,
                    logTbPrev,
                    logTbPrevSE,
                    study.country, 
                    WHO.region, 
                    study.end.year,
                    # `Adjusted Prevalence`,
                    # sex.analysis.indicator,
                    # XpertCultRequired,
                    year_z) %>%
             rename(country = study.country,
                    year = study.end.year) %>%
             left_join(notifStrat, by=c("country", "year", "Sex")) %>%
             rename(study.country = country,
                    study.end.year = year) %>%
             mutate(notifRate = ifelse(notifRate==0, NA, notifRate))

smrNotifStrat <- bactModelNatDF1  %>%
             select(id,
                    Sex,
                    logTbPrev,
                    logTbPrevSE,
                    study.country, 
                    WHO.region, 
                    study.end.year,
                    year_z) %>%
             rename(country = study.country,
                    year = study.end.year) %>%
             left_join(notifStrat, by=c("country", "year", "Sex")) %>%
             rename(study.country = country,
                    study.end.year = year) %>%
             mutate(notifRate = ifelse(notifRate==0, NA, notifRate))


npRatioDF <- left_join(bactModelNatDF1  %>%
                         select(id, Sex,
                                tbPrev,
                                logTbPrev, 
                                logTbPrevSE,
                                study.country, 
                                WHO.region,
                                study.end.year, 
                                year_z), 
                       notifStrat) 

notifStratNaRm <- bactNotifStrat %>% na.omit()

```


## Model 1: Sex-stratified TB prevalence only

```{r Sex-stratified Prevalence Model, warning=FALSE, message=FALSE, results="hide", fig.height = 8, fig.width =11}
tbPrevMod <- brms::brm(formula = logTbPrev | se(logTbPrevSE, sigma=TRUE) ~ 
                           1 + Sex + (1 + Sex | study.country) + (1 + Sex | id),
                       data = bactModelNatDF1,
                       prior = 
                               # prior(normal(0,10), class = Intercept) +
                               # prior(exponential(2), class = sd) +
                               # prior(normal(0,10), class = b) +
                               # prior(exponential(4), class = sigma),
                       
                               prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(normal(0, 1), class = b) +
                               prior(exponential(2), class = sd) +
                               prior(exponential(4), class = sigma) +
                               prior(lkj(4), class=cor),
                       
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 10000, warmup = 2500, thin=3)

# tbPrevSimpMod <- brms::brm(formula = logTbPrev | se(logTbPrevSE, sigma=TRUE) ~ 
#                            1 + Sex + (1 | study.country) + (1 + Sex | id),
#                        data = bactModelNatDF1,
#                        prior = prior(student_t(7, 0, 1.5), class = Intercept) +
#                                prior(exponential(2), class = sd),
#                             
#                        family = gaussian(),
#                        control = list(adapt_delta = 0.99),
#                        chains = 4,
#                        cores = 4,
#                        iter = 10000, warmup = 2500, thin=3)

### check the fit 
plot(tbPrevMod)

### posterior predictive check 
pp_check(tbPrevMod, ndraws=100)

### coefficients
summary(tbPrevMod)

### create model estimates 
survey_tb_prev_mod_est <-  tbPrevMod %>% 
    add_epred_draws(newdata = bactModelNatDF1 %>%  
                              mutate(logTbPrevSE = 0),
                    re_formula =  ~ 1 + Sex + (1 + Sex | study.country) + (1 + Sex | id), 
                    allow_new_levels=TRUE) %>%
    ungroup() %>% 
    mutate(label = paste(study.country, study.end.year)) %>% 
    dplyr::select(label, study.country, id, Sex, .draw, .epred) %>%
    group_by(Sex, id, study.country) %>%  
    mutate(tbPrev = exp(.epred) * 1e5)

### Create summary statistics 
survey_tb_prev_mod_summary <- 
    survey_tb_prev_mod_est %>%
    group_by(Sex, study.country, id, label) %>%
    mean_qi(tbPrev, na.rm=T) %>% 
    mutate(across(c(tbPrev, .lower, .upper), 
                  round, 2)); survey_tb_prev_mod_summary
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 8, fig.width =11}

ggplot() +
    geom_point(data=survey_tb_prev_mod_summary, 
              aes(x=reorder(label, tbPrev), y=tbPrev, group=Sex, color=Sex), size=2) +
    geom_errorbar(data=survey_tb_prev_mod_summary, 
              aes(x=reorder(label, tbPrev), ymin=.lower, ymax=.upper, group=Sex, color=Sex)) +
    theme_minimal(base_size=14) + theme(legend.position = "none",
                            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    facet_grid() + ggtitle("Sex-stratified TB prevalence by survey") +
    coord_flip() + 
    scale_color_manual(values = c("darkred", "dodgerblue4"))+
    ylab("Bacteriologically confirmed TB prevalence (per 100,000)") + 
    xlab("") + 
    geom_point(data=bactModelNatDF1 %>%
               mutate(label = paste(study.country, study.end.year)),
               aes(x=reorder(label, tbPrev), y=tbPrev*1e5,
                    group=Sex, color=Sex), alpha = .3, size=3) +
    facet_wrap(facets = vars(Sex))

```

### Country level prevalence 

```{r}
cntry_nd <- crossing(study.country = unique(bactModelNatDF1$study.country),
                            Sex = unique(bactModelNatDF1$Sex),
                            logTbPrevSE = 0)


cntry_tb_prev_mod_est <-  tbPrevMod %>% 
                          add_epred_draws(newdata = cntry_nd,
                                          re_formula =  ~ 1 + (1 + Sex | study.country),
                                          allow_new_levels=TRUE) %>%
                            ungroup() %>% 
                            dplyr::select(study.country, Sex, .draw, .epred) %>%
                            group_by(Sex, study.country) %>%  
                            mutate(tbPrev = exp(.epred) * 1e5)

cntry_tb_prev_mod_summary <- 
    cntry_tb_prev_mod_est %>%
    group_by(Sex, study.country) %>%
    mean_qi(tbPrev, na.rm=T) %>% 
    mutate(across(c(tbPrev, .lower, .upper), 
                  round, 2)); cntry_tb_prev_mod_summary
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 8, fig.width =11}
ggplot() +
    geom_point(data=cntry_tb_prev_mod_summary, 
              aes(x=reorder(study.country, tbPrev), y=tbPrev, group=Sex, color=Sex), size=2) +
    geom_errorbar(data=cntry_tb_prev_mod_summary, 
              aes(x=reorder(study.country, tbPrev), ymin=.lower, ymax=.upper, group=Sex, color=Sex)) +
    theme_minimal(base_size=14) + theme(legend.position = "none",
                            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    facet_grid() + ggtitle("Sex-stratified TB prevalence by country") +
    coord_flip() + 
    scale_color_manual(values = c("darkred", "dodgerblue4"))+
    ylab("Bacteriologically confirmed TB prevalence (per 100,000)") + 
    xlab("") + 
    geom_point(data=bactModelNatDF1,
               aes(x=reorder(study.country, tbPrev), y=tbPrev*1e5,
                    group=Sex, color=Sex), alpha = .3, size=3) +
    facet_wrap(facets = vars(Sex), scales = "free_x")
```


### Model 2: Notification as a covariate

```{r Sex-stratified Prevalence with Notifications Model, warning=FALSE, message=FALSE, results="hide", fig.height = 8, fig.width =11}

model2 <- brms::brm(formula = logTbPrev | se(logTbPrevSE, sigma=TRUE) ~ 
                           1 + Sex + (1 + Sex | study.country) + (1 + Sex | id) + notifRate,
                       data = notifStratNaRm %>% mutate("id" = row_number()),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(normal(0, 1), class = b) +
                               prior(exponential(2), class = sd) +
                               prior(exponential(4), class = sigma) +
                               prior(lkj(4), class=cor),
                       
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 10000, warmup = 2500, thin=3)

# tbPrevSimpMod <- brms::brm(formula = logTbPrev | se(logTbPrevSE, sigma=TRUE) ~ 
#                            1 + Sex + (1 | study.country) + (1 + Sex | id),
#                        data = bactModelNatDF1,
#                        prior = prior(student_t(7, 0, 1.5), class = Intercept) +
#                                prior(exponential(2), class = sd),
#                             
#                        family = gaussian(),
#                        control = list(adapt_delta = 0.99),
#                        chains = 4,
#                        cores = 4,
#                        iter = 10000, warmup = 2500, thin=3)

### check the fit 
# plot(model2)

### posterior predictive check 
pp_check(model2, ndraws=100)

### coefficients
summary(model2)

### create model estimates 
survey_mod2_est <-  model2 %>% 
    add_epred_draws(newdata = notifStratNaRm %>%  
                              mutate(logTbPrevSE = 0),
                    re_formula =  ~ 1 + Sex + (1 + Sex | study.country) + (1 + Sex | id) + notifRate, 
                    allow_new_levels=TRUE) %>%
    ungroup() %>% 
    mutate(label = paste(study.country, study.end.year)) %>% 
    dplyr::select(label, study.country, id, Sex, notifRate, .draw, .epred) %>%
    group_by(Sex, id, study.country) %>%  
    mutate(tbPrev = exp(.epred) * 1e5, 
           pnRatio = exp(.epred) / notifRate) %>% 
    select(!notifRate)

### Create summary statistics 
survey_mod2_summary <- 
    survey_mod2_est %>%
    group_by(Sex, study.country, id, label) %>%
    mean_qi(tbPrev, na.rm=T) %>% 
    mutate(across(c(tbPrev, .lower, .upper), 
                  round, 2)); survey_mod2_summary

survey_mod2_pn_summary <- 
    survey_mod2_est %>%
    group_by(Sex, study.country, id, label) %>%
    mean_qi(pnRatio, na.rm=T) %>% 
    mutate(across(c(pnRatio, .lower, .upper), 
                  round, 2)); survey_mod2_pn_summary
```

```{r Survey estimates, warning=FALSE, message=FALSE, results="hide", fig.height = 8, fig.width =11}
tmp <- left_join(survey_mod2_summary, notifStratNaRm %>% mutate(rawTBPrev = exp(logTbPrev) * 1e5) %>% select(!c(study.country)), 
                 by = c("id", "Sex"))

ggplot() +
    geom_point(data=survey_mod2_summary,
              aes(x=reorder(label, tbPrev), y=tbPrev, group=Sex, color=Sex), size=2) +
    geom_errorbar(data=survey_mod2_summary,
              aes(x=reorder(label, tbPrev), ymin=.lower, ymax=.upper, group=Sex, color=Sex)) +
    theme_minimal(base_size=14) + theme(legend.position = "none",
                            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    facet_grid() + ggtitle("Sex-stratified TB prevalence by survey") +
    coord_flip() +
    scale_color_manual(values = c("darkred", "dodgerblue4"))+
    ylab("Bacteriologically confirmed TB prevalence (per 100,000)") +
    xlab("") +
    geom_point(data=tmp, 
               aes(x=reorder(label, tbPrev), y=rawTBPrev,
                    group=Sex, color=Sex), alpha = .3, size=3) +
    facet_wrap(facets = vars(Sex))

```

### Country level prevalence 

```{r}
cntry_mod2_nd <- tmp %>% mutate(logTbPrevSE = 0)


cntry_mod2_est <-  model2 %>%
                          add_epred_draws(newdata = cntry_mod2_nd,
                                          re_formula =  ~ 1 + (1 + Sex | study.country) + notifRate,
                                          allow_new_levels=TRUE) %>%
                            ungroup() %>%
                            dplyr::select(study.country, Sex, notifRate, .draw, .epred) %>%
                            mutate(tbPrev = exp(.epred) * 1e5, 
                                   pnRatio = exp(.epred) / notifRate) %>% 
                            select(!notifRate)

cntry_mod2_summary <-
    cntry_mod2_est %>%
    group_by(Sex, study.country) %>%
    mean_qi(tbPrev, na.rm=T) %>%
    mutate(across(c(tbPrev, .lower, .upper),
                  round, 2)); cntry_mod2_summary

cntry_mod2_pn_summary <-
    cntry_mod2_est %>%
    group_by(Sex, study.country) %>%
    mean_qi(pnRatio, na.rm=T) %>%
    mutate(across(c(pnRatio, .lower, .upper),
                  round, 2)); cntry_mod2_pn_summary
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 8, fig.width =11}
### take order from last plot 
# cntry_mod2_summary$order <- arrange(cntry_tb_prev_mod_summary, tbPrev) %>% select(study.country) %>% unlist() 
# tmp <- left_join(tmp, cntry_mod2_summary %>% select(study.country, order))
# ggplot() +
#     geom_point(data=cntry_mod2_summary,
#               aes(x=reorder(study.country, tbPrev), y=tbPrev, group=Sex, color=Sex), size=2) +
#     geom_errorbar(data=cntry_mod2_summary,
#               aes(x=reorder(study.country, tbPrev), ymin=.lower, ymax=.upper, group=Sex, color=Sex)) +
#     theme_minimal(base_size=14) + theme(legend.position = "none",
#                             axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
#     facet_grid() + ggtitle("Sex-stratified TB prevalence by country") +
#     coord_flip() +
#     scale_color_manual(values = c("darkred", "dodgerblue4"))+
#     ylab("Bacteriologically confirmed TB prevalence (per 100,000)") +
#     xlab("") +
#     geom_point(data=tmp,
#                aes(x=reorder(study.country, rawTBPrev), y=rawTBPrev,
#                     group=Sex, color=Sex), alpha = .5, size=3) +
#     facet_wrap(facets = vars(Sex), scales = "free_x")

ggplot() +
    geom_point(data=cntry_mod2_summary,
              aes(x=reorder(study.country, tbPrev), y=tbPrev, group=Sex, color=Sex), size=2) +
    geom_errorbar(data=cntry_mod2_summary,
              aes(x=reorder(study.country, tbPrev), ymin=.lower, ymax=.upper, group=Sex, color=Sex)) +
    theme_minimal(base_size=14) + theme(legend.position = "none",
                            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    facet_grid() + ggtitle("Sex-stratified TB prevalence by country") +
    coord_flip() +
    scale_color_manual(values = c("darkred", "dodgerblue4"))+
    ylab("Bacteriologically confirmed TB prevalence (per 100,000)") +
    xlab("") +
    geom_point(data=tmp,
               aes(x=reorder(study.country, tbPrev), y=rawTBPrev,
                    group=Sex, color=Sex), alpha = .3, size=3) +
    facet_wrap(facets = vars(Sex), scales = "free_x")
```

### Plot the pn ratio 

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height = 8, fig.width =11}
ggplot() +
    geom_point(data=cntry_mod2_pn_summary,
              aes(x=reorder(study.country, pnRatio), y=pnRatio, group=Sex, color=Sex), size=2) +
    geom_errorbar(data=cntry_mod2_pn_summary,
              aes(x=reorder(study.country, pnRatio), ymin=.lower, ymax=.upper, group=Sex, color=Sex)) +
    theme_minimal(base_size=14) + theme(legend.position = "none",
                            axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    facet_grid() + ggtitle("Sex-stratified TB P:N ratio by country") +
    coord_flip() +
    scale_color_manual(values = c("darkred", "dodgerblue4"))+
    ylab("Bacteriologically confirmed TB prevalence (per 100,000)") +
    xlab("") +
    facet_wrap(facets = vars(Sex), scales = "free_x")
```

### Recreate WHO plot
```{r}
survey_mod2_pn_summary <- 
    survey_mod2_pn_summary %>% 
 select(label, Sex, pnRatio) %>% 
 pivot_wider(id_cols = label, names_from = Sex, values_from = pnRatio) %>% 
 rowwise() %>% 
  mutate(mean = mean(c(female,male) )) %>% 
  arrange(mean) %>% 
  mutate(label=factor(label, label))

ggplot(survey_mod2_pn_summary) + 
      geom_segment( aes(x=label, xend=label, y=female, yend=male), color="grey") +
  geom_point( aes(x=label, y=female), color="#6A5ACD", size=3 ) +
  geom_point( aes(x=label, y=male), color="#FFB302", size=3 ) +
  coord_flip() + theme_minimal() + 
  scale_y_continuous(breaks=seq(0, 26,2)) + 
  ylab("prevalence:notification ratio") + 
    xlab("") + 
    ggtitle("bact conf with pooling")


```

### without pooling
```{r}
lolliNoPoolDF <- 
    notifStrat %>% 
 select(study.country, study.end.year, Sex, logTbPrev, notifRate) %>%
 mutate(label = paste(study.country, study.end.year), pnRatio = exp(logTbPrev)/notifRate) %>% select(label, Sex,  pnRatio) %>% 
 pivot_wider(id_cols = label, names_from = Sex, values_from = pnRatio) %>% 
 rowwise() %>% 
  mutate(mean = mean(c(female,male) )) %>% 
  arrange(mean) %>% 
  mutate(label=factor(label, label)) %>% na.omit()

ggplot(lolliNoPoolDF) + 
      geom_segment( aes(x=label, xend=label, y=female, yend=male), color="grey") +
  geom_point( aes(x=label, y=female), color="#6A5ACD", size=3 ) +
  geom_point( aes(x=label, y=male), color="#FFB302", size=3 ) +
  coord_flip() + theme_minimal() + 
  scale_y_continuous(breaks=seq(0, 26,2)) + 
  ylab("prevalence:notification ratio") + 
    xlab("") + 
    ggtitle("bact conf without pooling")


```

### who uses smear positive so let's try that 

## SMEAR POSITIVE ANALYSIS 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

cleanSexSmrDF0 <- smrPosIndicator("sex") %>% 
                  filter(sex.smr.analysis.indicator !="none",
                  study.geography == "Nationally representative")  

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexSmrDF <- cleanSexSmrDF0 %>% 
                 mutate("id" = row_number(),
                        ### We have a lower ci of zero so replace with super small non-zero number
                        adj.prev100k.ci.lower.smear.positive.tb.female = ifelse(
                                            adj.prev100k.ci.lower.smear.positive.tb.female == 0, 
                                            1e-4, 
                                            adj.prev100k.ci.lower.smear.positive.tb.female
                        )) %>%
                 ### Scale all prevalence estimates back to decimals 
                 mutate_at(vars(matches('prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>% 
                 select(id,
                       covidence.id, 
                       figure.id.yr, 
                       study.geography, 
                       title.extracted,
                       study.country,
                       WHO.region,
                       study.start.year, 
                       study.end.year, 
                       publication.year,
                       study.quality.summary,
                       screening.tests,
                       bacteriological.tb.definition, 
                       xpert.samples, xpert.used, other.xpert.samples,
                       culture.samples, culture.used, other.culture.samples,
                       participants.sputum.collected,
                       ends_with("male"), ends_with("sex"),
                       sex.smr.analysis.indicator) %>%
    
    mutate("tbPrevSmr.male" =  case_when(sex.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.smr.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                        sex.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tbPrevSmr.female" = case_when(sex.smr.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.smr.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female),
           logTbPrevSmr.male = log(tbPrevSmr.male), 
           logTbPrevSmr.female = log(tbPrevSmr.female)) %>%

    #### Create a confidence interval for those without 
    #### these are using the wilson method with continuity correction
    #### cannot seem to get case_when to work so this is going to be verbose bc deadline
    
    rowwise() %>%
    mutate(calc.ci.lower.male = ifelse(sex.smr.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.male, n.participants.male))$conf.int[1],
                                        ifelse(sex.smr.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.male*n.participants.male, n.participants.male))$conf.int[1], 
                                              NA)),
            calc.ci.upper.male = ifelse(sex.smr.analysis.indicator == "n.smear.positive.tb",
                                        (prop.test(n.smear.positive.tb.male, n.participants.male))$conf.int[2],
                                  ifelse(sex.smr.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.male*n.participants.male, n.participants.male))$conf.int[2], 
                                              NA)),
           calc.ci.lower.female = ifelse(sex.smr.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.female, n.participants.female))$conf.int[1],
                                  ifelse(sex.smr.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.female*n.participants.female, n.participants.female))$conf.int[1], 
                                              NA)),
            calc.ci.upper.female = ifelse(sex.smr.analysis.indicator == "n.smear.positive.tb",
                                              (prop.test(n.smear.positive.tb.female, n.participants.female))$conf.int[2],
                                   ifelse(sex.smr.analysis.indicator == "prev100k.smear.positive.tb",
                                              (prop.test(prev100k.smear.positive.tb.female*n.participants.female, n.participants.female))$conf.int[2], 
                                              NA))) %>%
        ungroup() %>%
        mutate("tbPrevSmrSE.male" = case_when(
                                        sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         (adj.prev100k.ci.upper.smear.positive.tb.male -
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            (prev100k.ci.upper.smear.positive.tb.male -
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator %in% c("n.smear.positive.tb", 
                                                                      "prev100k.smear.positive.tb")  ~ 
                                            (calc.ci.upper.male - calc.ci.lower.male)/3.92),
           "tbPrevSmrSE.female" =  case_when(
                                        sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         (adj.prev100k.ci.upper.smear.positive.tb.female -
                                                                 adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,

                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            (prev100k.ci.upper.smear.positive.tb.female - 
                                                                 prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.smr.analysis.indicator %in% c("n.smear.positive.tb", 
                                                                      "prev100k.smear.positive.tb")  ~ 
                                                        (calc.ci.upper.male - calc.ci.lower.male)/3.92),
           "logTbPrevSmrSE.male" = case_when(
                                        sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         log(adj.prev100k.ci.upper.smear.positive.tb.male /
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            log(prev100k.ci.upper.smear.positive.tb.male /
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.smr.analysis.indicator %in% c("n.smear.positive.tb", 
                                                                      "prev100k.smear.positive.tb")  ~ 
                                            log(calc.ci.upper.male / calc.ci.lower.male)/3.92),
           "logTbPrevSmrSE.female" =  case_when(
                                        sex.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb"  ~ 
                                         log(adj.prev100k.ci.upper.smear.positive.tb.female /
                                                                 adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,

                                        sex.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" ~ 
                                            log(prev100k.ci.upper.smear.positive.tb.female / 
                                                                 prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.smr.analysis.indicator %in% c("n.smear.positive.tb", 
                                                                      "prev100k.smear.positive.tb")  ~ 
                                                        log(calc.ci.upper.male / calc.ci.lower.male)/3.92)) %>% 
     pivot_longer(cols = c(tbPrevSmr.male, tbPrevSmr.female,
                              logTbPrevSmr.male, logTbPrevSmr.female, 
                              logTbPrevSmrSE.male, logTbPrevSmrSE.female), 
                 names_to = c("Measure","Sex"), 
                 names_pattern = "(.*)\\.(.*)$", 
                 ) %>% 
        pivot_wider(names_from = Measure, values_from = value) %>% 
                                mutate(natRep = ifelse(is.na(study.geography), 
                                                       NA, 
                                                       ifelse(study.geography == "Nationally representative", 
                                                       "Yes", "No")),
                                       ## preserve the extracted study end year
                                       study.end.year.ext = study.end.year, 
                                       ## update missing study end year to be publication year - mean lag
                                       study.end.year = ifelse(!is.na(study.end.year), 
                                                               study.end.year.ext, 
                                                               publication.year - meanLag),
                                       ## Mean center year for temporal analysis
                                       year_z = study.end.year - round(mean(study.end.year, na.rm=TRUE)),

                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       
                                       symptomScreen = ifelse(is.na(screening.tests), 
                                                                    NA, 
                                                                    ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No")),
                                       symptomsRequired = ifelse(is.na(participants.sputum.collected) | grepl("unknown", 
                                                                                                              participants.sputum.collected,ignore.case=TRUE), 
                                                                       NA, 
                                                                       ifelse(!grepl("x-ray|all participants|other",
                                                                        participants.sputum.collected,ignore.case=TRUE),
                                                                 "Yes", "No")),

                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region)) 

##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdjPR <- varianceAdjust("sex")


smrModelNatDF1 <- cleanSexSmrDF %>%  
    select(id, WHO.region, study.country, study.end.year, 
sex.smr.analysis.indicator,
Sex, tbPrevSmr, 
           logTbPrevSmr, logTbPrevSmrSE,
           symptomsRequired,
           bias, 
           n.participants.male, n.participants.female, 
           bias, publication.year, year_z) 

smrModelNatDF <- cleanSexSmrDF %>%  
    select(id, WHO.region, study.country, study.end.year, 
           sex.smr.analysis.indicator, Sex, tbPrevSmr,
           logTbPrevSmr, logTbPrevSmrSE,
           symptomsRequired,
           bias, 
           n.participants.male, n.participants.female, 
           bias, publication.year, year_z) %>%
    pivot_wider(id_cols = c(id, WHO.region, study.country, study.end.year, 
           symptomsRequired, sex.smr.analysis.indicator,
            bias, 
           n.participants.male, n.participants.female, 
           bias, publication.year, year_z), 
                names_from = Sex,
                values_from = c(tbPrevSmr, logTbPrevSmr, logTbPrevSmrSE),
                names_sep = "_") %>%
    mutate(logPR = logTbPrevSmr_male - logTbPrevSmr_female,
           unadj_logPR_se = sqrt(logTbPrevSmrSE_male^2 + logTbPrevSmrSE_female^2), 
           log_PR_se = ifelse(grepl("adj", sex.smr.analysis.indicator), 
                                     unadj_logPR_se,
                                     ifelse(grepl("smear", sex.smr.analysis.indicator),
                                         sqrt(`unadj_logPR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logPR_se`^2 + varAdj["Bact Variance Median"]))))


#### Add in notifications
smrNotifStrat2 <- left_join(smrModelNatDF1, smrNotifStrat, 
                            by = c("study.country", "study.end.year", "Sex")) %>% 
    mutate(id = )

View(smrNotifStrat2)

```

### Compare smear to total notifications
```{r, warnings=FALSE}

ggplot(data=smrNotifStrat) + 
    geom_point(aes(x=study.country, y=notifRate, shape=Sex),  color = "springgreen3", size=3) + 
    geom_point(aes(x=study.country, y=smrNotifRate, shape=Sex),  color = "orange", size =3) + 
    theme_minimal()  + coord_flip() + theme(legend.position = "bottom")
    # facet_wrap(facets=vars(study.country))
    

```



### Model 3: prevalence with notification as a covariate 

```{r}
model3 <- brms::brm(formula = logTbPrevSmr | se(logTbPrevSmrSE, sigma=TRUE) ~ 
                           1 + Sex + (1 + Sex | study.country) + (1 + Sex | id) + smrNotifRate,
                       data = smrNotifStrat2 %>% filter(!is.na(smrNotifRate)) %>% 
                              mutate("id" = row_number()),
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                               prior(normal(0, 1), class = b) +
                               prior(exponential(2), class = sd) +
                               prior(exponential(4), class = sigma) +
                               prior(lkj(4), class=cor),
                       
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 10000, warmup = 2500, thin=3)


### coefficients
summary(model3)

### create model estimates 
survey_mod3_est <-  model3 %>% 
    add_epred_draws(newdata = smrNotifStrat2 %>% filter(!is.na(smrNotifRate)) %>% 
                              mutate("id" = row_number()) %>%  
                              mutate(logTbPrevSE = 0),
                    re_formula =  ~ 1 + Sex + (1 + Sex | study.country) + (1 + Sex | id) + smrNotifRate, 
                    allow_new_levels=TRUE) %>%
    ungroup() %>% 
    mutate(label = paste(study.country, study.end.year)) %>% 
    dplyr::select(label, study.country, id, Sex, smrNotifRate, .draw, .epred) %>%
    group_by(Sex, id, study.country) %>%  
    mutate(tbPrevSmr = exp(.epred) * 1e5, 
           pnRatio = exp(.epred) / smrNotifRate) %>% 
    select(!smrNotifRate)

### Create summary statistics 
survey_mod3_summary <- 
    survey_mod3_est %>%
    group_by(Sex, study.country, id, label) %>%
    mean_qi(tbPrevSmr, na.rm=T) %>% 
    mutate(across(c(tbPrevSmr, .lower, .upper), 
                  round, 2)); survey_mod3_summary

survey_mod3_pn_summary <- 
    survey_mod3_est %>%
    group_by(Sex, study.country, id, label) %>%
    mean_qi(pnRatio, na.rm=T) %>% 
    mutate(across(c(pnRatio, .lower, .upper), 
                  round, 2)); survey_mod3_pn_summary
```


### Recreate WHO plot
```{r}
survey_mod3_pn_summary <- 
    survey_mod3_pn_summary %>% 
 select(label, Sex, pnRatio) %>% 
 pivot_wider(id_cols = label, names_from = Sex, values_from = pnRatio) %>% 
 rowwise() %>% 
  mutate(mean = mean(c(female,male) )) %>% 
  arrange(mean) %>% 
  mutate(label=factor(label, label))

ggplot(survey_mod3_pn_summary) + 
      geom_segment( aes(x=label, xend=label, y=female, yend=male), color="grey") +
  geom_point( aes(x=label, y=female), color="#6A5ACD", size=3 ) +
  geom_point( aes(x=label, y=male), color="#FFB302", size=3 ) +
  coord_flip() + theme_minimal() + 
  scale_y_continuous(breaks=seq(0, 26,2)) + 
  ylab("smr positive prevalence:notification ratio") + 
    xlab("") + 
    ggtitle("Smr positive with pooling")

```

```{r}
lolliNoPoolDF <- 
    smrNotifStrat2 %>% 
 select(study.country, study.end.year, Sex, logTbPrevSmr, smrNotifRate) %>%
 mutate(label = paste(study.country, study.end.year), pnRatio = exp(logTbPrevSmr)/smrNotifRate) %>% select(label, Sex,  pnRatio) %>% 
 pivot_wider(id_cols = label, names_from = Sex, values_from = pnRatio) %>% 
 rowwise() %>% 
  mutate(mean = mean(c(female,male) )) %>% 
  arrange(mean) %>% 
  mutate(label=factor(label, label)) %>% na.omit()

ggplot(lolliNoPoolDF) + 
      geom_segment( aes(x=label, xend=label, y=female, yend=male), color="grey") +
  geom_point( aes(x=label, y=female), color="#6A5ACD", size=3 ) +
  geom_point( aes(x=label, y=male), color="#FFB302", size=3 ) +
  coord_flip() + theme_minimal() + 
  scale_y_continuous(breaks=seq(0, 26,2)) + 
  ylab("smr positive prevalence:notification ratio") + 
    xlab("") + 
    ggtitle("Smr positive without pooling")
```