---
title: "Sex Stratified TB Prevalence Estimates for LMICs" 
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
classoption: landscape
geometry: margin=.5cm
header-includes:
  - \pagenumbering{gobble}
---

# About this script 

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multilevel model on the studies that report sex stratified prevalence results. This model aims to estimate sex stratified prevalence. 

# Setup

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
# library(posterior)
library(tidybayes)
library(wesanderson)
library(ggplot2)
library(patchwork)

expit <- function (x) {exp(x)/(1 + exp(x))}

myPal <- c("black", "violet")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))

### Set seed 
set.seed(500)

##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################
cleanSexDF0 <- readRDS("data/sex-stratified-clean-data.rds")
### Calculate mean publication lag 
meanLag <- round(mean(cleanSexDF0$publication.year - cleanSexDF0$study.end.year, na.rm=TRUE))
```

# Bacteriologically-confirmed TB prevalence 

## Format data for modeling 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanSexDF <-  cleanSexDF0 %>% 
               filter(sex.analysis.indicator != "none") %>% 
               mutate("id" = row_number()) %>% 
    ### Scale all prevalence estimates back to decimals 
    mutate_at(vars(matches('prev100k.bacteriological.tb.|prev100k.ci.([a-z]{3})er.bacteriological.tb.|prev100k.smear.positive.tb.|prev100k.ci.([a-z]{3})er.smear.positive.tb.')),  ~ (. / 1e5)) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           study.rurality, 
           WHO.region,
           screening.tests,
           participants.sputum.collected,
           bacteriological.tb.definition,
           xpert.used,
           xpert.samples,
           other.xpert.samples,
           culture.used,
           culture.samples,
           other.culture.samples,
           study.start.year, 
           study.end.year, 
           publication.year,
           study.quality.summary,
           ends_with("male"), ends_with("sex"),
           sex.analysis.indicator) %>%
    
    mutate("tempPrevMale" =  case_when(sex.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.male,
                                       sex.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.male, 
                                       sex.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.male, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.male/n.participants.male, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.male/n.participants.male,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.male/n.participants.male,
                                        sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.male),
           "tempPrevFemale" = case_when(sex.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.female,
                                        sex.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.female, 
                                        sex.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.female, 
                                       sex.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.female/n.participants.female, 
                                       sex.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.female/n.participants.female,
                                       sex.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.female)) %>%
    pivot_longer(cols = c(tempPrevFemale,tempPrevMale), 
                 names_to = "Sex", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.bacteriological.tb.male-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.bacteriological.tb.female-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (adj.prev100k.ci.upper.smear.positive.tb.male-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (adj.prev100k.ci.upper.smear.positive.tb.female-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.bacteriological.tb.male-
                                                                 prev100k.ci.lower.bacteriological.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.bacteriological.tb.female-
                                                                   prev100k.ci.lower.bacteriological.tb.female)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Male" ~ (prev100k.ci.upper.smear.positive.tb.male-
                                                                 prev100k.ci.lower.smear.positive.tb.male)/3.92,
                                        sex.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Sex == "Female" ~ (prev100k.ci.upper.smear.positive.tb.female-
                                                                   prev100k.ci.lower.smear.positive.tb.female)/3.92,
                                        sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.male),
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.female)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           
           "LogOddsStandardError" = case_when(sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.male /
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.male)) -
                                                                   log(adj.prev100k.ci.lower.bacteriological.tb.male/
                                                                           (1-adj.prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.female/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Male" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.male/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.male)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.male /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                              sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Sex == "Female" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.female/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.female)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.female /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.female)))/3.92,
                                              sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Male" ~ (log(prev100k.ci.upper.bacteriological.tb.male/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.male)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.male/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.male))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Sex == "Female" ~ (log(prev100k.ci.upper.bacteriological.tb.female/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.female)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.female /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.female))) /3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Male" ~ (log(prev100k.ci.upper.smear.positive.tb.male/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.male)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.male /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.male)))/3.92,
                                                 sex.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Sex == "Female" ~ (log(prev100k.ci.upper.smear.positive.tb.female/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.female)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.female /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.female)))/3.92, 
                                            sex.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Male" ~ 
                                            log(1/(n.participants.male*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        sex.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Sex == "Female" ~ 
                                            log(1/(n.participants.female*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(is.na(study.geography), 
                                                       NA, 
                                                       ifelse(study.geography == "Nationally representative", 
                                                       "Yes", "No")),
                                       ## preserve the extracted study end year
                                       study.end.year.ext = study.end.year, 
                                       ## update missing study end year to be publication year - mean lag
                                       study.end.year = ifelse(!is.na(study.end.year), 
                                                               study.end.year.ext, 
                                                               publication.year - meanLag),
                                       ## Mean center year for temporal analysis
                                       year_z = study.end.year - round(mean(study.end.year, na.rm=TRUE)),

                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       
                                       symptomScreen = ifelse(is.na(screening.tests), 
                                                                    NA, 
                                                                    ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No")),
                                       symptomsRequired = ifelse(is.na(participants.sputum.collected) | grepl("unknown", 
                                                                                                              participants.sputum.collected,ignore.case=TRUE), 
                                                                       NA, 
                                                                       ifelse(!grepl("x-ray|all participants|other",
                                                                        participants.sputum.collected,ignore.case=TRUE),
                                                                 "Yes", "No")),

                                       XpertCultRequired = ifelse(is.na(bacteriological.tb.definition),
                                                                  NA, ifelse(!grepl("smear|micros|field|NTP", 
                                                                              bacteriological.tb.definition,
                                                                              ignore.case=TRUE), 
                                                                  "Yes", 
                                                                   "No")),
                                       ### Fix those that were missed with the above attempt 
                                       XpertCultRequired = ifelse(covidence.id %in% c(26989,26928,26920,26918,26892,26861,26857), 
                                                                  "Yes", XpertCultRequired), 
                                       XpertCultRequired = ifelse(covidence.id %in% c(28002), 
                                                                  "No", XpertCultRequired),
                                       #### if we used the culture positive count, then xpert culture was requried
                                       XpertCultRequired = ifelse(grepl("culture", sex.analysis.indicator), 
                                                                  "Yes", XpertCultRequired),
                                       #### if we used the smear positive prevs, then xpert culture was not requried
                                       XpertCultRequired = ifelse(grepl("smear", sex.analysis.indicator), 
                                                                  "No", XpertCultRequired),
                                       XpertCultUsed = ifelse((grepl("Yes", xpert.used) & 
                                                                 (grepl("All samples", xpert.samples) |
                                                                  !grepl("mear|NAAT", other.xpert.samples))) |
                                                                  (grepl("Yes", culture.used) & 
                                                                  grepl("All samples|Subset", culture.samples) & 
                                                                  !grepl("HIV", other.culture.samples)), 
                                                                 "Yes", "No"),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))

##############################################################################|
#### Dataset for modeling #####################################################
##############################################################################|
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("sex")

bactModelDF <- cleanSexDF %>%
    select(id, covidence.id, title.extracted, WHO.region, study.country, study.rurality,
           Sex, year_z, 
           natRep, bias, symptomScreen, symptomsRequired, 
           XpertCultRequired,  XpertCultUsed,
           LogOdds, LogOddsStandardError, sex.analysis.indicator, 
           n.participants.male, n.participants.female, 
           publication.year, study.end.year, study.end.year.ext) %>%
    pivot_wider(names_from = Sex,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(logOR = LogOdds_Male - LogOdds_Female,
           unadj_logOR_se = sqrt(LogOddsStandardError_Male^2 + LogOddsStandardError_Female^2), 
           logOR_se = ifelse(grepl("adj", sex.analysis.indicator), 
                                     unadj_logOR_se,
                                     ifelse(grepl("smear", sex.analysis.indicator),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Bact Variance Median"]))))

```


## Model 1: TB prevalence by sex

Model specification

```{r}

tbPrevMod <- brms::brm(formula = `LogOdds`|se(`LogOddsStandardError`, 
                                           sigma=TRUE) ~ 1 + (1|study.country) + (1|id) + Sex,
                    data = cleanSexDF,
                    prior = prior(normal(0,10), class=Intercept) +
                            prior(normal(0, 10), class = b) +
                            prior(exponential(1), class=sd), 
                    family = "gaussian",
                    control = list(adapt_delta = 0.99),
                    cores = 4,
                    chains = 4,
                    iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(tbPrevMod)
plot(tbPrevMod)
```

Prediction

```{r}
tbPrevMod_est <- tbPrevMod %>% 
                 as_draws_df() %>%
                 select(b_Intercept, b_SexMale) %>%
                 transmute(Female = expit(b_Intercept)*1e5, 
                           Male   = expit(b_Intercept + b_SexMale)*1e5) 
```

Summary values 
```{r}
### Female
tbPrevMod_est %>% 
    select(Female) %>%
    mean_qi() %>% 
    mutate("Mean prevalence (per 100K)" = round(Female,2), 
           "95% credible interval" = paste0(round(.lower,2), "-", round(.upper,2))) %>%
    select(`Mean prevalence (per 100K)`, `95% credible interval`)

### Male
tbPrevMod_est %>% 
    select(Male) %>%
    mean_qi() %>% 
    mutate("Mean prevalence (per 100K)" = round(Male,2), 
           "95% credible interval" = paste0(round(.lower,2), "-", round(.upper,2))) %>%
    select(`Mean prevalence (per 100K)`, `95% credible interval`)
```

Plot

```{r}
ggplot(data = tbPrevMod_est %>% pivot_longer(c("Female", "Male"))) +
    stat_halfeye(aes(x=value, fill=name), normalize="panels", alpha=0.5, size =10) +
    # facet_wrap(facets=vars(name)) + 
    labs(x="", y="") +
    theme_minimal() + theme(legend.position = "bottom") + 
    ggtitle("Estimated overall TB prevalence by sex from 102 prevalence surveys")
```

## Model 2: TB prevalence by sex over time

Model specification

```{r}
tbPrevTimeMod <- brms::brm(formula = `LogOdds`|se(`LogOddsStandardError`, sigma=TRUE) ~
                           1 + year_z*Sex + (1 + year_z*Sex | study.country) + (1 | id),
                           data = cleanSexDF,
                           prior = prior(normal(0,10), class=Intercept) +
                                   prior(normal(0, 10), class = b) +
                                   prior(exponential(1), class=sd), 
                           family = "gaussian",
                           control = list(adapt_delta = 0.99),
                           cores = 4,
                           chains = 4,
                           iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE, fig.height = 20, fig.width = 28}
summary(tbPrevTimeMod)
plot(tbPrevTimeMod)
```

Prediction

```{r}
ndPrevTime <- cleanSexDF %>%
              select(study.country) %>%
              crossing(year_z = seq(min(cleanSexDF$year_z, na.rm = TRUE), 
                                    max(cleanSexDF$year_z, na.rm = TRUE), 1), Sex = c("Male", "Female")) %>%
              mutate(LogOddsStandardError = 1)

tbPrevTimeModEst <- add_epred_draws(object = tbPrevTimeMod,
                                    newdata=na.omit(ndPrevTime),
                                    re_formula = ~ 1 + year_z*Sex,
                                    allow_new_levels = TRUE) %>%
                                    mutate(mfRatio = exp(.epred),
                                    type="Posterior mean and 95% credible interval")
```

Summary values 
```{r}
### Female
tbPrevTimeMod_est %>% 
    select(Female) %>%
    group_by(year_z) %>% 
    mean_qi() %>% 
    mutate(Year = year_z + round(mean(cleanSexDF$study.end.year, na.rm=TRUE)))
    # mutate("Mean prevalence (per 100K)" = round(Female,2), 
    #        "95% credible interval" = paste0(round(.lower,2), "-", round(.upper,2))) %>%
    # select(`Mean prevalence (per 100K)`, `95% credible interval`)

### Male
tbPrevTimeMod_est %>% 
    select(Male) %>%
    mean_qi() %>% 
    mutate("Mean prevalence (per 100K)" = round(Male,2), 
           "95% credible interval" = paste0(round(.lower,2), "-", round(.upper,2))) %>%
    select(`Mean prevalence (per 100K)`, `95% credible interval`)
```
