---
title: "Urban/rural meta-analysis"
output: html_notebook
---

### Description

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multivariate multilevel model on the studies that report urban-rural stratified prevalence results. This model aims to estimate urban-rural stratified prevalence adjusted for geographic factors. 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
library(posterior)
library(tidybayes)
library(ggplot2)
```

### Functions and options

```{r}
expit <- function (x) {exp(x)/(1 + exp(x))}

myPal <- c("black", "mediumseagreen")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))
source(here("code/varianceComponentModel.R"))
```

### Data import and prep

Load data and filter to studies reporting urban-rural stratification

```{r}
cleanRuralityDF0 <- bactPostIndicator("rurality") %>% filter(rurality.analysis.indicator !="none")

```

### Descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(cleanRuralityDF0)[1], "unique surveys with urban.rural-stratified bacteriological positive TB"))

# sort(unique(cleanRuralityDF0$title.extracted))
print(paste(length(unique(cleanRuralityDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanRuralityDF0$n.participants.rural)), "studies report participant numbers totaling", 
            sum(cleanRuralityDF0$n.participants.rural, cleanRuralityDF0$n.participants.urban, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanRuralityDF0$n.participants.urban, na.rm=TRUE)/sum(cleanRuralityDF0$n.participants.rural, cleanRuralityDF0$n.participants.urban, na.rm = TRUE)*100, 2), "% of participants were male" ))
```

### Risk of bias statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE}

table(cleanRuralityDF0$study.quality.summary)
```


### Prepare dataset for bacteriologically-positive modelling

```{r, echo = FALSE,  warning = FALSE, message=FALSE}
##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanRuralityDF <-  cleanRuralityDF0 %>% 
               filter(rurality.analysis.indicator != "none") %>% 
               mutate("id" = row_number(),
               adj.prev100k.bacteriological.tb.urban = adj.prev100k.bacteriological.tb.urban/1e5,
               adj.prev100k.ci.upper.bacteriological.tb.urban = adj.prev100k.ci.upper.bacteriological.tb.urban/1e5,
               adj.prev100k.ci.lower.bacteriological.tb.urban = adj.prev100k.ci.lower.bacteriological.tb.urban/1e5,
               
               adj.prev100k.bacteriological.tb.rural = adj.prev100k.bacteriological.tb.rural/1e5,
               adj.prev100k.ci.upper.bacteriological.tb.rural = adj.prev100k.ci.upper.bacteriological.tb.rural/1e5,
               adj.prev100k.ci.lower.bacteriological.tb.rural = adj.prev100k.ci.lower.bacteriological.tb.rural/1e5,
               
               adj.prev100k.smear.positive.tb.urban = adj.prev100k.smear.positive.tb.urban/1e5,
               adj.prev100k.ci.upper.smear.positive.tb.urban = adj.prev100k.ci.upper.smear.positive.tb.urban/1e5,
               adj.prev100k.ci.lower.smear.positive.tb.urban = adj.prev100k.ci.lower.smear.positive.tb.urban/1e5,
               
               adj.prev100k.smear.positive.tb.rural = adj.prev100k.smear.positive.tb.rural/1e5,
               adj.prev100k.ci.upper.smear.positive.tb.rural = adj.prev100k.ci.upper.smear.positive.tb.rural/1e5,
               adj.prev100k.ci.lower.smear.positive.tb.rural = adj.prev100k.ci.lower.smear.positive.tb.rural/1e5,
               
               prev100k.bacteriological.tb.urban = prev100k.bacteriological.tb.urban/1e5,
               prev100k.ci.upper.bacteriological.tb.urban = prev100k.ci.upper.bacteriological.tb.urban/1e5,
               prev100k.ci.lower.bacteriological.tb.urban = prev100k.ci.lower.bacteriological.tb.urban/1e5,
               
               prev100k.bacteriological.tb.rural = prev100k.bacteriological.tb.rural/1e5,
               prev100k.ci.upper.bacteriological.tb.rural = prev100k.ci.upper.bacteriological.tb.rural/1e5,
               prev100k.ci.lower.bacteriological.tb.rural = prev100k.ci.lower.bacteriological.tb.rural/1e5,
               
               prev100k.smear.positive.tb.urban = prev100k.smear.positive.tb.urban/1e5,
               prev100k.ci.upper.smear.positive.tb.urban = prev100k.ci.upper.smear.positive.tb.urban/1e5,
               prev100k.ci.lower.smear.positive.tb.urban = prev100k.ci.lower.smear.positive.tb.urban/1e5,
               
               prev100k.smear.positive.tb.rural = prev100k.smear.positive.tb.rural/1e5,
               prev100k.ci.upper.smear.positive.tb.rural = prev100k.ci.upper.smear.positive.tb.rural/1e5,
               prev100k.ci.lower.smear.positive.tb.rural = prev100k.ci.lower.smear.positive.tb.rural/1e5) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           WHO.region,
           screening.tests,
           study.start.year, 
           study.end.year, 
           study.quality.summary,
           ends_with("urban"), ends_with("rural"),
           rurality.analysis.indicator) %>%
    
    mutate("tempPrevUrban" =  case_when(rurality.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.urban, 
                                       rurality.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.urban,
                                       rurality.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.urban, 
                                       rurality.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.urban, 
                                       rurality.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.urban/n.participants.urban, 
                                       rurality.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.urban/n.participants.urban,
                                       rurality.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.urban/n.participants.urban,
                                        rurality.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.urban),
           "tempPrevRural" = case_when(rurality.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.rural, 
                                        rurality.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.rural,
                                        rurality.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.rural, 
                                        rurality.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.rural, 
                                       rurality.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.rural/n.participants.rural, 
                                       rurality.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.rural/n.participants.rural,
                                       rurality.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.rural/n.participants.rural,
                                       rurality.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.rural)) %>%
    pivot_longer(cols = c(tempPrevRural,tempPrevUrban), 
                 names_to = "Rurality", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(rurality.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Rurality == "Urban" ~ (adj.prev100k.ci.upper.bacteriological.tb.urban-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.urban)/3.92,
                                        rurality.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Rurality == "Rural" ~ (adj.prev100k.ci.upper.bacteriological.tb.rural-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.rural)/3.92,
                                        rurality.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Urban" ~ (adj.prev100k.ci.upper.smear.positive.tb.urban-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.urban)/3.92,
                                        rurality.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Rural" ~ (adj.prev100k.ci.upper.smear.positive.tb.rural-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.rural)/3.92,
                                        rurality.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Rurality == "Urban" ~ (prev100k.ci.upper.bacteriological.tb.urban-
                                                                 prev100k.ci.lower.bacteriological.tb.urban)/3.92,
                                        rurality.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Rurality == "Rural" ~ (prev100k.ci.upper.bacteriological.tb.rural-
                                                                   prev100k.ci.lower.bacteriological.tb.rural)/3.92,
                                        rurality.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Urban" ~ (prev100k.ci.upper.smear.positive.tb.urban-
                                                                 prev100k.ci.lower.smear.positive.tb.urban)/3.92,
                                        rurality.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Rural" ~ (prev100k.ci.upper.smear.positive.tb.rural-
                                                                   prev100k.ci.lower.smear.positive.tb.rural)/3.92,
                                        rurality.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Urban" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.urban),
                                        rurality.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Rural" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.rural)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(rurality.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Rurality == "Urban" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.urban/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.urban)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.urban/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.urban))) /3.92,
                                              rurality.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Rurality == "Rural" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.rural/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.rural)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.rural /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.rural))) /3.92,
                                              rurality.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Rurality == "Urban" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.urban/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.urban)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.urban /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.urban)))/3.92,
                                              rurality.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Rurality == "Rural" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.rural/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.rural)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.rural /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.rural)))/3.92,
                                              rurality.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Rurality == "Urban" ~ (log(prev100k.ci.upper.bacteriological.tb.urban/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.urban)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.urban/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.urban))) /3.92,
                                                 rurality.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Rurality == "Rural" ~ (log(prev100k.ci.upper.bacteriological.tb.rural/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.rural)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.rural /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.rural))) /3.92,
                                                 rurality.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Rurality == "Urban" ~ (log(prev100k.ci.upper.smear.positive.tb.urban/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.urban)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.urban /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.urban)))/3.92,
                                                 rurality.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Rurality == "Rural" ~ (log(prev100k.ci.upper.smear.positive.tb.rural/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.rural)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.rural /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.rural)))/3.92, 
                                            rurality.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Urban" ~ 
                                            log(1/(n.participants.urban*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        rurality.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Rural" ~ 
                                            log(1/(n.participants.rural*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       year_z = study.start.year - round(mean(study.start.year, na.rm=TRUE)),
                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       symptomScreen = ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No"),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))
```

```{r}
#### Calculate the variance adjustment for non-adjusted studies
varAdj <- varianceAdjust("rurality")
```

```{r}
#one study has missing study.end.date
#reading the manuscript, not clearly stated, but most likely date of prevalence survey is 2010

cleanRuralityDF <- cleanRuralityDF %>%
    mutate(year_z = case_when(
        is.na(year_z) ~ 2010 - round(mean(cleanRuralityDF$study.start.year, na.rm=TRUE)),
              TRUE ~ year_z)) %>%
    mutate(study.end.year = case_when(
        is.na(study.end.year) ~ 2010,
              TRUE ~ study.end.year))


```

```{r}
cleanRuralityDF <- cleanRuralityDF %>%
    select(id, WHO.region, study.country, Rurality, study.end.year, year_z, natRep, bias, symptomScreen, LogOdds, LogOddsStandardError, rurality.analysis.indicator) %>%
    pivot_wider(names_from = Rurality,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_Urban - LogOdds_Rural,
        unadj_logOR_se = sqrt(LogOddsStandardError_Urban^2 + LogOddsStandardError_Rural^2), 
        logOR_se = ifelse(grepl("adj", rurality.analysis.indicator), 
                                     unadj_logOR_se,
                                     ifelse(grepl("smear", rurality.analysis.indicator),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Smear Variance Median"]),
                                         sqrt(`unadj_logOR_se`^2 + varAdj["Bact Variance Median"]))))
```

```{r}
###### Incidence rates and models with incidence rates
###### Get the incidence rates
incidence <- read.csv(here("data/TB_burden_countries_2025-09-05.csv"))
# 1. Standardize country names in cleanRuralityDF
incidence <- incidence %>%
  mutate(country = case_when(
    country == "Viet Nam" ~ "Vietnam",
    country == "Lao People's Democratic Republic" ~ "Laos",
    country == "Democratic People's Republic of Korea" ~ "North Korea",
    country == "United Republic of Tanzania" ~ "Tanzania",
    TRUE ~ country
  ))


# 2. Perform the join
cleanRuralityDF <- cleanRuralityDF %>%
  left_join(
    incidence %>% 
      select(country, year, incidence = e_inc_100k),
    by = c("study.country" = "country", "study.end.year" = "year")
  )
```


Import the urbanisation data, and check
https://ourworldindata.org/urbanization


```{r}

urbanization <- read.csv("data/share-of-population-urban.csv")

countries_in_review <- cleanRuralityDF %>%
    select(study.country) %>%
    distinct() %>%
    pull()

urbanizationClean <- urbanization %>%
    filter(Entity %in% countries_in_review) %>%
    janitor::clean_names() %>%
    mutate(pct_urban = urban_population_of_total_population/100) %>%
    filter(year >= 2000)

urbanizationClean %>%
    ggplot() +
    geom_line(aes(x = year, y=pct_urban, colour = entity)) +
    scale_y_continuous(labels = scales::percent) +
    facet_wrap(.~entity)


urban_change <- urbanizationClean %>%
  filter(year %in% c(2000, 2023)) %>%
  select(entity, year, pct_urban) %>%
  pivot_wider(names_from = year, values_from = pct_urban,
              names_prefix = "yr_") %>%
  mutate(
    abs_change = yr_2023 - yr_2000,
    pct_change = (yr_2023 - yr_2000) / yr_2000 * 100
  )

urban_change %>% arrange(desc(abs_change)) %>% print(n=26)



```

Now get the percentage of population that was urban in the year of the survey

```{r}

urbanizationClean <- urbanizationClean %>%
    mutate(year_z = year - round(mean(cleanRuralityDF0$study.start.year, na.rm=TRUE))) %>%
    select(study.country = entity,
           year_z,
           pct_urban)
    
bactModelDF <- left_join(cleanRuralityDF, urbanizationClean)


```


We will drop the Eastern Med studies for the meta-analysis

```{r}

bactModelDF_noEM <- bactModelDF %>%
    filter(WHO.region !="Eastern Mediterranean region") 


```




#### Bacteriologically-positive: Model 1: Main effect model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|study.country) + (1|id),
                       data = bactModelDF_noEM,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       backend = "cmdstanr",
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 


```{r}
summary(mainEffectMod)
plot(mainEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
nd <- bactModelDF_noEM %>%
      select(study.country, LogOdds_Rural) %>%
      mutate(logOR_se=1)

mainEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=nd,
                                 re_formula = ~ 1, 
                                 allow_new_levels = TRUE) %>% 
                    mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectModEst %>% ungroup %>% mean_qi(urRatio)
```

Survey-level effects

```{r}

surveyEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=bactModelDF_noEM,
                                 re_formula = ~ 1  + (1|study.country) + (1|id), 
                                 allow_new_levels = TRUE) %>% 
                    mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 

surveyEffectModEst %>%
    group_by(id) %>%
    mean_qi(urRatio)

```



#### Bacteriologically-positive: Model 2: Effect by WHO region model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regionEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = bactModelDF_noEM,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       backend = "cmdstanr",
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(regionEffectMod)
plot(regionEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndRegion <- bactModelDF_noEM %>%
      select(WHO.region, LogOdds_Rural) %>%
      mutate(logOR_se=1)

regionEffectModEst <- add_epred_draws(object = regionEffectMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region),
                                 allow_new_levels = TRUE) %>% 
                       mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectModEst %>% group_by(WHO.region) %>% mean_qi(urRatio)
```


```{r}
### Get the empirical values 
empirical_prev <- bactModelDF_noEM %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        bias = bias,
        symptomScreen = symptomScreen, 
        p_rural_empirical = plogis(LogOdds_Rural),
        p_urban_empirical = plogis(LogOdds_Urban),
        or_empirical = p_urban_empirical / p_rural_empirical,
        or_empirical = (p_urban_empirical / (1 - p_urban_empirical)) / 
            (p_rural_empirical / (1 - p_rural_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific TB prevalence")

    ggplot() +
    geom_vline(aes(xintercept = 1), linetype=2) +
    geom_point(data = empirical_prev, aes(y=WHO.region, x=or_empirical, color = type2),alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectModEst, aes(x=urRatio, y=WHO.region, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Urban-to-rural prevalence ratio",
         y = "WHO region") +
    theme_bw() + theme(legend.title = element_blank(), legend.position = "bottom", 
          panel.grid.minor=element_blank())
    
```


### Figure 2: Forest plot

Bring estimates from regional model together to make a forest plot for the manuscript.

```{r}
# --- Combine modelled data ---
forestDF <- bind_rows(survey_sum, region_sum, global_sum) %>%
  mutate(
    display_label_raw = ifelse(is.na(figure.id.yr), grouping, figure.id.yr),
    facet_group = case_when(
      grouping == "survey"   ~ WHO.region,
      grouping == "regional" ~ figure.id.yr,
      TRUE                   ~ "Overall"
    )
  )

# ---- Ensure display_label text (bolding for region/overall) ----
forestDF <- forestDF %>%
  mutate(display_label = case_when(
    grouping == "Overall"  ~ "<b>Overall</b>",
    grouping == "regional" ~ paste0("<b>", display_label_raw, "</b>"),
    TRUE                   ~ display_label_raw
  ))

# --- Set facet order manually for plotting (Overall last at bottom) ---
group_order <- c(
  "Overall",
  "Western Pacific region",
  "South-East Asia region",
  "Eastern Mediterranean region",
  "Africa region"
)

forestDF <- forestDF %>%
  mutate(facet_group = factor(facet_group, levels = rev(group_order)))

# --- Build per-facet ordered factor for plotting ---
# 1) ordered survey-level labels per facet (excluding Overall)
survey_order <- forestDF %>%
  filter(grouping == "survey", facet_group != "Overall") %>%
  arrange(facet_group, desc(urRatio)) %>%
  mutate(display_label_facet = paste(facet_group, display_label, sep = "___")) %>%
  pull(display_label_facet)

# 2) append regional summaries per facet (excluding Overall)
regional_order <- unlist(lapply(setdiff(group_order, "Overall"), function(f) {
  survey_lvls <- survey_order[startsWith(survey_order, paste0(f, "___"))]
  summary_lvls <- forestDF %>%
    filter(grouping != "survey", as.character(facet_group) == f) %>%
    mutate(display_label_facet = paste(facet_group, display_label, sep = "___")) %>%
    pull(display_label_facet)
  c(survey_lvls, summary_lvls)
}))

# 3) append the Overall summary last
overall_label <- forestDF %>%
  filter(facet_group == "Overall") %>%
  mutate(display_label_facet = paste(facet_group, display_label, sep = "___")) %>%
  pull(display_label_facet)

final_levels <- c(overall_label, regional_order)

# 4) apply factor levels to forestDF and empirical points
forestDF <- forestDF %>%
  mutate(
    display_label_facet = paste(facet_group, display_label, sep = "___"),
    display_label_facet = factor(display_label_facet, levels = rev(final_levels))  # reverse for coord_flip
  )

empirical_prev_labels <- empirical_prev_labels %>%
  mutate(
    display_label_facet = paste(facet_group, display_label, sep = "___"),
    display_label_facet = factor(display_label_facet, levels = rev(final_levels))
  )

# --- Plot ---
ggplot(forestDF) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "darkred") +
  geom_pointinterval(
    aes(x = display_label_facet, y = urRatio, ymin = .lower, ymax = .upper),
    size = 0.25
  ) +
  geom_point(
    data = empirical_prev_labels,
    aes(x = display_label_facet, y = or_empirical),
    colour = "mediumseagreen", shape = 1
  ) +
  coord_flip() +
  facet_grid(facet_group ~ ., scales = "free", space = "free") +
  scale_x_discrete(labels = function(x) gsub("^.*___", "", x)) +  # drop facet prefix
  scale_y_log10(
    breaks = c(0.25, 0.5, 1, 2, 4),
    labels = c("0.25", "0.5", "1", "2", "4")
  ) +
  labs(
    x = NULL,
    y = "Urban-to-rural bacteriologically-confirmed TB prevalence ratio\n(Posterior mean, 95% credible interval)"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    strip.text.y = element_blank(),
    strip.background = element_blank(),
    panel.spacing.y = unit(1, "lines"),
    axis.text.y = ggtext::element_markdown(hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.background = element_rect(fill = "white")
  )

# Save
ggsave("output/urFigure2.png", height = 10, width=8)

```


#### Bacteriologically-positive: Model 3: Effect over time 


Note we include random intercepts for study country and survey id, and fixed effects term for centered year, as well as allowing the slope of centered year to vary by survey

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
timeEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z + (1 | study.country) + (1 + year_z | id),
                       data = bactModelDF_noEM,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       backend = "cmdstanr",
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r}
summary(timeEffectMod)
plot(timeEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndTime <- bactModelDF %>%
    select(logOR_se) %>%
    crossing(year_z = seq(min(cleanRuralityDF$year_z, na.rm = TRUE), max(cleanRuralityDF$year_z, na.rm = TRUE), 1))

timeEffectModEst <- add_epred_draws(object = timeEffectMod,
                                 newdata=ndTime,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```


Main estimate

```{r}
timeEffectModEst %>% 
    group_by(year_z) %>% 
    mean_qi(urRatio) %>% 
    mutate(Year = year_z + 2011) %>%
    select(Year, urRatio, .lower, .upper)

```


```{r}
    timeEffectModEst %>%
    ggplot(aes(x = year_z + 2011,
               y = urRatio)) +
    stat_lineribbon(.width = 0.95, alpha = 0.5, fill="cadetblue") +
    geom_point(data = empirical_prev, 
               aes(x=year_z + 2011, y=or_empirical, size = 1/logOR_se), 
               colour = "black", shape=1) +
    geom_hline(yintercept = 1, linetype = 2) +
    scale_y_log10(
    breaks = c(0.25, 0.5, 1, 2, 4),
    labels = c("0.25", "0.5", "1", "2", "4")
  ) +
    scale_x_continuous(
    limits = c(2000, 2020),
    breaks = c(2000, 2005, 2010, 2015, 2020),
    labels = c("2000", "2005", "2010", "2015", "2020")
  ) +
    scale_size_continuous(
      name = "Inverse standard error (log-odds)",
      range = c(1, 6),
      guide = "legend"
  ) +
    labs(y = "Urban-to-rural prevalence ratio",
         x = "Year") +
    theme_bw()  + theme(legend.position = "bottom")

ggsave("output/urFigure3.png", height = 5, width = 7)
```

Annual percentage change

```{r}

timeEffectMod_posterior <- as_draws_df(timeEffectMod)

timeEffectMod_beta_year_draws <- timeEffectMod_posterior$b_year_z

# Annual % change in OR
timeEffectMod_annual_pct_draws <- (exp(timeEffectMod_beta_year_draws) - 1) * 100

# Summarize
timeEffectMod_annual_pct_draws %>% 
  mean_qi()



```

What is the posterior probability that the urban-to-rural prevalence ratio exceeded 1 by 2019?

```{r}

year_2019_z <- 2019 - 2011

#get posterior draws for the linear predictor in 2019
post_2019 <-  add_epred_draws(object = timeEffectMod,
                              newdata = tibble(year_z = year_2019_z,
                                               logOR_se =1), 
                                               re_formula = NA)

post_2019 <- post_2019 %>%
  mutate(OR = exp(.epred))

#compute posterior probability that OR > 1
prob_OR_gt1 <- mean(post_2019$OR > 1)

prob_OR_gt1



```


#### Bacteriologically-positive: Model 3b: Region effect over time 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
timeEffectRegionMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                              1 + year_z + (1 + year_z | WHO.region) + (1+ year_z |id), 
                       data = bactModelDF_noEM,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       backend = "cmdstanr",
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r}
summary(timeEffectRegionMod)
plot(timeEffectRegionMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndTimeReg <- bactModelDF_noEM %>%
             select(logOR_se, WHO.region) %>%
             distinct() %>%
             crossing(year_z = seq(min(cleanRuralityDF$year_z, na.rm = TRUE),
                                   max(cleanRuralityDF$year_z, na.rm = TRUE), 1))

timeEffectRegionModEst <- add_epred_draws(object = timeEffectRegionMod,
                                 newdata=ndTimeReg,
                                 re_formula = ~(1 + year_z |WHO.region),
                                 allow_new_levels = TRUE) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
timeEffectRegionModEst %>% 
    group_by(year_z, WHO.region) %>% 
    mean_qi(urRatio) %>% 
    mutate(Year = year_z + 2011) %>%
    select(WHO.region, Year, urRatio, .lower, .upper) %>%
    arrange(WHO.region, Year)


```


```{r}
    timeEffectRegionModEst %>%
    ggplot(aes(x = year_z + 2011,
               y = urRatio,
               fill = WHO.region, 
               color = WHO.region)) +
    stat_lineribbon(aes(color = WHO.region, fill = WHO.region), .width = 0.95, alpha = 0.25) +
    geom_hline(yintercept = 1, linetype = 2) +
    facet_wrap(~ WHO.region) +
    labs(y = "Urban-to-rural odds ratio!",
         x = "Year",
         fill = "WHO region", 
         color = "WHO region") +
    theme_bw()  + theme(legend.position = "bottom")
```

Percentage change per year

```{r}
newdata <- expand.grid(
  year_z = seq(min(bactModelDF_noEM$year_z), max(bactModelDF_noEM$year_z), length.out = 2), 
  WHO.region = unique(bactModelDF_noEM$WHO.region),
  id = NA 
) %>%
  mutate(logOR_se =1)

pred_draws <- add_epred_draws(timeEffectRegionMod, newdata = newdata, re_formula = ~(1 + year_z | WHO.region))

region_pct_change <- pred_draws %>%
  group_by(.draw, WHO.region) %>%
  summarise(
    logOR_diff = diff(.epred),
    n_years = diff(range(year_z)),  
    .groups = "drop"
  ) %>%
  mutate(
    slope_per_year = logOR_diff / n_years,
    annual_pct_change = (exp(slope_per_year) - 1) * 100
  ) %>%
  group_by(WHO.region) %>%
  mean_qi(annual_pct_change)

region_pct_change

```

Posterior probability >1 in 2019

```{r}

#get posterior draws for the linear predictor in 2019
post_2019_region <-  add_epred_draws(object = timeEffectRegionMod,
                              newdata = tibble(year_z = year_2019_z,
                                               logOR_se =1)%>%
                                                expand_grid(WHO.region = unique(bactModelDF_noEM$WHO.region)),
                                               re_formula = ~(1 + year_z| WHO.region)) 

post_2019_region <- post_2019_region %>%
  mutate(OR = exp(.epred))

#compute posterior probability that OR > 1
prob_OR_gt1_region <-  post_2019_region %>%
        group_by(WHO.region) %>%
        summarise(post_prob = mean(OR > 1))

prob_OR_gt1_region

```


### Bacteriologically-positive: Univariate analysis of covariates 


First sort out the other two covariates (pct_urban, and TB incidence)

```{r}

#convert TB incidence so that we can estimate per 100 per 100,000 per year change, rather than per 1 per 100,000 per year change (not a meaningful unit)

bactModelDF_noEM <- bactModelDF_noEM %>%
  mutate(incidence_100 = incidence/100) %>%
  mutate(pct_unit = pct_urban*100)


```

Write functions for running models


Firstly write a list of formulae that we will run

```{r}

formulas <- list(
  
  #Nationally representative
  natRep_region = bf(logOR | se(logOR_se, sigma=TRUE) ~ 1 + natRep + (1 + natRep | WHO.region) + (1 | id)),
  
  #risk of bias
  bias_region = bf(logOR | se(logOR_se, sigma=TRUE) ~ 1 + bias + (1 + bias | WHO.region) + (1 | id)),
  
  #symptom screening
  symptom_region = bf(logOR | se(logOR_se, sigma=TRUE) ~ 1 + symptomScreen + (1 + symptomScreen | WHO.region) + (1 | id)),
  
  #TB incidence
  incidence_region = bf(logOR | se(logOR_se, sigma=TRUE) ~ 1 + incidence_100 + (1 + incidence_100 | WHO.region) + (1 | id)),
  
  #Percent of population living in urban areas
  pct_urban_region = bf(logOR | se(logOR_se, sigma=TRUE) ~ 1 + pct_unit + (1 + pct_unit | WHO.region) + (1 | id))
  )
```


Function to fit the univariate models

```{r}

fit_univariate_models <- function(formula, data, iter = 4000, chains = 4, cores = 4){
  
  fit <- brm(
    formula = formula,
    data = data,
    family = gaussian(),
    backend = "cmdstanr",
    prior = prior(student_t(7, 0, 1.5), class = Intercept) +
            prior(normal(0, 10), class = b) +
            prior(exponential(2), class = sd),
    control = list(adapt_delta = 0.99),
    iter = iter,
    chains = chains,
    cores = cores
  )

  lst(univariate_fits = fit)
}

# Fit all univariate models
fit_all_univariate <- function(formula_list, data, iter = 4000, chains = 4, cores = 4) {
  
  #If no names, create generic ones
  if (is.null(names(formula_list))) {
    names(formula_list) <- paste0("model_", seq_along(formula_list))
  }
  
  #Fit each model and store in a named list
  fits <- map(formula_list, ~ fit_univariate_models(.x, data, iter, chains, cores)$univariate_fits)
  
  # Return named list of fitted models
  fits
}

```

Function to take posterior draws and summarise global associations with covariates.

```{r}

tidy_region_effect_with_global <- function(fit) {
  draws <- as_draws_df(fit)
  
  # Identify the single predictor slope (exclude intercept)
  b_cols <- grep("^b_", names(draws), value = TRUE)
  b_cols <- setdiff(b_cols, "b_Intercept")
  if (length(b_cols) != 1) stop("Expected exactly one predictor column")
  slope_name <- b_cols
  coef_name <- str_remove(slope_name, "^b_")
  
  # Extract region-specific random slopes
  r_cols <- grep(paste0("^r_WHO.region\\[.*,", coef_name, "\\]"), names(draws), value = TRUE)
  regions <- str_match(r_cols, paste0("\\[(.*),", coef_name, "\\]"))[,2]
  
  # Compute regional ORs (equivalent to prevalence ratios, as prevalence is small)
  region_tbl <- map2_dfr(r_cols, regions, function(col, region) {
    draws %>%
      rowwise() %>%
      transmute(logOR = sum(c_across(all_of(c(slope_name, col))))) %>%
      ungroup() %>%
      mutate(or = exp(logOR)) %>%
      mean_qi(or) %>%
      mutate(coefficient = coef_name, area = region)
  }) %>%
    transmute(
      coefficient,
      area,
      `odds_ratio_and95%cri` = paste0(round(or,2), " (", round(.lower,2), "-", round(.upper,2), ")")
    )
  
  # Compute global OR from the mean of the regional random slopes
  global_tbl <- draws %>%
    rowwise() %>%
    transmute(logOR = sum(c_across(all_of(slope_name))) +
                      mean(c_across(all_of(r_cols)))) %>%
    ungroup() %>%
    mutate(or = exp(logOR)) %>%
    mean_qi(or) %>%
    mutate(coefficient = coef_name, area = "global") %>%
    transmute(coefficient, area, `odds_ratio_and95%cri` = paste0(round(or,2), " (", round(.lower,2), "-", round(.upper,2), ")"))
  
  bind_rows(global_tbl, region_tbl)
}



```

Summarise all posterior draws

```{r}
summarise_all_univariate <- function(fit_list) {
  
  # Identify regional models
  region_names <- names(fit_list)[str_detect(names(fit_list), "_region$")]
  
  # Extract both global + regional summaries
  all_tbl <- map_df(region_names, function(nm) {
    
    # global effects
    global <- tidy_global_effect(fit_list[[nm]]) %>%
      mutate(model = nm,
             area = "Global")   # tag clearly
    
    # regional effects
    regional <- tidy_region_effect(fit_list[[nm]]) %>%
      mutate(model = nm)
    
    bind_rows(global, regional)
  })
  
  # Arrange consistently
  all_tbl %>%
    arrange(coefficient, area, model)
}

```

Now fit all the models. Note we exclude the two studies from EM

```{r}
res <- fit_all_univariate(formulas, data = bactModelDF_noEM)
```

Summarise the univariate associations, and arrange for output table

```{r}
all_univariate_out <- summarise_all_univariate(res)

all_univariate_out_wide <- all_univariate_out %>%
  select(coefficient, area, `odds_ratio_and95%cri`) %>%
  pivot_wider(
    names_from = area,
    values_from = `odds_ratio_and95%cri`
  ) %>%
  arrange(coefficient) %>%
  select(coefficient,
         Global,
         `Africa region` = `Africa.region`,
         `South-East Asia region` = `South-East.Asia.region`,
         `Western Pacific region` = Western.Pacific.region) %>%
  mutate(
      coefficient = case_when(
        str_detect(coefficient, "natRep") ~ "Nationally representative (Yes vs No)",
        str_detect(coefficient, "bias")   ~ "Bias (not low vs low)",
        str_detect(coefficient, "symptomScreen") ~ "Symptom screening (Yes vs No)",
        str_detect(coefficient, "incidence_100") ~ "National TB incidence (per 100/100,000 increase)",
        str_detect(coefficient, "pct_unit") ~ "Percent of population in urban areas (per % point increase)",
        TRUE ~ NA_character_
      ))

all_univariate_out_wide

```

Adjusted main effects. This is our multivariable model. We allow effects of coefficients (slopes) to vary by 3 global regions.

```{r}

multi_region_varying <- brms::brm(
  logOR | se(logOR_se, sigma = TRUE) ~ 
    1 + natRep + bias + symptomScreen + incidence_100 + pct_unit + (1 | id) + (1 + year_z + natRep + bias + symptomScreen + incidence_100 + pct_unit | WHO.region), 
  data = bactModelDF_noEM,
  family = gaussian(),
  prior = prior(student_t(7, 0, 1.5), class = Intercept) +
          prior(normal(0, 10), class = b) +
          prior(exponential(2), class = sd) +   
          prior(lkj(2), class = cor, group = "WHO.region"),        
  control = list(adapt_delta = 0.99),
  backend = "cmdstanr",
  iter = 4000,
  chains = 4,
  cores = 4
)

summary(multi_region_varying)

```

Now function for tudying up the regional effects

```{r}
tidy_multi_region_effects <- function(fit, region_order = NULL) {
  draws <- as_draws_df(fit)
  
  # fixed-effect columns (excluding intercept)
  fe_cols <- grep("^b_", names(draws), value = TRUE)
  fe_cols <- setdiff(fe_cols, "b_Intercept")
  if (length(fe_cols) < 1) stop("No fixed-effect b_ columns found (excluding intercept).")
  coef_names <- sub("^b_", "", fe_cols)
  
  # global table
  global_tbl <- map_dfr(seq_along(fe_cols), function(i) {
    col <- fe_cols[i]
    coef <- coef_names[i]
    vals <- draws[[col]]
    med <- median(vals)
    lo  <- quantile(vals, 0.025)
    hi  <- quantile(vals, 0.975)
    or  <- exp(med); or_lo <- exp(lo); or_hi <- exp(hi)
    tibble(
      coefficient = coef,
      area = "Global",
      `odds_ratio_and95%cri` = sprintf("%.2f (%.2f-%.2f)", or, or_lo, or_hi)
    )
  })
  
  # region-specific table
  region_tbl <- map_dfr(seq_along(fe_cols), function(i) {
    col <- fe_cols[i]
    coef <- coef_names[i]
    # rcols like "r_WHO.region[Region.name,coef]"
    r_pattern <- paste0("^r_WHO.region\\[([^,]+),", coef, "\\]")
    rcols <- grep(r_pattern, names(draws), value = TRUE)
    if (length(rcols) == 0) return(tibble())  # no regional slopes for this coef
    
    map_dfr(rcols, function(rcol) {
      region <- sub(paste0("^r_WHO.region\\[([^,]+),", coef, "\\]$"), "\\1", rcol)
      vals <- draws[[col]] + draws[[rcol]]   # vector add: fixed + region deviation
      med <- median(vals)
      lo  <- quantile(vals, 0.025)
      hi  <- quantile(vals, 0.975)
      or  <- exp(med); or_lo <- exp(lo); or_hi <- exp(hi)
      tibble(
        coefficient = coef,
        area = region,
        `odds_ratio_and95%cri` = sprintf("%.2f (%.2f-%.2f)", or, or_lo, or_hi)
      )
    })
  })
  
  combined <- bind_rows(global_tbl, region_tbl)
  
  # pivot wider
  wide <- combined %>%
    pivot_wider(names_from = area, values_from = `odds_ratio_and95%cri`) %>%
    arrange(coefficient) %>%
  mutate(
      coefficient = case_when(
        str_detect(coefficient, "natRep") ~ "Nationally representative (Yes vs No)",
        str_detect(coefficient, "bias")   ~ "Bias (not low vs low)",
        str_detect(coefficient, "symptomScreen") ~ "Symptom screening (Yes vs No)",
        str_detect(coefficient, "incidence_100") ~ "National TB incidence (per 100/100,000 increase)",
        str_detect(coefficient, "pct_unit") ~ "Percent of population in urban areas (per % point increase)",
        TRUE ~ NA_character_
      ))
  
  if (!is.null(region_order)) {
    region_order <- unique(c(region_order, "Global"))
    cols <- c("coefficient", intersect(region_order, colnames(wide)))
    other_cols <- setdiff(colnames(wide), cols)
    wide <- wide %>% select(all_of(cols), all_of(other_cols))
  }
  
  wide
}


```

Run the function to get and tidy the multivariable associations.

```{r}

multi_region_out <- tidy_multi_region_effects(multi_region_varying)

```

Join together in a single table

```{r}

#first, standardise column names across the two tables
all_uni <- all_univariate_out_wide %>%
  rename(
    coefficient = coefficient,
    Global_univariate            = Global,
    Africa_univariate            = `Africa region`,
    `South-East Asia_univariate` = `South-East Asia region`,
    `Western Pacific_univariate` = `Western Pacific region`
  )

multi <- multi_region_out %>%
  rename(
    coefficient = coefficient,
    Global_multivariate            = Global,
    Africa_multivariate            = Africa.region,
    `South-East Asia_multivariate` = `South-East.Asia.region`,
    `Western Pacific_multivariate` = Western.Pacific.region
  ) 

#join on coefficient
combined <- full_join(all_uni, multi, by = "coefficient") %>%
    select(coefficient,
            Global_univariate, `Global_multivariate`,
            `Africa_univariate`, `Africa_multivariate`,
            `South-East Asia_univariate`, `South-East Asia_multivariate`,
            `Western Pacific_univariate`, `Western Pacific_multivariate`)

combined

```




### Smear positive analysis 

Prepare the dataset for smear positive modelling.

```{r, echo=FALSE, warning = FALSE, message=FALSE}
cleanRuralitySmrDF0 <- smrPosIndicator("rurality") %>% filter(rurality.smr.analysis.indicator !="none")

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanRuralitySmrDF <- cleanRuralitySmrDF0 %>% 
              filter(rurality.smr.analysis.indicator != "none") %>% 
              mutate("id" = row_number(),

           adj.prev100k.smear.positive.tb.urban = adj.prev100k.smear.positive.tb.urban/1e5,
           adj.prev100k.ci.upper.smear.positive.tb.urban = adj.prev100k.ci.upper.smear.positive.tb.urban/1e5,
           adj.prev100k.ci.lower.smear.positive.tb.urban = adj.prev100k.ci.lower.smear.positive.tb.urban/1e5,
           
           adj.prev100k.smear.positive.tb.rural = adj.prev100k.smear.positive.tb.rural/1e5,
           adj.prev100k.ci.upper.smear.positive.tb.rural = adj.prev100k.ci.upper.smear.positive.tb.rural/1e5,
           adj.prev100k.ci.lower.smear.positive.tb.rural = adj.prev100k.ci.lower.smear.positive.tb.rural/1e5,
           
           prev100k.smear.positive.tb.urban = prev100k.smear.positive.tb.urban/1e5,
           prev100k.ci.upper.smear.positive.tb.urban = prev100k.ci.upper.smear.positive.tb.urban/1e5,
           prev100k.ci.lower.smear.positive.tb.urban = prev100k.ci.lower.smear.positive.tb.urban/1e5,
           
           prev100k.smear.positive.tb.rural = prev100k.smear.positive.tb.rural/1e5,
           prev100k.ci.upper.smear.positive.tb.rural = prev100k.ci.upper.smear.positive.tb.rural/1e5,
           prev100k.ci.lower.smear.positive.tb.rural = prev100k.ci.lower.smear.positive.tb.rural/1e5) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           WHO.region,
           study.start.year, 
           study.end.year, 
           ends_with("urban"), ends_with("rural"),
           rurality.smr.analysis.indicator) %>%
    
    mutate("tempPrevUrban" =  case_when(rurality.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.urban,
                                       rurality.smr.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.urban, 
                                       rurality.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.urban/n.participants.urban,
                                        rurality.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.urban),
           "tempPrevRural" = case_when(rurality.smr.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.rural,
                                        rurality.smr.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.rural, 
                                       rurality.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.rural/n.participants.rural,
                                       rurality.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.rural)) %>%
    pivot_longer(cols = c(tempPrevRural,tempPrevUrban), 
                 names_to = "Rurality", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(rurality.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Urban" ~ (adj.prev100k.ci.upper.smear.positive.tb.urban-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.urban)/3.92,
                                        rurality.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Rural" ~ (adj.prev100k.ci.upper.smear.positive.tb.rural-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.rural)/3.92,
                                       
                                        rurality.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Urban" ~ (prev100k.ci.upper.smear.positive.tb.urban-
                                                                 prev100k.ci.lower.smear.positive.tb.urban)/3.92,
                                        rurality.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Rural" ~ (prev100k.ci.upper.smear.positive.tb.rural-
                                                                   prev100k.ci.lower.smear.positive.tb.rural)/3.92,
                                        rurality.smr.analysis.indicator %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Rurality == "Urban" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.urban),
                                        rurality.smr.analysis.indicator  %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Rurality == "Rural" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.rural)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(rurality.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Rurality == "Urban" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.urban/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.urban)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.urban /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.urban)))/3.92,
                                              rurality.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Rurality == "Rural" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.rural/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.rural)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.rural /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.rural)))/3.92,
                                                 rurality.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Rurality == "Urban" ~ (log(prev100k.ci.upper.smear.positive.tb.urban/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.urban)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.urban /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.urban)))/3.92,
                                                 rurality.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Rurality == "Rural" ~ (log(prev100k.ci.upper.smear.positive.tb.rural/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.rural)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.rural /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.rural)))/3.92, 
                                            rurality.smr.analysis.indicator %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Urban" ~ 
                                            log(1/(n.participants.urban*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        rurality.smr.analysis.indicator  %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Rural" ~ 
                                            log(1/(n.participants.rural*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       year_z = study.start.year - round(mean(study.start.year, na.rm=TRUE)),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))  %>% filter(LogOddsStandardError < Inf)

```

Prepare the dataset for smear-positive analysis

```{r}
smrModelDF <- cleanRuralitySmrDF %>%
    select(id, WHO.region, study.country, Rurality, year_z, natRep, LogOdds, LogOddsStandardError) %>%
    pivot_wider(names_from = Rurality,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_Urban - LogOdds_Rural,
        logOR_se = sqrt(LogOddsStandardError_Urban^2 + LogOddsStandardError_Rural^2) 
    ) 
```


### Smear positive descriptive statistics

```{r}

print(paste(dim(cleanRuralitySmrDF0)[1], "unique surveys with urban/rural-stratified smear positive TB"))

# sort(unique(cleanRuralitySmrDF0$title.extracted))
print(paste(length(unique(cleanRuralitySmrDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanRuralitySmrDF0$n.participants.rural)), "studies report participant numbers totaling", 
            sum(cleanRuralitySmrDF0$n.participants.rural, cleanRuralitySmrDF0$n.participants.urban, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanRuralitySmrDF0$n.participants.urban, na.rm=TRUE)/sum(cleanRuralitySmrDF0$n.participants.rural, cleanRuralitySmrDF0$n.participants.urban, na.rm = TRUE)*100, 2), "% of participants were urban" ))
```

We will also exclude the Eastern Mediterranean surveys from the main model.

```{r}

smrModelDF_noEM <- smrModelDF %>%
    filter(WHO.region !="Eastern Mediterranean region") 

```

#### Smear positive: Model 1: Smear positive main effect 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | study.country) + (1|id),
                       data = smrModelDF_noEM,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       backend = "cmdstanr",
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```


Summarize and check 

```{r}
summary(mainEffectSmrMod)
plot(mainEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}

nd <- smrModelDF_noEM %>%
      select(study.country, LogOdds_Rural) %>%
      mutate(logOR_se=1)

mainEffectSmrModEst <- add_epred_draws(object = mainEffectSmrMod,
                                 newdata=nd,
                                 re_formula = ~ 1) %>% 
                    mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectSmrModEst %>% ungroup %>% mean_qi(urRatio)
```

#### Smear positive:  Model 2: Smear positive effect by world region model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regionEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = smrModelDF_noEM,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       backend = "cmdstanr",
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r}
summary(regionEffectSmrMod)
plot(regionEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndRegion <- smrModelDF_noEM %>%
      select(WHO.region, LogOdds_Rural) %>%
      mutate(logOR_se=1)

regionEffectSmrModEst <- add_epred_draws(object = regionEffectSmrMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region)) %>% 
                       mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectSmrModEst %>% group_by(WHO.region) %>% mean_qi(urRatio)
```

```{r}
### Get the empirical values 
empirical_smr_prev <- smrModelDF_noEM %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        p_rural_empirical = plogis(LogOdds_Rural),
        p_urban_empirical = plogis(LogOdds_Urban),
        or_empirical = p_urban_empirical / p_rural_empirical,
        or_empirical = (p_urban_empirical / (1 - p_urban_empirical)) / 
            (p_rural_empirical / (1 - p_rural_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific smear positive TB prevalence")

ggplot() +
    geom_vline(aes(xintercept = 1), linetype=2) +
    geom_point(data = empirical_smr_prev, aes(y=WHO.region, x=or_empirical, color = type2),alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectSmrModEst, aes(x=urRatio, y=WHO.region, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Urban-to-rural prevalence ratio",
         y = "World region") +
    theme_bw() + theme(legend.title = element_blank(), 
          panel.grid.minor=element_blank(),
          legend.position = "bottom")
    
```

Change in smear-positive urban-to-rural prevalence ratio over time?

```{r}
smrtimeEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z + (1 | study.country) + (1 + year_z | id),
                       data = smrModelDF_noEM,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       backend = "cmdstanr",
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

```{r}

summary(smrtimeEffectMod)
plot(smrtimeEffectMod)
conditional_effects(smrtimeEffectMod)

```

```{r}

smrtimeEffectModEst <- add_epred_draws(object = smrtimeEffectMod,
                                 newdata=ndTime,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate

```{r}
smrtimeEffectModEst %>% 
    group_by(year_z) %>% 
    mean_qi(urRatio) %>% 
    mutate(Year = year_z + 2011) %>%
    select(Year, urRatio, .lower, .upper)

```


```{r}
    smrtimeEffectModEst %>%
    ggplot(aes(x = year_z + 2011,
               y = urRatio)) +
    stat_lineribbon(.width = 0.95, alpha = 0.5, fill="cadetblue") +
    scale_y_log10(
    breaks = c(0.25, 0.5, 1, 2, 4),
    labels = c("0.25", "0.5", "1", "2", "4")
  ) +
    scale_x_continuous(
    limits = c(2000, 2020),
    breaks = c(2000, 2005, 2010, 2015, 2020),
    labels = c("2000", "2005", "2010", "2015", "2020")
  ) +
    geom_hline(yintercept = 1, linetype = 2) +
    labs(y = "Urban-to-rural prevalence ratio",
         x = "Year",
         size = "Log odds of standard error of TB prevalence") +
    theme_bw()  + theme(legend.position = "bottom")
```

Annual percentage change

```{r}

smrtimeEffectMod_posterior <- as_draws_df(smrtimeEffectMod)

smrtimeEffectMod_beta_year_draws <- smrtimeEffectMod_posterior$b_year_z

# Annual % change in OR
smrtimeEffectMod_annual_pct_draws <- (exp(smrtimeEffectMod_beta_year_draws) - 1) * 100

# Summarize
smrtimeEffectMod_annual_pct_draws %>% 
  mean_qi()



```

What is the posterior probability that the urban-to-rural prevalence ratio exceeded 1 by 2019?

```{r}

#get posterior draws for the linear predictor in 2019
smrpost_2019 <-  add_epred_draws(object = smrtimeEffectMod,
                              newdata = tibble(year_z = year_2019_z,
                                               logOR_se =1), 
                                               re_formula = NA)

smrpost_2019 <- smrpost_2019 %>%
  mutate(OR = exp(.epred))

#compute posterior probability that OR > 1
smrprob_OR_gt1 <- mean(smrpost_2019$OR > 1)

smrprob_OR_gt1



```
