---
title: "Urban/rural meta-analysis"
output: html_notebook
---

### Description

This script reads in the clean dataset from data folder and then runs fits a Bayesian generalized multivariate multilevel model on the studies that report urban-rural stratified prevalence results. This model aims to estimate urban-rural stratified prevalence adjusted for geographic factors. 

```{r, echo = FALSE,  warning = FALSE, message=FALSE}

library(dplyr)
library(forcats)
library(tidyr)
library(reshape2)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(scales)
library(posterior)
library(tidybayes)
library(ggplot2)
```

### Functions and options

```{r}
expit <- function (x) {exp(x)/(1 + exp(x))}

myPal <- c("black", "mediumseagreen")

source(here("code/bacterialPositiveIndicator.R"))
source(here("code/smearPositiveIndicator.R"))

```

### Data import and prep

Load data and filter to studies reporting urban-rural stratification

```{r}
cleanRuralityDF0 <- bactPostIndicator("rurality") %>% filter(rurality.analysis.indicator !="none")

```

### Descriptive statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE, echo=FALSE}

print(paste(dim(cleanRuralityDF0)[1], "unique surveys with urban.rural-stratified bacteriological positive TB"))

# sort(unique(cleanRuralityDF0$title.extracted))
print(paste(length(unique(cleanRuralityDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanRuralityDF0$n.participants.rural)), "studies report participant numbers totaling", 
            sum(cleanRuralityDF0$n.participants.rural, cleanRuralityDF0$n.participants.urban, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanRuralityDF0$n.participants.urban, na.rm=TRUE)/sum(cleanRuralityDF0$n.participants.rural, cleanRuralityDF0$n.participants.urban, na.rm = TRUE)*100, 2), "% of participants were male" ))
```

### Risk of bias statistics

```{r, warning = FALSE, message=FALSE, cache=TRUE}

table(cleanRuralityDF0$study.quality.summary)
```


### Prepare dataset for bacteriologically-positive modelling

```{r, echo = FALSE,  warning = FALSE, message=FALSE}
##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanRuralityDF <-  cleanRuralityDF0 %>% 
               filter(rurality.analysis.indicator != "none") %>% 
               mutate("id" = row_number(),
               adj.prev100k.bacteriological.tb.urban = adj.prev100k.bacteriological.tb.urban/1e5,
               adj.prev100k.ci.upper.bacteriological.tb.urban = adj.prev100k.ci.upper.bacteriological.tb.urban/1e5,
               adj.prev100k.ci.lower.bacteriological.tb.urban = adj.prev100k.ci.lower.bacteriological.tb.urban/1e5,
               
               adj.prev100k.bacteriological.tb.rural = adj.prev100k.bacteriological.tb.rural/1e5,
               adj.prev100k.ci.upper.bacteriological.tb.rural = adj.prev100k.ci.upper.bacteriological.tb.rural/1e5,
               adj.prev100k.ci.lower.bacteriological.tb.rural = adj.prev100k.ci.lower.bacteriological.tb.rural/1e5,
               
               adj.prev100k.smear.positive.tb.urban = adj.prev100k.smear.positive.tb.urban/1e5,
               adj.prev100k.ci.upper.smear.positive.tb.urban = adj.prev100k.ci.upper.smear.positive.tb.urban/1e5,
               adj.prev100k.ci.lower.smear.positive.tb.urban = adj.prev100k.ci.lower.smear.positive.tb.urban/1e5,
               
               adj.prev100k.smear.positive.tb.rural = adj.prev100k.smear.positive.tb.rural/1e5,
               adj.prev100k.ci.upper.smear.positive.tb.rural = adj.prev100k.ci.upper.smear.positive.tb.rural/1e5,
               adj.prev100k.ci.lower.smear.positive.tb.rural = adj.prev100k.ci.lower.smear.positive.tb.rural/1e5,
               
               prev100k.bacteriological.tb.urban = prev100k.bacteriological.tb.urban/1e5,
               prev100k.ci.upper.bacteriological.tb.urban = prev100k.ci.upper.bacteriological.tb.urban/1e5,
               prev100k.ci.lower.bacteriological.tb.urban = prev100k.ci.lower.bacteriological.tb.urban/1e5,
               
               prev100k.bacteriological.tb.rural = prev100k.bacteriological.tb.rural/1e5,
               prev100k.ci.upper.bacteriological.tb.rural = prev100k.ci.upper.bacteriological.tb.rural/1e5,
               prev100k.ci.lower.bacteriological.tb.rural = prev100k.ci.lower.bacteriological.tb.rural/1e5,
               
               prev100k.smear.positive.tb.urban = prev100k.smear.positive.tb.urban/1e5,
               prev100k.ci.upper.smear.positive.tb.urban = prev100k.ci.upper.smear.positive.tb.urban/1e5,
               prev100k.ci.lower.smear.positive.tb.urban = prev100k.ci.lower.smear.positive.tb.urban/1e5,
               
               prev100k.smear.positive.tb.rural = prev100k.smear.positive.tb.rural/1e5,
               prev100k.ci.upper.smear.positive.tb.rural = prev100k.ci.upper.smear.positive.tb.rural/1e5,
               prev100k.ci.lower.smear.positive.tb.rural = prev100k.ci.lower.smear.positive.tb.rural/1e5) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           WHO.region,
           screening.tests,
           study.start.year, 
           study.end.year, 
           study.quality.summary,
           ends_with("urban"), ends_with("rural"),
           rurality.analysis.indicator) %>%
    
    mutate("tempPrevUrban" =  case_when(rurality.analysis.indicator == 
                                           "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.urban, 
                                       rurality.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.urban,
                                       rurality.analysis.indicator == 
                                           "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.urban, 
                                       rurality.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.urban, 
                                       rurality.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.urban/n.participants.urban, 
                                       rurality.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.urban/n.participants.urban,
                                       rurality.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.urban/n.participants.urban,
                                        rurality.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.urban),
           "tempPrevRural" = case_when(rurality.analysis.indicator == 
                                            "adj.prev100k.ci.bacteriological.tb" ~ adj.prev100k.bacteriological.tb.rural, 
                                        rurality.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.rural,
                                        rurality.analysis.indicator == 
                                            "prev100k.ci.bacteriological.tb" ~ prev100k.bacteriological.tb.rural, 
                                        rurality.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.rural, 
                                       rurality.analysis.indicator == 
                                           "n.bacteriological.tb" ~ n.bacteriological.tb.rural/n.participants.rural, 
                                       rurality.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.rural/n.participants.rural,
                                       rurality.analysis.indicator == 
                                           "n.culture.positive.tb" ~ n.culture.positive.tb.rural/n.participants.rural,
                                       rurality.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.rural)) %>%
    pivot_longer(cols = c(tempPrevRural,tempPrevUrban), 
                 names_to = "Rurality", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(rurality.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Rurality == "Urban" ~ (adj.prev100k.ci.upper.bacteriological.tb.urban-
                                                                 adj.prev100k.ci.lower.bacteriological.tb.urban)/3.92,
                                        rurality.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" & 
                                            Rurality == "Rural" ~ (adj.prev100k.ci.upper.bacteriological.tb.rural-
                                                                   adj.prev100k.ci.lower.bacteriological.tb.rural)/3.92,
                                        rurality.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Urban" ~ (adj.prev100k.ci.upper.smear.positive.tb.urban-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.urban)/3.92,
                                        rurality.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Rural" ~ (adj.prev100k.ci.upper.smear.positive.tb.rural-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.rural)/3.92,
                                        rurality.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Rurality == "Urban" ~ (prev100k.ci.upper.bacteriological.tb.urban-
                                                                 prev100k.ci.lower.bacteriological.tb.urban)/3.92,
                                        rurality.analysis.indicator == "prev100k.ci.bacteriological.tb" & 
                                            Rurality == "Rural" ~ (prev100k.ci.upper.bacteriological.tb.rural-
                                                                   prev100k.ci.lower.bacteriological.tb.rural)/3.92,
                                        rurality.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Urban" ~ (prev100k.ci.upper.smear.positive.tb.urban-
                                                                 prev100k.ci.lower.smear.positive.tb.urban)/3.92,
                                        rurality.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Rural" ~ (prev100k.ci.upper.smear.positive.tb.rural-
                                                                   prev100k.ci.lower.smear.positive.tb.rural)/3.92,
                                        rurality.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Urban" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.urban),
                                        rurality.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Rural" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.rural)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(rurality.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Rurality == "Urban" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.urban/
                                                                           (1-adj.prev100k.ci.upper.bacteriological.tb.urban)) -
                                                                       log(adj.prev100k.ci.lower.bacteriological.tb.urban/
                                                                               (1-adj.prev100k.ci.lower.bacteriological.tb.urban))) /3.92,
                                              rurality.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" &
                                                  Rurality == "Rural" ~ (log(adj.prev100k.ci.upper.bacteriological.tb.rural/
                                                                             (1-adj.prev100k.ci.upper.bacteriological.tb.rural)) -
                                                                         log(adj.prev100k.ci.lower.bacteriological.tb.rural /
                                                                                 (1-adj.prev100k.ci.lower.bacteriological.tb.rural))) /3.92,
                                              rurality.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Rurality == "Urban" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.urban/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.urban)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.urban /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.urban)))/3.92,
                                              rurality.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Rurality == "Rural" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.rural/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.rural)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.rural /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.rural)))/3.92,
                                              rurality.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Rurality == "Urban" ~ (log(prev100k.ci.upper.bacteriological.tb.urban/
                                                                     (1-prev100k.ci.upper.bacteriological.tb.urban)) -
                                                                 log(prev100k.ci.lower.bacteriological.tb.urban/
                                                                     (1-prev100k.ci.lower.bacteriological.tb.urban))) /3.92,
                                                 rurality.analysis.indicator == "prev100k.ci.bacteriological.tb" &
                                                 Rurality == "Rural" ~ (log(prev100k.ci.upper.bacteriological.tb.rural/
                                                                        (1-prev100k.ci.upper.bacteriological.tb.rural)) -
                                                                    log(prev100k.ci.lower.bacteriological.tb.rural /
                                                                        (1-prev100k.ci.lower.bacteriological.tb.rural))) /3.92,
                                                 rurality.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Rurality == "Urban" ~ (log(prev100k.ci.upper.smear.positive.tb.urban/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.urban)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.urban /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.urban)))/3.92,
                                                 rurality.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Rurality == "Rural" ~ (log(prev100k.ci.upper.smear.positive.tb.rural/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.rural)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.rural /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.rural)))/3.92, 
                                            rurality.analysis.indicator %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Urban" ~ 
                                            log(1/(n.participants.urban*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        rurality.analysis.indicator  %in% c("n.bacteriological.tb", "n.smear.positive.tb", 
                                                                      "n.culture.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Rural" ~ 
                                            log(1/(n.participants.rural*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       year_z = study.start.year - round(mean(study.start.year, na.rm=TRUE)),
                                       bias= ifelse(grepl("Low", study.quality.summary), "Low", "Not low"),
                                       symptomScreen = ifelse(grepl("Symptom", 
                                                                    screening.tests, 
                                                                    ignore.case=TRUE), "Yes", "No"),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))
```


```{r}
bactModelDF <- cleanRuralityDF %>%
    select(id, WHO.region, study.country, Rurality, year_z, natRep, bias, symptomScreen, LogOdds, LogOddsStandardError) %>%
    pivot_wider(names_from = Rurality,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_Urban - LogOdds_Rural,
        logOR_se = sqrt(LogOddsStandardError_Urban^2 + LogOddsStandardError_Rural^2)
    )
```



#### Bacteriologically-positive: Model 1: Main effect model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|study.country) + (1|id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 


```{r}
summary(mainEffectMod)
plot(mainEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
nd <- bactModelDF %>%
      select(study.country, LogOdds_Rural) %>%
      mutate(logOR_se=1)

mainEffectModEst <- add_epred_draws(object = mainEffectMod,
                                 newdata=nd,
                                 re_formula = ~ 1, 
                                 allow_new_levels = TRUE) %>% 
                    mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectModEst %>% ungroup %>% mean_qi(urRatio)
```

#### Bacteriologically-positive: Model 2: Effect by world region model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regionEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(regionEffectMod)
plot(regionEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndRegion <- bactModelDF %>%
      select(WHO.region, LogOdds_Rural) %>%
      mutate(logOR_se=1)

regionEffectModEst <- add_epred_draws(object = regionEffectMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region),
                                 allow_new_levels = TRUE) %>% 
                       mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectModEst %>% group_by(WHO.region) %>% mean_qi(urRatio)
```


```{r}
### Get the empirical values 
empirical_prev <- bactModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        bias = bias,
        symptomScreen = symptomScreen, 
        p_rural_empirical = plogis(LogOdds_Rural),
        p_urban_empirical = plogis(LogOdds_Urban),
        or_empirical = p_urban_empirical / p_rural_empirical,
        or_empirical = (p_urban_empirical / (1 - p_urban_empirical)) / 
            (p_rural_empirical / (1 - p_rural_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific TB prevalence")

    ggplot() +
    geom_vline(aes(xintercept = 1), linetype=2) +
    geom_point(data = empirical_prev, aes(y=WHO.region, x=or_empirical, color = type2),alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectModEst, aes(x=urRatio, y=WHO.region, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Urban-to-rural prevalence ratio",
         y = "WHO region") +
    theme_bw() + theme(legend.title = element_blank(), legend.position = "bottom", 
          panel.grid.minor=element_blank())
    
```

#### Bacteriologically-positive: Model 3: Effect over time 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
timeEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r}
summary(timeEffectMod)
plot(timeEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndTime <- bactModelDF %>%
    select(logOR_se) %>%
    crossing(year_z = seq(min(cleanRuralityDF$year_z, na.rm = TRUE), max(cleanRuralityDF$year_z, na.rm = TRUE), 1))

timeEffectModEst <- add_epred_draws(object = timeEffectMod,
                                 newdata=ndTime,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```


Main estimate

```{r}
timeEffectModEst %>% 
    group_by(year_z) %>% 
    mean_qi(urRatio) %>% 
    mutate(Year = year_z + round(mean(cleanRuralityDF$study.end.year, na.rm=TRUE))) %>%
    select(Year, urRatio, .lower, .upper)

```


```{r}
    timeEffectModEst %>%
    ggplot(aes(x = year_z + round(mean(cleanRuralityDF$study.end.year, na.rm=TRUE)),
               y = urRatio)) +
    stat_lineribbon(.width = 0.95, alpha = 0.25, fill="mediumseagreen") +
    geom_point(data = empirical_prev, 
               aes(x=year_z + round(mean(cleanRuralityDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = logOR_se), 
               colour = "black") +
    geom_hline(yintercept = 1, linetype = 2) +
    labs(y = "Urban-to-rural odds ratio",
         x = "Year") +
    theme_bw()  + theme(legend.position = "bottom")
```

#### Bacteriologically-positive: Model 8: Change in prevalence over time.

Only for countries with more than one survey

```{r}

multipleSurveysBactModelDF <- bactModelDF %>%
    group_by(study.country) %>%
    mutate(surveys = n()) %>%
    filter(surveys >1)

#one study has missing study.end.date
#reading the manuscript, not clearly stated, but most likely date of prevalence survey is 2010

multipleSurveysBactModelDF <- multipleSurveysBactModelDF %>%
    mutate(year_z = case_when(
        is.na(year_z) ~ 2010 - round(mean(cleanRuralityDF$study.start.year, na.rm=TRUE)),
              TRUE ~ year_z)
    )

```

fit the model

```{r}

timeEffectMod2 <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + year_z + (1 + year_z | study.country) + (1 | id),
                       data = multipleSurveysBactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)


```

Summarize and check 

```{r}
summary(timeEffectMod2)
plot(timeEffectMod2)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}

timeEffectModEst2 <- add_epred_draws(object = timeEffectMod2,
                                 newdata=multipleSurveysBactModelDF,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```


Effect by country

```{r}
    timeEffectModEst2 %>%
    ggplot(aes(x = year_z + round(mean(cleanRuralityDF$study.end.year, na.rm=TRUE)),
               y = urRatio, fill=study.country, group=study.country)) +
    stat_lineribbon(.width = 0.95, alpha = 0.25) +
    geom_point(data = empirical_prev %>% 
                   inner_join(multipleSurveysBactModelDF), 
                aes(x=year_z + round(mean(cleanRuralityDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = logOR_se), 
                colour = "black", shape=1) +
    geom_hline(yintercept = 1, linetype = 2) +
    facet_wrap(study.country~., scales = "free_y") +
    labs(y = "Urban-to-rural odds ratio",
         x = "Year") +
    theme_bw()  + 
    theme(legend.position = "none")
```





#### Bacteriologically-positive: Model 3b: Region effect over time 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
timeEffectRegionMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                              1 + year_z * WHO.region + (1 | study.country) + (1 | id), 
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r}
summary(timeEffectRegionMod)
plot(timeEffectRegionMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndTimeReg <- bactModelDF %>%
             select(logOR_se, WHO.region) %>%
             distinct() %>%
             crossing(year_z = seq(min(cleanRuralityDF$year_z, na.rm = TRUE),
                                   max(cleanRuralityDF$year_z, na.rm = TRUE), 1))

timeEffectRegionModEst <- add_epred_draws(object = timeEffectRegionMod,
                                 newdata=ndTimeReg,
                                 re_formula = NA,
                                 allow_new_levels = TRUE) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
timeEffectRegionModEst %>% 
    group_by(year_z, WHO.region) %>% 
    mean_qi(urRatio) %>% 
    mutate(Year = year_z + round(mean(cleanRuralityDF$study.end.year, na.rm=TRUE))) %>%
    select(WHO.region, Year, urRatio, .lower, .upper) %>%
    arrange(WHO.region, Year)

```


```{r}
    timeEffectRegionModEst %>%
    filter(WHO.region !="Americas region",
           WHO.region !="Eastern Mediterranean region") %>%
    ggplot(aes(x = year_z + round(mean(cleanRuralityDF$study.end.year, na.rm = TRUE)),
               y = urRatio,
               fill = WHO.region, 
               color = WHO.region)) +
    stat_lineribbon(aes(color = WHO.region, fill = WHO.region), .width = 0.95, alpha = 0.25) +
    geom_point(data = empirical_prev %>%
                   filter(WHO.region !="Americas region",
                          WHO.region !="Eastern Mediterranean region"), 
               aes(x=year_z + round(mean(cleanRuralityDF$study.end.year, na.rm=TRUE)), y=or_empirical, size = logOR_se), 
               colour = "black") +
    geom_hline(yintercept = 1, linetype = 2) +
    facet_wrap(~ WHO.region) +
    labs(y = "Urban-to-rural odds ratio!",
         x = "Year",
         fill = "WHO region", 
         color = "WHO region") +
    theme_bw()  + theme(legend.position = "bottom")
```

### Bacteriologically-positive: Univariate analysis of other covariates 

#### Bacteriologically-positive: Model 6: National versus subnational 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
natEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + natRep + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r, echo=FALSE, warning = FALSE, message=FALSE}
summary(natEffectMod)
plot(natEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndNat <- bactModelDF %>%
    select(logOR_se, natRep)

natEffectModEst <- add_epred_draws(object = natEffectMod,
                                 newdata=ndNat,
                                 re_formula = NA) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
natEffectModEst %>% 
    group_by(natRep) %>% 
    mean_qi(urRatio) %>% 
    select(natRep, urRatio, .lower, .upper)
```


```{r}
    ggplot() +
    # geom_hline(aes(yintercept = 1), linetype=2) +
    geom_point(data = empirical_prev, aes(y=natRep, x=or_empirical, color = type2, size=logOR_se), alpha = 0.5) +
    stat_pointinterval(data=natEffectModEst, aes(x=urRatio, y=natRep, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Urban-to-rural prevalence ratio", color = "",
         y = "Survey representativeness") +
    scale_size_continuous(range = c(8,16)) +
    theme_bw() + theme(
                                     legend.position = c(.83,.2), legend.box.just = "right",
                                     # legend.direction = "horizontal",
          panel.grid.minor=element_blank())
```

#### Bacteriologically-positive: Model 6: Study quality (High vs. not high)

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}

biasEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + bias + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r}
summary(biasEffectMod)
plot(biasEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndBias <- bactModelDF %>%
    select(logOR_se, bias)

biasEffectModEst <- add_epred_draws(object = biasEffectMod,
                                 newdata=ndBias,
                                 re_formula = NA) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
biasEffectModEst %>% 
    group_by(bias) %>% 
    mean_qi(urRatio) %>% 
    select(bias, urRatio, .lower, .upper)
```


```{r}
    ggplot() +
    geom_point(data = empirical_prev, aes(y=bias, x=or_empirical, color = type2, size=logOR_se), alpha = 0.5) +
    stat_pointinterval(data=biasEffectModEst, aes(x=urRatio, y=bias, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio", color = "",
         y = "Low probability of survey bias vs. not low probability of survey bias") +
    scale_size_continuous(range = c(8,16)) +
    theme_bw() + theme(
                                     legend.position = c(.83,.2), legend.box.just = "right",
                                     # legend.direction = "horizontal",
          panel.grid.minor=element_blank())
```

#### Bacteriologically-positive: Model 7: Symptom screen vs. no symptom screen

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}

sympEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + symptomScreen + (1 | study.country) + (1 | id),
                       data = bactModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

Summarize and check 

```{r}
summary(sympEffectMod)
plot(sympEffectMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndSymp <- bactModelDF %>%
    select(logOR_se, symptomScreen)

sympEffectModEst <- add_epred_draws(object = sympEffectMod,
                                 newdata=ndSymp,
                                 re_formula = NA) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
sympEffectModEst %>% 
    group_by(symptomScreen) %>% 
    mean_qi(urRatio) %>% 
    select(symptomScreen, urRatio, .lower, .upper)
```


```{r, echo=FALSE, warning = FALSE, message=FALSE}
    ggplot() +
    geom_point(data = empirical_prev, aes(y=symptomScreen, x=or_empirical, color = type2, size=logOR_se), alpha = 0.5) +
    stat_pointinterval(data=sympEffectModEst, aes(x=urRatio, y=symptomScreen, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Male-to-female prevalence ratio", color = "",
         y = "Symptom screening required") +
    scale_size_continuous(range = c(8,16)) +
    theme_bw(base_size = 30) + theme(
                                     legend.position = c(.83,.2), legend.box.just = "right",
                                     # legend.direction = "horizontal",
          panel.grid.minor=element_blank(),
          plot.title = element_text(size=36))
```
 
#### Bacteriologically-positive: Model 8: country level urbanisation

Import the urbanisation data, and check
https://ourworldindata.org/urbanization


```{r}

urbanization <- read.csv("data/share-of-population-urban.csv")

countries_in_review <- bactModelDF %>%
    select(study.country) %>%
    distinct() %>%
    pull()

urbanizationClean <- urbanization %>%
    filter(Entity %in% countries_in_review) %>%
    janitor::clean_names() %>%
    mutate(pct_urban = urban_population_of_total_population/100)

urbanizationClean %>%
    ggplot() +
    geom_line(aes(x = year, y=pct_urban, colour = entity)) +
    scale_y_continuous(labels = scales::percent)

```

Now get the percentage of population that was urban in the year of the survey

```{r}

urbanizationClean <- urbanizationClean %>%
    mutate(year_z = year - round(mean(cleanRuralityDF0$study.start.year, na.rm=TRUE))) %>%
    select(study.country = entity,
           year_z,
           pct_urban)
    
urbanizationModelDF <- left_join(bactModelDF, urbanizationClean)


```

Construct a model to look at the effect of urbanization in country at time of survey

```{r}

urbanizationEffectMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + pct_urban + (1 | study.country) + (1 | id),
                       data = urbanizationModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
    


```


```{r}
summary(urbanizationEffectMod)
plot(urbanizationEffectMod)



```
Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndUrbanization <- crossing(
    pct_urban = modelr::seq_range(urbanizationModelDF$pct_urban, n=30),
    logOR_se = 1
)

urbanizationEffectModEst <- add_epred_draws(object = urbanizationEffectMod,
                                 newdata=ndUrbanization,
                                 re_formula = NA) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}
    ggplot() +
    geom_hline(aes(yintercept = 1), linetype=2) +
    stat_lineribbon(data=urbanizationEffectModEst, aes(y=urRatio, x=pct_urban), .width = 0.95, alpha=0.5) +
    scale_color_manual(values = myPal) +
    labs(y = "Male-to-female prevalence ratio", color = "",
         x = "Percent of population living in urban areas") +
    theme_bw() + theme(panel.grid.minor=element_blank())
```


#### Bacteriologically-positive: Model 8: country level urbanisation by region

```{r}

urbanizationEffectMod2 <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + pct_urban*WHO.region + (1 | study.country) + (1 | id),
                       data = urbanizationModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```


```{r}
summary(urbanizationEffectMod2)
plot(urbanizationEffectMod2)



```
Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndUrbanization2 <- crossing(
    pct_urban = modelr::seq_range(urbanizationModelDF$pct_urban, n=30),
    WHO.region = unique(urbanizationModelDF$WHO.region),
    logOR_se = 1
)

urbanizationEffectModEst2 <- add_epred_draws(object = urbanizationEffectMod2,
                                 newdata=ndUrbanization2,
                                 re_formula = NA) %>% 
                           mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

```{r, echo=FALSE, warning = FALSE, message=FALSE}

empirical_prev_urban <- urbanizationModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        bias = bias,
        pct_urban = pct_urban,
        p_rural_empirical = plogis(LogOdds_Rural),
        p_urban_empirical = plogis(LogOdds_Urban),
        or_empirical = p_urban_empirical / p_rural_empirical,
        or_empirical = (p_urban_empirical / (1 - p_urban_empirical)) / 
            (p_rural_empirical / (1 - p_rural_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific TB prevalence")


urbanizationEffectModEst2 %>%
    filter(WHO.region !="Eastern Mediterranean region") %>%
    ggplot() +
    geom_hline(aes(yintercept = 1), linetype=2) +
    stat_lineribbon(aes(y=urRatio, x=pct_urban, fill=WHO.region), 
                    .width = 0.95, alpha=0.5) +
    geom_point(data = empirical_prev_urban %>% filter(WHO.region !="Eastern Mediterranean region"),
               aes(x=pct_urban, y=or_empirical, size = logOR_se), shape = 1) +
    facet_wrap(WHO.region ~.) +
    scale_x_continuous(labels = scales::percent) +
    scale_fill_brewer(palette = "Set1") +
    labs(y = "Urban-to-rural prevalence ratio", color = "",
         x = "Percent of population living in urban areas") +
    theme_bw() + theme(panel.grid.minor=element_blank())
```

#### Multlivariate model, with all covariates


```{r}

mvMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1 + pct_urban + WHO.region + year_z + natRep + bias + symptomScreen +
                       (1 | study.country) + (1 | id),
                       data = urbanizationModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)

```

```{r}

summary(mvMod)
plot(mvMod)



```

```{r}

summarize_draws(mvMod) %>%
    filter(stringr::str_detect(variable, "b_"),
           variable !="b_Intercept") %>%
    select(variable, mean, q5, q95) %>%
    mutate(across(is.numeric, exp)) %>%
    ggplot() +
    geom_hline(aes(yintercept = 1), linetype=2) +
    geom_pointinterval(aes(y=mean, ymin=q5, ymax =q95, x=variable)) +
    coord_flip() +
    scale_y_log10() +
    theme(panel.background = element_rect(colour="grey78"))


```



### Smear positive analysis 

Prepare the dataset for smear positive modelling.

```{r, echo=FALSE, warning = FALSE, message=FALSE}
cleanRuralitySmrDF0 <- smrPosIndicator("rurality") %>% filter(rurality.smr.analysis.indicator !="none")

##############################################################################|
##### CREATE A TIDY DATASET OF RELEVANT VARIABLES #############################
cleanRuralitySmrDF <- cleanRuralitySmrDF0 %>% 
              filter(rurality.smr.analysis.indicator != "none") %>% 
              mutate("id" = row_number(),

           adj.prev100k.smear.positive.tb.urban = adj.prev100k.smear.positive.tb.urban/1e5,
           adj.prev100k.ci.upper.smear.positive.tb.urban = adj.prev100k.ci.upper.smear.positive.tb.urban/1e5,
           adj.prev100k.ci.lower.smear.positive.tb.urban = adj.prev100k.ci.lower.smear.positive.tb.urban/1e5,
           
           adj.prev100k.smear.positive.tb.rural = adj.prev100k.smear.positive.tb.rural/1e5,
           adj.prev100k.ci.upper.smear.positive.tb.rural = adj.prev100k.ci.upper.smear.positive.tb.rural/1e5,
           adj.prev100k.ci.lower.smear.positive.tb.rural = adj.prev100k.ci.lower.smear.positive.tb.rural/1e5,
           
           prev100k.smear.positive.tb.urban = prev100k.smear.positive.tb.urban/1e5,
           prev100k.ci.upper.smear.positive.tb.urban = prev100k.ci.upper.smear.positive.tb.urban/1e5,
           prev100k.ci.lower.smear.positive.tb.urban = prev100k.ci.lower.smear.positive.tb.urban/1e5,
           
           prev100k.smear.positive.tb.rural = prev100k.smear.positive.tb.rural/1e5,
           prev100k.ci.upper.smear.positive.tb.rural = prev100k.ci.upper.smear.positive.tb.rural/1e5,
           prev100k.ci.lower.smear.positive.tb.rural = prev100k.ci.lower.smear.positive.tb.rural/1e5) %>%
    select(id,
           covidence.id, 
           figure.id.yr, 
           study.geography, 
           title.extracted,
           study.country,
           WHO.region,
           study.start.year, 
           study.end.year, 
           ends_with("urban"), ends_with("rural"),
           rurality.smr.analysis.indicator) %>%
    
    mutate("tempPrevUrban" =  case_when(rurality.smr.analysis.indicator == 
                                           "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.urban,
                                       rurality.smr.analysis.indicator == 
                                           "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.urban, 
                                       rurality.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.urban/n.participants.urban,
                                        rurality.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.urban),
           "tempPrevRural" = case_when(rurality.smr.analysis.indicator == 
                                            "adj.prev100k.ci.smear.positive.tb" ~ adj.prev100k.smear.positive.tb.rural,
                                        rurality.smr.analysis.indicator == 
                                            "prev100k.ci.smear.positive.tb" ~ prev100k.smear.positive.tb.rural, 
                                       rurality.smr.analysis.indicator == 
                                           "n.smear.positive.tb" ~ n.smear.positive.tb.rural/n.participants.rural,
                                       rurality.smr.analysis.indicator == 
                                           "prev100k.smear.positive.tb" ~ prev100k.smear.positive.tb.rural)) %>%
    pivot_longer(cols = c(tempPrevRural,tempPrevUrban), 
                 names_to = "Rurality", 
                 values_to = "Adjusted Prevalence", 
                 names_prefix = "tempPrev") %>% 
    mutate("Standard Error" = case_when(rurality.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Urban" ~ (adj.prev100k.ci.upper.smear.positive.tb.urban-
                                                                 adj.prev100k.ci.lower.smear.positive.tb.urban)/3.92,
                                        rurality.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Rural" ~ (adj.prev100k.ci.upper.smear.positive.tb.rural-
                                                                   adj.prev100k.ci.lower.smear.positive.tb.rural)/3.92,
                                       
                                        rurality.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Urban" ~ (prev100k.ci.upper.smear.positive.tb.urban-
                                                                 prev100k.ci.lower.smear.positive.tb.urban)/3.92,
                                        rurality.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" & 
                                            Rurality == "Rural" ~ (prev100k.ci.upper.smear.positive.tb.rural-
                                                                   prev100k.ci.lower.smear.positive.tb.rural)/3.92,
                                        rurality.smr.analysis.indicator %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Rurality == "Urban" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.urban),
                                        rurality.smr.analysis.indicator  %in% c("n.smear.positive.tb","prev100k.smear.positive.tb") & Rurality == "Rural" ~ 
                                            sqrt((`Adjusted Prevalence`*(1-`Adjusted Prevalence`))/n.participants.rural)),
  
           "Phi" = (`Adjusted Prevalence`/`Standard Error`^2) - (`Adjusted Prevalence`/`Standard Error`)^2 - 1, 
           "LogOdds" = log(`Adjusted Prevalence`/(1-`Adjusted Prevalence`)), 
           "LogOddsStandardError" = case_when(rurality.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Rurality == "Urban" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.urban/
                                                                           (1-adj.prev100k.ci.upper.smear.positive.tb.urban)) -
                                                                       log(adj.prev100k.ci.lower.smear.positive.tb.urban /
                                                                               (1-adj.prev100k.ci.lower.smear.positive.tb.urban)))/3.92,
                                              rurality.smr.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" &
                                                  Rurality == "Rural" ~ (log(adj.prev100k.ci.upper.smear.positive.tb.rural/
                                                                             (1-adj.prev100k.ci.upper.smear.positive.tb.rural)) -
                                                                         log(adj.prev100k.ci.lower.smear.positive.tb.rural /
                                                                                 (1-adj.prev100k.ci.lower.smear.positive.tb.rural)))/3.92,
                                                 rurality.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Rurality == "Urban" ~ (log(prev100k.ci.upper.smear.positive.tb.urban/
                                                                          (1-prev100k.ci.upper.smear.positive.tb.urban)) -
                                                                      log(prev100k.ci.lower.smear.positive.tb.urban /
                                                                          (1-prev100k.ci.lower.smear.positive.tb.urban)))/3.92,
                                                 rurality.smr.analysis.indicator == "prev100k.ci.smear.positive.tb" &
                                                     Rurality == "Rural" ~ (log(prev100k.ci.upper.smear.positive.tb.rural/
                                                                            (1-prev100k.ci.upper.smear.positive.tb.rural)) -
                                                                        log(prev100k.ci.lower.smear.positive.tb.rural /
                                                                            (1-prev100k.ci.lower.smear.positive.tb.rural)))/3.92, 
                                            rurality.smr.analysis.indicator %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Urban" ~ 
                                            log(1/(n.participants.urban*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2,
                                        rurality.smr.analysis.indicator  %in% c("n.smear.positive.tb", "prev100k.smear.positive.tb") & Rurality == "Rural" ~ 
                                            log(1/(n.participants.rural*`Adjusted Prevalence`*(1-`Adjusted Prevalence`)))^2)) %>% 
                                mutate(natRep = ifelse(study.geography == "Nationally representative", "Yes", "No"),
                                       year_z = study.start.year - round(mean(study.start.year, na.rm=TRUE)),
                                       WHO.region = gsub("\\(WHO\\)", "region", WHO.region))  %>% filter(LogOddsStandardError < Inf)

```


```{r}
smrModelDF <- cleanRuralitySmrDF %>%
    select(id, WHO.region, study.country, Rurality, year_z, natRep, LogOdds, LogOddsStandardError) %>%
    pivot_wider(names_from = Rurality,
                values_from = c(LogOdds, LogOddsStandardError),
                names_sep = "_") %>%
    mutate(
        logOR = LogOdds_Urban - LogOdds_Rural,
        logOR_se = sqrt(LogOddsStandardError_Urban^2 + LogOddsStandardError_Rural^2) 
    ) 
```


### Smear positive descriptive statistics

```{r}

print(paste(dim(cleanRuralitySmrDF0)[1], "unique surveys with urban/rural-stratified smear positive TB"))

# sort(unique(cleanRuralitySmrDF0$title.extracted))
print(paste(length(unique(cleanRuralitySmrDF0$title.extracted))-1, "unique publications"))

print(paste(sum(!is.na(cleanRuralitySmrDF0$n.participants.rural)), "studies report participant numbers totaling", 
            sum(cleanRuralitySmrDF0$n.participants.rural, cleanRuralitySmrDF0$n.participants.urban, na.rm = TRUE), "participants"))

print(paste0(round(sum(cleanRuralitySmrDF0$n.participants.urban, na.rm=TRUE)/sum(cleanRuralitySmrDF0$n.participants.rural, cleanRuralitySmrDF0$n.participants.urban, na.rm = TRUE)*100, 2), "% of participants were urban" ))
```

#### Smear positive: Model 1: Smear positive main effect 

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
mainEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1 | study.country) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```


Summarize and check 

```{r}
summary(mainEffectSmrMod)
plot(mainEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
nd <- smrModelDF %>%
      select(study.country, LogOdds_Rural) %>%
      mutate(logOR_se=1)

mainEffectSmrModEst <- add_epred_draws(object = mainEffectSmrMod,
                                 newdata=nd,
                                 re_formula = ~ 1) %>% 
                    mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
mainEffectSmrModEst %>% ungroup %>% mean_qi(urRatio)
```
#### Smear positive:  Model 2: Smear positive effect by world region model

Model specification
```{r, warning = FALSE, message=FALSE, cache=TRUE}
regionEffectSmrMod <- brms::brm(formula = logOR | se(logOR_se, sigma=TRUE) ~ 
                           1  + (1|WHO.region) + (1|id),
                       data = smrModelDF,
                       prior = prior(student_t(7, 0, 1.5), class = Intercept) +
                           # prior(normal(0, 10), class = b) +
                           prior(exponential(2), class = sd),
                       family = gaussian(),
                       control = list(adapt_delta = 0.99),
                       chains = 4,
                       cores = 4,
                       iter = 4000)
```

Summarize and check 

```{r}
summary(regionEffectSmrMod)
plot(regionEffectSmrMod)
```

Prediction

```{r, echo=FALSE, warning = FALSE, message=FALSE}
ndRegion <- smrModelDF %>%
      select(WHO.region, LogOdds_Rural) %>%
      mutate(logOR_se=1)

regionEffectSmrModEst <- add_epred_draws(object = regionEffectSmrMod,
                                 newdata=ndRegion,
                                 re_formula = ~ 1 + (1|WHO.region)) %>% 
                       mutate(urRatio = exp(.epred),
                           type="Posterior mean and 95% credible interval") 
```

Main estimate
```{r}
regionEffectSmrModEst %>% group_by(WHO.region) %>% mean_qi(urRatio)
```

```{r}
### Get the empirical values 
empirical_smr_prev <- smrModelDF %>%
    transmute(
        id,
        WHO.region = WHO.region,
        year_z = year_z,
        logOR_se,
        natRep = natRep,
        p_rural_empirical = plogis(LogOdds_Rural),
        p_urban_empirical = plogis(LogOdds_Urban),
        or_empirical = p_urban_empirical / p_rural_empirical,
        or_empirical = (p_urban_empirical / (1 - p_urban_empirical)) / 
            (p_rural_empirical / (1 - p_rural_empirical))  # odds ratio
    ) %>% mutate(type = "Survey-specific ratios",
                 type2 = "Survey-specific smear positive TB prevalence")

ggplot() +
    geom_vline(aes(xintercept = 1), linetype=2) +
    geom_point(data = empirical_smr_prev, aes(y=WHO.region, x=or_empirical, color = type2),alpha = 0.5, size=8) +
    stat_pointinterval(data=regionEffectSmrModEst, aes(x=urRatio, y=WHO.region, color=type), .width = 0.95, size=20) +
    scale_color_manual(values = myPal) +
    scale_y_discrete(limits=rev, 
                     labels = label_wrap(10)) +
    labs(x = "Urban-to-rural prevalence ratio",
         y = "World region") +
    theme_bw() + theme(legend.title = element_blank(), 
          panel.grid.minor=element_blank(),
          legend.position = "bottom")
    
```


