---
title: "Horton Analysis Replica"
output: rmarkdown::pdf_document
---

```{r, warning=FALSE, message=FALSE}
##### ABOUT THIS SCRIPT #######################################################
### This script reads in the clean dataset from data folder and then 
### replicates the methods in Horton et. al to determine change.  
##############################################################################|
##### LOAD IN NECESSARY PACKAGES ##############################################
library(dplyr)
library(here)
library(magrittr)
library(DT)
library(metafor)
library(meta)
library(brms)
library(posterior)
library(tidybayes)
##############################################################################|
##### LOAD CLEAN DATA SET AND FILTER TO SURVEYS REPORTING SEX #################
source(here("code/bacterialPositiveIndicator.R"))
cleanSexDF0 <- bactPostIndicator("sex", FALSE) %>% filter(sex.analysis.indicator !="none")
```

## Bacteriologically positive TB 

Number of surveys reporting adjusted bacteriologically confirmed prevalence with confidence intervals: 
```{r, echo=FALSE}
table(cleanSexDF0$sex.analysis.indicator)[1]
```


Number of surveys reporting adjusted smear positive prevalence with confidence intervals")
```{r, echo=FALSE}
table(cleanSexDF0$sex.analysis.indicator)[2]
```

Number of surveys reporting crude bacteriologically positive prevalence with confidence intervals
```{r, echo=FALSE}
table(cleanSexDF0$sex.analysis.indicator)[7]
```


Number of surveys reporting crude smear positive prevalence with confidence intervals
```{r, echo=FALSE}
table(cleanSexDF0$sex.analysis.indicator)[8]
```
Number of surveys reporting number of bacteriologically positive TB cases and participants. 
```{r, echo=FALSE}
table(cleanSexDF0$sex.analysis.indicator)[3]
```
Number of surveys reporting number of culture positive TB cases and participants. 
```{r, echo=FALSE}
table(cleanSexDF0$sex.analysis.indicator)[4]
```

Number of surveys reporting number of smear positive TB cases and participants. 
```{r, echo=FALSE}
table(cleanSexDF0$sex.analysis.indicator)[5]
```

Number of surveys reporting crude prevalence of smear positive TB and number of participants. 
```{r, echo=FALSE}
table(cleanSexDF0$sex.analysis.indicator)[9]
```

Number of surveys disqualified from quantitative analysis.
```{r, echo=FALSE}
table(cleanSexDF0$sex.analysis.indicator)[6]
```

## Sex stratified prevalence of bacteriologically positive TB

```{r, echo=FALSE, warning=FALSE, message = FALSE}
### Open a PDF 
# pdf(file = here("output/HortonAnalysis.pdf"), width = 8.5, height = 11.5)

### Calculation of the bacteriological TB prevalence
### Random effects weighted model 
prevSexBact <- cleanSexDF0 %>% 
               mutate(n.bacteriological.tb.male.all = case_when(sex.analysis.indicator == "n.bacteriological.tb" ~ 
                                                                n.bacteriological.tb.male, 
                                                            sex.analysis.indicator == "n.smear.positive.tb" ~ 
                                                                n.smear.positive.tb.male, 
                                                            sex.analysis.indicator == "n.culture.positive.tb" ~ 
                                                                n.culture.positive.tb.male, 
                                                            sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~
                                                                prev100k.smear.positive.tb.male*n.participants.male/1e5, 
                                                            ################################################################
                                                            ##### NEED TO BE UPDATED TO CALCULATE BASED ON CONFIDENCE INTERVAL
                                                            ################################################################
                                                            sex.analysis.indicator == "prev100k.ci.bacteriological.tb" ~
                                                                prev100k.bacteriological.tb.male*n.participants.male/1e5, 
                                                            sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~
                                                                prev100k.smear.positive.tb.male*n.participants.male/1e5, 
                                                            sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" ~
                                                                adj.prev100k.bacteriological.tb.male*n.participants.male/1e5, 
                                                            sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" ~
                                                                adj.prev100k.smear.positive.tb.male*n.participants.male/1e5
                                                            ),
                      n.bacteriological.tb.female.all = case_when(sex.analysis.indicator == "n.bacteriological.tb" ~ 
                                                                n.bacteriological.tb.female, 
                                                            sex.analysis.indicator == "n.smear.positive.tb" ~ 
                                                                n.smear.positive.tb.female, 
                                                            sex.analysis.indicator == "n.culture.positive.tb" ~ 
                                                                n.culture.positive.tb.female, 
                                                            sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~
                                                                prev100k.smear.positive.tb.female*n.participants.female/1e5, 
                                                            ################################################################
                                                            ##### NEED TO BE UPDATED TO CALCULATE BASED ON CONFIDENCE INTERVAL
                                                            ################################################################
                                                            sex.analysis.indicator == "prev100k.ci.bacteriological.tb" ~
                                                                prev100k.bacteriological.tb.female*n.participants.female/1e5, 
                                                            sex.analysis.indicator == "prev100k.ci.smear.positive.tb" ~
                                                                prev100k.smear.positive.tb.female*n.participants.female/1e5, 
                                                            sex.analysis.indicator == "adj.prev100k.ci.bacteriological.tb" ~
                                                                adj.prev100k.bacteriological.tb.female*n.participants.female/1e5, 
                                                            sex.analysis.indicator == "adj.prev100k.ci.smear.positive.tb" ~
                                                                adj.prev100k.smear.positive.tb.female*n.participants.female/1e5)) %>% 
                    filter(is.na(n.bacteriological.tb.male.all)== FALSE &
                           is.na(n.participants.male)== FALSE)

### Males
propBactMale <-metaprop(n.bacteriological.tb.male.all, n.participants.male, figure.id.yr, 
                         data=prevSexBact, random=TRUE, 
                         pscale = 1e5)

summary(propBactMale)

### Females
propBactFemale <-metaprop(n.bacteriological.tb.female.all, n.participants.female, figure.id.yr, 
                         data=prevSexBact, random=TRUE, 
                         pscale = 1e5)

summary(propBactFemale)
```


### Male to female ratio of bacteriologically positive TB prevalence
```{r, echo=FALSE, fig.width=8.5, fig.height=11, warning=FALSE, message = FALSE}
##############################################################################|
##### MALE TO FEMALE RATIO OF BACTERIOLOGICAL PREVALENCE  #####################
##############################################################################|

bactPrevMF <- rma(ai=n.bacteriological.tb.male.all, 
                  bi=n.participants.male - n.bacteriological.tb.male.all, 
                  ci=n.bacteriological.tb.female.all, 
                  di=n.participants.female - n.bacteriological.tb.female.all, 
                  data=prevSexBact, measure="RR", 
                  slab=paste(figure.id.yr), method="REML")

forest(bactPrevMF,
       atransf=exp, cex=.75,  xlab="M:F ratio (Bacteriologically-Positive TB)", mlab="Overall summary", 
       psize=1, col="black", border="black")

```

#COUNTRIES WITH MORE THAN ONE PREVALENCE SURVEY 

```{r, echo=FALSE, warning=FALSE, message = FALSE}
##############################################################################|
##### COUNTRIES WITH MORE THAN ONE PREVALENCE SURVEY  #########################
##############################################################################|

countries <- as.vector(unlist(prevSexBact %>% count(study.country) %>% 
                              filter(n > 1) %>%
                              select(study.country)))

for (country in countries){
    tmpBactPrevMF <- rma(ai=n.bacteriological.tb.male.all, 
                       bi=n.participants.male - n.bacteriological.tb.male.all, 
                       ci=n.bacteriological.tb.female.all, 
                       di=n.participants.female - n.bacteriological.tb.female.all, 
                       data=prevSexBact, measure="RR", 
                       subset = (study.country == country),
                       slab=paste(figure.id.yr), method="REML")
    par(cex=1, font=2)
    print(forest(tmpBactPrevMF,
           atransf=exp, cex=.75,  xlab="M:F ratio (Bacteriologically-Positive TB)", 
           mlab="Overall summary", 
           psize=1, col="black", border="black" ))
}
```

## Univariate meta-regression of male-to-female prevalence ratios

```{r}
##############################################################################|
##### UNIVARIATE META REGRESSION OF MF PREVALENCE RATIOS ######################
##############################################################################|
print("Univariate meta regression")
### BY WHO REGION 
# prevSexBact$WHO.region <- as.factor(prevSexBact$WHO.region)
#     
# metab1 <- rma(ai=n.bacteriological.tb.male, 
#               bi=n.participants.male - n.bacteriological.tb.male, 
#               ci=n.bacteriological.tb.female, 
#               di=n.participants.female - n.bacteriological.tb.female, 
#               mods = regionWHO,
#               data=prevSexBact, measure="RR", 
#               slab=paste(figure.id.yr), method="REML")
# metab1$b <- exp(metab1$b)
# metab1$ci.lb <- exp(metab1$ci.lb)
# metab1$ci.ub <- exp(metab1$ci.ub)
# summary(metab1)

### BY NATIONAL VS. SUBNATIONAL SURVEYS 
prevSexBact %<>% mutate(national=ifelse(study.geography == "Nationally representative",
                        1, 0))

metab2 <- rma(ai=n.bacteriological.tb.male.all, 
              bi=n.participants.male - n.bacteriological.tb.male.all, 
              ci=n.bacteriological.tb.female.all, 
              di=n.participants.female - n.bacteriological.tb.female.all, 
              mods = ~national,
              data=prevSexBact, measure="RR", 
              slab=paste(figure.id.yr), method="REML")

print("Horton intrcpt=2.1930 , se=0.0893")
print(paste("Our intrcpt =", round(exp(metab2$b[1]),4), "se=", round(exp(metab2$se[1]),3)))
print("Horton national=1.0142 ,se=0.1533")
print(paste("Our national =", round(exp(metab2$b[2]),4), "se=", round(exp(metab2$se[2]),3)))

summary(metab2)

### BY SURVEY START YEAR
prevSexBact %<>% mutate(start.year = study.start.year - 2001)
metab3 <- rma(ai=n.bacteriological.tb.male.all, 
              bi=n.participants.male - n.bacteriological.tb.male.all, 
              ci=n.bacteriological.tb.female.all, 
              di=n.participants.female - n.bacteriological.tb.female.all, 
              mods = ~start.year,
              data=prevSexBact, measure="RR", 
              slab=paste(figure.id.yr), method="REML")

print("Horton intrcpt=5205946.3416, se=37.7545")
print(paste("Our intrcpt =", round(exp(metab3$b[1]),4), "se=", round(exp(metab3$se[1]),3)))
print("Horton start.year=0.9927, se=0.0188")
print(paste("Our start.year =", round(exp(metab3$b[2]),4), "se=", round(exp(metab3$se[2]),3)))

summary(metab3)

### BY LOW VS MODERATE/HIGH RISK OF BIAS 

prevSexBact %<>% mutate(low.risk=ifelse(study.quality.summary ==
"Low risk of bias - Further research is very unlikely to change our confidence in the estimate",
                        1, 0))

metab7 <- rma(ai=n.bacteriological.tb.male.all, 
              bi=n.participants.male - n.bacteriological.tb.male.all, 
              ci=n.bacteriological.tb.female.all, 
              di=n.participants.female - n.bacteriological.tb.female.all, 
              mods = ~low.risk,
              data=prevSexBact, measure="RR", 
              slab=paste(figure.id.yr), method="REML")

print("Horton intrcpt=2.4240 , se=0.1155")
print(paste("Our intrcpt =", round(exp(metab7$b[1]),4), "se=", round(exp(metab7$se[1]),3)))
print("Horton low.risk=0.8481, se=0.1480")
print(paste("Our low.risk =", round(exp(metab7$b[2]),4), "se=", round(exp(metab7$se[2]),3)))

summary(metab7)


### surveys with initial screening procedures requiring self-report of
### signs/symptoms vs.broader initial screening procedures.

prevSexBact %<>% mutate(symp.only = ifelse(screening.tests == "Symptom screen only",
                                        1, 0))

metab8 <- rma(ai=n.bacteriological.tb.male.all, 
              bi=n.participants.male - n.bacteriological.tb.male.all, 
              ci=n.bacteriological.tb.female.all, 
              di=n.participants.female - n.bacteriological.tb.female.all, 
              mods = ~symp.only,
              data=prevSexBact, measure="RR", 
              slab=paste(figure.id.yr), method="REML")

print("Horton intrcpt=2.3384 , se=0.0829")
print(paste("Our intrcpt =", round(exp(metab8$b[1]),4), "se=", round(exp(metab8$se[1]),3)))
print("Horton low.risk=0.8005, se=0.1622")
print(paste("Our low.risk =", round(exp(metab8$b[2]),4), "se=", round(exp(metab8$se[2]),3)))

summary(metab8)

### Surveys with diagnosis by smear microscopy vs.other diagnostic measures.

# prevSexBact %<>% mutate(smear.only = ifelse(screening.tests == "Symptom screen only",
#                                            1, 0))
# 
# metab8 <- rma(ai=n.bacteriological.tb.male, 
#               bi=n.participants.male - n.bacteriological.tb.male, 
#               ci=n.bacteriological.tb.female, 
#               di=n.participants.female - n.bacteriological.tb.female, 
#               mods = ~symp.only,
#               data=prevSexBact, measure="RR", 
#               slab=paste(figure.id.yr), method="REML")
# 
# exp(metab8$b)
# exp(metab8$ci.lb)
# exp(metab8$ci.ub)
# summary(metab8)
```


## Smear positive TB 
```{r,echo=FALSE}

print("smear.positive POSITIVE TB")
print("Number of surveys reporting adjusted prevalence with confidence intervals")
dim(cleanSexDF0 %>% filter(!is.na(adj.prev100k.ci.smear.positive.tb.male)))[1]

print("Number of surveys reporting crude prevalence with confidence intervals")
dim(cleanSexDF0 %>% filter(is.na(adj.prev100k.ci.smear.positive.tb.male) & 
                           !is.na(prev100k.ci.smear.positive.tb.male)))[1]

print("Number of surveys reporting numerator and denominator")
dim(cleanSexDF0 %>% filter(is.na(adj.prev100k.ci.smear.positive.tb.male) & 
                           is.na(prev100k.ci.smear.positive.tb.male) &
                           !is.na(n.smear.positive.tb.male) &
                           !is.na(n.participants.male)))[1]

print("Number of surveys reporting numerator/denominator and adjusted prevalence")
dim(cleanSexDF0 %>% filter(is.na(adj.prev100k.ci.smear.positive.tb.male) & 
                           is.na(prev100k.ci.smear.positive.tb.male) &
                           xor(!is.na(n.smear.positive.tb.male), !is.na(n.participants.male)) &
                           !is.na(adj.prev100k.smear.positive.tb.male)))[1]

print("Number of surveys reporting numerator and crude prevalence")
dim(cleanSexDF0 %>% filter(is.na(adj.prev100k.ci.smear.positive.tb.male) & 
                           is.na(prev100k.ci.smear.positive.tb.male) &
                           xor(!is.na(n.smear.positive.tb.male), !is.na(n.participants.male)) &
                           is.na(adj.prev100k.ci.smear.positive.tb.male) &
                           !is.na(prev100k.smear.positive.tb.male)))[1]


##############################################################################|
##### SEX STRATIFIED SMEAR POSITIVE TB PREVALENCE   ###########################
##############################################################################|

### Calculation of smear positive TB prevalence among males 
### Random effects weighted model 
prevSexSmear <- cleanSexDF0 %>% filter(is.na(n.smear.positive.tb.male)== FALSE &
                                           is.na(n.participants.male)== FALSE )

propSmearMale <-metaprop(n.smear.positive.tb.male, n.participants.male, figure.id.yr, 
                         data=prevSexSmear, random=TRUE, pscale = 1e5)

summary(propSmearMale)

### Calculation of smear positive TB prevalence among females 
### Random effects weighted model 
propSmearFemale <-metaprop(n.smear.positive.tb.female, n.participants.female, figure.id.yr, 
                           data=prevSexSmear, random=TRUE, pscale = 1e5)

summary(propSmearFemale)

##############################################################################|
##### MALE TO FEMALE RATIO OF SMEAR POSITIVE PREVALENCE  ######################
##############################################################################|

### calculate order 
prevSexSmear %<>% mutate(smear.order = n.smear.positive.tb.male/n.participants.male/
                             n.smear.positive.tb.female/n.participants.female)
smearPrevMF <- rma(ai=n.smear.positive.tb.male, 
                  bi=n.participants.male - n.smear.positive.tb.male, 
                  ci=n.smear.positive.tb.female, 
                  di=n.participants.female - n.smear.positive.tb.female, 
                  data=prevSexSmear, measure="RR", 
                  slab=paste(figure.id.yr), method="REML")

par(cex=1, font=2)
metafor::forest(smearPrevMF, 
                order = prevSexSmear$WHO.region,
       atransf=exp, cex=.85,  xlab="M:F ratio (Smear-Positive TB)", 
       mlab="Overall summary", 
       psize=1, col="black", border="black",
       rows=c(3:20, 23, 26:49, 52:61))
par(cex=0.85, font=2)
text(-8.65, 50.5, "South-East Asian Region")
text(-8.85, 24.5, "Region of the Americas")
text(-9.65, 21.5, "African Region")
```