---
title: "New variables"
output: html_notebook
---

Load required libraries

```{r}
library(tidyverse)
library(here)
```

import the clean dataset

```{r}
source(here("code/standardizeData.R"))
stndDF <- standardData(fullDataRaw)$cleanDF


```

#Standardise column names

```{r}
stndDF <- stndDF %>%
  rename_with(str_to_lower)


```


#Separate confidence intervals out into columns

https://github.com/nswartwo/tb-lmic-prevalence-review/issues/8

We will use consistence naming throughout for all variables

First make a dataset containing only those variables with want to separate out, with the covidence ID.

Also need to swap out some non-standard seprarator characters

Also need to remove some "n/a"s etc and recode as missing

```{r}

confvars <- stndDF %>%
    select(covidence.id,
           contains(".ci.")) %>%
    mutate(across(
        .cols = -covidence.id,
        .fns = ~ str_replace(., "–", "-"))) %>%
    mutate(across(
        .cols = -covidence.id,
        .fns = ~ case_when(.=="n/a" | .=="NR" | .=="N/AN/A" ~ NA_character_,
                           TRUE ~ .)
    ))

```

There is one study where confidence intervals are presented for multiple strata in the same cell (covidence.id = 26972) -pull this one out currently until fixed.


```{r}

confvars2 <- confvars %>%
    filter(covidence.id !=26972)


```

Now separate the confidence intervals into `.lower`, and `.higher`

```{r}

# Identify all columns containing ".CI."
ci_cols <- names(confvars2)[str_detect(names(confvars2), "\\.ci\\.")]

# Use purrr::reduce() to iteratively apply separate()
confvars_cleaned <- reduce(
  ci_cols,
  ~ .x %>%
    separate(
      col = .y,
      into = c(
        str_replace(.y, "\\.ci\\.", ".ci.lower."),
        str_replace(.y, "\\.ci\\.", ".ci.upper.")
      ),
      sep = "-"
    ),
  .init = confvars2
)

```

Convert all confidence interval columns to numeric

```{r}

confvars_cleaned <- confvars_cleaned %>%
    mutate(across(contains(".ci."), as.numeric))


```

Now join back onto the main dataset

```{r}

cleanDF <- left_join(stndDF, confvars_cleaned)

```


### TB symptom variables

Now we will separate out the TB symptom variables https://github.com/nswartwo/tb-lmic-prevalence-review/issues/9

Select out the columns that we need

```{r}

symptomvars <- cleanDF %>%
    select(covidence.id, screening.symptoms)


```

Separate the rows out into multiple columns

```{r}

symptomvars_clean <- symptomvars %>%
  # Separate rows by symptoms
  separate_rows(screening.symptoms, sep = ";") %>%
  # Clean up whitespace and standardize text
  mutate(screening.symptoms = str_trim(screening.symptoms)) %>%
  # Create a column for "Other" symptoms and separate them from the rest
  mutate(
    other_symptom = if_else(str_starts(screening.symptoms, "Other:"), screening.symptoms, NA_character_),
    screening.symptoms = if_else(!str_starts(screening.symptoms, "Other:"), screening.symptoms, NA_character_)
  ) %>%
  # Combine "Other" symptoms into a single character column
  group_by(covidence.id) %>%
  summarise(
    other_symptom = paste(na.omit(other_symptom), collapse = "; "),
    screening.symptoms = list(na.omit(screening.symptoms)),
    .groups = "drop"
  ) %>%
  # Flatten symptom lists and unnest
  unnest_longer(screening.symptoms) %>%
  mutate(present = 1) %>%
  pivot_wider(names_from = screening.symptoms, values_from = present, values_fill = 0) %>%
  #reorder the columns for ease of use
  select(covidence.id, `Symptom screen was not performed`, `Cough ≥ 2 weeks`, `Cough ≥ 3 weeks`, `Cough of other or unknown duration`,
         Haemoptysis, `Chest pain`, Fever, `Night sweats`, `Weight loss`, `Sputum production`, `other_symptom`) %>%
  #now sort out column names to standard format
  rename_with(str_to_lower) %>% 
  rename_with(~ gsub(" ",".", .x), contains(" ")) %>% 
  rename_with(~ gsub("_",".", .x), contains("_")) %>%
  #remove the ≥ symbols from variable names
  rename_with(~ gsub("≥", "greater.or.equal.to", .x), contains("≥")) %>%
  #remove unnecessary text from `other.symptom` column 
  mutate(other.symptom = str_remove(other.symptom, "Other: ")) %>%
  #convert boolean screening variables to yes/no factors
  mutate(across(c(symptom.screen.was.not.performed:sputum.production), ~
                    case_when(.==0 ~ "No",
                              .==1 ~ "Yes")))
  
```

Now join back onto the clean dataset

```{r}

cleanDF <- left_join(cleanDF, symptomvars_clean)


```

