---
title: "R Notebook"
output: html_notebook
---

Load required libraries

```{r}
library(tidyverse)
library(here)
```

import the clean dataset

```{r}
source(here("code/standardizeData.R"))
stndDF <- standardData(fullDataRaw)$cleanDF


```


#Separate confidence intervals out into columns

We will use consistence naming throughout for all variables

First make a dataset containing only those variables with want to separate out, with the covidence ID.

Also need to swap out some non-standard seprarator characters

Also need to remove some "n/a"s and recode as missing

```{r}

confvars <- stndDF %>%
    select(covidence.id,
           contains(".ci.")) %>%
    mutate(across(
        .cols = -covidence.id,
        .fns = ~ str_replace(., "–", "-"))) %>%
    mutate(across(
        .cols = -covidence.id,
        .fns = ~ case_when(.=="n/a" | .=="NR" | .=="N/AN/A" ~ NA_character_,
                           TRUE ~ .)
    ))

```

There is one study where confidence intervals are presented for multiple strata in the same cell (covidence.id = 26972) -pull this one out currently until fixed.


```{r}

confvars2 <- confvars %>%
    filter(covidence.id !=26972)


```

Now separate the confidence intervals into `.lower`, and `.higher`

```{r}


confvars3 <- confvars2 %>%
    separate_wider_delim(-covidence.id, delim = "-", names_sep = "_", cols_remove = TRUE) %>% #separate columns
    rename_at(vars(contains("_1")), ~ paste0(., "lower"))


```

