---
title: "R Notebook"
output: html_notebook
---

Load required libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
```

import the clean dataset

```{r}
source(here("code/standardizeData.R"))
stndDF <- standardData(fullDataRaw)$cleanDF


```

#Standardise column names

```{r}
stndDF <- stndDF %>%
  rename_with(str_to_lower)


```


#Separate confidence intervals out into columns

We will use consistence naming throughout for all variables

First make a dataset containing only those variables with want to separate out, with the covidence ID.

Also need to swap out some non-standard seprarator characters

Also need to remove some "n/a"s etc and recode as missing

```{r}

confvars <- stndDF %>%
    select(covidence.id,
           contains(".ci.")) %>%
    mutate(across(
        .cols = -covidence.id,
        .fns = ~ str_replace(., "–", "-"))) %>%
    mutate(across(
        .cols = -covidence.id,
        .fns = ~ case_when(.=="n/a" | .=="NR" | .=="N/AN/A" ~ NA_character_,
                           TRUE ~ .)
    ))

```

There is one study where confidence intervals are presented for multiple strata in the same cell (covidence.id = 26972) -pull this one out currently until fixed.


```{r}

confvars2 <- confvars %>%
    filter(covidence.id !=26972)


```

Now separate the confidence intervals into `.lower`, and `.higher`

```{r}

# Identify all columns containing ".CI."
ci_cols <- names(confvars2)[str_detect(names(confvars2), "\\.ci\\.")]

# Use purrr::reduce() to iteratively apply separate()
confvars_cleaned <- reduce(
  ci_cols,
  ~ .x %>%
    separate(
      col = .y,
      into = c(
        str_replace(.y, "\\.ci\\.", ".ci.lower."),
        str_replace(.y, "\\.ci\\.", ".ci.upper.")
      ),
      sep = "-"
    ),
  .init = confvars2
)

```

Convert all confidence interval columns to numeric

```{r}

confvars_cleaned <- confvars_cleaned %>%
    mutate(across(contains(".ci."), as.numeric))


```

Now join back onto the main dataset

```{r}

cleanDF <- left_join(stndDF, confvars_cleaned)

```