---
title: "urban-rural-model"
author: "Alireza Mortazavi"
date: "2025-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}

expit <- function (x) {1/(1 + exp(-x))} #convert back to the response scale

```
## Subset of data
```{r}
SDS <- bactPosRuralityData %>% 
    select(c(1,2,3,55,130:132,148,149,150,151:153,160:195,902:907))

```
## Filterout the studies that dont have bacterilogical result based on rurality.analysis.indicator
```{r}
SDS_filtered <- SDS %>%
  filter(!rurality.analysis.indicator %in% c(
    "adj.prev100k.ci.smear.positive.tb",
    "n.smear.positive.tb",
    "prev100k.ci.smear.positive.tb"
  ))
```
Hierchy of the model
```{r}
SDS_filtered <- SDS_filtered %>%
  mutate(
    urban_p = case_when(
      !is.na(adj.prev100k.bacteriological.tb.urban) ~ adj.prev100k.bacteriological.tb.urban,
      !is.na(prev100k.bacteriological.tb.urban) ~ prev100k.bacteriological.tb.urban,
      !is.na(n.bacteriological.tb.urban) & !is.na(n.participants.urban) ~
        (n.bacteriological.tb.urban / n.participants.urban) * 100000,
      TRUE ~ NA_real_
    ),
    urban_ci = case_when(
      !is.na(adj.prev100k.ci.bacteriological.tb.urban) ~ as.character(adj.prev100k.ci.bacteriological.tb.urban),
      !is.na(prev100k.ci.bacteriological.tb.urban) ~ as.character(prev100k.ci.bacteriological.tb.urban),
      TRUE ~ NA_character_
    )
  )
```
```{r}
SDS_filtered <- SDS_filtered %>%
  mutate(
    rural_p = case_when(
      !is.na(adj.prev100k.bacteriological.tb.rural) ~ adj.prev100k.bacteriological.tb.rural,
      !is.na(prev100k.bacteriological.tb.rural) ~ prev100k.bacteriological.tb.rural,
      !is.na(n.bacteriological.tb.rural) & !is.na(n.participants.rural) ~
        (n.bacteriological.tb.rural / n.participants.rural) * 100000,
      TRUE ~ NA_real_
    ),
    rural_ci = case_when(
      !is.na(adj.prev100k.ci.bacteriological.tb.rural) ~ as.character(adj.prev100k.ci.bacteriological.tb.rural),
      !is.na(prev100k.ci.bacteriological.tb.rural) ~ as.character(prev100k.ci.bacteriological.tb.rural),
      TRUE ~ NA_character_
    )
  )
```

```{r}
SDS_filtered <- SDS_filtered %>%
  mutate(
    total_prev = case_when(
      !is.na(adj.prev100k.bacteriological.tb.rurality.total) ~ adj.prev100k.bacteriological.tb.rurality.total,
      !is.na(prev100k.bacteriological.tb.rurality.total) ~ prev100k.bacteriological.tb.rurality.total,
      !is.na(n.bacteriological.tb.rurality.total) & !is.na(n.participants.rurality.total) ~
        (n.bacteriological.tb.rurality.total / n.participants.rurality.total) * 100000,
      TRUE ~ NA_real_
    ),
    total_ci = case_when(
      !is.na(adj.prev100k.ci.bacteriological.tb.rurality.total) ~ as.character(adj.prev100k.ci.bacteriological.tb.rurality.total),
      !is.na(prev100k.ci.bacteriological.tb.rurality.total) ~ as.character(prev100k.ci.bacteriological.tb.rurality.total),
      TRUE ~ NA_character_
    )
  )
```

Seperate confidence intervals into upper and lower intervals
```{r}
library(tidyr)

SDS_filtered <- SDS_filtered %>%
  separate(urban_ci, into = c("urban_lo", "urban_hi"), sep = "-", remove = FALSE, convert = TRUE) %>%
  separate(rural_ci, into = c("rural_lo", "rural_hi"), sep = "-", remove = FALSE, convert = TRUE) %>%
  separate(total_ci, into = c("total_lo", "total_hi"), sep = "-", remove = FALSE, convert = TRUE)
```
Transfer to logodds scale
```{r}
SDS_log <- SDS_filtered %>%
  mutate(across(c(urban_p,urban_lo,urban_hi,rural_p,rural_lo,rural_hi,total_prev,total_lo,total_hi), ~./100000)) %>%
    mutate(id = row_number()) %>%
    mutate(
    logodds_urban = log(urban_p/(1-urban_p)),
    logodds_rural = log(rural_p/(1-rural_p)),
    logodds_total = log(total_prev/(1-total_prev))
  ) %>%
  mutate(
    logodds_urban_lower = log(urban_lo/(1-urban_lo)),
    logodds_urban_upper = log(urban_hi/(1-urban_hi)),
    logodds_rural_lower = log(rural_lo/(1-rural_lo)),
    logodds_rural_upper = log(rural_hi/(1-rural_hi)),
    logodds_total_lower = log(total_lo/(1-total_lo)),
    logodds_total_upper = log(total_hi/(1-total_hi))
  ) %>%
  mutate(logodds_se_urban = (logodds_urban_upper-logodds_urban_lower)/3.92,
         logodds_se_rural = (logodds_rural_upper-logodds_rural_lower)/3.92,
         logodds_se_total = (logodds_total_upper-logodds_total_lower)/3.92
         )%>%
  select(id, WHO.region, logodds_urban, logodds_rural, logodds_se_urban, logodds_se_rural)  %>%
  pivot_longer(
    cols = starts_with("logodds"),
    names_to = c("metric", "type"),
    names_pattern = "logodds(_se)?_(.+)"
  ) %>%
  pivot_wider(
    names_from = metric,
    values_from = value,
    names_prefix = "logodds"
  )
  )
```

```{r}
SDS_log <- SDS_filtered %>%
  mutate(across(c(urban_p:rural_hi, -total_ci, -urban_ci, -rural_ci), ~./100000)) %>%
  mutate(id = row_number()) %>%
  mutate(
    logodds_urban = log(urban_p/(1-urban_p)),
    logodds_rural = log(rural_p/(1-rural_p))
  ) %>%
    mutate(
    logodds_urban_lower = log(urban_lo/(1-urban_lo)),
    logodds_urban_upper = log(urban_hi/(1-urban_hi)),
    logodds_rural_lower = log(rural_lo/(1-rural_lo)),
    logodds_rural_upper = log(rural_hi/(1-rural_hi))
  ) %>%
  mutate(logodds_se_urban = (logodds_urban_upper-logodds_urban_lower)/3.92,
         logodds_se_rural = (logodds_rural_upper-logodds_rural_lower)/3.92
  ) %>%
  select(id, WHO.region, logodds_urban, logodds_rural, logodds_se_urban, logodds_se_rural)  %>%
  pivot_longer(
    cols = starts_with("logodds"),
    names_to = c("metric", "type"),
    names_pattern = "logodds(_se)?_(.+)"
  ) %>%
  pivot_wider(
    names_from = metric,
    values_from = value,
    names_prefix = "logodds"
  )

```

Fitting the model
```{r}
library(brms)
m3 <- brm(
  logodds | se(logodds_se, sigma=TRUE) ~ 1 + type + (1|id),
  data = SDS_log,
  family = "gaussian",
  control = list(adapt_delta = 0.99),
  cores = 4,
  chains = 4,
  iter = 4000
)

summary(m3)
posterior_summary(m3)
plot(m3)
pp_check(m3, ndraws =100)
```
```{r}
m3_draws <- m3 %>% 
  as_draws_df() %>%
  select(b_Intercept, b_typeurban) %>%
  transmute(rural = b_Intercept, urban = b_Intercept+b_typeurban) %>%
  mutate(across(c(urban, rural), ~expit(.))) %>%
  mutate(u_to_rural_ratio = urban/rural) %>%
  mutate(model = "m3")
```

```{r}
library(ggplot2)
library(bayesplot)

m3_drawsxxx <- as_draws_df(m3)
colnames(m3_drawsxxx)

mcmc_areas(m3_drawsxxx, pars = c("b_Intercept", "b_typeurban"))
m3_draws %>%
  pivot_longer(cols = c(urban, rural), names_to = "area", values_to = "prevalence") %>%
  ggplot(aes(x = prevalence, fill = area)) +
  geom_density(alpha = 0.6) +
  labs(title = "Posterior Distribution of Urban and Rural TB Prevalence",
       x = "Estimated TB Prevalence", y = "Density") +
  theme_minimal()

m3_drawsxxx %>%
  summarise(
    OR_median = median(exp(b_typeurban)),
    OR_95lo = quantile(exp(b_typeurban), 0.025),
    OR_95hi = quantile(exp(b_typeurban), 0.975)
  )

ggplot(m3_draws, aes(x = u_to_rural_ratio)) +
  geom_density(fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "darkred") +
  labs(
    title = "Posterior Distribution of Urban-to-Rural TB Prevalence Ratio",
    x = "Urban : Rural Ratio",
    y = "Density"
  ) +
  theme_minimal()

m3_draws %>%
  summarise(
    median_ratio = median(u_to_rural_ratio),
    CI_lower = quantile(u_to_rural_ratio, 0.025),
    CI_upper = quantile(u_to_rural_ratio, 0.975)
  )
```
Model with regional effects
```{r}
m4 <- brm(
  logodds | se(logodds_se, sigma=TRUE) ~ 1 + type*WHO.region + (1|id),
  data = SDS_log,
  family = "gaussian",
  control = list(adapt_delta = 0.99),
  cores = 4,
  chains = 4,
  iter = 4000
)
summary(m4)
plot(m4)
pp_check(m4, ndraws =100)
```

