---
title: "urban-rural-model"
author: "Alireza Mortazavi"
date: "2025-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidybayes)
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}

expit <- function (x) {1/(1 + exp(-x))} #convert back to the response scale

```
## Subset of data
```{r}
SDS <- bactPosRuralityData %>% 
    select(c(1,2,3,7,14,55,130:132,148,149,150,151:153,160:195,902:907))

```
## Filterout the studies that dont have bacterilogical result based on rurality.analysis.indicator
```{r}
SDS_filtered <- SDS %>%
  filter(!rurality.analysis.indicator %in% c(
    "adj.prev100k.ci.smear.positive.tb",
    "n.smear.positive.tb",
    "prev100k.ci.smear.positive.tb"
  ))
```
Hierchy of the model
```{r}
SDS_filtered <- SDS_filtered %>%
  mutate(
    urban_p = case_when(
      !is.na(adj.prev100k.bacteriological.tb.urban) ~ adj.prev100k.bacteriological.tb.urban,
      !is.na(prev100k.bacteriological.tb.urban) ~ prev100k.bacteriological.tb.urban,
      !is.na(n.bacteriological.tb.urban) & !is.na(n.participants.urban) ~
        (n.bacteriological.tb.urban / n.participants.urban) * 100000,
      TRUE ~ NA_real_
    ),
    urban_ci = case_when(
      !is.na(adj.prev100k.ci.bacteriological.tb.urban) ~ as.character(adj.prev100k.ci.bacteriological.tb.urban),
      !is.na(prev100k.ci.bacteriological.tb.urban) ~ as.character(prev100k.ci.bacteriological.tb.urban),
      TRUE ~ NA_character_
    )
  )
```
```{r}
SDS_filtered <- SDS_filtered %>%
  mutate(
    rural_p = case_when(
      !is.na(adj.prev100k.bacteriological.tb.rural) ~ adj.prev100k.bacteriological.tb.rural,
      !is.na(prev100k.bacteriological.tb.rural) ~ prev100k.bacteriological.tb.rural,
      !is.na(n.bacteriological.tb.rural) & !is.na(n.participants.rural) ~
        (n.bacteriological.tb.rural / n.participants.rural) * 100000,
      TRUE ~ NA_real_
    ),
    rural_ci = case_when(
      !is.na(adj.prev100k.ci.bacteriological.tb.rural) ~ as.character(adj.prev100k.ci.bacteriological.tb.rural),
      !is.na(prev100k.ci.bacteriological.tb.rural) ~ as.character(prev100k.ci.bacteriological.tb.rural),
      TRUE ~ NA_character_
    )
  )
```

```{r}
SDS_filtered <- SDS_filtered %>%
  mutate(
    total_prev = case_when(
      !is.na(adj.prev100k.bacteriological.tb.rurality.total) ~ adj.prev100k.bacteriological.tb.rurality.total,
      !is.na(prev100k.bacteriological.tb.rurality.total) ~ prev100k.bacteriological.tb.rurality.total,
      !is.na(n.bacteriological.tb.rurality.total) & !is.na(n.participants.rurality.total) ~
        (n.bacteriological.tb.rurality.total / n.participants.rurality.total) * 100000,
      TRUE ~ NA_real_
    ),
    total_ci = case_when(
      !is.na(adj.prev100k.ci.bacteriological.tb.rurality.total) ~ as.character(adj.prev100k.ci.bacteriological.tb.rurality.total),
      !is.na(prev100k.ci.bacteriological.tb.rurality.total) ~ as.character(prev100k.ci.bacteriological.tb.rurality.total),
      TRUE ~ NA_character_
    )
  )
```

Seperate confidence intervals into upper and lower intervals
```{r}
library(tidyr)

SDS_filtered <- SDS_filtered %>%
  separate(urban_ci, into = c("urban_lo", "urban_hi"), sep = "-", remove = FALSE, convert = TRUE) %>%
  separate(rural_ci, into = c("rural_lo", "rural_hi"), sep = "-", remove = FALSE, convert = TRUE) %>%
  separate(total_ci, into = c("total_lo", "total_hi"), sep = "-", remove = FALSE, convert = TRUE)
```
Transfer to logodds scale

```{r}
SDS_log <- SDS_filtered %>%
  mutate(across(c(urban_p:rural_hi, -total_ci, -urban_ci, -rural_ci), ~./100000)) %>%
  mutate(id = row_number()) %>%
  mutate(
    logodds_urban = log(urban_p/(1-urban_p)),
    logodds_rural = log(rural_p/(1-rural_p))
  ) %>%
    mutate(
    logodds_urban_lower = log(urban_lo/(1-urban_lo)),
    logodds_urban_upper = log(urban_hi/(1-urban_hi)),
    logodds_rural_lower = log(rural_lo/(1-rural_lo)),
    logodds_rural_upper = log(rural_hi/(1-rural_hi))
  ) %>%
  mutate(logodds_se_urban = (logodds_urban_upper-logodds_urban_lower)/3.92,
         logodds_se_rural = (logodds_rural_upper-logodds_rural_lower)/3.92
  ) %>%
  select(id, WHO.region, study.country, title.extracted, logodds_urban, logodds_rural, logodds_se_urban, logodds_se_rural)  %>%
  pivot_longer(
    cols = starts_with("logodds"),
    names_to = c("metric", "type"),
    names_pattern = "logodds(_se)?_(.+)"
  ) %>%
  pivot_wider(
    names_from = metric,
    values_from = value,
    names_prefix = "logodds"
  )

```

Fitting the model
```{r}
library(brms)
m3 <- brm(
  logodds | se(logodds_se, sigma=TRUE) ~ 1 + type + (1|id),
  data = SDS_log,
  family = "gaussian",
  control = list(adapt_delta = 0.99),
  cores = 4,
  chains = 4,
  iter = 4000
)

summary(m3)
posterior_summary(m3)
plot(m3)
pp_check(m3, ndraws =100)
```
```{r}
m3_draws <- m3 %>% 
  as_draws_df() %>%
  select(b_Intercept, b_typeurban) %>%
  transmute(rural = b_Intercept, urban = b_Intercept+b_typeurban) %>%
  mutate(across(c(urban, rural), ~expit(.))) %>%
  mutate(u_to_rural_ratio = urban/rural) %>%
  mutate(model = "m3")
```

```{r}
library(ggplot2)
library(bayesplot)

m3_drawsxxx <- as_draws_df(m3)
colnames(m3_drawsxxx)

mcmc_areas(m3_drawsxxx, pars = c("b_Intercept", "b_typeurban"))
m3_draws %>%
  pivot_longer(cols = c(urban, rural), names_to = "area", values_to = "prevalence") %>%
  ggplot(aes(x = prevalence, fill = area)) +
  geom_density(alpha = 0.6) +
  labs(title = "Posterior Distribution of Urban and Rural TB Prevalence",
       x = "Estimated TB Prevalence", y = "Density") +
  theme_minimal()

m3_drawsxxx %>%
  summarise(
    OR_median = median(exp(b_typeurban)),
    OR_95lo = quantile(exp(b_typeurban), 0.025),
    OR_95hi = quantile(exp(b_typeurban), 0.975)
  )

ggplot(m3_draws, aes(x = u_to_rural_ratio)) +
  geom_density(fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "darkred") +
  labs(
    title = "Posterior Distribution of Urban-to-Rural TB Prevalence Ratio",
    x = "Urban : Rural Ratio",
    y = "Density"
  ) +
  theme_minimal()

m3_draws %>%
  summarise(
    median_ratio = median(u_to_rural_ratio),
    CI_lower = quantile(u_to_rural_ratio, 0.025),
    CI_upper = quantile(u_to_rural_ratio, 0.975)
  )
```
# Prevalence ratio from M3 model
```{r}
# Extract posterior samples
post_m3 <- posterior_samples(m3)

# Get urban/rural log-odds difference
logodds_diff <- post_m3$b_typeurban

# Convert to prevalence ratio (PR)
PR <- exp(logodds_diff)

# Summarise
PR_summary <- tibble(
  Mean = mean(PR),
  Lower_95CrI = quantile(PR, 0.025),
  Upper_95CrI = quantile(PR, 0.975)
)

# Print
print(PR_summary)
```

Model with regional effects
```{r}
m4 <- brm(
  logodds | se(logodds_se, sigma=TRUE) ~ 1 + type*WHO.region + (1|id),
  data = SDS_log,
  family = "gaussian",
  control = list(adapt_delta = 0.99),
  cores = 4,
  chains = 4,
  iter = 4000
)
summary(m4)
pp_check(m4, ndraws =100)
```
```{r}
library(broom.mixed)
library(ggplot2)

tidy_m4 <- broom.mixed::tidy(m4, effects = "fixed", conf.int = TRUE)

ggplot(tidy_m4, aes(x = estimate, y = term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = "Posterior Estimates (Fixed Effects)",
       x = "Estimate (log-odds scale)", y = "") +
  theme_minimal()



draws <- as_draws_df(m4) %>%
  transmute(
    AFR     = b_typeurban,
    EMR       = b_typeurban + `b_typeurban:WHO.regionEasternMediterraneanWHO`,
    SEAR      = b_typeurban + `b_typeurban:WHO.regionSouthMEastAsiaWHO`,
    WPR       = b_typeurban + `b_typeurban:WHO.regionWesternPacificWHO`
  ) %>%
  pivot_longer(everything(), names_to = "region", values_to = "logodds_diff")

ggplot(draws, aes(x = logodds_diff, fill = region)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = "Urban vs Rural TB Log-Odds Difference by WHO Region",
       x = "Urban - Rural (log-odds)", y = "Density") +
  theme_minimal()


### Prevalence ratio plot
draws %>%
  mutate(prevalence_ratio = exp(logodds_diff)) %>%
  ggplot(aes(x = prevalence_ratio, fill = region)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(title = "Urban vs Rural TB Prevalence Ratio by Region",
       x = "Prevalence Ratio (Urban / Rural)", y = "Density") +
  theme_minimal()
```
```{r}
library(posterior)


# Calculate log-odds for each region & setting
m4_draws <- as_draws_df(m4)

region_prevalence_draws <- m4_draws %>%
  transmute(
    Africa_rural  = b_Intercept,
    Africa_urban  = b_Intercept + b_typeurban,

    EMR_rural    = b_Intercept + `b_WHO.regionEasternMediterraneanWHO`,
    EMR_urban    = b_Intercept + b_typeurban +
                    `b_WHO.regionEasternMediterraneanWHO` + `b_typeurban:WHO.regionEasternMediterraneanWHO`,

    SEAR_rural   = b_Intercept + `b_WHO.regionSouthMEastAsiaWHO`,
    SEAR_urban   = b_Intercept + b_typeurban +
                    `b_WHO.regionSouthMEastAsiaWHO` + `b_typeurban:WHO.regionSouthMEastAsiaWHO`,

    WPR_rural    = b_Intercept + `b_WHO.regionWesternPacificWHO`,
    WPR_urban    = b_Intercept + b_typeurban +
                    `b_WHO.regionWesternPacificWHO` + `b_typeurban:WHO.regionWesternPacificWHO`
  ) %>%
  pivot_longer(cols = everything(), names_to = "group", values_to = "logodds") %>%
  separate(group, into = c("region", "type"), sep = "_") %>%
  mutate(prevalence_per_100k = plogis(logodds) * 100000)


prevalence_summary <- region_prevalence_draws %>%
  group_by(region, type) %>%
  summarise(
    median = median(prevalence_per_100k),
    lower = quantile(prevalence_per_100k, 0.025),
    upper = quantile(prevalence_per_100k, 0.975),
    .groups = "drop"
  )

ggplot(prevalence_summary, aes(x = median, y = region, color = type)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbarh(aes(xmin = lower, xmax = upper),
                 position = position_dodge(width = 0.6), height = 0.2) +
  scale_x_continuous(labels = scales::comma_format()) +
  labs(title = "Estimated TB Prevalence per 100,000 by WHO Region and Setting",
       x = "TB Prevalence per 100,000 population (posterior median with 95% CrI)",
       y = "WHO Region", color = "Setting") +
  theme_minimal()

```
# Calculating prevalence ratios base on M4 model
```{r}
post <- posterior_samples(m4)
# Define baseline (AFR): just 'typeurban'
logodds_AFR <- post$b_typeurban
PR_AFR <- exp(logodds_AFR)

# For other regions: add interaction terms
logodds_EMR <- post$b_typeurban + post$`b_typeurban:WHO.regionEasternMediterraneanWHO`
PR_EMR <- exp(logodds_EMR)

logodds_SEAR <- post$b_typeurban + post$`b_typeurban:WHO.regionSouthMEastAsiaWHO`
PR_SEAR <- exp(logodds_SEAR)

logodds_WPR <- post$b_typeurban + post$`b_typeurban:WHO.regionWesternPacificWHO`
PR_WPR <- exp(logodds_WPR)

# Function to summarize posterior distributions
summarise_posterior <- function(x) {
  tibble(
    Mean = mean(x),
    Lower_95CrI = quantile(x, 0.025),
    Upper_95CrI = quantile(x, 0.975)
  )
}

# Create summary table
PR_summary <- bind_rows(
  summarise_posterior(PR_AFR) %>% mutate(Region = "AFR"),
  summarise_posterior(PR_EMR) %>% mutate(Region = "EMR"),
  summarise_posterior(PR_SEAR) %>% mutate(Region = "SEAR"),
  summarise_posterior(PR_WPR) %>% mutate(Region = "WPR")
) %>% 
  select(Region, everything())

# View the result
print(PR_summary)

```

### model with the urban/rural population ratio
```{r}
## calculate urban/rural population ratio
SDS_filtered <- SDS_filtered %>%
  mutate(
    urban_pop = case_when(
      !is.na(n.participants.urban) ~ n.participants.urban,
      TRUE ~ NA_real_
    ),
    rural_pop = case_when(
      !is.na(n.participants.rural) ~ n.participants.rural,
      TRUE ~ NA_real_
    ),
    total_pop = case_when(
      !is.na(n.participants.rurality.total) ~ n.participants.rurality.total,
      TRUE ~ NA_real_
    ),
    urban_to_rural_ratio = urban_pop / rural_pop
  )

## fit the model (to be completed)
```
New models based on our latest group discussion
# (1 + Sex |WHO.region/study.country) + (1 + Sex|id/title.extracted)
```{r}
    library(brms)
    m5 <- brm(
      formula = logodds | se(logodds_se, sigma=TRUE) ~ 1 + type + 
        (1 + type | WHO.region/study.country) + 
        (1 + type | id/title.extracted),
      
      data = SDS_log,
      family = gaussian(),
      
      control = list(adapt_delta = 0.99),
      cores = 4,
      chains = 4,
      iter = 4000,
      
      prior = prior(normal(0, 10), class = Intercept) +
              prior(normal(0, 10), class = b) +
              prior(exponential(1), class = sd) +
              prior(lkj(4), class = cor)
    )

summary(m5)
conditional_effects(m5)
pp_check(m5, ndraws = 100)

# Prepare new data
nd <- SDS_log %>%
  select(WHO.region, study.country, type) %>%
  distinct() %>%
  mutate(logodds_se = 1)


m5_draws_country <- m5 %>%
  add_epred_draws(newdata = nd, re_formula = ~(1 + type | WHO.region/study.country)) %>%
  ungroup() %>%
  mutate(value = expit(.epred) * 1e5) %>%
  select(study.country, WHO.region, type, value, .draw)

m5_draws_country <- m5_draws_country %>%
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(urRatio = urban / rural) %>%
  pivot_longer(cols = c(urban, rural, urRatio)) %>%
  rename(type = name)

m5_draws_region <- m5 %>%
  add_epred_draws(newdata = nd, re_formula = ~(1 + type | WHO.region)) %>%
  ungroup() %>%
  mutate(value = expit(.epred) * 1e5) %>%
  select(study.country, WHO.region, type, value, .draw)

m5_draws_region <- m5_draws_region %>%
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(urRatio = urban / rural) %>%
  pivot_longer(cols = c(urban, rural, urRatio)) %>%
  rename(type = name)

# Overall
m5_draws_country %>%
  group_by(type) %>%
  mean_qi(value)

# By region
m5_draws_region %>%
  group_by(WHO.region, type) %>%
  mean_qi(value)

# By country
m5_draws_country %>%
  group_by(study.country, type) %>%
  mean_qi(value)

#EMPERICAL VALUES
m5_empirical <- SDS_filtered %>%
  select(covidence.id, WHO.region, study.country, urban = urban_p, rural = rural_p) %>%
  mutate(
    urRatio = urban / rural
  ) %>%
  pivot_longer(cols = c(urban, rural, urRatio)) %>%
  rename(type = name, value = value)

# Country-level plot
a <- m5_draws_country %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = study.country), .width = 0.95) +
  geom_point(data = m5_empirical, aes(y = study.country, x = value),
             colour = "mediumseagreen", shape = 1) +
  facet_grid(. ~ fct_relevel(type, "urban", "rural"), scale = "free_x") +
  theme_tidybayes() +
  theme(panel.background = element_rect(fill = NA, colour = "grey78"))

# Region-level plot
b <- m5_draws_region %>%
  ggplot() +
  stat_pointinterval(aes(x = value, y = WHO.region), .width = 0.95) +
  geom_point(data = m5_empirical, aes(y = WHO.region, x = value),
             colour = "mediumseagreen", shape = 1) +
  facet_grid(. ~ fct_relevel(type, "urban", "rural"), scale = "free_x") +
  theme_tidybayes() +
  theme(panel.background = element_rect(fill = NA, colour = "grey78"))

# Show plots
a
b
```

```{r}
nd <- expand.grid(
  WHO.region = unique(SDS_log$WHO.region),
  type = c("urban", "rural")
) %>%
  mutate(logodds_se = 1)

m5_preds <- m5 %>%
  add_epred_draws(newdata = nd, re_formula = ~(1 + type | WHO.region)) %>%
  ungroup() %>%
  mutate(prevalence_per100k = expit(.epred) * 1e5)

urban_rural_summary <- m5_preds %>%
  group_by(WHO.region, type) %>%
  mean_qi(prevalence_per100k)

print(urban_rural_summary)


# Pivot to wide to calculate ratios
pr_ratio_draws <- m5_preds %>%
  select(.draw, WHO.region, type, prevalence_per100k) %>%
  pivot_wider(names_from = type, values_from = prevalence_per100k) %>%
  mutate(PR_urban_rural = urban / rural)

# Summarise ratios
pr_summary <- pr_ratio_draws %>%
  group_by(WHO.region) %>%
  mean_qi(PR_urban_rural)

print(pr_summary)

pr_summary_tbl <- pr_summary %>%
  ungroup() %>%
  select(
    Region = WHO.region,
    Mean = PR_urban_rural,
    Lower_95CrI = .lower,
    Upper_95CrI = .upper
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

# Print as tibble
print(pr_summary_tbl)
```